### 2023-03-06 README.Rmd
init repo

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# simulating-meta-analysis

<!-- badges: start -->
<!-- badges: end -->

### 2023-03-07 README.Rmd
updating readme

@@
   collapse = TRUE,
-  comment = "#>"
+  comment = "#>",
+  fig.path = "man/figures/README-",
+  out.width = "100%"
 )
@@
 
-# simulating-meta-analysis
+# Understanding meta-analysis through data simulation, with applications to power analysis
 
 <!-- badges: start -->
+[<img alt="alt_text" src="https://img.shields.io/badge/OSF-https%3A%2F%2Fosf.io%2F54djn%2F-blue" />](https://osf.io/54djn/)
 <!-- badges: end -->
+
+This repository contains the code to reproduce the analysis of the paper **Understanding meta-analysis through data simulation, with applications to power analysis** by Gambarota Filippo and Gianmarco Altoè.
+The project can also be found on the Open Science Framework https://osf.io/54djn/.
+
+## Repository
+
+The project is organized as follows:
+
+- `R`: contains the simulation functions used in the paper and supplementary materials
+- `docs`: contains the `.Rmd` files to reproduce the paper and supplementary materials manuscripts
+- `scripts`: contains the script used to produce complex objects and figures
+  - `scripts\r-img.R`: create the paper figures.
+Figures that are not produced in this script using `ggplot2` are created with the open-source program [*Inkscape*](https://inkscape.org/) and contained in the `../img/` folders.
+- `objects`: contains the result of objects created by the `scripts/` scripts.
+
+The project is organized and structured as an R package, thus after opening the `simulating-meta-analysis.Rproj` file and installing the `devtools` package, calling `devtools::load_all()` will load each function within the `R/` folder.

### 2023-03-28 README.Rmd
adding files

--- 2023-03-06 README.Rmd
@@
   collapse = TRUE,
-  comment = "#>"
+  comment = "#>",
+  fig.path = "man/figures/README-",
+  out.width = "100%"
 )
+
+devtools::load_all()
 ```
 
-# simulating-meta-analysis
+# Understanding meta-analysis through data simulation with applications to power analysis
 
 <!-- badges: start -->
+[<img alt="alt_text" src="https://img.shields.io/badge/OSF-https%3A%2F%2Fosf.io%2F54djn%2F-blue" />](https://osf.io/54djn/)
 <!-- badges: end -->
+
+This repository contains the code to reproduce the analysis of the paper **Understanding meta-analysis through data simulation with applications to power analysis** by Gambarota Filippo and Gianmarco Altoè.
+The project can also be found on the Open Science Framework https://osf.io/54djn/ with a preprint uploaded on PsyArXiv [https://psyarxiv.com/br6vy/](https://psyarxiv.com/br6vy/).
+
+## Usage
+
+Functions used in the paper and supplementary materials are documented and can be used opening the `simulating-meta-analysis.Rproj` file and using the `devtools::load_all()` function (this requires the `devtools` package installed).
+Then each function will be available as using a standard R package.
+
+## Documents
+
+The repository (`documents/` folder) contains the paper and supplementary materials created using R Markdown and the [`papaja`](https://cran.r-project.org/web/packages/papaja/index.html) package.
+
+- Paper [PDF](documents/paper/paper.pdf)
+- Supplementary Materials [PDF](documents/supplementary/supplementary.pdf)
+
+## Repository
+
+The project is organized as follows:
+
+- `R`: contains the simulation and utility functions used in the paper and supplementary materials
+- `documents`: contains the `.Rmd` files to reproduce the paper and supplementary materials manuscripts
+- `scripts`: contains the script used to produce complex objects and figures
+  - `scripts\r-img.R`: create the paper figures.
+Figures that are not created in this script using `ggplot2` are created with the open-source program [*Inkscape*](https://inkscape.org/) and contained in the `../img/` folders.
+- `objects`: contains the result of objects created by the `scripts/` scripts.
+
+# Session
+
+## Session info
+
+```{r, echo=FALSE}
+sinfo <- sessioninfo::session_info()$platform
+sinfo <- sinfo[names(sinfo) != "date"]
+paste0(
+  cli::style_bold(cli::col_blue(names(sinfo))), ": ", sinfo
+) |> 
+  cat(sep = "\n")
+```
+
+## Packages
+
+```{r, echo = FALSE, message=FALSE, warning=FALSE, results='asis'}
+pkgs <- get_packages()
+sprintf("- `%s` (%s)", pkgs$pkg, pkgs$version) |> 
+  cat(sep = "\n")
+```

### 2023-03-28 documents/paper/paper.Rmd
adding files

---
title             : "Understanding meta-analysis through data simulation with applications to power analysis"
shorttitle        : "Simulating meta-analysis"
author: 
  - name          : "Filippo Gambarota \\orcidlink{0000-0002-6666-1747}"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Via Venezia 8, 35131 Padova (PD)"
    email         : "filippo.gambarota@unipd.it"
    role:         
      - "Conceptualization"
      - "Methodology"
      - "Formal Analysis"
      - "Software"
      - "Writing – Original Draft"
  - name          : "Gianmarco Altoè \\orcidlink{0000-0003-1154-9528}"
    affiliation   : "1"
    role:
      - "Conceptualization"
      - "Methodology"
      - "Writing - Review & Editing"
      - "Supervision"
affiliation:
  - id            : "1"
    institution   : "Department of Developmental and Social Psychology, University of Padova, Italy"
abstract: |
  Meta-analysis is a powerful tool to combine evidence from existing literature. Despite several introductory and advanced materials about organizing, conducting, and reporting a meta-analysis, to our knowledge, there are no introductive materials about simulating the most common meta-analysis models. Data simulation is essential for developing and validating new statistical models and procedures. Furthermore, data simulation is a powerful educational tool for understanding a statistical method. In this tutorial, we show how to simulate fixed-effects, random-effects, and meta-regression models and illustrate how to estimate statistical power. In the supplementary materials, we extended these simulations also for multilevel and multivariate models. All materials associated with this article can be accessed on Open Science Framework ([https://osf.io/54djn/](https://osf.io/54djn/)).

keywords          : "meta-analysis, monte carlo simulations, power analysis"
wordcount         : "7877"
bibliography      : "../files/references.bib"
floatsintext      : yes
figsintext        : yes
linenumbers       : "`r params$linenumbers`"
draft             : no
mask              : no
#toc               : yes
figurelist        : no
tablelist         : no
footnotelist      : no
classoption       : "man"
numbersections    : true
nocite: | 
  @r-lang
header-includes:
  - \usepackage{setspace}
  - \captionsetup[figure]{font={stretch=1,footnotesize}}
  - \usepackage{float}
  - \usepackage{orcidlink}

#csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa6"
output            :
  papaja::apa6_pdf:
    latex_engine: xelatex
params: 
  linenumbers: "yes"
  draft: "no"
  finalsection: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      out.width = "80%",
                      fig.width = 9,
                      size = "scriptsize",
                      fig.asp = 0.618,
                      warning = FALSE,
                      message = FALSE,
                      fig.pos = 'H',
                      dev = "tikz")

# chunk size
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r packages, include=FALSE}
library(papaja)
library(metafor)
library(dplyr)
library(tidyr)
library(kableExtra)
library(here)
library(latex2exp)
library(ltxplot)

rcit <- sprintf("(version %s.%s; R Core Team, %s)", R.version$major, R.version$minor, R.version$year)
ltxplot::load_theme_ltx()
devtools::load_all()
seed <- 2023 # general seed for all simulations
```

```{r functions, include = FALSE}
# theme for plots
theme_paper <- function(base_size = 15){
  ltxplot::theme_latex(base_size = base_size)
}

# quick table
qtab <- function(data, caption, ...){
  tab <- data |> 
    apa_table(caption = caption,
              placement = "H",
              font_size = "small",
              ...)
  return(tab)
}
```

```{r data, include=FALSE}
# creating figures
source(here("scripts", "r-img.R"), local = knitr::knit_global())
r_imgs <- readRDS(here("objects", "r-imgs.rds"))
simfuns <- get_funs(here("R", "sim-utils.R")) # getting all funs
```

# Introduction

> "If you do not simulate it, you have not understood it."

A meta-analysis is an essential tool for combining knowledge from multiple studies quantitatively. Meta-analysis is commonly used together with a systematic review of the literature. The meta-analysis has several advantages. Firstly, it allows combining evidence from multiple studies assigning more weight to studies with lower estimation variability. Then using meta-regression, it is possible to assess the impact of study-level variables that could impact the phenomenon. Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias. Despite the advantages, meta-analysis implementation is not always straightforward, especially for complex data structures. Additionally, the fact that there are several introductory and advanced resources to understand meta-analysis [@Borenstein2009-mo; @Schmid2022-xe; @Harrer2021-bz], to our knowledge, there are no introductory resources about how to simulate realistic meta-analytic data. Simulating data has several advantages because it requires understanding the statistical method and the data-generation process. Furthermore, data simulation is the primary tool when it comes to evaluating a new analysis method, estimating the statistical power, or understanding the long-run behavior of our data generation process [@Ingalls2011-xt; @Gelman2020-tg, pp. 69-76; @Gelman2006-pc, pp. 155-176]. A recent paper by DeBruine and Barr [-@DeBruine2021-id], which deeply inspired the current work, proposed a stimulating way to understand linear mixed effects via data simulation. Simulating data is also a powerful educational tool within this framework. For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for fixed and random-effects, and meta-regression models with applications also for statistical power calculation. In the first section, we will introduce basic concepts of the meta-analysis that are useful for setting up the simulation. We will evaluate the effect size, variance calculation, and the fixed vs. random effects model distinction. Then we will describe how to simulate data for these models and simulate a meta-regression model with a categorical and numerical predictor. Finally, we will introduce the power analysis extending the previous examples to estimate the statistical power. We used the R statistical programming language `r rcit`. We assume the reader is familiar with the basic concepts of R, but core functions will be explained. Code and materials are available on the OSF repository ([https://osf.io/54djn/](https://osf.io/54djn/)). Multivariate and multilevel models’ simulation examples and more details about the coding approach are available in the supplementary materials.

## Meta-analysis introduction

The meta-analysis is a statistical procedure to combine multiple studies (i.e., *primary studies*) into a single statistical analysis [@Borenstein2009-mo]. The idea is that combining numerous preliminary studies improve the estimation of a particular phenomenon more efficiently compared to conducting a single study. In statistical terms, the concept of the meta-analysis is to switch the statistical unit from the single participant or observation (i.e., *level 1*) to the study (i.e., *level 2*). Given that some studies will give more information because their estimation variability is smaller (e.g., higher sample size), the meta-analysis combines the studies assigning more weight as a function of the precision (i.e., the inverse of the variance).

As an example that will be used throughout the paper, we consider the efficacy of memory training in improving memory performance during a cognitive task. The typical primary study will collect data from a group of participants receiving the memory training (*experimental group*) and another group receiving a control treatment (*control group*). The focus of the meta-analysis is collecting multiple studies with similar aims and methods and estimating the average effect of memory training. Despite differences in the type of cognitive task or experimental setup, each primary study collects an *experimental group* ($n_{C}$) and *control group* ($n_{C}$) and computes the average performance (*experimental group* ${\bar y}_{T}$ and *control group* ${\bar y}_{C}$) and standard deviations (*experimental group* $s_{T}$ and *control group*
$s_{C}$).

### Effect size and variance

The first step of a meta-analysis is to standardize information from each included study to convert each effect into a standard measure. This common measure should give an immediate idea of the direction (i.e., the treatment improves or reduces performance) and the size of the effect. Standardized effect size measures [see @Lakens2013-eq for an overview], such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's $d$ [@Cohen1988-ad] or the Pearson correlation coefficient $\rho$ which could be calculated using the associated sample estimator *r* are commonly used to compare heterogeneous outcome variables. If all studies used the same raw measure, such as reaction times, it is possible to directly meta-analyze the studies without standardizing. In practice, studies often use heterogeneous measures, therefore requiring standardized measures.

As reported in the previous section, beyond the effect size of each study, we need to assign a weight according to the precision. For this reason, we need to calculate the sampling variability of the effect size or the raw measure that will represent the estimation precision. Each raw or standardized effect size measure (e.g., raw mean difference, Cohen's $d$, and Pearson's correlation) has a different formula to calculate the sampling variability. The idea is to choose the appropriate measure considering the study design (e.g., between vs. within-subjects) and available information and find the proper formula to compute the sampling variability. In addition, there are formulas and approaches to convert from one effect size measure to another [@Borenstein2009-mo; @Lipsey2001-vp; @Lakens2013-eq].^[See also the blog post by James Pustejovsky ([link](https://www.jepusto.com/alternative-formulas-for-the-smd/)) for a general method to compute the sampling variability for standardized mean difference measures.] Usually, the effect size sampling variability depends mainly on the sample size that determines the weight assigned during the meta-analytic estimation.

### Fixed vs random-effects model

The core of a meta-analysis is combining the results of multiple studies giving more weight to studies that provide a more precise effect estimation. We can divide meta-analysis models into two categories: the *fixed-effects* and the *random-effects model*. If we imagine combining multiple studies replicating the same experiment, we assume the real treatment effect $\theta_{f}$ is *fixed,* and each study is a more or less precise estimation of this parameter. This is an essential and often inappropriate assumption because it ignores that study-level characteristics may impact the actual treatment effect.

Suppose some study-level characteristics (e.g., participants' age, sex, or socioeconomic status) or the experimental paradigm (e.g., type of memory task or difficulty) could impact the treatment effect. In that case, the assumption of the *fixed-effects model* is no longer plausible. There is no unique real effect but a distribution of effects with mean $\theta_{r}$ and variance $\tau^{2}$. The mean is the *average* effect, and the variance is the heterogeneity of the phenomenon. The *random-effects model* will estimate the mean ($\theta_{r}$) and the heterogeneity $\tau^{2}$ will help generalize to the population of effects beyond the pool of included studies.

Most importantly, the parameters estimated with the *fixed* and *random-effects* do not represent the same quantity. The *random-effects model* estimates the *average* effect across a distribution of true effects, while the *fixed-effects model* estimates a single true effect. Figure 1 depicts the idea of the fixed vs. random-effects model. The *fixed-effects model* can be considered a particular type of *random-effects model* where $\tau^{2} = 0$ thus, the average effect is also the unique true effect.

```{r img-fixed-vs-random, echo=FALSE, fig.asp=1, fig.cap=get_caption()}
r_imgs$fixed_vs_random
```

### Multilevel and multivariate models

The most straightforward situation for a meta-analysis is when each included study contributes with a single effect size. However, it is common to have multiple effect sizes belonging to the same study creating a situation of dependency between statistical units. There are numerous sources of dependency for a meta-analysis [@Cheung2014-fg; @Cheung2015-bw, pp. 121-122]. The first type of dependency (*multilevel*) consists of multiple effect sizes from independent participants. Despite being collected with different participants, effect sizes from the same study could be correlated. We obtain a multilevel data structure with independent effect sizes in the same study. There are multiple ways to model this data structure [see @Cheung2014-fg for an overview]. Still, the most common way is to use a *multilevel meta-analysis model,* considering the heterogeneity between and within studies. The standard *random-effects* (but also the *fixed-effects model*) model is also called a *two-level* model because it combines data after aggregating the participants-level (*first-level*) data. The *three-level* model is commonly used to manage the multilevel data structure [@Cheung2014-fg; @Cheung2019-po]. Compared to the *two-level* model, the *three-level* model estimates two heterogeneity parameters ($\tau^{2}$ and $\omega^{2}$), one for the heterogeneity between studies and the other for the heterogeneity within studies.

A second type of dependency (*multivariate*) consists of multiple effect sizes collected from the same pool of participants. For example, when using two cognitive tasks on the same pool of participants. This is commonly known as *multivariate meta-analysis* because the effect sizes sampling errors are correlated given that participants are assessed multiple times. A *multivariate meta-analysis model* could be used to
consider the correlation between sampling errors.

A third type of dependency arises when true effect sizes are correlated at the population level [@Cheung2015-bw, pp. 121-122]. For example, two different memory tasks could be correlated because the latent psychological constructs are intrinsically correlated. This correlation is present even when other groups of participants perform the two memory tasks (i.e., no *multivariate* dependency) or are administered by different researchers (i.e., no *multilevel* dependency). This type of dependency could be managed by estimating the correlation between outcomes using a *multivariate random-effects model* [@Cheung2015-bw, pp. 127-130]. Figures \@ref(fig:img-multilevel) and \@ref(fig:img-multivariate) depict the conceptual idea of *multilevel* and *multivariate* models.

```{r img-multilevel, echo=FALSE, out.width="80%", fig.cap = get_caption()}
knitr::include_graphics("img/multilevel.pdf")
```

```{r img-multivariate, echo=FALSE, out.width="80%", fig.cap = get_caption()}
knitr::include_graphics("img/multivariate.pdf ")
```

# Simulation

## Monte Carlo simulations

The Monte Carlo methods are controlled experiments [@Gentle2009-cj]. Given a set of fixed parameters, probability distributions, and the possibility of generating random numbers, it is possible to simulate the behavior of an empirical system. Monte Carlo simulations are used for statistical and mathematical problems that cannot be solved analytically.

A straightforward example regards estimating the sampling variability of the mean difference. When calculating the mean difference between two samples, we estimate the true mean difference at the population level with a certain degree of error (i.e., the standard error of the mean difference). The central limit theorem states that the difference between the means of two random samples ($\overline{X}$ and $\overline{Y}$) is approximately normally distributed with mean $\mu_{x} - \mu_{y}$ and standard error $\sqrt{\frac{\sigma^2_x}{{n_x}} + \frac{\sigma^2_y}{{n_y}}}$. The same results can be obtained using Monte Carlo simulations using the following procedure:

1. Generating two random samples from two normal distributions with a fixed mean difference
2. Calculating the mean difference
3. Repeating the same process, many times
4. Calculate the standard deviation of the simulated values

Using the method, we are estimating via simulation the standard deviation of the sampling distribution of the mean difference (i.e., the standard error of the mean difference). Increasing the number of simulations will produce more stable results.

```{r monte-carlo-simulation}
set.seed(123) # seed for simulation to ensure the reproducibility of results
theta <- 0.5 # real mean difference in the populations
sigma <- 1 # real standard deviations in the population
n <- 30 # sample size for both groups
nsims <- 1e5 # number of simulations = 100000

# simulates the sampling distribution of the mean difference
d_mc <- replicate(nsims, expr = {
  g1 <- rnorm(n = n, mean = theta, sd = sigma)
  g2 <- rnorm(n = n, mean = 0, sd = sigma)
  mean(g1) - mean(g2) # calculates the sample mean difference and returns the value
})

# analytical standard error of the mean difference
se_an <- sqrt(sigma^2/n + sigma^2/n)

# monte-carlo standard error of the mean difference
se_mc <- sd(d_mc)
```

The standard error estimated solving analytically is `r round(se_an, 4)` and using the Monte Carlo simulation we arrive at the same result (i.e., `r round(se_mc, 4)`).

## Simulation setup

We load relevant packages and set the `seed` for reproducibility to prepare the R environment for the simulations.

```{r sim-setup, eval = FALSE}
library(dplyr) # for data manipulation
library(tidyr) # for data manipulation
library(metafor) # for meta-analysis models fitting
seed <- 2023 # general seed for all simulations
```

Before diving into the specific simulations, in this section, we define the common aspects of all simulations in the following sections. The current paper focuses on the *two-level*, *fixed,* and *random-effects model*s. All the examples refer to primary studies that assess the efficacy of a treatment by comparing a control and an experimental group. Equation \@ref(eq:datagen) formalizes the assumption about generating participants' data within a single study. The two groups (*experimental* $y_{T}$ and *control* $y_{C}$) are generated from normal distributions with $\mu = 0$ for the control group and $\mu = \theta$ (where $\theta$ is the effect size) for the experimental group. Both groups are assumed to be sampled from populations with homogeneous variances $\sigma^{2} = 1$.

\begin{align}
\begin{gathered}
y_{Ti} \sim \mathcal{N}(\theta, 1) \\
y_{Ci} \sim \mathcal{N}(0, 1)
(\#eq:datagen)
\end{gathered}
\end{align}

The true Standardized Mean Difference can be estimated using Cohen's $d$ as reported in Equation \@ref(eq:effsize). The difference between the two means is divided by the pooled standard deviation of the two groups. It is common to apply a small-sample correction version of this effect size called Hedges' $g$ [@Hedges1981-za]

\begin{align}
\begin{gathered}
d = \frac{\bar y_T - \bar y_C}{S_p} \\
S_p = \sqrt{\frac{S^2_{y_T}(n_{y_T} - 1) + S^2_{y_C}(n_{C} - 1)}{n_{T} + n_{C} - 2}} \\
\sigma^2_d = \frac{n_{T} + n_{C}}{n_{T}n_{C}} + \frac{d^2}{2(n_{T} + n_{C} - 2)}
(\#eq:effsize)
\end{gathered}
\end{align}

We can use the following algorithm implemented in the `sim_study()` function to simulate a single study. Before using the `sim_study()`
function, we can create a data frame for the simulation using the `make_data()` function^[This function is a simple wrapper of `data.frame()` that given the number of studies and other variables create the data structure for the simulation.]. Table \@ref(tab:sim-data-example-tab) depicts an example of the `make_data()` output.

1. Choose a $\theta$, $n_{T}$, and $n_{C}$ value.
2. Simulate $n_{T}$ observations from a Gaussian distribution with $\mu = \theta$ and $\sigma^{2} = 1$ and $n_{T}$ observations from a Gaussian distribution with $\mu = 0$ and $\sigma^{2} = 1$. This way, the expected difference between groups will be $d$ and the variance is fixed at 1.
3. Calculate the pooled standard deviation $S_{p}$ that is an estimation of the population standard deviation ($\sigma$) to standardize the mean difference.
4. Calculate the observed effect size $d$ and the sampling variance $\sigma_{d}^{2}$.

```{r fun-sim-study, results='asis', echo=FALSE}
print_fun(simfuns$make_data)
```

```{r fun-make-data, results='asis', echo=FALSE}
print_fun(c(simfuns$sim_study, simfuns$sim_studies))
```

The `theta` is the true effect size, `nc` and `nt` are the sample size for the control and experimental group, `aggregate` controls if returning the effect size and the corresponding variance or the participant-level data, and `hedges` specify whether applying the correction. We can generate a single study with the desired parameters with this function. A suggestion for each simulation step is to generate a large $n$ to reduce the sampling error and check the recovery of simulated parameters. This is a general strategy that can be applied to every simulation. The `sim_studies()` function will iterate through variables in the `...` argument creating the meta-analysis data frame. The `mapply()` function is clearly explained in the supplementary materials.

The following code simulates a single study with $n = 10000$ and check the estimated mean and standard deviation.

```{r sim-single-study, eval = FALSE}
set.seed(seed)
sim_study(theta = 0.3, nc = 10000, nt = 10000, aggregate = FALSE)
```

```{r sim-single-study-tab, echo = FALSE}
set.seed(seed)
sim <- sim_study(theta = 0.3, nc = 10000, nt = 10000, aggregate = FALSE) |> 
  group_by(group) |> 
  summarise(mean = mean(y),
            sd = sd(y))
```

The control group has a mean of `r sprintf("%.3f (SD = %.3f)", sim$mean[1], sim$sd[1])` and the experimental group has a mean of `r sprintf("%.3f (SD = %.3f)", sim$mean[2], sim$sd[2])` which are remarkably close to the simulated values.

Using the `sim_study()` function multiple times (with the appropriate adjustments) we can generate a series of studies simulating a data set for a meta-analysis (using the `sim_studies()` function). After each example, we will compute the appropriate model (e.g., *fixed* or *random-effects*) using the `metafor` package [@Viechtbauer2010-xz] to check the recovery of simulated parameters. Table \@ref(tab:notation-tab) summarizes the simulation parameters notation in equations and code.

```{r sim-data-example}
k <- 30 # number of studies
n <- 20 # number of participants per group, per study
theta <- 0.3 # real effect size
sim <- make_data(k = k, nt = n, nc = n, theta = theta)
```

```{r sim-data-example-tab, echo = FALSE}
sim |> 
  trim_df(2) |> 
  qtab(caption = get_caption())
```

```{r notation-tab, echo=FALSE, result = "asis"}
notation_table <- tibble::tribble(
     ~`Equation`,     ~`Code`,                                               ~`Description`,
       "$\\theta_f$",   "theta\\_f",                  "Fixed-effects model real effect size",
       "$\\theta_r$",   "theta\\_r",                 "Random-effects model real effect size",
         "$\\tau^2$",    "tau2",                            "Effect size heterogeneity",
       "$\\tau^2_r$",   "tau2r",                   "Residual effect size heterogeneity",
  "$\\bar y_{T, C}$",        "",               "Mean of the experimental/control group",
      "$s^2_{T, C}$",        "", "Standard deviation of the experimental/control group",
        "$n_{T, C}$",  "nt, nc",        "Sample size of the experimental/control group",
             "$d_i$",      "yi",                                 "Observed effect size",
      "$\\sigma^2_i$",      "vi",                          "Observed sampling variance",
       "$\\theta_i$", "theta\\_i",              "Random effect for the study i",
        "$\\beta_0$",      "b0",                            "Meta-regression intercept",
         "$\\beta_1$",      "b1",                                "Meta-regression slope",
               "$k$",       "k",                                    "Number of studies",
     "$\\epsilon_i$",        "",             "Sampling error for the study i"
  )

notation_table$`Code` <- sprintf("\\texttt{%s}", notation_table$`Code`)
notation_table[is.na(notation_table)] <- ""

notation_table |> 
  apa_table(align = c("l", "l", "l"),
            caption = "Variables in the data-generating model and associated R code.",
            escape = FALSE,
            placement = "H",
            note = "The \\texttt{yi} and \\texttt{vi} notation for the observed effect size ($d_i$) and sampling variance ($\\sigma^2_i$) has been used to be consistent with the \\texttt{metafor} notation.")
```

## Fixed-effects model

The most basic model to simulate is the *fixed-effects model* (FE). As reported in the previous sections, the *fixed-effects model* assumes the presence of a single true effect ($\theta_{f}$) and the observed variability is because each effect size is an imprecise estimation of the true effect. In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies
(i.e., the sample size). Equation \@ref(eq:fixed-effect-model) depicts the *fixed-effect* equation.

\begin{align}
\begin{gathered}
d_i = \theta_f + \epsilon_i \\
\epsilon_i \sim \mathcal{N}(0,\sigma^2_i)
(\#eq:fixed-effect-model)
\end{gathered}
\end{align}

Each observed effect size ($d_{i}$) is composed of the real effect size $\theta_{f}$ plus an error term ($\epsilon_{i}$) that is sampled from a normal distribution with $\mu = 0$ and $\sigma^{2} = \sigma_{i}^{2}$ (i.e., the known sampling variance of the study $i$). As demonstrated in
Equation \@ref(eq:effsize), the increase in sample size will decrease the sampling variability. An exact study will essentially have $\theta_{f}$ as the observed effect size. Since we are sampling participants' level data, the error component is already included in the `sim_study()` function. To simulate this model in R we can just call the `sim_study()` multiple times according to the number of desired studies ($k$). In addition, we simulate that each primary study will have a sample size of $n = 20$ for both groups. We will discuss later the appropriateness of this assumption.

```{r sim-fixed-effect}
set.seed(seed)
k <- 30 # number of studies
n <- 20 # number of participants per group, per study
theta_f <- 0.3 # real effect size

sim <- make_data(k = k, nt = n, nc = n, theta_f = theta_f)
sim <- sim_studies(sim$theta_f, sim$nc, sim$nt, data = sim) # simulated data
res <- rma(yi = yi, vi = vi, method = "FE", data = sim)
```

We are using the $\theta_{f}$ parameter to generate random data using the `sim_study()` function that will introduce the random error component $\epsilon_{i}$. Then we can fit the *fixed-effects* meta-analysis model with the `rma` function and `method = "FE"` of the `metafor` package. The parameter we are estimating is $\theta_{f}$ which is close to our simulation value (see Table \@ref(tab:res-fixed-effect)). Increasing the number of studies ($k$) and/or the number of participants in each study ($n$) will improve estimation reducing the standard error.

```{r res-fixed-effect, echo = FALSE}
res_tidy <- rma_tidy(res)
rma_table(res_tidy$res, res_tidy$notes, caption = get_caption())
```

Increasing the sample size for each study will increase the estimation precision thus the variability among studies will be reduced. This can be easily demonstrated by simulating studies with a high sample size as reported in Figure \@ref(fig:fixed-effect-high-low-precision). As the sample size increases the only source of variability (i.e., the error component $\epsilon_{i}$) is close to zero so each study is closer to the true simulated value.

```{r fixed-effect-high-low-precision, echo = FALSE, fig.asp=0.8, fig.cap=get_caption()}
set.seed(seed)
k <- 10 # number of studies
theta_f <- 0.3 # real effect size
n <- c(30, 500)

sim <- make_data(k = k, theta_f = theta_f)

sim <- expand_grid(sim, nc = n)
sim$nt <- sim$nc

sim <- sim_studies(sim$theta_f, sim$nc, sim$nt, data = sim)

sim <- escalc(yi = yi, vi = vi, data = sim)
sim <- summary(sim)

fit_lp <- rma(yi, vi, method = "FE", data = filter(sim, nt == 30 & nc == 30))
fit_hp <- rma(yi, vi, method = "FE", data = filter(sim, nt == 500 & nc == 500))

xlim <- c(min(sim$ci.lb) -0.5, max(sim$ci.ub) + 0.5)

par(mfrow = c(1,2))
forest(fit_lp, main = "N = 30", xlim = c(-2, 3), alim = c(-1, 2))
forest(fit_hp, main = "N = 500", xlim = c(-2, 3), alim = c(-1, 2))
```

## Random-effects model {#re-model}

The *random-effects model* (RE) can be considered an extension of the *fixed-effects model*. The *fixed-effects model* assumes that the real effect is a single value. The random-effects model relaxes this assumption allowing $\theta_{f}$ (now $\theta_{r}$) to vary across studies. For example, the difference between groups we are simulating could be influenced by the type of experiment or the participants' age.
Now $\theta_{r}$ is no longer a single value but a distribution of values. Due to the effect size being a distribution, we need two parameters $\mu$ and $\sigma^{2}$ to describe the phenomenon. The parameter $\theta_{r}$ average true effect is now interpreted as the average effect size across different true effect sizes and $\sigma^{2}$ usually known as $\tau^{2}$ is the effect size variability or heterogeneity. In practical terms, we now have two sources of variability: $\tau^{2}$, which express the real difference among effect sizes, and $\sigma_{i}^{2}$ which is the known sampling variance of each study as in the *fixed-effects model*. We can easily extend Equation \@ref(eq:fixed-effect-model) with Equation \@ref(eq:random-effect-model).

\begin{align}
\begin{aligned}
d_i = \theta_r + \theta_i + \epsilon_i \\
\theta_i \sim \mathcal{N}(0, \tau^2) \\
\epsilon_i \sim \mathcal{N}(0,\sigma^2_i)
(\#eq:random-effect-model)
\end{aligned}
\end{align}

Compared to the *fixed-effects model* we need to generate another adjustment to the overall effect $\theta_{r}$ from a normal distribution with mean $\theta_{r}$ and variance $\tau^{2}$. The real effect size for the study $i$ will be $\theta_{r} + \theta_{i}$ where $\theta_{i}$ are the random-effects regulated by the $\tau^{2}$ parameter.

```{r sim-random-effect}
set.seed(seed)
k <- 30 # number of studies
n <- 20 # number of participants per group, per study
theta_r <- 0.3 # real effect size
tau2 <- 0.2 # the effect size heterogeneity

sim <- make_data(k = k, nc = n, nt = n, theta_r = theta_r)

# simulate the random-effects adjustment
sim$theta_i <- rnorm(k, 0, sqrt(tau2))

# adding the by-study adjustment to the average effect
sim$theta_r_theta_i <- sim$theta_r + sim$theta_i

# now we are using theta_r_theta_i and no longer theta_f
sim <- sim_studies(sim$theta_r_theta_i, sim$nc, sim$nt, data = sim)
res <- rma(yi = yi, vi = vi, method = "REML", data = sim)
```

Then we can fit the *random-effects* meta-analysis model with the `rma` function and `method = "REML"`. Table \@ref(tab:res-random-effect) depicts the model results^[The `method = "REML"` is not the only method to estimate a random-effects model. See the `rma` documentation (https://wviechtb.github.io/metafor/reference/rma.uni.html#specifying-the-model) for an overview of other estimation methods] of the metafor package. We are now estimating two parameters: $\theta_{r}$ and $\tau^{2}$. As for the *fixed-effects model*, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.

```{r res-random-effect, echo = FALSE}
res_tidy <- rma_tidy(res)
rma_table(res_tidy$res, res_tidy$notes, caption = get_caption())
```

An important aspect of the *random-effects model* is the interplay between the heterogeneity ($\tau^{2}$) and the sampling variability ($\sigma_{i}^{2}$). As the number of studies $k$ increases, the estimation of $\theta_{r}$ and $\tau^{2}$ becomes more precise. Additionally, as each study\'s sample size decreases, each $\theta_{i}$ will be estimated with higher precision but as long as $\tau^{2} \neq 0$
there will be variability among effect sizes. In other words, increasing the sample size of each study or the number of studies will not affect the value of $\tau^{2}$ but only the estimation precision [see @Borenstein2009-mo, Chapter 16, Figure 16.6 for a clear explanation]. This can be easily demonstrated using the previous simulation, increasing each study\'s sample size. Figure \@ref(fig:random-effect-high-low-precision) depicts the same meta-analysis but with different precision in estimating the study $d_{i}$. Compared to Figure \@ref(fig:fixed-effect-high-low-precision), increasing the sample size of primary studies improves the estimation of each study without reducing the between-studies heterogeneity.

```{r random-effect-high-low-precision, echo = FALSE, fig.asp=0.8, fig.cap=get_caption()}
set.seed(seed)
k <- 10 # number of studies
theta_r <- 0.3 # real effect size
tau2 <- 0.2
n <- c(30, 500)

sim <- make_data(k = k, theta_r = theta_r)

sim <- expand_grid(sim, nc = n)
sim$nt <- sim$nc

theta_i <- rnorm(k, 0, sqrt(tau2))

sim$theta_i <- rep(theta_i, each = 2)

sim$theta_r_theta_i <- sim$theta_r + sim$theta_i
sim <- sim_studies(sim$theta_r_theta_i, sim$nc, sim$nt, data = sim)

fit_lp <- rma(yi, vi, method = "REML", data = filter(sim, nc == 30 & nt == 30))
fit_hp <- rma(yi, vi, method = "REML", data = filter(sim, nc == 500 & nt == 500))

par(mfrow = c(1, 2))
forest(fit_lp, main = "N = 30", xlim = c(-3, 4), alim = c(-2, 2))
forest(fit_hp, main = "N = 500", xlim = c(-3, 4), alim = c(-2, 2))
```

The relationship between the sampling error and the heterogeneity can be expressed using the $I^{2}$ statistics [e.g., @Higgins2002-fh] that is the percentage of the total variability $\tau^{2} + \tilde{v}$ that is attributable to real heterogeneity between studies ($\tau^{2}$). In Equation \@ref(eq:isquared), $\tilde{v}$ is the "typical" within-study sampling variability [e.g., @Higgins2002-fh]^[see also here https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate#fn__1 for an overview of the $I^2$ statistics also for complex models].

\begin{align}
\begin{gathered}
I^2 = \frac{\hat{\tau}^2}{\hat{\tau}^2 + \tilde{v}}
(\#eq:isquared)
\end{gathered}
\end{align}

From Equation \@ref(eq:isquared) and Figure \@ref(fig:random-effect-high-low-precision) is clear that if each included study has a considerable sample size the total variability will be reduced, mainly driven by real heterogeneity ($\tau^2$). This is the crucial difference between the fixed-effects and the random-effects model [see also @Borenstein2009-mo, pp. 117–122].

Given the interpretation of $I^{2}$ it is possible to simulate a meta-analysis fixing a certain $I^{2}$ value. The only caveat is fixing the $\tilde{v}$. When the sample size of each study is the same, $\tilde{v} = \sigma^2_i$. In the other case, $\tilde{v}$ needs to be calculated from sampling variances reported in Higgins and Thompson [-@Higgins2002-fh] and cannot be easily fixed a priori. However, with the assumption of homogeneous sample size across studies^[The homogeneous sample size of primary studies assumption is commonly used to calculate the statistical power [see @Borenstein2009-mo, Chapter 29]], we can solve Equation \@ref(eq:isquared) for $\tau^{2}$ obtaining the heterogeneity value associated with a certain $I^{2}$ as reported in Equation \@ref(eq:tau2-from-isquared). Table \@ref(tab:res-random-effect-i2) depicts the results of the random-effects model fixing the $I^{2}$ value.

\begin{align}
\begin{gathered}
\tau^2 = - \frac{I^2\tilde{v}}{I^2 - 1}
(\#eq:tau2-from-isquared)
\end{gathered}
\end{align}

```{r sim-random-effect-i2}
set.seed(seed)
k <- 30 # number of studies
theta_r <- 0.3 # real effect size
n <- 30 # sample size per group, per study
I2 <- 0.6 # desired I2 value

v <- (n + n)/(n * n) + theta_r^2/(2 * (n + n - 2)) # typical within study variance
tau2 <- -((I2*v)/(I2 - 1))

sim <- make_data(k = k, nc = n, nt = n, theta_r = theta_r)

sim$theta_i <- rnorm(k, 0, sqrt(tau2))
sim$theta_r_theta_i <- sim$theta_r + sim$theta_i 
sim <- sim_studies(sim$theta_r_theta_i, sim$nc, sim$nt, data = sim)
res <- rma(yi, vi, method = "REML", data = sim)
```

```{r res-random-effect-i2, echo = FALSE}
res_tidy <- rma_tidy(res)
rma_table(res_tidy$res, res_tidy$notes, caption = get_caption())
```

## Meta-regression {#metareg}

From a linear regression perspective, both the FE and the RE models can be seen as *intercept-only* models where only the mean (i.e., the linear regression intercept) is estimated Borenstein et al. (2009). As reported in the meta-analysis introduction, the between-study heterogeneity usually represents the true variability of the effect due to differences among primary studies. A natural extension of the *intercept-only* analysis is a model that includes variables (i.e., *moderators*) that could explain the observed heterogeneity among effect sizes. For example, a group of studies could use a particular memory task where the expected effect is higher than another. In this way, considering the
type of task will explain part of the observed heterogeneity as in standard regression models. Figures \@ref(fig:img-metaregression-bin) and \@ref(fig:img-metaregression-num) depict a random-effects meta-regression model for a categorical and numerical predictor.

```{r img-metaregression-bin, echo = FALSE, fig.cap=get_caption(), fig.asp=1}
r_imgs$plot_metareg_bin
```

```{r img-metaregression-num, echo = FALSE, fig.cap=get_caption(), fig.asp=1}
r_imgs$plot_metareg_cont
```

### Meta-regression with a categorical moderator

A common example of meta-regression is by including a categorical predictor, including information about study-level features. In our example, a group of studies uses an online memory task while others use a standard lab-based task. The Equation \@ref(eq:random-effect-model) can be easily extended for a meta-regression model by including a variable encoding the type of task (online vs. lab-based) and the expected difference between the two levels of the moderator (i.e., the *lab vs. online effect*). In regression terms, we could use a dummy variable ($X_{1}$) that takes the value of 0 for the *lab-based task* (*l*) and a value of 1 for an *online task* (*o*)^[The *dummy coding* (also known as treatment coding) is the default in R and `metafor` but other coding schemes could be used [see @Schad2020-ht for an overview of contrast coding schemes]]. Now we fix $\beta_{1}$ to be the *lab vs. online effect* Equation \@ref(eq:random-effect-model) into Equation \@ref(eq:meta-regression). Now the product between $\beta_{1}X_{1}$ will consider the *lab vs. online effect*. To simulate this meta-regression, we need to fix the $\beta_{1}$ effect. Due to the fact that we are on a standardized scale, we can fix $\beta_{1}$ to be the expected difference in Cohen's $d$ scale between the two groups of studies. Crucially, despite $\tau^{2}$ is still the heterogeneity between effect sizes, now we need to fix $\tau^{2}$ considering that we included a moderator. In other terms, $\tau_{r}^{2}$ is the residual heterogeneity after explaining the
variance due to $X_{1}$. We can describe our model using two equations according to the value of $X_{1}$ as reported in Equation \@ref(eq:meta-regression-equations).

\begin{align}
\begin{gathered}
d_i = \beta_0 + \theta_i + \beta_1 X_1 + \epsilon_i \\
\theta_i \sim \mathcal{N}(0, \tau^2) \\
\epsilon_i \sim \mathcal{N}(0,\sigma^2_i)
(\#eq:meta-regression)
\end{gathered}
\end{align}

\begin{align}
\begin{gathered}
d_i = \beta_0 + \theta_i + \beta_1 X_1 + \epsilon_i \\
d_{il} = \beta_0 + \theta_i + \beta_1 \times 0 + \epsilon_i \\
d_{io} = \beta_0 + \theta_i + \beta_1 \times 1 + \epsilon_i \\
(\#eq:meta-regression-equations)
\end{gathered}
\end{align}

We can simulate the same scenario of the random-effects model with $k_o = 15$ (*online tasks*) and $k_l = 15$ (*lab-based tasks*). Then we fix the $\beta_1 = 0.2$ and $\tau_r^2 = 0.1$ ($r$ for residual).

```{r sim-meta-reg-dummy}
set.seed(seed)
k <- 30 # the total number of studies
b0 <- 0.1 # intercept, the effect size of the lab-based studies
b1 <- 0.2 # the difference between the two levels of the moderator
tau2r <- 0.1 # the residual heterogeneity
n <- 30 # the sample size per group, per study

sim <- make_data(k = k, nc = n, nt = n, exp = rep(c("lab", "online"), each = k/2))

sim$theta_i <- rnorm(k, 0, sqrt(tau2r)) # the by-study residual adjustment
sim$b0_theta_i_x <- b0 + sim$theta_i + b1*ifelse(sim$exp == "lab", 0, 1)

sim <- sim_studies(sim$b0_theta_i_x, sim$nc, sim$nt, data = sim)
res <- rma(yi, vi, mods = ~exp, method = "REML", data = sim)
```

Now we can fit the meta-regression model with the `rma` function as for the random-effects model with the addition of `mods = ~ exp` that indicates which variable/s to consider as moderators. The results are presented in Table \@ref(tab:res-meta-reg-dummy). Now the model will estimate an *intercept* parameter (i.e., $\beta_0$) that is the value of $y$ when $X_1$ is zero (i.e., for lab-based studies) or in other terms the expected value for lab-based studies. Then the $\beta_1$ parameter represents the estimated difference in $y$ between the values of $X_1$ (i.e., lab-based vs online experiments). As said before, $\tau^2$ is now the residual heterogeneity that is interpreted as the variability between effect sizes after controlling for the moderator $X_1$.

Now we can fit the meta-regression model with the `rma` function as for the random-effects model with the addition of mods = `~ exp` that indicates which variable/s to consider as moderators. The results are presented in Table \@ref(tab:res-meta-reg-dummy). Next, the model will estimate an *intercept* parameter (i.e., $\beta_{0}$) that is the value of $y$ when $X_{1}$ is zero (i.e., for lab-based studies) or, in other terms, the expected value for lab-based studies. Then the $\beta_{1}$ parameter represents the estimated difference in $y$ between the values of $X_{1}$ (i.e., lab-based vs. online experiments). As mentioned above, $\tau^{2}$ the residual heterogeneity is now interpreted as the variability between effect sizes after controlling for the moderator $X_{1}$.

```{r res-meta-reg-dummy, echo = FALSE}
res_tidy <- rma_tidy(res)
rma_table(res_tidy$res, res_tidy$notes, caption = get_caption())
```

### Meta-regression with a numerical moderator

The same approach can be used for a continuous predictor. For example, we can simulate that the average participant's age within each study could explain part of the observed heterogeneity. Now the $X_1$ is a continuous predictor representing the average age for each study and $\beta_1$ is the effect size increase for a unit increase (i.e., 1 year) of average age. Sometimes guessing a plausible $\beta_1$ value with a continuous predictor is not straightforward. A first strategy could be to use values estimated from the literature. Another approach consists in setting up the model and simulating several expected $d_i$ and calculating the range of simulated values. A third possibility is fixing the proportion of explained heterogeneity calculating the $\beta_1$ value accordingly. As in standard regression analysis, we can use the $R^2$ statistic to describe the amount of heterogeneity explained by the included moderators. Equation \@ref(eq:rsquared) reports how to calculate the $R^2$ for a meta-regression model. The $\tau^2_{r}$ is the residual heterogeneity after considering the moderators and $\tau^2_{f}$ is the heterogeneity estimated without considering the moderators. In the next sections, we will present an example of the simulation-based and the $R^2$ based approaches.

\begin{align}
\begin{aligned}
R^2 = 1 - \frac{\tau^2_{r}}{\tau^2_{f}}
(\#eq:rsquared)
\end{aligned}
\end{align}

In terms of regression parameters now the intercept ($\beta_0$) is no longer the overall effect or the average of one category as in the previous example but the estimated value for a specific $X_1$ thus for a specific age. If $X_1$ is a variable representing the average age for each study, then the $\beta_0$ is the average effect size when the age is zero. Depending on the moderator, the intercept is interpreted in different ways. For example, with the age, the intercept has no empirical meaning given that no studies could have a participant average age of zero. A strategy could be to mean-center the age (i.e., subtracting from each study age the average age across the study). Now the intercept is still the average effect size when the age is zero but now zero is the average age. Importantly, the contrast coding for categorical predictors or centering numerical variables does not affect the overall model estimation but only parameters values and interpretation.

#### Assessing the impact of $\beta_1$

As reported in the previous section, a strategy to guess plausible values for $\beta_{1}$ is by simulating several expected $y_{i}$ given the meta-regression equation and summarizing or plotting the effect size range. The range of simulated $y_{i}$ values are also affected by the simulated age values across studies. However, it is probably more intuitive to guess a plausible range of moderator values compared to the $\beta_{1}$ value. In this specific example, if all studies target a specific population (e.g., adults below 50 years), the expected average age range can be easily simulated. In our case, we simulated $k$ average age values from a uniform distribution $\text{ag}e_{i} \sim U\left( 20,40 \right)$. Then we can plot the distribution of $d_{i}$ values to check the plausibility of simulated values. As shown in Figure \@ref(fig:plot-meta-reg-plausible-implausible), with the same range for the moderator, a $\beta_{1} = 0.1$ gives a plausible effect sizes range while a $\beta_{1} = 0.7$ predicts very extreme values.

```{r sim-meta-reg-numerical}
set.seed(seed)
k <- 1000 # number of studies
b0 <- 0.3 # the intercept i.e., average yi when x is 0
b1 <- c(0.1, 0.7) # the beta1 i.e., the increase in yi for an increase in 1 year
tau2r <- 0.1 # the residual heterogeneity after including x1
n <- 30 # number of participants per study, per group
x1 <- runif(k, 20, 40) # random mean-age for each study
x10 <- x1 - mean(x1) # centering the age

sim <- tidyr::expand_grid(id = 1:k, nc = n, nt = n, b1 = b1)

sim$age <- rep(x1, each = 2)
sim$age0 <- rep(x10, each = 2)

theta_i <- rnorm(k, 0, sqrt(tau2r))
sim$theta_i <- rep(theta_i, each = 2)

sim$b0_theta_i_x <- b0 + sim$theta_i + sim$b1*sim$age0

sim <- sim_studies(sim$b0_theta_i_x, sim$nc, sim$nt, data = sim)
```

```{r plot-meta-reg-plausible-implausible, echo = FALSE, fig.cap=get_caption()}
siml <- split(sim, sim$b1)

plausible_b <- siml[[1]] |> 
  ggplot(aes(y = yi, x = age)) +
  geom_point(alpha = 0.5) +
  ylab(latex2exp::TeX("$d_i$")) +
  xlab("Age") +
  ggtitle(TeX("$\\beta_1 = 0.1$")) +
  #ggthemes::theme_par(base_size = 15) +
  ltxplot::theme_latex(base_size = 15) +
  theme(aspect.ratio = 1,
        plot.title = element_text(hjust = 0.5))

implausible_b <- siml[[2]] |> 
  ggplot(aes(y = yi, x = age)) +
  geom_point(alpha = 0.5) +
  ylab(latex2exp::TeX("$d_i$")) +
  xlab("Age") +
  ggtitle(TeX("$\\beta_1 = 0.7$")) +
  ggthemes::theme_par(base_size = 15) +
  #theme(axis.title.y = element_blank())
  ltxplot::theme_latex(base_size = 15) +
  coord_fixed(ratio = 1) +
  theme(aspect.ratio = 1,
        plot.title = element_text(hjust = 0.5))

plausible_bm <- ggExtra::ggMarginal(plausible_b, type = "hist")
implausible_bm <- ggExtra::ggMarginal(implausible_b, type = "hist")
gridExtra::grid.arrange(plausible_bm, implausible_bm, ncol = 2)
```

#### Simulating using $R^2$

A more intuitive way to simulate a continuous predictor is fixing the desired $R^{2}$ value and finding the coefficient that produces the desired value. This approach has been implemented by Lopez-Lopez and colleagues [-@Lopez-Lopez2014-it]. We can use Equation \@ref(eq:beta-from-r2) to find the $\beta_{1}$ value that is associated with a certain $R^{2}$.

\begin{align}
\begin{aligned}
\beta^2_1 = \tau^2R^2 \\
\tau^2_{r} = \tau^2 - \beta^2_1
(\#eq:beta-from-r2)
\end{aligned}
\end{align}

Now we can simulate the regression model using $\sqrt{\beta^2_1}$ as coefficient and $\tau_{r}^{2}$ as residual heterogeneity. Results from the fitted model fixing the $R^{2}$ values are presented in Table \@ref(tab:res-meta-reg-numerical-r2) and Figure \@ref(fig:plot-meta-reg-numerical-r2). As Lopez-Lopez and colleagues [-@Lopez-Lopez2014-it] demonstrated, to reliably estimate $R^{2}$ the number of studies needs to be large^[see also https://www.metafor-project.org/doku.php/tips:ci_for_r2 for a discussion about confidence intervals for the $R^2$ statistic]. Lopez-Lopez and colleagues [-@Lopez-Lopez2014-it] generated the moderator ($X_{1}$) values from a standard normal distribution. In the following example, we standardized the moderator (`scale()`) after simulating values on the *age* scale
(e.g., `runif(k, 20, 40)`).

```{r sim-meta-reg-numerical-r2}
set.seed(seed)
k <- 100 # the number of studies
r2 <- 0.2 # the desired r2 value
tau2 <- 0.3 # the overall tau2
b0 <- 0.3 # the intercept i.e., average yi when x1 is 0
b1_2 <- tau2 * r2 # the beta1^2 i.e., the increase in yi for an increase in 1 year
b1 <- sqrt(b1_2) # b1_2 is squared, back to the original scale
tau2r <- tau2 - b1_2 # the residual heterogeneity after including x1
n <- 30 # number of participants per study, per group
x1 <- runif(k, 20, 40) # random mean-age for each study

sim <- make_data(k = k, nt = n, nc = n, age = x1)

sim$theta_i <- rnorm(k, 0, sqrt(tau2r))
sim$age0 <- scale(sim$age, center = TRUE, scale = TRUE) # standardize the moderator
sim$b0_theta_i_x <- b0 + sim$theta_i + b1*sim$age0

sim <- sim_studies(sim$b0_theta_i_x, sim$nt, sim$nc, data = sim)
res <- rma(yi, vi, mods = ~age0, method = "REML", data = sim)
```

```{r res-meta-reg-numerical-r2, echo = FALSE}
res_tidy <- rma_tidy(res)
rma_table(res_tidy$res, res_tidy$notes, caption = get_caption())
```

```{r plot-meta-reg-numerical-r2, echo = FALSE, fig.cap=get_caption()}
res <- rma(yi, vi, mods = ~age, data = sim)
sim$wi <- weights(res) # meta-analysis weights

age_p <- seq(20, 40, 0.01) # age values to compute predictions
preds <- predict(res, newmods = age_p)
preds$age <- age_p
preds <- data.frame(preds)

ggplot() +
  geom_line(data = preds,
            aes(x = age, y = pred)) +
  geom_ribbon(data = preds,
              alpha = 0.2,
              aes(x = age, y = pred, 
                  ymin = ci.lb, ymax = ci.ub)) +
  geom_point(data = sim, 
             aes(x = age, y = yi, size = wi),
             show.legend = FALSE) +
  ylab(latex2exp::TeX("$d_i$")) +
  xlab("Age") +
  theme_paper()
```

# Power Analysis

The previous simulation examples can be easily implemented for multiple purposes. For example, we can use different effect sizes and variance estimators when using the `sim_study()` function to check the impact on the fitted meta-analysis model. However, one of the most critical applications is estimating the power of a specific statistical model.

As explained in the introduction, there are several approaches and tools to estimate the power of *fixed* and *random-effects models* [see @Borenstein2009-mo, Chapter 29; @Harrer2019-rl, Chapter 14]. These methods are easy to implement but made strong assumptions, such as the homogeneity of sample size, and did not consider the uncertainty in estimating $\tau^{2}$. Jackson and colleagues [-@Jackson2017-dv]
partially solve the issue by developing an interesting method that takes into account the uncertainty in estimating $\tau^{2}$ without using simulations. However, only Monte Carlo simulations can consider complex scenarios and extra parameters.

A general Monte Carlo simulation for the power analysis can be implemented with the following steps:

1.  Choose the model that generates the data (e.g., fixed, or random-effects model).
2.  Fix the relevant parameters (e.g., $\tau^{2}$ and $\theta$).
3.  Simulate a dataset.
4.  Fit the appropriate model.
5.  Store the p-value associated with the parameter of interest
6.  Repeat 3-5 a large number of times (e.g., 10000).
7.  Calculate the power as the proportion of p-values below the $\alpha$ level.

For example, we can estimate the power of a *random-effects model* by repeating the simulation presented in Section \@ref(re-model) many times. We simulated heterogeneity of sample sizes sampling $n_{T}$ and $n_{C}$ values from a Poisson distribution with $\lambda = 20$. In this way, on average, the sample size is 20 for primary studies with a certain amount of heterogeneity. Usually, it is more informative to simulate different scenarios according to the relevant parameters, such as sample sizes, number of studies, or heterogeneity. For example, we can estimate the power with a different number of studies $k$. We define the `do_sim()` function that, according to the input parameter, repeats the simulation a certain number of times (i.e., `nsim`)^[We are using the `purrr::pmap()` function that can be considered very similar to `mapply()` but less verbose. Compared to `mapply()`, the `sim_grid` columns are directly passed to the `do_sim()` function without specifying the order.]. Increasing the number of simulations will increase the power analysis estimation precision. Then the `summary_sim()` function analyzes each simulation returning the relevant values. We repeat the simulation of the *random-effects model* several times with different parameters.

```{r fpower-analysis-functions, results='asis', echo=FALSE}
print_fun(c(simfuns$do_sim, simfuns$summary_sim))
```

Simulation results are presented in Figure \@ref(fig:plot-power-analysis) and Table \@ref(tab:tab-power-analysis-example) showing that to reach 80% power (usually considered an appropriate level) with $\alpha = 0.05$ we need ~35 studies. The same approach could be used to estimate the power of a meta-regression by simply modifying the `do_sim()` function simulating the effect of a moderator and extracting the relevant p-value.

```{r power-analysis-example, cache = TRUE}
set.seed(seed)
nsim <- 5000 # number of simulations per condition (5000, higher is better)
k <- c(5, 15, 25, 35, 50) # number of studies
theta <- 0.3 # the average effect size
tau2 <- 0.3 # the heterogeneity
navg <- 20 # average sample size per study
nmin <- 10 # minimum sample size per study
alpha <- 0.05 # the alpha level

# creating all combinations
sim_grid <- tidyr::expand_grid(k, theta, tau2, navg, nmin, nsim)

# apply the simulation to all combinations
res <- purrr::pmap(sim_grid, do_sim)

# combine the results
res <- dplyr::bind_rows(res)
sim_grid <- cbind(sim_grid, res)
```

```{r tab-power-analysis-example, echo = FALSE}
sim_grid$navg <- as.integer(sim_grid$navg)
sim_grid$k <- as.integer(sim_grid$k)
sim_grid$nmin <- as.integer(sim_grid$nmin)
sim_grid$nsim <- as.integer(sim_grid$nsim)

sim_grid_t <- sim_grid
names(sim_grid_t) <- c("k", "$\\theta$", "$\\tau^2$", "$n_{avg}$", "$n_{min}$", "nsim", "Power")

sim_grid_t |> 
  qtab(caption = get_caption(), escape = FALSE)
```

```{r plot-power-analysis, echo = FALSE, fig.cap=get_caption()}
title <- latex2exp::TeX(sprintf("$n_{avg} = %.1f$, $\\tau^2 = %s$, $\\theta_r = %s$, $\\alpha = 0.05$", sim_grid$navg[1], tau2, theta_r, 0.05))

sim_grid |> 
  ggplot(aes(x = k, y = power)) +
  geom_hline(yintercept = 0.8, linetype = "dashed", color = "firebrick4") +
  geom_line() +
  geom_point(size = 4, shape = 21, color = "white", fill = "black", stroke = 3) +
  theme_paper() +
  ylim(c(0, 1)) +
  xlab("Number of studies (k)") +
  ylab("Power") +
  ggtitle(title)
```

# Conclusions

The present work introduced the basic concepts of the meta-analysis regarding fixed, random-effects models and meta-regression with a simulation-based approach. We believe the presented examples can help implement alternative or more complex models. For example, the `sim_study()` function can be easily modified to simulate another effect sizes index, such as a paired sample Cohen's $d$ or a Pearson's
correlation. In addition, more complex models, such as *multivariate* or *multilevel* models, can be implemented following the same approach (see supplementary materials). For example, the *three-level* model estimates another heterogeneity component representing the variability of multiple effect sizes in the same study. Similarly, the *multivariate* model could include the correlation between multiple outcomes and the correlation between sampling errors.

The present work did have a few limitations. First, we only introduced basic concepts about meta-analysis and Monte Carlo simulations, while setting up complex simulations requires more knowledge and complexity of the simulation setup. We decided to give the foundations to understand meta-analyses with a simulation approach because more complex models are still based on the same principles. Second, there are limitations
concerning simulating participant-level data. We decided to simulate the meta-analysis dataset starting from the participants' level to maximize the flexibility and clearness of each step. The downside concerns the efficiency and scalability of the simulation setup. For large-scale simulations (e.g., many conditions, iterations, or complex models), simulating from aggregated statistics is probably more efficient (see
Heuvel et al., 2020 for an example) to improve the computation speed^[see https://www.jepusto.com/simulating-correlated-smds/ for a very clear example of the participant-level vs aggregated data simulation].

In conclusion, data simulation is a very powerful tool for each step of a data analysis process. Starting from the learning phase, where simulating data can be used to understand the statistical model in terms of assumptions and the data generation process, to the estimation of statistical power. Moreover, we believe that data simulation as part of a standard research workflow could improve the overall research quality. Data simulation requires understanding the statistical model, setting appropriate and reasoned parameters, and realizing how the chosen analysis method behaves across different scenarios.

\newpage

### Acknowledgments {-}

We thank the [R-sig-meta-analysis](https://stat.ethz.ch/mailman/listinfo/r-sig-meta-analysis) mailing-list where suggestions and clarifications significantly improved our simulations approach and the R code.

### Conflicts of Interest {-}

The author(s) declare that there were no conflicts of interest with respect to the authorship or the publication of this article.

### Data, materials, and online resources {-}

The code to reproduce simulations, figures and tables can be found on Open Science Framework (https://osf.io/54djn/)

### Supplemental Material {-}

### Prior versions {-}

The manuscript preprint has been uploaded on PsyArXiv [https://psyarxiv.com/br6vy/](https://psyarxiv.com/br6vy/)

\newpage

# References {-}

::: {#refs custom-style="Bibliography"}
:::

<!-- FIGURE AND TABLES CAPTIONS -->

(ref:img-fixed-vs-random) The difference between the assumptions of the *fixed* and *random-effects model*. Each distribution depicts the sampling distribution of $k = 4$ hypothetical studies ($i = 1$, $2$, ..., $4$) with a certain observed effect size $d_{i}$, sampling variability $\sigma_{i}^{2}$ and the 95% confidence interval (black segment). The *fixed-effects* plot on the left suggests that each observed effect size has the same underlying true effect $\theta_{f}$ (each distribution has the same mean) with a different degree of precision (e.g., Study 1 is more precise than Study 3). In practice, each study has a different observed effect size where studies with high precision (i.e., narrow sampling distributions) will be remarkably close to the real effect ($\theta_{f}$). The *random-effects model* on the right suggests that beyond the error term ($\epsilon_{i}$), each real effect size is composed of a fixed part (now $\theta_{r}$) and a random part ($\theta_{i}$, black segments below the sampling distributions). The random part is determined by $\tau^{2}$, which represents the variability of the effect size distribution at the population level. When $\tau^{2}$ is zero, the random-effects reduces to a fixed-effects model.

(ref:img-multilevel) Graphical representation of a *multilevel* (three-level) model. On the top, the effect sizes are distributed at the population level. Then, each study is defined as the overall effect and the random effects determined by $\tau^{2}$, $\theta_{r} + \theta_{i}$. Finally, each experiment (effect size) within a study is composed of the study-specific effect $\theta_{r} + \theta_{i}$ and the experiment-specific random-effects $\theta_{r} + \theta_{i} + \theta_{\text{ij}}$ determined by $\omega^{2}$.

(ref:img-multivariate) Graphical representation of a *multivariate* model. In this example, we have two outcomes (magenta and green) with an average effect, heterogeneity parameters, and a certain true correlation value ($\rho$, third type of dependency). Each study at level-2 (as in the standard random-effects model) has two effect sizes with corresponding sampling variances. The two effect sizes are correlated ($\rho_{s}$, with $s$ for sampling errors) because each effect is collected on the same participants.

(ref:sim-data-example-tab) Example of data generated with the `make_data()` function. The `id` column is the identifier for each study.

(ref:notation-tab) Overview of math notation and the corresponding variables names used in the code examples.

(ref:res-fixed-effect) Summary of the simulated fixed-effects model. The only estimated parameter is the average effect* $\theta_{f}$ *with the standard error,
95% confidence interval, and the Wald z-test. The z-test evaluates the null hypothesis that the real effect equals zero.

(ref:fixed-effect-high-low-precision) Forest plots of two simulated *fixed-effects models*. On the left, the simulated model has $nt,nc = 30$ for each included study while on the right the sample size for each study is $nt,nc = 500$. Given that the data were generated under a *fixed-effects model*, when the sample size is high (on the right) each study is aligned on the real effect size because the error component ($\epsilon_{i}$) is almost entirely reduced. The average effect size is similar (depending on the random numbers generation) between the two scenarios while the estimation precision (the width of the black diamond) is narrower on the right.

(ref:res-random-effect) Summary of the simulated random-effects model: Compared to the fixed-effects model, there are more parameters. The $\beta$ is the average effect ($\theta_{r}$) with the standard error, 95% confidence interval, and the Wald z-test. The $\tau^{2}$ is the estimated heterogeneity and the $I^{2}$ (explained in Section \@ref(re-model)) represents the percentage of total variability due to between-study heterogeneity.

(ref:random-effect-high-low-precision) Forest plots of two simulated *random-effects model*s. On the left, the simulated model has $nt,nc = 30$ for each included study, while on the right, the sample size for each study is $nt,nc = 500$. Compared to Figure \@ref(fig:fixed-effect-high-low-precision), data were generated under a *random-effects model*. The estimated average effect is similar between the two scenarios regarding average effect and precision. However, compared to the *fixed-effect* simulation, increasing the sample size of primary studies only affects the precision without reducing the true heterogeneity (i.e., $\tau^{2} \neq 0$).

(ref:res-random-effect-i2) Summary of the *random-effect* model fixing the $I^2$ value. Parameters are the same as described in Table \@ref(tab:res-random-effect).

(ref:img-metaregression-bin) Graphical representation of a *random-effects meta-regression* model with a categorical predictor (Condition A and Condition B). Each grey distribution represents the sampling distribution of included studies. The dotted line is the average effect (i.e., random-effects model without moderators). The effect size differs between conditions A and B, including the *condition* moderator explaining part of the total heterogeneity (red plus green segments). The green segments depict the explained heterogeneity, and the red segments the residual (unexplained) heterogeneity.

(ref:img-metaregression-num) Graphical representation of a *random-effects meta-regression* model with a numerical predictor ($x$). Each grey distribution represents the sampling distribution of included studies. The dotted line is the average effect (i.e., random-effects model without moderators). The effect size increases as a function of the $x$ variable. Therefore, including $x$ as a predictor explains the
heterogeneity. The rest of the Figure follows the same logic as Figure \@ref(fig:img-metaregression-bin).

(ref:res-meta-reg-dummy) Summary of the *random-effect* model with a categorical predictor (lab vs online experiments). The `intercept` is the average effect for lab-based experiments and `exponline` is the difference between lab-based and online experiments. The $R^2$ is the percentage of explained heterogeneity and $\tau^2_r$ is the estimated residual heterogeneity. Other parameters are the same as the standard *random-effect* model (see Table \@ref(tab:res-random-effect)).

(ref:res-meta-reg-numerical-r2) Summary of the *random-effects* model fixing the $R^{2}$ value. The `intercept` is the effect size for the average `age` (given that age is mean-centered). The `age0` parameter is the slope between `age` and the effect size interpreted as an increase in effect size for a unit increase in the average age. Other parameters are the same as described in Table \@ref(tab:res-random-effect) and \@ref(tab:res-meta-reg-dummy).

(ref:plot-meta-reg-plausible-implausible) Scatter plots with marginal histograms for the range of simulated $d_i$ with two $\beta_1$ values. The x-axis depicts the average age of simulated studies and the y-axis depicts the simulated effect size. On the left, the majority of simulated values range between -1.5 and 1.5 thus $\beta_1 = 0.1$ can be considered a plausible value. On the right, ($\beta = 0.7$) values range between -10 to 10 that despite theoretically possible is highly implausible values in real meta-analysis for psychological data.

(ref:plot-meta-reg-numerical-r2) Meta-regression results for the *random-effects model* with a numerical moderator. Each effect size is represented with a black dot where the dimension represents the weight according to the inverse of the precision. The line represents the estimated meta-regression slope with the 95% confidence interval (grey bands).

(ref:plot-power-analysis) Results from the *random-effects model* power analysis. The x-axis depicts the number of studies ($k$), and the y-axis the estimated power. The red dotted line is the 80% power level, usually considered a good value for power analysis.

(ref:tab-power-analysis-example) The results from the power analysis simulation. The table depicts the simulation parameters, the estimated power using the `summary_sim()` function, and the average sample size ($n$) across the simulations.

### 2023-03-28 documents/paper/paper.tex
adding files

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{meta-analysis, monte carlo simulations, power analysis\newline\indent Word count: 7877}
\usepackage{csquotes}
\usepackage{setspace}
\captionsetup[figure]{font={stretch=1,footnotesize}}
\usepackage{float}
\usepackage{orcidlink}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Understanding meta-analysis through data simulation with applications to power analysis},
  pdfauthor={Filippo Gambarota 1 \& Gianmarco Altoè 1},
  pdflang={en-EN},
  pdfkeywords={meta-analysis, monte carlo simulations, power analysis},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Understanding meta-analysis through data simulation with applications to power analysis}
\author{Filippo Gambarota \orcidlink{0000-0002-6666-1747}\textsuperscript{1} \& Gianmarco Altoè \orcidlink{0000-0003-1154-9528}\textsuperscript{1}}
\date{}

\shorttitle{Simulating meta-analysis}

\authornote{

The authors made the following contributions. Filippo Gambarota \orcidlink{0000-0002-6666-1747}: Conceptualization, Methodology, Formal Analysis, Software, Writing -- Original Draft; Gianmarco Altoè \orcidlink{0000-0003-1154-9528}: Conceptualization, Methodology, Writing - Review \& Editing, Supervision.

Correspondence concerning this article should be addressed to Filippo Gambarota \orcidlink{0000-0002-6666-1747}, Via Venezia 8, 35131 Padova (PD). E-mail: \href{mailto:filippo.gambarota@unipd.it}{\nolinkurl{filippo.gambarota@unipd.it}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Developmental and Social Psychology, University of Padova, Italy}

\abstract{%
Meta-analysis is a powerful tool to combine evidence from existing literature. Despite several introductory and advanced materials about organizing, conducting, and reporting a meta-analysis, to our knowledge, there are no introductive materials about simulating the most common meta-analysis models. Data simulation is essential for developing and validating new statistical models and procedures. Furthermore, data simulation is a powerful educational tool for understanding a statistical method. In this tutorial, we show how to simulate fixed-effects, random-effects, and meta-regression models and illustrate how to estimate statistical power. In the supplementary materials, we extended these simulations also for multilevel and multivariate models. All materials associated with this article can be accessed on Open Science Framework (\url{https://osf.io/54djn/}).
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\begin{quote}
``If you do not simulate it, you have not understood it.''
\end{quote}

A meta-analysis is an essential tool for combining knowledge from multiple studies quantitatively. Meta-analysis is commonly used together with a systematic review of the literature. The meta-analysis has several advantages. Firstly, it allows combining evidence from multiple studies assigning more weight to studies with lower estimation variability. Then using meta-regression, it is possible to assess the impact of study-level variables that could impact the phenomenon. Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias. Despite the advantages, meta-analysis implementation is not always straightforward, especially for complex data structures. Additionally, the fact that there are several introductory and advanced resources to understand meta-analysis (Borenstein, Hedges, Higgins, \& Rothstein, 2009; Harrer, Cuijpers, Furukawa, \& Ebert, 2021; Schmid, Stijnen, \& White, 2022), to our knowledge, there are no introductory resources about how to simulate realistic meta-analytic data. Simulating data has several advantages because it requires understanding the statistical method and the data-generation process. Furthermore, data simulation is the primary tool when it comes to evaluating a new analysis method, estimating the statistical power, or understanding the long-run behavior of our data generation process (Gelman \& Hill, 2006, pp. 155--176; Gelman, Hill, \& Vehtari, 2020, pp. 69--76; Ingalls, 2011). A recent paper by DeBruine and Barr (2021), which deeply inspired the current work, proposed a stimulating way to understand linear mixed effects via data simulation. Simulating data is also a powerful educational tool within this framework. For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for fixed and random-effects, and meta-regression models with applications also for statistical power calculation. In the first section, we will introduce basic concepts of the meta-analysis that are useful for setting up the simulation. We will evaluate the effect size, variance calculation, and the fixed vs.~random effects model distinction. Then we will describe how to simulate data for these models and simulate a meta-regression model with a categorical and numerical predictor. Finally, we will introduce the power analysis extending the previous examples to estimate the statistical power. We used the R statistical programming language (version 4.2.3; R Core Team, 2023). We assume the reader is familiar with the basic concepts of R, but core functions will be explained. Code and materials are available on the OSF repository (\url{https://osf.io/54djn/}). Multivariate and multilevel models' simulation examples and more details about the coding approach are available in the supplementary materials.

\hypertarget{meta-analysis-introduction}{%
\subsection{Meta-analysis introduction}\label{meta-analysis-introduction}}

The meta-analysis is a statistical procedure to combine multiple studies (i.e., \emph{primary studies}) into a single statistical analysis (Borenstein et al., 2009). The idea is that combining numerous preliminary studies improve the estimation of a particular phenomenon more efficiently compared to conducting a single study. In statistical terms, the concept of the meta-analysis is to switch the statistical unit from the single participant or observation (i.e., \emph{level 1}) to the study (i.e., \emph{level 2}). Given that some studies will give more information because their estimation variability is smaller (e.g., higher sample size), the meta-analysis combines the studies assigning more weight as a function of the precision (i.e., the inverse of the variance).

As an example that will be used throughout the paper, we consider the efficacy of memory training in improving memory performance during a cognitive task. The typical primary study will collect data from a group of participants receiving the memory training (\emph{experimental group}) and another group receiving a control treatment (\emph{control group}). The focus of the meta-analysis is collecting multiple studies with similar aims and methods and estimating the average effect of memory training. Despite differences in the type of cognitive task or experimental setup, each primary study collects an \emph{experimental group} (\(n_{C}\)) and \emph{control group} (\(n_{C}\)) and computes the average performance (\emph{experimental group} \({\bar y}_{T}\) and \emph{control group} \({\bar y}_{C}\)) and standard deviations (\emph{experimental group} \(s_{T}\) and \emph{control group}
\(s_{C}\)).

\hypertarget{effect-size-and-variance}{%
\subsubsection{Effect size and variance}\label{effect-size-and-variance}}

The first step of a meta-analysis is to standardize information from each included study to convert each effect into a standard measure. This common measure should give an immediate idea of the direction (i.e., the treatment improves or reduces performance) and the size of the effect. Standardized effect size measures (see Lakens, 2013 for an overview), such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's \(d\) (Cohen, 1988) or the Pearson correlation coefficient \(\rho\) which could be calculated using the associated sample estimator \emph{r} are commonly used to compare heterogeneous outcome variables. If all studies used the same raw measure, such as reaction times, it is possible to directly meta-analyze the studies without standardizing. In practice, studies often use heterogeneous measures, therefore requiring standardized measures.

As reported in the previous section, beyond the effect size of each study, we need to assign a weight according to the precision. For this reason, we need to calculate the sampling variability of the effect size or the raw measure that will represent the estimation precision. Each raw or standardized effect size measure (e.g., raw mean difference, Cohen's \(d\), and Pearson's correlation) has a different formula to calculate the sampling variability. The idea is to choose the appropriate measure considering the study design (e.g., between vs.~within-subjects) and available information and find the proper formula to compute the sampling variability. In addition, there are formulas and approaches to convert from one effect size measure to another (Borenstein et al., 2009; Lakens, 2013; Lipsey \& Wilson, 2001).\footnote{See also the blog post by James Pustejovsky (\href{https://www.jepusto.com/alternative-formulas-for-the-smd/}{link}) for a general method to compute the sampling variability for standardized mean difference measures.} Usually, the effect size sampling variability depends mainly on the sample size that determines the weight assigned during the meta-analytic estimation.

\hypertarget{fixed-vs-random-effects-model}{%
\subsubsection{Fixed vs random-effects model}\label{fixed-vs-random-effects-model}}

The core of a meta-analysis is combining the results of multiple studies giving more weight to studies that provide a more precise effect estimation. We can divide meta-analysis models into two categories: the \emph{fixed-effects} and the \emph{random-effects model}. If we imagine combining multiple studies replicating the same experiment, we assume the real treatment effect \(\theta_{f}\) is \emph{fixed,} and each study is a more or less precise estimation of this parameter. This is an essential and often inappropriate assumption because it ignores that study-level characteristics may impact the actual treatment effect.

Suppose some study-level characteristics (e.g., participants' age, sex, or socioeconomic status) or the experimental paradigm (e.g., type of memory task or difficulty) could impact the treatment effect. In that case, the assumption of the \emph{fixed-effects model} is no longer plausible. There is no unique real effect but a distribution of effects with mean \(\theta_{r}\) and variance \(\tau^{2}\). The mean is the \emph{average} effect, and the variance is the heterogeneity of the phenomenon. The \emph{random-effects model} will estimate the mean (\(\theta_{r}\)) and the heterogeneity \(\tau^{2}\) will help generalize to the population of effects beyond the pool of included studies.

Most importantly, the parameters estimated with the \emph{fixed} and \emph{random-effects} do not represent the same quantity. The \emph{random-effects model} estimates the \emph{average} effect across a distribution of true effects, while the \emph{fixed-effects model} estimates a single true effect. Figure 1 depicts the idea of the fixed vs.~random-effects model. The \emph{fixed-effects model} can be considered a particular type of \emph{random-effects model} where \(\tau^{2} = 0\) thus, the average effect is also the unique true effect.

\scriptsize

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/img-fixed-vs-random-1}

}

\caption{The difference between the assumptions of the \emph{fixed} and \emph{random-effects model}. Each distribution depicts the sampling distribution of \(k = 4\) hypothetical studies (\(i = 1\), \(2\), \ldots, \(4\)) with a certain observed effect size \(d_{i}\), sampling variability \(\sigma_{i}^{2}\) and the 95\% confidence interval (black segment). The \emph{fixed-effects} plot on the left suggests that each observed effect size has the same underlying true effect \(\theta_{f}\) (each distribution has the same mean) with a different degree of precision (e.g., Study 1 is more precise than Study 3). In practice, each study has a different observed effect size where studies with high precision (i.e., narrow sampling distributions) will be remarkably close to the real effect (\(\theta_{f}\)). The \emph{random-effects model} on the right suggests that beyond the error term (\(\epsilon_{i}\)), each real effect size is composed of a fixed part (now \(\theta_{r}\)) and a random part (\(\theta_{i}\), black segments below the sampling distributions). The random part is determined by \(\tau^{2}\), which represents the variability of the effect size distribution at the population level. When \(\tau^{2}\) is zero, the random-effects reduces to a fixed-effects model.}\label{fig:img-fixed-vs-random}
\end{figure}

\normalsize

\hypertarget{multilevel-and-multivariate-models}{%
\subsubsection{Multilevel and multivariate models}\label{multilevel-and-multivariate-models}}

The most straightforward situation for a meta-analysis is when each included study contributes with a single effect size. However, it is common to have multiple effect sizes belonging to the same study creating a situation of dependency between statistical units. There are numerous sources of dependency for a meta-analysis (Cheung, 2014; 2015, pp. 121--122). The first type of dependency (\emph{multilevel}) consists of multiple effect sizes from independent participants. Despite being collected with different participants, effect sizes from the same study could be correlated. We obtain a multilevel data structure with independent effect sizes in the same study. There are multiple ways to model this data structure (see Cheung, 2014 for an overview). Still, the most common way is to use a \emph{multilevel meta-analysis model,} considering the heterogeneity between and within studies. The standard \emph{random-effects} (but also the \emph{fixed-effects model}) model is also called a \emph{two-level} model because it combines data after aggregating the participants-level (\emph{first-level}) data. The \emph{three-level} model is commonly used to manage the multilevel data structure (Cheung, 2014, 2019). Compared to the \emph{two-level} model, the \emph{three-level} model estimates two heterogeneity parameters (\(\tau^{2}\) and \(\omega^{2}\)), one for the heterogeneity between studies and the other for the heterogeneity within studies.

A second type of dependency (\emph{multivariate}) consists of multiple effect sizes collected from the same pool of participants. For example, when using two cognitive tasks on the same pool of participants. This is commonly known as \emph{multivariate meta-analysis} because the effect sizes sampling errors are correlated given that participants are assessed multiple times. A \emph{multivariate meta-analysis model} could be used to
consider the correlation between sampling errors.

A third type of dependency arises when true effect sizes are correlated at the population level (Cheung, 2015, pp. 121--122). For example, two different memory tasks could be correlated because the latent psychological constructs are intrinsically correlated. This correlation is present even when other groups of participants perform the two memory tasks (i.e., no \emph{multivariate} dependency) or are administered by different researchers (i.e., no \emph{multilevel} dependency). This type of dependency could be managed by estimating the correlation between outcomes using a \emph{multivariate random-effects model} (Cheung, 2015, pp. 127--130). Figures \ref{fig:img-multilevel} and \ref{fig:img-multivariate} depict the conceptual idea of \emph{multilevel} and \emph{multivariate} models.

\scriptsize

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{img/multilevel}

}

\caption{Graphical representation of a \emph{multilevel} (three-level) model. On the top, the effect sizes are distributed at the population level. Then, each study is defined as the overall effect and the random effects determined by \(\tau^{2}\), \(\theta_{r} + \theta_{i}\). Finally, each experiment (effect size) within a study is composed of the study-specific effect \(\theta_{r} + \theta_{i}\) and the experiment-specific random-effects \(\theta_{r} + \theta_{i} + \theta_{\text{ij}}\) determined by \(\omega^{2}\).}\label{fig:img-multilevel}
\end{figure}

\normalsize

\scriptsize

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{img/multivariate.pdf }

}

\caption{Graphical representation of a \emph{multivariate} model. In this example, we have two outcomes (magenta and green) with an average effect, heterogeneity parameters, and a certain true correlation value (\(\rho\), third type of dependency). Each study at level-2 (as in the standard random-effects model) has two effect sizes with corresponding sampling variances. The two effect sizes are correlated (\(\rho_{s}\), with \(s\) for sampling errors) because each effect is collected on the same participants.}\label{fig:img-multivariate}
\end{figure}

\normalsize

\hypertarget{simulation}{%
\section{Simulation}\label{simulation}}

\hypertarget{monte-carlo-simulations}{%
\subsection{Monte Carlo simulations}\label{monte-carlo-simulations}}

The Monte Carlo methods are controlled experiments (Gentle, 2009). Given a set of fixed parameters, probability distributions, and the possibility of generating random numbers, it is possible to simulate the behavior of an empirical system. Monte Carlo simulations are used for statistical and mathematical problems that cannot be solved analytically.

A straightforward example regards estimating the sampling variability of the mean difference. When calculating the mean difference between two samples, we estimate the true mean difference at the population level with a certain degree of error (i.e., the standard error of the mean difference). The central limit theorem states that the difference between the means of two random samples (\(\overline{X}\) and \(\overline{Y}\)) is approximately normally distributed with mean \(\mu_{x} - \mu_{y}\) and standard error \(\sqrt{\frac{\sigma^2_x}{{n_x}} + \frac{\sigma^2_y}{{n_y}}}\). The same results can be obtained using Monte Carlo simulations using the following procedure:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Generating two random samples from two normal distributions with a fixed mean difference
\item
  Calculating the mean difference
\item
  Repeating the same process, many times
\item
  Calculate the standard deviation of the simulated values
\end{enumerate}

Using the method, we are estimating via simulation the standard deviation of the sampling distribution of the mean difference (i.e., the standard error of the mean difference). Increasing the number of simulations will produce more stable results.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{) }\CommentTok{\# seed for simulation to ensure the reproducibility of results}
\NormalTok{theta }\OtherTok{\textless{}{-}} \FloatTok{0.5} \CommentTok{\# real mean difference in the populations}
\NormalTok{sigma }\OtherTok{\textless{}{-}} \DecValTok{1} \CommentTok{\# real standard deviations in the population}
\NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{30} \CommentTok{\# sample size for both groups}
\NormalTok{nsims }\OtherTok{\textless{}{-}} \FloatTok{1e5} \CommentTok{\# number of simulations = 100000}

\CommentTok{\# simulates the sampling distribution of the mean difference}
\NormalTok{d\_mc }\OtherTok{\textless{}{-}} \FunctionTok{replicate}\NormalTok{(nsims, }\AttributeTok{expr =}\NormalTok{ \{}
\NormalTok{  g1 }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\AttributeTok{n =}\NormalTok{ n, }\AttributeTok{mean =}\NormalTok{ theta, }\AttributeTok{sd =}\NormalTok{ sigma)}
\NormalTok{  g2 }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\AttributeTok{n =}\NormalTok{ n, }\AttributeTok{mean =} \DecValTok{0}\NormalTok{, }\AttributeTok{sd =}\NormalTok{ sigma)}
  \FunctionTok{mean}\NormalTok{(g1) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(g2) }\CommentTok{\# calculates the sample mean difference and returns the value}
\NormalTok{\})}

\CommentTok{\# analytical standard error of the mean difference}
\NormalTok{se\_an }\OtherTok{\textless{}{-}} \FunctionTok{sqrt}\NormalTok{(sigma}\SpecialCharTok{\^{}}\DecValTok{2}\SpecialCharTok{/}\NormalTok{n }\SpecialCharTok{+}\NormalTok{ sigma}\SpecialCharTok{\^{}}\DecValTok{2}\SpecialCharTok{/}\NormalTok{n)}

\CommentTok{\# monte{-}carlo standard error of the mean difference}
\NormalTok{se\_mc }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(d\_mc)}
\end{Highlighting}
\end{Shaded}

\normalsize

The standard error estimated solving analytically is 0.26 and using the Monte Carlo simulation we arrive at the same result (i.e., 0.26).

\hypertarget{simulation-setup}{%
\subsection{Simulation setup}\label{simulation-setup}}

We load relevant packages and set the \texttt{seed} for reproducibility to prepare the R environment for the simulations.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(dplyr) }\CommentTok{\# for data manipulation}
\FunctionTok{library}\NormalTok{(tidyr) }\CommentTok{\# for data manipulation}
\FunctionTok{library}\NormalTok{(metafor) }\CommentTok{\# for meta{-}analysis models fitting}
\NormalTok{seed }\OtherTok{\textless{}{-}} \DecValTok{2023} \CommentTok{\# general seed for all simulations}
\end{Highlighting}
\end{Shaded}

\normalsize

Before diving into the specific simulations, in this section, we define the common aspects of all simulations in the following sections. The current paper focuses on the \emph{two-level}, \emph{fixed,} and \emph{random-effects model}s. All the examples refer to primary studies that assess the efficacy of a treatment by comparing a control and an experimental group. Equation \eqref{eq:datagen} formalizes the assumption about generating participants' data within a single study. The two groups (\emph{experimental} \(y_{T}\) and \emph{control} \(y_{C}\)) are generated from normal distributions with \(\mu = 0\) for the control group and \(\mu = \theta\) (where \(\theta\) is the effect size) for the experimental group. Both groups are assumed to be sampled from populations with homogeneous variances \(\sigma^{2} = 1\).

\begin{align}
\begin{gathered}
y_{Ti} \sim \mathcal{N}(\theta, 1) \\
y_{Ci} \sim \mathcal{N}(0, 1)
\label{eq:datagen}
\end{gathered}
\end{align}

The true Standardized Mean Difference can be estimated using Cohen's \(d\) as reported in Equation \eqref{eq:effsize}. The difference between the two means is divided by the pooled standard deviation of the two groups. It is common to apply a small-sample correction version of this effect size called Hedges' \(g\) (Hedges, 1981)

\begin{align}
\begin{gathered}
d = \frac{\bar y_T - \bar y_C}{S_p} \\
S_p = \sqrt{\frac{S^2_{y_T}(n_{y_T} - 1) + S^2_{y_C}(n_{C} - 1)}{n_{T} + n_{C} - 2}} \\
\sigma^2_d = \frac{n_{T} + n_{C}}{n_{T}n_{C}} + \frac{d^2}{2(n_{T} + n_{C} - 2)}
\label{eq:effsize}
\end{gathered}
\end{align}

We can use the following algorithm implemented in the \texttt{sim\_study()} function to simulate a single study. Before using the \texttt{sim\_study()}
function, we can create a data frame for the simulation using the \texttt{make\_data()} function\footnote{This function is a simple wrapper of \texttt{data.frame()} that given the number of studies and other variables create the data structure for the simulation.}. Table \ref{tab:sim-data-example-tab} depicts an example of the \texttt{make\_data()} output.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Choose a \(\theta\), \(n_{T}\), and \(n_{C}\) value.
\item
  Simulate \(n_{T}\) observations from a Gaussian distribution with \(\mu = \theta\) and \(\sigma^{2} = 1\) and \(n_{T}\) observations from a Gaussian distribution with \(\mu = 0\) and \(\sigma^{2} = 1\). This way, the expected difference between groups will be \(d\) and the variance is fixed at 1.
\item
  Calculate the pooled standard deviation \(S_{p}\) that is an estimation of the population standard deviation (\(\sigma\)) to standardize the mean difference.
\item
  Calculate the observed effect size \(d\) and the sampling variance \(\sigma_{d}^{2}\).
\end{enumerate}

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]

\NormalTok{make\_data }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(k, }\AttributeTok{nt =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{nc =} \ConstantTok{NULL}\NormalTok{, ...)\{}
\NormalTok{  params }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{as.list}\NormalTok{(}\FunctionTok{environment}\NormalTok{()), }\FunctionTok{list}\NormalTok{(...))}
\NormalTok{  cols }\OtherTok{\textless{}{-}}\NormalTok{ params[}\SpecialCharTok{!}\FunctionTok{sapply}\NormalTok{(params, is.null) }\SpecialCharTok{\&} \FunctionTok{names}\NormalTok{(params) }\SpecialCharTok{!=} \StringTok{"k"}\NormalTok{]}
\NormalTok{  dat }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
    \AttributeTok{id =} \DecValTok{1}\SpecialCharTok{:}\NormalTok{params}\SpecialCharTok{$}\NormalTok{k}
\NormalTok{  )}
  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{length}\NormalTok{(cols) }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{)\{}
\NormalTok{    dat }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(dat, cols)}
\NormalTok{  \}}
  \FunctionTok{return}\NormalTok{(dat)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\normalsize

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]

\NormalTok{sim\_study }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(theta, nt, }\AttributeTok{nc =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{hedges =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{aggregate =} \ConstantTok{TRUE}\NormalTok{)\{}
  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{is.null}\NormalTok{(nc)) nc }\OtherTok{\textless{}{-}}\NormalTok{ nt}
  \CommentTok{\# generate from normal distribution}
\NormalTok{  yc }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(nc, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)}
\NormalTok{  yt }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(nt, theta, }\DecValTok{1}\NormalTok{)}
  \CommentTok{\# pooled variance}
\NormalTok{  sp }\OtherTok{\textless{}{-}} \FunctionTok{sqrt}\NormalTok{((}\FunctionTok{var}\NormalTok{(yc)}\SpecialCharTok{*}\NormalTok{(nc }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{var}\NormalTok{(yt)}\SpecialCharTok{*}\NormalTok{(nt }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{)) }\SpecialCharTok{/}\NormalTok{ (nc }\SpecialCharTok{+}\NormalTok{ nt }\SpecialCharTok{{-}} \DecValTok{2}\NormalTok{))}
  \CommentTok{\# effect size}
\NormalTok{  yi }\OtherTok{\textless{}{-}}\NormalTok{ (}\FunctionTok{mean}\NormalTok{(yt) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(yc)) }\SpecialCharTok{/}\NormalTok{ sp}

  \CommentTok{\# transform to hedges\textquotesingle{} g}
  \ControlFlowTok{if}\NormalTok{(hedges)\{}
\NormalTok{    yi }\OtherTok{\textless{}{-}}\NormalTok{ yi }\SpecialCharTok{*}\NormalTok{ metafor}\SpecialCharTok{:::}\FunctionTok{.cmicalc}\NormalTok{(nc }\SpecialCharTok{+}\NormalTok{ nt }\SpecialCharTok{{-}} \DecValTok{2}\NormalTok{)}
\NormalTok{  \}}

  \CommentTok{\# sampling variance}
\NormalTok{  vi }\OtherTok{\textless{}{-}}\NormalTok{ (nc }\SpecialCharTok{+}\NormalTok{ nt)}\SpecialCharTok{/}\NormalTok{(nc }\SpecialCharTok{*}\NormalTok{ nt) }\SpecialCharTok{+}\NormalTok{ yi}\SpecialCharTok{\^{}}\DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\DecValTok{2} \SpecialCharTok{*}\NormalTok{ (nc }\SpecialCharTok{+}\NormalTok{ nt }\SpecialCharTok{{-}} \DecValTok{2}\NormalTok{))}

  \ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\NormalTok{aggregate)\{}
    \CommentTok{\# return raw data}
    \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{id =} \DecValTok{1}\SpecialCharTok{:}\NormalTok{(nc }\SpecialCharTok{+}\NormalTok{ nt),}
               \AttributeTok{group =} \FunctionTok{rep}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"c"}\NormalTok{, }\StringTok{"t"}\NormalTok{), }\FunctionTok{c}\NormalTok{(nc, nt)),}
               \AttributeTok{y =} \FunctionTok{c}\NormalTok{(yc, yt))}
\NormalTok{  \}}\ControlFlowTok{else}\NormalTok{\{}
    \CommentTok{\# compute effect size}
    \FunctionTok{data.frame}\NormalTok{(yi, vi)}
\NormalTok{  \}}
\NormalTok{\}}

\NormalTok{sim\_studies }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(..., }\AttributeTok{data =} \ConstantTok{NULL}\NormalTok{)\{}
  \CommentTok{\# ... (dots) are the arguments passed to mapply, with the order required by sim\_study}
\NormalTok{  res }\OtherTok{\textless{}{-}} \FunctionTok{mapply}\NormalTok{(sim\_study, ..., }\AttributeTok{SIMPLIFY =} \ConstantTok{FALSE}\NormalTok{)}
  \CommentTok{\# everything to a dataframe}
\NormalTok{  res }\OtherTok{\textless{}{-}}\NormalTok{ dplyr}\SpecialCharTok{::}\FunctionTok{bind\_rows}\NormalTok{(res)}

  \ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.null}\NormalTok{(data))\{}
    \CommentTok{\# attach to the original dataset}
    \FunctionTok{cbind}\NormalTok{(data, res)}
\NormalTok{  \}}\ControlFlowTok{else}\NormalTok{\{}
\NormalTok{    res}
\NormalTok{  \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\normalsize

The \texttt{theta} is the true effect size, \texttt{nc} and \texttt{nt} are the sample size for the control and experimental group, \texttt{aggregate} controls if returning the effect size and the corresponding variance or the participant-level data, and \texttt{hedges} specify whether applying the correction. We can generate a single study with the desired parameters with this function. A suggestion for each simulation step is to generate a large \(n\) to reduce the sampling error and check the recovery of simulated parameters. This is a general strategy that can be applied to every simulation. The \texttt{sim\_studies()} function will iterate through variables in the \texttt{...} argument creating the meta-analysis data frame. The \texttt{mapply()} function is clearly explained in the supplementary materials.

The following code simulates a single study with \(n = 10000\) and check the estimated mean and standard deviation.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(seed)}
\FunctionTok{sim\_study}\NormalTok{(}\AttributeTok{theta =} \FloatTok{0.3}\NormalTok{, }\AttributeTok{nc =} \DecValTok{10000}\NormalTok{, }\AttributeTok{nt =} \DecValTok{10000}\NormalTok{, }\AttributeTok{aggregate =} \ConstantTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\normalsize

\scriptsize

\normalsize

The control group has a mean of -0.008 (SD = 0.996) and the experimental group has a mean of 0.306 (SD = 1.013) which are remarkably close to the simulated values.

Using the \texttt{sim\_study()} function multiple times (with the appropriate adjustments) we can generate a series of studies simulating a data set for a meta-analysis (using the \texttt{sim\_studies()} function). After each example, we will compute the appropriate model (e.g., \emph{fixed} or \emph{random-effects}) using the \texttt{metafor} package (Viechtbauer, 2010) to check the recovery of simulated parameters. Table \ref{tab:notation-tab} summarizes the simulation parameters notation in equations and code.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{k }\OtherTok{\textless{}{-}} \DecValTok{30} \CommentTok{\# number of studies}
\NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{20} \CommentTok{\# number of participants per group, per study}
\NormalTok{theta }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# real effect size}
\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{theta =}\NormalTok{ theta)}
\end{Highlighting}
\end{Shaded}

\normalsize

\scriptsize

\begin{table}[H]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:sim-data-example-tab}Example of data generated with the \texttt{make\_data()} function. The \texttt{id} column is the identifier for each study.}

\small{

\begin{tabular}{llll}
\toprule
id & \multicolumn{1}{c}{nt} & \multicolumn{1}{c}{nc} & \multicolumn{1}{c}{theta}\\
\midrule
1 & 20 & 20 & 0.3\\
2 & 20 & 20 & 0.3\\
... & ... & ... & ...\\
29 & 20 & 20 & 0.3\\
30 & 20 & 20 & 0.3\\
\bottomrule
\end{tabular}

}

\end{threeparttable}
\end{center}

\end{table}

\normalsize

\scriptsize

\begin{table}[H]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:notation-tab}Variables in the data-generating model and associated R code.}

\begin{tabular}{lll}
\toprule
Equation & \multicolumn{1}{c}{Code} & \multicolumn{1}{c}{Description}\\
\midrule
$\theta_f$ & \texttt{theta\_f} & Fixed-effects model real effect size\\
$\theta_r$ & \texttt{theta\_r} & Random-effects model real effect size\\
$\tau^2$ & \texttt{tau2} & Effect size heterogeneity\\
$\tau^2_r$ & \texttt{tau2r} & Residual effect size heterogeneity\\
$\bar y_{T, C}$ & \texttt{} & Mean of the experimental/control group\\
$s^2_{T, C}$ & \texttt{} & Standard deviation of the experimental/control group\\
$n_{T, C}$ & \texttt{nt, nc} & Sample size of the experimental/control group\\
$d_i$ & \texttt{yi} & Observed effect size\\
$\sigma^2_i$ & \texttt{vi} & Observed sampling variance\\
$\theta_i$ & \texttt{theta\_i} & Random effect for the study i\\
$\beta_0$ & \texttt{b0} & Meta-regression intercept\\
$\beta_1$ & \texttt{b1} & Meta-regression slope\\
$k$ & \texttt{k} & Number of studies\\
$\epsilon_i$ & \texttt{} & Sampling error for the study i\\
\bottomrule
\addlinespace
\end{tabular}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} The \texttt{yi} and \texttt{vi} notation for the observed effect size ($d_i$) and sampling variance ($\sigma^2_i$) has been used to be consistent with the \texttt{metafor} notation.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\normalsize

\hypertarget{fixed-effects-model}{%
\subsection{Fixed-effects model}\label{fixed-effects-model}}

The most basic model to simulate is the \emph{fixed-effects model} (FE). As reported in the previous sections, the \emph{fixed-effects model} assumes the presence of a single true effect (\(\theta_{f}\)) and the observed variability is because each effect size is an imprecise estimation of the true effect. In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies
(i.e., the sample size). Equation \eqref{eq:fixed-effect-model} depicts the \emph{fixed-effect} equation.

\begin{align}
\begin{gathered}
d_i = \theta_f + \epsilon_i \\
\epsilon_i \sim \mathcal{N}(0,\sigma^2_i)
\label{eq:fixed-effect-model}
\end{gathered}
\end{align}

Each observed effect size (\(d_{i}\)) is composed of the real effect size \(\theta_{f}\) plus an error term (\(\epsilon_{i}\)) that is sampled from a normal distribution with \(\mu = 0\) and \(\sigma^{2} = \sigma_{i}^{2}\) (i.e., the known sampling variance of the study \(i\)). As demonstrated in
Equation \eqref{eq:effsize}, the increase in sample size will decrease the sampling variability. An exact study will essentially have \(\theta_{f}\) as the observed effect size. Since we are sampling participants' level data, the error component is already included in the \texttt{sim\_study()} function. To simulate this model in R we can just call the \texttt{sim\_study()} multiple times according to the number of desired studies (\(k\)). In addition, we simulate that each primary study will have a sample size of \(n = 20\) for both groups. We will discuss later the appropriateness of this assumption.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(seed)}
\NormalTok{k }\OtherTok{\textless{}{-}} \DecValTok{30} \CommentTok{\# number of studies}
\NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{20} \CommentTok{\# number of participants per group, per study}
\NormalTok{theta\_f }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# real effect size}

\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{theta\_f =}\NormalTok{ theta\_f)}
\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{theta\_f, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim) }\CommentTok{\# simulated data}
\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(}\AttributeTok{yi =}\NormalTok{ yi, }\AttributeTok{vi =}\NormalTok{ vi, }\AttributeTok{method =} \StringTok{"FE"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
\end{Highlighting}
\end{Shaded}

\normalsize

We are using the \(\theta_{f}\) parameter to generate random data using the \texttt{sim\_study()} function that will introduce the random error component \(\epsilon_{i}\). Then we can fit the \emph{fixed-effects} meta-analysis model with the \texttt{rma} function and \texttt{method\ =\ "FE"} of the \texttt{metafor} package. The parameter we are estimating is \(\theta_{f}\) which is close to our simulation value (see Table \ref{tab:res-fixed-effect}). Increasing the number of studies (\(k\)) and/or the number of participants in each study (\(n\)) will improve estimation reducing the standard error.

\scriptsize

\begin{table}[H]

\caption{\label{tab:res-fixed-effect}Summary of the simulated fixed-effects model. The only estimated parameter is the average effect* \(\theta_{f}\) *with the standard error,
95\% confidence interval, and the Wald z-test. The z-test evaluates the null hypothesis that the real effect equals zero.}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{ccccc}
\toprule
 & $\beta$ & 95\% CI & z & p\\
\midrule
overall & 0.334 (SE = 0.059) & $[0.219, 0.449]$ & 5.701 & < 0.001\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} $k = 30$}\\
\end{tabular}
\end{table}

\normalsize

Increasing the sample size for each study will increase the estimation precision thus the variability among studies will be reduced. This can be easily demonstrated by simulating studies with a high sample size as reported in Figure \ref{fig:fixed-effect-high-low-precision}. As the sample size increases the only source of variability (i.e., the error component \(\epsilon_{i}\)) is close to zero so each study is closer to the true simulated value.

\scriptsize

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/fixed-effect-high-low-precision-1}

}

\caption{Forest plots of two simulated \emph{fixed-effects models}. On the left, the simulated model has \(nt,nc = 30\) for each included study while on the right the sample size for each study is \(nt,nc = 500\). Given that the data were generated under a \emph{fixed-effects model}, when the sample size is high (on the right) each study is aligned on the real effect size because the error component (\(\epsilon_{i}\)) is almost entirely reduced. The average effect size is similar (depending on the random numbers generation) between the two scenarios while the estimation precision (the width of the black diamond) is narrower on the right.}\label{fig:fixed-effect-high-low-precision}
\end{figure}

\normalsize

\hypertarget{re-model}{%
\subsection{Random-effects model}\label{re-model}}

The \emph{random-effects model} (RE) can be considered an extension of the \emph{fixed-effects model}. The \emph{fixed-effects model} assumes that the real effect is a single value. The random-effects model relaxes this assumption allowing \(\theta_{f}\) (now \(\theta_{r}\)) to vary across studies. For example, the difference between groups we are simulating could be influenced by the type of experiment or the participants' age.
Now \(\theta_{r}\) is no longer a single value but a distribution of values. Due to the effect size being a distribution, we need two parameters \(\mu\) and \(\sigma^{2}\) to describe the phenomenon. The parameter \(\theta_{r}\) average true effect is now interpreted as the average effect size across different true effect sizes and \(\sigma^{2}\) usually known as \(\tau^{2}\) is the effect size variability or heterogeneity. In practical terms, we now have two sources of variability: \(\tau^{2}\), which express the real difference among effect sizes, and \(\sigma_{i}^{2}\) which is the known sampling variance of each study as in the \emph{fixed-effects model}. We can easily extend Equation \eqref{eq:fixed-effect-model} with Equation \eqref{eq:random-effect-model}.

\begin{align}
\begin{aligned}
d_i = \theta_r + \theta_i + \epsilon_i \\
\theta_i \sim \mathcal{N}(0, \tau^2) \\
\epsilon_i \sim \mathcal{N}(0,\sigma^2_i)
\label{eq:random-effect-model}
\end{aligned}
\end{align}

Compared to the \emph{fixed-effects model} we need to generate another adjustment to the overall effect \(\theta_{r}\) from a normal distribution with mean \(\theta_{r}\) and variance \(\tau^{2}\). The real effect size for the study \(i\) will be \(\theta_{r} + \theta_{i}\) where \(\theta_{i}\) are the random-effects regulated by the \(\tau^{2}\) parameter.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(seed)}
\NormalTok{k }\OtherTok{\textless{}{-}} \DecValTok{30} \CommentTok{\# number of studies}
\NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{20} \CommentTok{\# number of participants per group, per study}
\NormalTok{theta\_r }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# real effect size}
\NormalTok{tau2 }\OtherTok{\textless{}{-}} \FloatTok{0.2} \CommentTok{\# the effect size heterogeneity}

\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{theta\_r =}\NormalTok{ theta\_r)}

\CommentTok{\# simulate the random{-}effects adjustment}
\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2))}

\CommentTok{\# adding the by{-}study adjustment to the average effect}
\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_r\_theta\_i }\OtherTok{\textless{}{-}}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_r }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_i }

\CommentTok{\# now we are using theta\_r\_theta\_i and no longer theta\_f}
\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{theta\_r\_theta\_i, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim)}
\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(}\AttributeTok{yi =}\NormalTok{ yi, }\AttributeTok{vi =}\NormalTok{ vi, }\AttributeTok{method =} \StringTok{"REML"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
\end{Highlighting}
\end{Shaded}

\normalsize

Then we can fit the \emph{random-effects} meta-analysis model with the \texttt{rma} function and \texttt{method\ =\ "REML"}. Table \ref{tab:res-random-effect} depicts the model results\footnote{The \texttt{method\ =\ "REML"} is not the only method to estimate a random-effects model. See the \texttt{rma} documentation (\url{https://wviechtb.github.io/metafor/reference/rma.uni.html\#specifying-the-model}) for an overview of other estimation methods} of the metafor package. We are now estimating two parameters: \(\theta_{r}\) and \(\tau^{2}\). As for the \emph{fixed-effects model}, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.

\scriptsize

\begin{table}[H]

\caption{\label{tab:res-random-effect}Summary of the simulated random-effects model: Compared to the fixed-effects model, there are more parameters. The \(\beta\) is the average effect (\(\theta_{r}\)) with the standard error, 95\% confidence interval, and the Wald z-test. The \(\tau^{2}\) is the estimated heterogeneity and the \(I^{2}\) (explained in Section \ref{re-model}) represents the percentage of total variability due to between-study heterogeneity.}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{ccccc}
\toprule
 & $\beta$ & 95\% CI & z & p\\
\midrule
overall & 0.381 (SE = 0.107) & $[0.172, 0.590]$ & 3.578 & < 0.001\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} $k = 30$}\\
\multicolumn{5}{l}{\textsuperscript{} $\tau^2 = 0.234$ ($SE = 0.089$)}\\
\multicolumn{5}{l}{\textsuperscript{} $I^2 = 68.842\%$}\\
\end{tabular}
\end{table}

\normalsize

An important aspect of the \emph{random-effects model} is the interplay between the heterogeneity (\(\tau^{2}\)) and the sampling variability (\(\sigma_{i}^{2}\)). As the number of studies \(k\) increases, the estimation of \(\theta_{r}\) and \(\tau^{2}\) becomes more precise. Additionally, as each study\textquotesingle s sample size decreases, each \(\theta_{i}\) will be estimated with higher precision but as long as \(\tau^{2} \neq 0\)
there will be variability among effect sizes. In other words, increasing the sample size of each study or the number of studies will not affect the value of \(\tau^{2}\) but only the estimation precision (see Borenstein et al., 2009, Chapter 16, Figure 16.6 for a clear explanation). This can be easily demonstrated using the previous simulation, increasing each study\textquotesingle s sample size. Figure \ref{fig:random-effect-high-low-precision} depicts the same meta-analysis but with different precision in estimating the study \(d_{i}\). Compared to Figure \ref{fig:fixed-effect-high-low-precision}, increasing the sample size of primary studies improves the estimation of each study without reducing the between-studies heterogeneity.

\scriptsize

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/random-effect-high-low-precision-1}

}

\caption{Forest plots of two simulated \emph{random-effects model}s. On the left, the simulated model has \(nt,nc = 30\) for each included study, while on the right, the sample size for each study is \(nt,nc = 500\). Compared to Figure \ref{fig:fixed-effect-high-low-precision}, data were generated under a \emph{random-effects model}. The estimated average effect is similar between the two scenarios regarding average effect and precision. However, compared to the \emph{fixed-effect} simulation, increasing the sample size of primary studies only affects the precision without reducing the true heterogeneity (i.e., \(\tau^{2} \neq 0\)).}\label{fig:random-effect-high-low-precision}
\end{figure}

\normalsize

The relationship between the sampling error and the heterogeneity can be expressed using the \(I^{2}\) statistics (e.g., Higgins \& Thompson, 2002) that is the percentage of the total variability \(\tau^{2} + \tilde{v}\) that is attributable to real heterogeneity between studies (\(\tau^{2}\)). In Equation \eqref{eq:isquared}, \(\tilde{v}\) is the ``typical'' within-study sampling variability (e.g., Higgins \& Thompson, 2002)\footnote{see also here \url{https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate\#fn__1} for an overview of the \(I^2\) statistics also for complex models}.

\begin{align}
\begin{gathered}
I^2 = \frac{\hat{\tau}^2}{\hat{\tau}^2 + \tilde{v}}
\label{eq:isquared}
\end{gathered}
\end{align}

From Equation \eqref{eq:isquared} and Figure \ref{fig:random-effect-high-low-precision} is clear that if each included study has a considerable sample size the total variability will be reduced, mainly driven by real heterogeneity (\(\tau^2\)). This is the crucial difference between the fixed-effects and the random-effects model (see also Borenstein et al., 2009, pp. 117--122).

Given the interpretation of \(I^{2}\) it is possible to simulate a meta-analysis fixing a certain \(I^{2}\) value. The only caveat is fixing the \(\tilde{v}\). When the sample size of each study is the same, \(\tilde{v} = \sigma^2_i\). In the other case, \(\tilde{v}\) needs to be calculated from sampling variances reported in Higgins and Thompson (2002) and cannot be easily fixed a priori. However, with the assumption of homogeneous sample size across studies\footnote{The homogeneous sample size of primary studies assumption is commonly used to calculate the statistical power (see Borenstein et al., 2009, Chapter 29)}, we can solve Equation \eqref{eq:isquared} for \(\tau^{2}\) obtaining the heterogeneity value associated with a certain \(I^{2}\) as reported in Equation \eqref{eq:tau2-from-isquared}. Table \ref{tab:res-random-effect-i2} depicts the results of the random-effects model fixing the \(I^{2}\) value.

\begin{align}
\begin{gathered}
\tau^2 = - \frac{I^2\tilde{v}}{I^2 - 1}
\label{eq:tau2-from-isquared}
\end{gathered}
\end{align}

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(seed)}
\NormalTok{k }\OtherTok{\textless{}{-}} \DecValTok{30} \CommentTok{\# number of studies}
\NormalTok{theta\_r }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# real effect size}
\NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{30} \CommentTok{\# sample size per group, per study}
\NormalTok{I2 }\OtherTok{\textless{}{-}} \FloatTok{0.6} \CommentTok{\# desired I2 value}

\NormalTok{v }\OtherTok{\textless{}{-}}\NormalTok{ (n }\SpecialCharTok{+}\NormalTok{ n)}\SpecialCharTok{/}\NormalTok{(n }\SpecialCharTok{*}\NormalTok{ n) }\SpecialCharTok{+}\NormalTok{ theta\_r}\SpecialCharTok{\^{}}\DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\DecValTok{2} \SpecialCharTok{*}\NormalTok{ (n }\SpecialCharTok{+}\NormalTok{ n }\SpecialCharTok{{-}} \DecValTok{2}\NormalTok{)) }\CommentTok{\# typical within study variance}
\NormalTok{tau2 }\OtherTok{\textless{}{-}} \SpecialCharTok{{-}}\NormalTok{((I2}\SpecialCharTok{*}\NormalTok{v)}\SpecialCharTok{/}\NormalTok{(I2 }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{))}

\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{theta\_r =}\NormalTok{ theta\_r)}

\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2))}
\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_r\_theta\_i }\OtherTok{\textless{}{-}}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_r }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_i }
\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{theta\_r\_theta\_i, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim)}
\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(yi, vi, }\AttributeTok{method =} \StringTok{"REML"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
\end{Highlighting}
\end{Shaded}

\normalsize

\scriptsize

\begin{table}[H]

\caption{\label{tab:res-random-effect-i2}Summary of the \emph{random-effect} model fixing the \(I^2\) value. Parameters are the same as described in Table \ref{tab:res-random-effect}.}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{ccccc}
\toprule
 & $\beta$ & 95\% CI & z & p\\
\midrule
overall & 0.298 (SE = 0.071) & $[0.159, 0.438]$ & 4.191 & < 0.001\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} $k = 30$}\\
\multicolumn{5}{l}{\textsuperscript{} $\tau^2 = 0.083$ ($SE = 0.040$)}\\
\multicolumn{5}{l}{\textsuperscript{} $I^2 = 54.830\%$}\\
\end{tabular}
\end{table}

\normalsize

\hypertarget{metareg}{%
\subsection{Meta-regression}\label{metareg}}

From a linear regression perspective, both the FE and the RE models can be seen as \emph{intercept-only} models where only the mean (i.e., the linear regression intercept) is estimated Borenstein et al.~(2009). As reported in the meta-analysis introduction, the between-study heterogeneity usually represents the true variability of the effect due to differences among primary studies. A natural extension of the \emph{intercept-only} analysis is a model that includes variables (i.e., \emph{moderators}) that could explain the observed heterogeneity among effect sizes. For example, a group of studies could use a particular memory task where the expected effect is higher than another. In this way, considering the
type of task will explain part of the observed heterogeneity as in standard regression models. Figures \ref{fig:img-metaregression-bin} and \ref{fig:img-metaregression-num} depict a random-effects meta-regression model for a categorical and numerical predictor.

\scriptsize

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/img-metaregression-bin-1}

}

\caption{Graphical representation of a \emph{random-effects meta-regression} model with a categorical predictor (Condition A and Condition B). Each grey distribution represents the sampling distribution of included studies. The dotted line is the average effect (i.e., random-effects model without moderators). The effect size differs between conditions A and B, including the \emph{condition} moderator explaining part of the total heterogeneity (red plus green segments). The green segments depict the explained heterogeneity, and the red segments the residual (unexplained) heterogeneity.}\label{fig:img-metaregression-bin}
\end{figure}

\normalsize

\scriptsize

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/img-metaregression-num-1}

}

\caption{Graphical representation of a \emph{random-effects meta-regression} model with a numerical predictor (\(x\)). Each grey distribution represents the sampling distribution of included studies. The dotted line is the average effect (i.e., random-effects model without moderators). The effect size increases as a function of the \(x\) variable. Therefore, including \(x\) as a predictor explains the
heterogeneity. The rest of the Figure follows the same logic as Figure \ref{fig:img-metaregression-bin}.}\label{fig:img-metaregression-num}
\end{figure}

\normalsize

\hypertarget{meta-regression-with-a-categorical-moderator}{%
\subsubsection{Meta-regression with a categorical moderator}\label{meta-regression-with-a-categorical-moderator}}

A common example of meta-regression is by including a categorical predictor, including information about study-level features. In our example, a group of studies uses an online memory task while others use a standard lab-based task. The Equation \eqref{eq:random-effect-model} can be easily extended for a meta-regression model by including a variable encoding the type of task (online vs.~lab-based) and the expected difference between the two levels of the moderator (i.e., the \emph{lab vs.~online effect}). In regression terms, we could use a dummy variable (\(X_{1}\)) that takes the value of 0 for the \emph{lab-based task} (\emph{l}) and a value of 1 for an \emph{online task} (\emph{o})\footnote{The \emph{dummy coding} (also known as treatment coding) is the default in R and \texttt{metafor} but other coding schemes could be used (see Schad, Vasishth, Hohenstein, \& Kliegl, 2020 for an overview of contrast coding schemes)}. Now we fix \(\beta_{1}\) to be the \emph{lab vs.~online effect} Equation \eqref{eq:random-effect-model} into Equation \eqref{eq:meta-regression}. Now the product between \(\beta_{1}X_{1}\) will consider the \emph{lab vs.~online effect}. To simulate this meta-regression, we need to fix the \(\beta_{1}\) effect. Due to the fact that we are on a standardized scale, we can fix \(\beta_{1}\) to be the expected difference in Cohen's \(d\) scale between the two groups of studies. Crucially, despite \(\tau^{2}\) is still the heterogeneity between effect sizes, now we need to fix \(\tau^{2}\) considering that we included a moderator. In other terms, \(\tau_{r}^{2}\) is the residual heterogeneity after explaining the
variance due to \(X_{1}\). We can describe our model using two equations according to the value of \(X_{1}\) as reported in Equation \eqref{eq:meta-regression-equations}.

\begin{align}
\begin{gathered}
d_i = \beta_0 + \theta_i + \beta_1 X_1 + \epsilon_i \\
\theta_i \sim \mathcal{N}(0, \tau^2) \\
\epsilon_i \sim \mathcal{N}(0,\sigma^2_i)
\label{eq:meta-regression}
\end{gathered}
\end{align}

\begin{align}
\begin{gathered}
d_i = \beta_0 + \theta_i + \beta_1 X_1 + \epsilon_i \\
d_{il} = \beta_0 + \theta_i + \beta_1 \times 0 + \epsilon_i \\
d_{io} = \beta_0 + \theta_i + \beta_1 \times 1 + \epsilon_i \\
\label{eq:meta-regression-equations}
\end{gathered}
\end{align}

We can simulate the same scenario of the random-effects model with \(k_o = 15\) (\emph{online tasks}) and \(k_l = 15\) (\emph{lab-based tasks}). Then we fix the \(\beta_1 = 0.2\) and \(\tau_r^2 = 0.1\) (\(r\) for residual).

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(seed)}
\NormalTok{k }\OtherTok{\textless{}{-}} \DecValTok{30} \CommentTok{\# the total number of studies}
\NormalTok{b0 }\OtherTok{\textless{}{-}} \FloatTok{0.1} \CommentTok{\# intercept, the effect size of the lab{-}based studies}
\NormalTok{b1 }\OtherTok{\textless{}{-}} \FloatTok{0.2} \CommentTok{\# the difference between the two levels of the moderator}
\NormalTok{tau2r }\OtherTok{\textless{}{-}} \FloatTok{0.1} \CommentTok{\# the residual heterogeneity}
\NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{30} \CommentTok{\# the sample size per group, per study}

\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{exp =} \FunctionTok{rep}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"lab"}\NormalTok{, }\StringTok{"online"}\NormalTok{), }\AttributeTok{each =}\NormalTok{ k}\SpecialCharTok{/}\DecValTok{2}\NormalTok{))}

\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2r)) }\CommentTok{\# the by{-}study residual adjustment}
\NormalTok{sim}\SpecialCharTok{$}\NormalTok{b0\_theta\_i\_x }\OtherTok{\textless{}{-}}\NormalTok{ b0 }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_i }\SpecialCharTok{+}\NormalTok{ b1}\SpecialCharTok{*}\FunctionTok{ifelse}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{exp }\SpecialCharTok{==} \StringTok{"lab"}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)}

\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{b0\_theta\_i\_x, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim)}
\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(yi, vi, }\AttributeTok{mods =} \SpecialCharTok{\textasciitilde{}}\NormalTok{exp, }\AttributeTok{method =} \StringTok{"REML"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
\end{Highlighting}
\end{Shaded}

\normalsize

Now we can fit the meta-regression model with the \texttt{rma} function as for the random-effects model with the addition of \texttt{mods\ =\ \textasciitilde{}\ exp} that indicates which variable/s to consider as moderators. The results are presented in Table \ref{tab:res-meta-reg-dummy}. Now the model will estimate an \emph{intercept} parameter (i.e., \(\beta_0\)) that is the value of \(y\) when \(X_1\) is zero (i.e., for lab-based studies) or in other terms the expected value for lab-based studies. Then the \(\beta_1\) parameter represents the estimated difference in \(y\) between the values of \(X_1\) (i.e., lab-based vs online experiments). As said before, \(\tau^2\) is now the residual heterogeneity that is interpreted as the variability between effect sizes after controlling for the moderator \(X_1\).

Now we can fit the meta-regression model with the \texttt{rma} function as for the random-effects model with the addition of mods = \texttt{\textasciitilde{}\ exp} that indicates which variable/s to consider as moderators. The results are presented in Table \ref{tab:res-meta-reg-dummy}. Next, the model will estimate an \emph{intercept} parameter (i.e., \(\beta_{0}\)) that is the value of \(y\) when \(X_{1}\) is zero (i.e., for lab-based studies) or, in other terms, the expected value for lab-based studies. Then the \(\beta_{1}\) parameter represents the estimated difference in \(y\) between the values of \(X_{1}\) (i.e., lab-based vs.~online experiments). As mentioned above, \(\tau^{2}\) the residual heterogeneity is now interpreted as the variability between effect sizes after controlling for the moderator \(X_{1}\).

\scriptsize

\begin{table}[H]

\caption{\label{tab:res-meta-reg-dummy}Summary of the \emph{random-effect} model with a categorical predictor (lab vs online experiments). The \texttt{intercept} is the average effect for lab-based experiments and \texttt{exponline} is the difference between lab-based and online experiments. The \(R^2\) is the percentage of explained heterogeneity and \(\tau^2_r\) is the estimated residual heterogeneity. Other parameters are the same as the standard \emph{random-effect} model (see Table \ref{tab:res-random-effect}).}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{ccccc}
\toprule
 & $\beta$ & 95\% CI & z & p\\
\midrule
intercept & 0.041 (SE = 0.101) & $[-0.157, 0.238]$ & 0.403 & 0.687\\
exponline & 0.321 (SE = 0.143) & $[0.040, 0.601]$ & 2.243 & 0.025\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} $k = 30$}\\
\multicolumn{5}{l}{\textsuperscript{} $\tau^2_r = 0.085$ ($SE = 0.041$)}\\
\multicolumn{5}{l}{\textsuperscript{} $R^2 = 19.902\%$}\\
\multicolumn{5}{l}{\textsuperscript{} $I^2 = 55.435\%$}\\
\end{tabular}
\end{table}

\normalsize

\hypertarget{meta-regression-with-a-numerical-moderator}{%
\subsubsection{Meta-regression with a numerical moderator}\label{meta-regression-with-a-numerical-moderator}}

The same approach can be used for a continuous predictor. For example, we can simulate that the average participant's age within each study could explain part of the observed heterogeneity. Now the \(X_1\) is a continuous predictor representing the average age for each study and \(\beta_1\) is the effect size increase for a unit increase (i.e., 1 year) of average age. Sometimes guessing a plausible \(\beta_1\) value with a continuous predictor is not straightforward. A first strategy could be to use values estimated from the literature. Another approach consists in setting up the model and simulating several expected \(d_i\) and calculating the range of simulated values. A third possibility is fixing the proportion of explained heterogeneity calculating the \(\beta_1\) value accordingly. As in standard regression analysis, we can use the \(R^2\) statistic to describe the amount of heterogeneity explained by the included moderators. Equation \eqref{eq:rsquared} reports how to calculate the \(R^2\) for a meta-regression model. The \(\tau^2_{r}\) is the residual heterogeneity after considering the moderators and \(\tau^2_{f}\) is the heterogeneity estimated without considering the moderators. In the next sections, we will present an example of the simulation-based and the \(R^2\) based approaches.

\begin{align}
\begin{aligned}
R^2 = 1 - \frac{\tau^2_{r}}{\tau^2_{f}}
\label{eq:rsquared}
\end{aligned}
\end{align}

In terms of regression parameters now the intercept (\(\beta_0\)) is no longer the overall effect or the average of one category as in the previous example but the estimated value for a specific \(X_1\) thus for a specific age. If \(X_1\) is a variable representing the average age for each study, then the \(\beta_0\) is the average effect size when the age is zero. Depending on the moderator, the intercept is interpreted in different ways. For example, with the age, the intercept has no empirical meaning given that no studies could have a participant average age of zero. A strategy could be to mean-center the age (i.e., subtracting from each study age the average age across the study). Now the intercept is still the average effect size when the age is zero but now zero is the average age. Importantly, the contrast coding for categorical predictors or centering numerical variables does not affect the overall model estimation but only parameters values and interpretation.

\hypertarget{assessing-the-impact-of-beta_1}{%
\paragraph{\texorpdfstring{Assessing the impact of \(\beta_1\)}{Assessing the impact of \textbackslash beta\_1}}\label{assessing-the-impact-of-beta_1}}

As reported in the previous section, a strategy to guess plausible values for \(\beta_{1}\) is by simulating several expected \(y_{i}\) given the meta-regression equation and summarizing or plotting the effect size range. The range of simulated \(y_{i}\) values are also affected by the simulated age values across studies. However, it is probably more intuitive to guess a plausible range of moderator values compared to the \(\beta_{1}\) value. In this specific example, if all studies target a specific population (e.g., adults below 50 years), the expected average age range can be easily simulated. In our case, we simulated \(k\) average age values from a uniform distribution \(\text{ag}e_{i} \sim U\left( 20,40 \right)\). Then we can plot the distribution of \(d_{i}\) values to check the plausibility of simulated values. As shown in Figure \ref{fig:plot-meta-reg-plausible-implausible}, with the same range for the moderator, a \(\beta_{1} = 0.1\) gives a plausible effect sizes range while a \(\beta_{1} = 0.7\) predicts very extreme values.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(seed)}
\NormalTok{k }\OtherTok{\textless{}{-}} \DecValTok{1000} \CommentTok{\# number of studies}
\NormalTok{b0 }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# the intercept i.e., average yi when x is 0}
\NormalTok{b1 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.7}\NormalTok{) }\CommentTok{\# the beta1 i.e., the increase in yi for an increase in 1 year}
\NormalTok{tau2r }\OtherTok{\textless{}{-}} \FloatTok{0.1} \CommentTok{\# the residual heterogeneity after including x1}
\NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{30} \CommentTok{\# number of participants per study, per group}
\NormalTok{x1 }\OtherTok{\textless{}{-}} \FunctionTok{runif}\NormalTok{(k, }\DecValTok{20}\NormalTok{, }\DecValTok{40}\NormalTok{) }\CommentTok{\# random mean{-}age for each study}
\NormalTok{x10 }\OtherTok{\textless{}{-}}\NormalTok{ x1 }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(x1) }\CommentTok{\# centering the age}

\NormalTok{sim }\OtherTok{\textless{}{-}}\NormalTok{ tidyr}\SpecialCharTok{::}\FunctionTok{expand\_grid}\NormalTok{(}\AttributeTok{id =} \DecValTok{1}\SpecialCharTok{:}\NormalTok{k, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{b1 =}\NormalTok{ b1)}

\NormalTok{sim}\SpecialCharTok{$}\NormalTok{age }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(x1, }\AttributeTok{each =} \DecValTok{2}\NormalTok{)}
\NormalTok{sim}\SpecialCharTok{$}\NormalTok{age0 }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(x10, }\AttributeTok{each =} \DecValTok{2}\NormalTok{)}

\NormalTok{theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2r))}
\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(theta\_i, }\AttributeTok{each =} \DecValTok{2}\NormalTok{)}

\NormalTok{sim}\SpecialCharTok{$}\NormalTok{b0\_theta\_i\_x }\OtherTok{\textless{}{-}}\NormalTok{ b0 }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_i }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{b1}\SpecialCharTok{*}\NormalTok{sim}\SpecialCharTok{$}\NormalTok{age0}

\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{b0\_theta\_i\_x, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim)}
\end{Highlighting}
\end{Shaded}

\normalsize

\scriptsize

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/plot-meta-reg-plausible-implausible-1}

}

\caption{Scatter plots with marginal histograms for the range of simulated \(d_i\) with two \(\beta_1\) values. The x-axis depicts the average age of simulated studies and the y-axis depicts the simulated effect size. On the left, the majority of simulated values range between -1.5 and 1.5 thus \(\beta_1 = 0.1\) can be considered a plausible value. On the right, (\(\beta = 0.7\)) values range between -10 to 10 that despite theoretically possible is highly implausible values in real meta-analysis for psychological data.}\label{fig:plot-meta-reg-plausible-implausible}
\end{figure}

\normalsize

\hypertarget{simulating-using-r2}{%
\paragraph{\texorpdfstring{Simulating using \(R^2\)}{Simulating using R\^{}2}}\label{simulating-using-r2}}

A more intuitive way to simulate a continuous predictor is fixing the desired \(R^{2}\) value and finding the coefficient that produces the desired value. This approach has been implemented by Lopez-Lopez and colleagues (2014). We can use Equation \eqref{eq:beta-from-r2} to find the \(\beta_{1}\) value that is associated with a certain \(R^{2}\).

\begin{align}
\begin{aligned}
\beta^2_1 = \tau^2R^2 \\
\tau^2_{r} = \tau^2 - \beta^2_1
\label{eq:beta-from-r2}
\end{aligned}
\end{align}

Now we can simulate the regression model using \(\sqrt{\beta^2_1}\) as coefficient and \(\tau_{r}^{2}\) as residual heterogeneity. Results from the fitted model fixing the \(R^{2}\) values are presented in Table \ref{tab:res-meta-reg-numerical-r2} and Figure \ref{fig:plot-meta-reg-numerical-r2}. As Lopez-Lopez and colleagues (2014) demonstrated, to reliably estimate \(R^{2}\) the number of studies needs to be large\footnote{see also \url{https://www.metafor-project.org/doku.php/tips:ci_for_r2} for a discussion about confidence intervals for the \(R^2\) statistic}. Lopez-Lopez and colleagues (2014) generated the moderator (\(X_{1}\)) values from a standard normal distribution. In the following example, we standardized the moderator (\texttt{scale()}) after simulating values on the \emph{age} scale
(e.g., \texttt{runif(k,\ 20,\ 40)}).

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(seed)}
\NormalTok{k }\OtherTok{\textless{}{-}} \DecValTok{100} \CommentTok{\# the number of studies}
\NormalTok{r2 }\OtherTok{\textless{}{-}} \FloatTok{0.2} \CommentTok{\# the desired r2 value}
\NormalTok{tau2 }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# the overall tau2}
\NormalTok{b0 }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# the intercept i.e., average yi when x1 is 0}
\NormalTok{b1\_2 }\OtherTok{\textless{}{-}}\NormalTok{ tau2 }\SpecialCharTok{*}\NormalTok{ r2 }\CommentTok{\# the beta1\^{}2 i.e., the increase in yi for an increase in 1 year}
\NormalTok{b1 }\OtherTok{\textless{}{-}} \FunctionTok{sqrt}\NormalTok{(b1\_2) }\CommentTok{\# b1\_2 is squared, back to the original scale}
\NormalTok{tau2r }\OtherTok{\textless{}{-}}\NormalTok{ tau2 }\SpecialCharTok{{-}}\NormalTok{ b1\_2 }\CommentTok{\# the residual heterogeneity after including x1}
\NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{30} \CommentTok{\# number of participants per study, per group}
\NormalTok{x1 }\OtherTok{\textless{}{-}} \FunctionTok{runif}\NormalTok{(k, }\DecValTok{20}\NormalTok{, }\DecValTok{40}\NormalTok{) }\CommentTok{\# random mean{-}age for each study}

\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{age =}\NormalTok{ x1)}

\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2r))}
\NormalTok{sim}\SpecialCharTok{$}\NormalTok{age0 }\OtherTok{\textless{}{-}} \FunctionTok{scale}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{age, }\AttributeTok{center =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{scale =} \ConstantTok{TRUE}\NormalTok{) }\CommentTok{\# standardize the moderator}
\NormalTok{sim}\SpecialCharTok{$}\NormalTok{b0\_theta\_i\_x }\OtherTok{\textless{}{-}}\NormalTok{ b0 }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_i }\SpecialCharTok{+}\NormalTok{ b1}\SpecialCharTok{*}\NormalTok{sim}\SpecialCharTok{$}\NormalTok{age0}

\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{b0\_theta\_i\_x, sim}\SpecialCharTok{$}\NormalTok{nt, sim}\SpecialCharTok{$}\NormalTok{nc, }\AttributeTok{data =}\NormalTok{ sim)}
\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(yi, vi, }\AttributeTok{mods =} \SpecialCharTok{\textasciitilde{}}\NormalTok{age0, }\AttributeTok{method =} \StringTok{"REML"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
\end{Highlighting}
\end{Shaded}

\normalsize

\scriptsize

\begin{table}[H]

\caption{\label{tab:res-meta-reg-numerical-r2}Summary of the \emph{random-effects} model fixing the \(R^{2}\) value. The \texttt{intercept} is the effect size for the average \texttt{age} (given that age is mean-centered). The \texttt{age0} parameter is the slope between \texttt{age} and the effect size interpreted as an increase in effect size for a unit increase in the average age. Other parameters are the same as described in Table \ref{tab:res-random-effect} and \ref{tab:res-meta-reg-dummy}.}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{ccccc}
\toprule
 & $\beta$ & 95\% CI & z & p\\
\midrule
intercept & 0.335 (SE = 0.051) & $[0.236, 0.435]$ & 6.586 & < 0.001\\
age0 & 0.163 (SE = 0.051) & $[0.063, 0.263]$ & 3.184 & 0.001\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} $k = 100$}\\
\multicolumn{5}{l}{\textsuperscript{} $\tau^2_r = 0.189$ ($SE = 0.037$)}\\
\multicolumn{5}{l}{\textsuperscript{} $R^2 = 11.283\%$}\\
\multicolumn{5}{l}{\textsuperscript{} $I^2 = 73.038\%$}\\
\end{tabular}
\end{table}

\normalsize

\scriptsize

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/plot-meta-reg-numerical-r2-1}

}

\caption{Meta-regression results for the \emph{random-effects model} with a numerical moderator. Each effect size is represented with a black dot where the dimension represents the weight according to the inverse of the precision. The line represents the estimated meta-regression slope with the 95\% confidence interval (grey bands).}\label{fig:plot-meta-reg-numerical-r2}
\end{figure}

\normalsize

\hypertarget{power-analysis}{%
\section{Power Analysis}\label{power-analysis}}

The previous simulation examples can be easily implemented for multiple purposes. For example, we can use different effect sizes and variance estimators when using the \texttt{sim\_study()} function to check the impact on the fitted meta-analysis model. However, one of the most critical applications is estimating the power of a specific statistical model.

As explained in the introduction, there are several approaches and tools to estimate the power of \emph{fixed} and \emph{random-effects models} (see Borenstein et al., 2009, Chapter 29; Harrer, Cuijpers, \& Ebert, 2019, Chapter 14). These methods are easy to implement but made strong assumptions, such as the homogeneity of sample size, and did not consider the uncertainty in estimating \(\tau^{2}\). Jackson and colleagues (2017)
partially solve the issue by developing an interesting method that takes into account the uncertainty in estimating \(\tau^{2}\) without using simulations. However, only Monte Carlo simulations can consider complex scenarios and extra parameters.

A general Monte Carlo simulation for the power analysis can be implemented with the following steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Choose the model that generates the data (e.g., fixed, or random-effects model).
\item
  Fix the relevant parameters (e.g., \(\tau^{2}\) and \(\theta\)).
\item
  Simulate a dataset.
\item
  Fit the appropriate model.
\item
  Store the p-value associated with the parameter of interest
\item
  Repeat 3-5 a large number of times (e.g., 10000).
\item
  Calculate the power as the proportion of p-values below the \(\alpha\) level.
\end{enumerate}

For example, we can estimate the power of a \emph{random-effects model} by repeating the simulation presented in Section \ref{re-model} many times. We simulated heterogeneity of sample sizes sampling \(n_{T}\) and \(n_{C}\) values from a Poisson distribution with \(\lambda = 20\). In this way, on average, the sample size is 20 for primary studies with a certain amount of heterogeneity. Usually, it is more informative to simulate different scenarios according to the relevant parameters, such as sample sizes, number of studies, or heterogeneity. For example, we can estimate the power with a different number of studies \(k\). We define the \texttt{do\_sim()} function that, according to the input parameter, repeats the simulation a certain number of times (i.e., \texttt{nsim})\footnote{We are using the \texttt{purrr::pmap()} function that can be considered very similar to \texttt{mapply()} but less verbose. Compared to \texttt{mapply()}, the \texttt{sim\_grid} columns are directly passed to the \texttt{do\_sim()} function without specifying the order.}. Increasing the number of simulations will increase the power analysis estimation precision. Then the \texttt{summary\_sim()} function analyzes each simulation returning the relevant values. We repeat the simulation of the \emph{random-effects model} several times with different parameters.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]

\NormalTok{do\_sim }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(k, theta, tau2, navg, nmin, nsim, }\AttributeTok{alpha =} \FloatTok{0.05}\NormalTok{, }\AttributeTok{summary =} \ConstantTok{TRUE}\NormalTok{)\{}
  \CommentTok{\# preallocate for computation speed}
\NormalTok{  p }\OtherTok{\textless{}{-}} \FunctionTok{vector}\NormalTok{(}\AttributeTok{mode =} \StringTok{"numeric"}\NormalTok{, }\AttributeTok{length =}\NormalTok{ nsim)}

  \CommentTok{\# start the simulation loop}
  \ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{nsim)\{}
\NormalTok{    theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2))}
\NormalTok{    theta\_theta\_i }\OtherTok{\textless{}{-}}\NormalTok{ theta }\SpecialCharTok{+}\NormalTok{ theta\_i}
    \CommentTok{\# simulate sample size}
\NormalTok{    n }\OtherTok{\textless{}{-}}\NormalTok{ nmin }\SpecialCharTok{+} \FunctionTok{rpois}\NormalTok{(k, navg }\SpecialCharTok{{-}}\NormalTok{ nmin)}
    \CommentTok{\# simulate the studies}
\NormalTok{    sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(theta\_theta\_i, n, n, }\AttributeTok{data =} \ConstantTok{NULL}\NormalTok{)}
\NormalTok{    res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(yi, vi, }\AttributeTok{method =} \StringTok{"REML"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
\NormalTok{    p[i] }\OtherTok{\textless{}{-}}\NormalTok{ res}\SpecialCharTok{$}\NormalTok{pval }\CommentTok{\# store the p value}
\NormalTok{  \}}
  \ControlFlowTok{if}\NormalTok{(summary)\{}
    \CommentTok{\# return directly the power}
    \FunctionTok{summary\_sim}\NormalTok{(p, alpha)}
\NormalTok{  \}}\ControlFlowTok{else}\NormalTok{\{}
    \CommentTok{\# return the list of pvalues}
    \FunctionTok{data.frame}\NormalTok{(p)}
\NormalTok{  \}}
\NormalTok{\}}

\NormalTok{summary\_sim }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(p, alpha)\{}
\NormalTok{  power }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(p }\SpecialCharTok{\textless{}=}\NormalTok{ alpha) }\CommentTok{\# compute power}
  \FunctionTok{data.frame}\NormalTok{(power)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\normalsize

Simulation results are presented in Figure \ref{fig:plot-power-analysis} and Table \ref{tab:tab-power-analysis-example} showing that to reach 80\% power (usually considered an appropriate level) with \(\alpha = 0.05\) we need \textasciitilde35 studies. The same approach could be used to estimate the power of a meta-regression by simply modifying the \texttt{do\_sim()} function simulating the effect of a moderator and extracting the relevant p-value.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(seed)}
\NormalTok{nsim }\OtherTok{\textless{}{-}} \DecValTok{5000} \CommentTok{\# number of simulations per condition (5000, higher is better)}
\NormalTok{k }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{15}\NormalTok{, }\DecValTok{25}\NormalTok{, }\DecValTok{35}\NormalTok{, }\DecValTok{50}\NormalTok{) }\CommentTok{\# number of studies}
\NormalTok{theta }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# the average effect size}
\NormalTok{tau2 }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# the heterogeneity}
\NormalTok{navg }\OtherTok{\textless{}{-}} \DecValTok{20} \CommentTok{\# average sample size per study}
\NormalTok{nmin }\OtherTok{\textless{}{-}} \DecValTok{10} \CommentTok{\# minimum sample size per study}
\NormalTok{alpha }\OtherTok{\textless{}{-}} \FloatTok{0.05} \CommentTok{\# the alpha level}

\CommentTok{\# creating all combinations}
\NormalTok{sim\_grid }\OtherTok{\textless{}{-}}\NormalTok{ tidyr}\SpecialCharTok{::}\FunctionTok{expand\_grid}\NormalTok{(k, theta, tau2, navg, nmin, nsim)}

\CommentTok{\# apply the simulation to all combinations}
\NormalTok{res }\OtherTok{\textless{}{-}}\NormalTok{ purrr}\SpecialCharTok{::}\FunctionTok{pmap}\NormalTok{(sim\_grid, do\_sim)}

\CommentTok{\# combine the results}
\NormalTok{res }\OtherTok{\textless{}{-}}\NormalTok{ dplyr}\SpecialCharTok{::}\FunctionTok{bind\_rows}\NormalTok{(res)}
\NormalTok{sim\_grid }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(sim\_grid, res)}
\end{Highlighting}
\end{Shaded}

\normalsize

\scriptsize

\begin{table}[H]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:tab-power-analysis-example}The results from the power analysis simulation. The table depicts the simulation parameters, the estimated power using the \texttt{summary\_sim()} function, and the average sample size (\(n\)) across the simulations.}

\small{

\begin{tabular}{lllllll}
\toprule
k & \multicolumn{1}{c}{$\theta$} & \multicolumn{1}{c}{$\tau^2$} & \multicolumn{1}{c}{$n_{avg}$} & \multicolumn{1}{c}{$n_{min}$} & \multicolumn{1}{c}{nsim} & \multicolumn{1}{c}{Power}\\
\midrule
5 & 0.30 & 0.30 & 20 & 10 & 5000 & 0.25\\
15 & 0.30 & 0.30 & 20 & 10 & 5000 & 0.46\\
25 & 0.30 & 0.30 & 20 & 10 & 5000 & 0.65\\
35 & 0.30 & 0.30 & 20 & 10 & 5000 & 0.79\\
50 & 0.30 & 0.30 & 20 & 10 & 5000 & 0.92\\
\bottomrule
\end{tabular}

}

\end{threeparttable}
\end{center}

\end{table}

\normalsize

\scriptsize

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/plot-power-analysis-1}

}

\caption{Results from the \emph{random-effects model} power analysis. The x-axis depicts the number of studies (\(k\)), and the y-axis the estimated power. The red dotted line is the 80\% power level, usually considered a good value for power analysis.}\label{fig:plot-power-analysis}
\end{figure}

\normalsize

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

The present work introduced the basic concepts of the meta-analysis regarding fixed, random-effects models and meta-regression with a simulation-based approach. We believe the presented examples can help implement alternative or more complex models. For example, the \texttt{sim\_study()} function can be easily modified to simulate another effect sizes index, such as a paired sample Cohen's \(d\) or a Pearson's
correlation. In addition, more complex models, such as \emph{multivariate} or \emph{multilevel} models, can be implemented following the same approach (see supplementary materials). For example, the \emph{three-level} model estimates another heterogeneity component representing the variability of multiple effect sizes in the same study. Similarly, the \emph{multivariate} model could include the correlation between multiple outcomes and the correlation between sampling errors.

The present work did have a few limitations. First, we only introduced basic concepts about meta-analysis and Monte Carlo simulations, while setting up complex simulations requires more knowledge and complexity of the simulation setup. We decided to give the foundations to understand meta-analyses with a simulation approach because more complex models are still based on the same principles. Second, there are limitations
concerning simulating participant-level data. We decided to simulate the meta-analysis dataset starting from the participants' level to maximize the flexibility and clearness of each step. The downside concerns the efficiency and scalability of the simulation setup. For large-scale simulations (e.g., many conditions, iterations, or complex models), simulating from aggregated statistics is probably more efficient (see
Heuvel et al., 2020 for an example) to improve the computation speed\footnote{see \url{https://www.jepusto.com/simulating-correlated-smds/} for a very clear example of the participant-level vs aggregated data simulation}.

In conclusion, data simulation is a very powerful tool for each step of a data analysis process. Starting from the learning phase, where simulating data can be used to understand the statistical model in terms of assumptions and the data generation process, to the estimation of statistical power. Moreover, we believe that data simulation as part of a standard research workflow could improve the overall research quality. Data simulation requires understanding the statistical model, setting appropriate and reasoned parameters, and realizing how the chosen analysis method behaves across different scenarios.

\newpage

\hypertarget{acknowledgments}{%
\subsubsection*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{subsubsection}{Acknowledgments}

We thank the \href{https://stat.ethz.ch/mailman/listinfo/r-sig-meta-analysis}{R-sig-meta-analysis} mailing-list where suggestions and clarifications significantly improved our simulations approach and the R code.

\hypertarget{conflicts-of-interest}{%
\subsubsection*{Conflicts of Interest}\label{conflicts-of-interest}}
\addcontentsline{toc}{subsubsection}{Conflicts of Interest}

The author(s) declare that there were no conflicts of interest with respect to the authorship or the publication of this article.

\hypertarget{data-materials-and-online-resources}{%
\subsubsection*{Data, materials, and online resources}\label{data-materials-and-online-resources}}
\addcontentsline{toc}{subsubsection}{Data, materials, and online resources}

The code to reproduce simulations, figures and tables can be found on Open Science Framework (\url{https://osf.io/54djn/})

\hypertarget{supplemental-material}{%
\subsubsection*{Supplemental Material}\label{supplemental-material}}
\addcontentsline{toc}{subsubsection}{Supplemental Material}

\hypertarget{prior-versions}{%
\subsubsection*{Prior versions}\label{prior-versions}}
\addcontentsline{toc}{subsubsection}{Prior versions}

The manuscript preprint has been uploaded on PsyArXiv \url{https://psyarxiv.com/br6vy/}

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Borenstein2009-mo}{}}%
Borenstein, M., Hedges, L. V., Higgins, J. P. T., \& Rothstein, H. R. (2009). \emph{Introduction to {Meta-Analysis}}. \url{https://doi.org/10.1002/9780470743386}

\leavevmode\vadjust pre{\hypertarget{ref-Cheung2014-fg}{}}%
Cheung, M. W.-L. (2014). Modeling dependent effect sizes with three-level meta-analyses: A structural equation modeling approach. \emph{Psychol. Methods}, \emph{19}(2), 211--229. \url{https://doi.org/10.1037/a0032968}

\leavevmode\vadjust pre{\hypertarget{ref-Cheung2015-bw}{}}%
Cheung, M. W.-L. (2015). \emph{{Meta-Analysis}: A structural equation modeling approach}. John Wiley \& Sons.

\leavevmode\vadjust pre{\hypertarget{ref-Cheung2019-po}{}}%
Cheung, M. W.-L. (2019). A guide to conducting a {Meta-Analysis} with {Non-Independent} effect sizes. \emph{Neuropsychol. Rev.}, \emph{29}(4), 387--396. \url{https://doi.org/10.1007/s11065-019-09415-6}

\leavevmode\vadjust pre{\hypertarget{ref-Cohen1988-ad}{}}%
Cohen, J. (1988). \emph{Statistical power analysis for the behavioral sciences} (2nd ed.). Routledge. \url{https://doi.org/10.4324/9780203771587}

\leavevmode\vadjust pre{\hypertarget{ref-DeBruine2021-id}{}}%
DeBruine, L. M., \& Barr, D. J. (2021). Understanding {Mixed-Effects} models through data simulation. \emph{Advances in Methods and Practices in Psychological Science}, \emph{4}(1), 2515245920965119. \url{https://doi.org/10.1177/2515245920965119}

\leavevmode\vadjust pre{\hypertarget{ref-Gelman2006-pc}{}}%
Gelman, A., \& Hill, J. (2006). \emph{Data analysis using regression and {Multilevel/Hierarchical} models}. Cambridge University Press. \url{https://doi.org/10.1017/CBO9780511790942}

\leavevmode\vadjust pre{\hypertarget{ref-Gelman2020-tg}{}}%
Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other stories}. Cambridge University Press. \url{https://doi.org/10.1017/9781139161879}

\leavevmode\vadjust pre{\hypertarget{ref-Gentle2009-cj}{}}%
Gentle, J. E. (2009). Monte carlo methods for statistical inference. In J. E. Gentle (Ed.), \emph{Computational statistics} (pp. 417--433). New York, NY: Springer New York. \url{https://doi.org/10.1007/978-0-387-98144-4/_11}

\leavevmode\vadjust pre{\hypertarget{ref-Harrer2019-rl}{}}%
Harrer, M., Cuijpers, P., \& Ebert, D. (2019). \emph{Doing {Meta-Analysis} in {R}}. \url{https://doi.org/10.5281/zenodo.2551803}

\leavevmode\vadjust pre{\hypertarget{ref-Harrer2021-bz}{}}%
Harrer, M., Cuijpers, P., Furukawa, T. A., \& Ebert, D. D. (2021). \emph{Doing {Meta-Analysis} with r: A {Hands-On} guide}. CRC Press.

\leavevmode\vadjust pre{\hypertarget{ref-Hedges1981-za}{}}%
Hedges, L. V. (1981). Distribution theory for glass's estimator of effect size and related estimators. \emph{J. Educ. Behav. Stat.}, \emph{6}(2), 107--128. \url{https://doi.org/10.3102/10769986006002107}

\leavevmode\vadjust pre{\hypertarget{ref-Higgins2002-fh}{}}%
Higgins, J. P. T., \& Thompson, S. G. (2002). Quantifying heterogeneity in a meta-analysis. \emph{Stat. Med.}, \emph{21}(11), 1539--1558. \url{https://doi.org/10.1002/sim.1186}

\leavevmode\vadjust pre{\hypertarget{ref-Ingalls2011-xt}{}}%
Ingalls, R. G. (2011). Introduction to simulation. \emph{Proceedings of the 2011 Winter Simulation Conference ({WSC})}, 1374--1388. ieeexplore.ieee.org. \url{https://doi.org/10.1109/WSC.2011.6147858}

\leavevmode\vadjust pre{\hypertarget{ref-Jackson2017-dv}{}}%
Jackson, D., \& Turner, R. (2017). Power analysis for random-effects meta-analysis. \emph{Res. Synth. Methods}, \emph{8}(3), 290--302. \url{https://doi.org/10.1002/jrsm.1240}

\leavevmode\vadjust pre{\hypertarget{ref-Lakens2013-eq}{}}%
Lakens, D. (2013). Calculating and reporting effect sizes to facilitate cumulative science: A practical primer for t-tests and {ANOVAs}. \emph{Front. Psychol.}, \emph{4}, 863. \url{https://doi.org/10.3389/fpsyg.2013.00863}

\leavevmode\vadjust pre{\hypertarget{ref-Lipsey2001-vp}{}}%
Lipsey, M. W., \& Wilson, D. B. (2001). Practical meta-analysis. \emph{Applied Social Research Methods Series; Vol 49.}, \emph{247}.

\leavevmode\vadjust pre{\hypertarget{ref-Lopez-Lopez2014-it}{}}%
López-López, J. A., Marı́n-Martı́nez, F., Sánchez-Meca, J., Van den Noortgate, W., \& Viechtbauer, W. (2014). Estimation of the predictive power of the model in mixed-effects meta-regression: A simulation study. \emph{Br. J. Math. Stat. Psychol.}, \emph{67}(1), 30--48. \url{https://doi.org/10.1111/bmsp.12002}

\leavevmode\vadjust pre{\hypertarget{ref-r-lang}{}}%
R Core Team. (2022). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\vadjust pre{\hypertarget{ref-Schad2020-ht}{}}%
Schad, D. J., Vasishth, S., Hohenstein, S., \& Kliegl, R. (2020). How to capitalize on a priori contrasts in linear (mixed) models: A tutorial. \emph{J. Mem. Lang.}, \emph{110}, 104038. \url{https://doi.org/10.1016/j.jml.2019.104038}

\leavevmode\vadjust pre{\hypertarget{ref-Schmid2022-xe}{}}%
Schmid, C. H., Stijnen, T., \& White, I. R. (2022). \emph{Handbook of {Meta-Analysis}}. Taylor \& Francis Limited.

\leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2010-xz}{}}%
Viechtbauer, W. (2010). Conducting meta-analyses in {R} with the {metafor} package. \url{https://doi.org/10.18637/jss.v036.i03}

\end{CSLReferences}

\end{document}

### 2023-03-28 documents/supplementary/supplementary.Rmd
adding files

---
title: "Understanding meta-analysis through data simulation with application to power analysis"
subtitle: "Supplementary materials"
bibliography: "../files/references.bib"
csl: "../files/apa.csl"
author:
  - Filippo Gambarota
  - Gianmarco Altoè
output:
  bookdown::pdf_document2:
    toc: true
    latex_engine: xelatex
header-includes:
  - \setcounter{table}{0} 
  - \renewcommand*{\thetable}{S\arabic{table}}
  - \setcounter{figure}{0} 
  - \renewcommand*{\thefigure}{S\arabic{figure}}
  - \setcounter{equation}{0}
  - \renewcommand{\theequation}{S\arabic{equation}}
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      eval = TRUE)
options(width = 60)
```

```{r packages, include = FALSE}
library(dplyr)
library(tidyr)
library(here)
devtools::load_all() # loading all custom functions
```

```{r funs, include=FALSE}
simfuns <- get_funs(here("R", "sim-utils.R"))
```

\newpage

# Introduction

In the main paper, we presented the basic meta-analysis models and simulations. However, in real-world meta-analyses, it is common to find complex data structures such as multilevel and multivariate. As reported in the paper, there could be a different source of dependencies when each study reports multiple effect sizes. Rephrasing the [@Cheung2015-bw, pp. 121--122] taxonomy, the first type of dependency, which we call *multilevel* dependency, arises when multiple independent experiments are associated with the same study. Level 1 (i.e., participants' level) is different, but the effects are likely correlated because the same research group conducted the experiments. The second type of dependency, *multivariate* dependency, arises when multiple effects are collected on the same pool of participants within each study. Now sampling errors within each study are correlated because the same participant is tested multiple times. Finally, we have a dependency on multiple effects at the population level. Two outcomes could be correlated (e.g., the type of
memory tasks) even if the pool of participants is different. The first two types of dependency are the reasons to use a multilevel and multivariate model. The following sections will show two examples to simulate these complex data structures.

# Simulations

```{r sim-setup}
library(dplyr)
library(tidyr)
seed <- 2023 # random seed for simulations
```

In the paper, we used the `mapply()` function, a particularly useful tool in R, to apply the same function (in parallel) with multiple parameters using multiple vectors. In our case, the `sim_study()` function has several parameters (`theta`, `nt`, among others), and we can apply the function multiple times with a combination of parameters. It can be considered essentially a `for` loop but more compact and less verbose.

The following code illustrates the idea of `mapply()` compared to a standard `for` loop. The sim data set contains our pool of $k$ studies, and we want to apply the `sim_study()` function to each dataset row. We can use a `for` loop that is verbose but more explicit or the `mapply()` function that is more compact but less explicit. The computation time and results `for` the two approaches are equivalent, given the appropriate setup.

```{r mapply-vs-for, eval = FALSE}
# using a for loop

# preallocate for speed (not necessary with mapply)
res <- vector(mode = "list", length = nrow(sim))
for(i in 1:length(res)){
  res[[i]] <- sim_study(theta = sim$theta[i], nc = sim$nc[i], nt = sim$nt[i])
}

# using mapply
sim <- mapply(sim_study, sim$theta, sim$nc, sim$nt)

# using sim_studies (with mapply inside)
sim <- sim_studies(sim$theta, sim$nc, sim$nt, data = sim)
```

# Three-level model

Compared to the standard two-level model, the three-level model estimates another source of heterogeneity ($\omega^{2}$) representing the variability within studies. We can easily extend the two-level *random-effects* equation into equation \@ref(eq:three-level-model) from the main paper adding another "adjustment" to the overall effect.

\begin{align}
\begin{gathered}
d_{ij} = \theta_r + \theta_i + \theta_{ij} + \epsilon_{ij} \\
\theta_i \sim N(0, \tau^2) \\
\theta_{ij} \sim N(0, \omega^2) \\
\epsilon_{ij} \sim N(0,\sigma^2_i)
(\#eq:three-level-model)
\end{gathered}
\end{align}

The notation suggests that each effect size belongs to an experiment (effect size) ($j$) nested within a study ($i$). The variability between studies (level 3) is handled by the $\tau^{2}$ parameter and the variability within studies (level 2) is handled by the $\omega^{2}$ parameter. Practically, each study is composed of the average effect size $\theta_{r}$ plus a study-specific adjustment determined by $\tau^{2}$. Then each effect size is composed of the study-specific
effect ($\theta_{r} + \theta_{i}$) plus the adjustment determined by $\omega^{2}$ thus $\theta_{r} + \theta_{i} + \theta_{\text{ij}}$. The number of effect sizes within each study will determine the estimation precision of the $\omega^{2}$ parameter and the number of studies will determine the estimation precision of the $\tau^{2}$ parameter. The
relationship between $\tau^{2}$ and $\omega^{2}$ can be expressed using the intraclass correlation coefficient (ICC) expressed as $ICC = \frac{\tau^{2}}{\tau^{2} + \omega^{2}}$ representing the proportion of total heterogeneity associated with level 2 or level 3. For example, an ICC of 0.5 suggests that 50% of heterogeneity is caused by between-studies variability. On the extreme, an ICC close to 1 suggests that the between-studies variability causes most heterogeneity
(i.e., the effect sizes within each study are very similar).

```{r sim-multilevel}
set.seed(seed)
i <- 50 # Number of studies (level 3)
j <- 5 # Number of effect sizes within each study (level 2)
tau2 <- 0.3 # heterogeneity between studies
omega2 <- 0.1 # heterogeneity within studies
icc <- tau2 / (tau2 + omega2) # real ICC
n <- 30 # sample size for each group, within each study
theta_r <- 0.3 # real average effect size

theta_i <- rnorm(i, 0, sqrt(tau2))
theta_ij <- rnorm(i*j, 0, sqrt(omega2))

sim <- tidyr::expand_grid(
  study = 1:i,
  effect = 1:j,
  nt = n,
  nc = n,
  theta_r = theta_r
)

sim$theta_i <- theta_i[sim$study]
sim$theta_ij <- theta_ij

head(sim)
```

Each study $i$ is repeated $j$ times^[To create a more realistic simulation, the number of effect sizes for each paper could be heterogeneous (e.g., sampled from a Poisson distribution)] along the $\theta_{i}$ while each effect has a unique $\theta_{\text{ij}}$. We can use the same approach as the main paper using the `sim_studies()` function.

```{r}
set.seed(seed)
sim$theta_r_i_ij <- sim$theta_r + sim$theta_i + sim$theta_ij
sim <- sim_studies(theta = sim$theta_r_i_ij, nt = sim$nt, nc = sim$nc, data = sim)
```

Finally we fit a *three-level models* using the `rma.mv` from the `metafor` package. The syntax differs slightly from the rma function and is clearly explained here https://www.metafor-project.org/doku.php/analyses:konstantopoulos2011. The essential part is adding the nested random effect using the `random = ` argument specifying that the `study` variable is nested within the `study` variable with `~ 1|study/experiment`.

```{r}
res <- rma.mv(yi, vi, random = ~1|study/effect, data = sim)
summary(res)
```

As explained above, the model estimates the average effect $\theta_{r}$ and the two heterogeneity parameters. To create a more realistic simulation, we can create some variability in the number of experiments within the same paper, similar to what we did for the number of participants in the main paper. We simulated a three-level model, but it is possible to extend to *n* levels for a more complicated nested structure. In this case, we need more heterogeneity parameters (i.e., more nested "adjustments"), but the overall simulation setup is the same.

# Multivariate model

Compared to a standard two-level random-effects model, the *multivariate model* must consider the correlation between effect sizes collected on the same pool of participants. Authors should report these correlations, but often we need to guess a plausible value for the meta-analysis. Compared to the *two-level* and *three-level* models, we now have multiple outcomes ($p$).

Following the example of the main paper, we have three memory tests collected on the same pool of participants. For this reason we have three real effects ($\theta_{r_{1}}$, $\theta_{r_{2}}$ and $\theta_{r_{3}}$), three heterogeneity ($\tau^2_1$, $\tau^2_2$ and $\tau^2_3$) and the population-level correlations between these outcomes ($\rho_{12}$, $\rho_{13}$ and $\rho_{23}$). Moreover, the sampling errors ($\epsilon_i$) are now correlated; thus each study will have three variances for each outcome ($\sigma^2_1$, $\sigma^2_2$, and $\sigma^2_3$), and the correlations between sampling errors ($\rho_{s12}$, $\rho_{s13}$ and $\rho_{s23}$)^[We used the subscript $s$ for sampling errors to distinguish between between-outcomes correlations and correlations between sampling errors]. Equation \@ref(eq:multivariate-model) formalize the multivariate model for a single study $i$ with three outcomes. The random-effects adjustments ($\theta_i$) to the overall effect and sampling errors are now sampled from a multivariate normal distribution, respectively $\theta_i \sim \mathcal{MVN}(0, \mathrm{T}^2)$ and $\theta_i \sim \mathcal{MVN}(0, \mathrm{V}^2)$.

\begin{align}
\begin{gathered}
\begin{bmatrix}
d_{i1} \\
d_{i2} \\
d_{i3}
\end{bmatrix} 
=
\begin{bmatrix}
\theta_{1} \\
\theta_{2} \\
\theta_{3}
\end{bmatrix}
+
\begin{bmatrix}
\theta_{i1} \\
\theta_{i2} \\
\theta_{i3}
\end{bmatrix}
+
\begin{bmatrix}
\epsilon_{i1} \\
\epsilon_{i2} \\
\epsilon_{i3}
\end{bmatrix} \\
\begin{bmatrix}
\theta_{i1} \\
\theta_{i2} \\
\theta_{i3}
\end{bmatrix}
\sim \mathcal{MVN}(0, \mathrm{T}^2) \\
\begin{bmatrix}
\epsilon_{i1} \\
\epsilon_{i2} \\
\epsilon_{i3}
\end{bmatrix}
\sim \mathcal{MVN}(0, \mathrm{V}^2) \\
\mathrm{T}^2 = \begin{bmatrix} 
\tau_1^2 & & & \\ 
\rho_{21}\tau_2\tau_1 & \tau_2^2 & & \\ 
\rho_{31}\tau_3\tau_1 & \rho_{32}\tau_3\tau_2 & \tau_3^2
\end{bmatrix} \\
\mathrm{V}^2 = \begin{bmatrix} 
\sigma_1^2 & & & \\ 
\rho_{s21}\sigma_2\sigma_1 & \sigma_2^2 & & \\ 
\rho_{s31}\sigma_3\sigma_1 & \rho_{s32}\sigma_3\sigma_2 & \sigma_3^2
\end{bmatrix} \\
(\#eq:multivariate-model)
\end{gathered}
\end{align}

We can use the `MASS::mvrnorm()` to generate data from a multivariate normal distribution in R. Compared to `rnorm()`, `MASS::mvrnorm()` requires the vector of means (in this case zero) and the variance-covariance matrix.

```{r sim-multivariate}
set.seed(seed)
p <- 3 # number of outcomes
taus2 <- c(0.3, 0.1, 0.2) # vector of taus
rho <- 0.6 # correlation between outcomes
Tcmat <- rho + diag(1 - rho, nrow = p) # correlation matrix
Tcmat

# correlation matrix to variance-covariance matrix
Tvcov <- diag(sqrt(taus2)) %*% Tcmat %*% diag(sqrt(taus2))
round(Tvcov, 3)

multi_sim <- MASS::mvrnorm(n = 100, mu = rep(0, p), Sigma = Tvcov)
head(multi_sim)

# check the result, close to the simulated values
apply(multi_sim, 2, mean) # mean
apply(multi_sim, 2, var) # variance
cor(multi_sim) # correlation
```

The previous code simulated a *compound symmetry* (CS) structure where all variances ($\tau^2$) and covariances ($\rho\tau_1\tau_2$) are the same. There could be different structures of the $T^2$ matrix that are clearly explained in the `metafor::rma.mv()` documentation (see https://wviechtb.github.io/metafor/reference/rma.mv.html#specifying-random-effects). For the sake of simplicity we will use the CS structure. We can use the `sim_T()` function that given the number of studies $k$, outcomes $p$, the vector of $\tau^2$ and the correlation $\rho$ generates the random-effect adjustments^[The function allow to specify heterogeneous $\tau^2$ values where the CS structure assume the same $\tau^2$ for each outcome. Using a fixed $\rho$ and different $\tau^2$ will create what `rma.mv()` calls a *heteroscedastic compound symmetric structure* thus a matrix where the correlation is the same between outcomes and each outcome has a different $\tau^2$ value].

```{r, echo = FALSE, results='asis'}
print_fun(simfuns$sim_T)
```

```{r}
set.seed(seed)
# compound symmetry
sim_T(k = 10, p = 3, taus2 = c(0.2, 0.2, 0.2), rho = 0.7)

# heteroscedastic compound symmetry
sim_T(k = 10, p = 3, taus2 = c(0.1, 0.3, 0.2), rho = 0.7)
```

For the sampling errors, the matrix $V^{2}$ for the study, $i$ is essentially a $p_{i} \times p_{i}$ matrix, where the diagonal contains the sampling variances, and off-diagonal elements are the covariances. Crucially, sampling errors are correlated within a study but uncorrelated between studies. For this reason, stacking all matrices
from $k$ studies will create a block variance-covariance matrix where the correlations between sampling errors of different studies are fixed to zero. Figure \@ref(fig:fig-block-vcov) depicts the idea of the block-variance covariance matrix for three studies with a different number of outcomes. The block variance-covariance matrix can be easily created using the `metafor::vcalc()` function.

```{r fig-block-vcov, echo=FALSE, fig.cap=get_caption(), eval = TRUE}
knitr::include_graphics("img/block-variance-covariance.pdf")
```

We have all pieces to simulate the multivariate model. We can use the `sim_study_m()` that follows the same idea of the `sim_study()` function but extends to multiple outcomes. We generate a single study with $p$ outcomes where sampling errors are correlated according to a compound symmetry structure^[Also for the $V^2$ matrix we can assume different variance-covariance structures. In this example we are using a CS structure]. In practical terms, we assume that the
correlation at the participant level between multiple outcomes is $r$ and fixed between studies.

```{r fun-sim-study-m, results='asis', echo = FALSE}
print_fun(simfuns$sim_study_m)
```

```{r}
set.seed(seed)
thetas <- c(0.1, 0, 0.7) # three real average outcomes
n <- 30 # number of participant per study
r <- 0.6 # correlation between sampling errors
sim_study_m(thetas, nt = n, nc = n, r = r)
```

As in the main paper, we can create a dataframe with simulation parameters and iterating the `sim_study_m()` producing a meta-analysis dataframe.

```{r}
set.seed(seed)
k <- 100 # number of studies
p <- 3 # number of outcomes per study
rho <- 0.7 # population level correlation between outcomes
r <- 0.5 # correlation between sampling errors
n <- 30 # number of participants per group per study

thetas_r <- c(0.1, 0, 1) # population level real effect sizes
taus2 <- c(0.3, 0.3, 0.3) # population level real tau2s

# creating the dataframe structure
sim <- expand_grid(
  id = 1:k,
  p = 1:p,
  nt = n,
  nc = n
)

head(sim)

# adding the real effects and the random-effect adjustments
sim$theta_r <- rep(thetas_r, k)
sim$theta_i <- sim_T(k, p, taus2, rho)

# average effect + random-effect adjustment
sim$theta_r_i <- sim$theta_r + sim$theta_i

# splitting to list of studies for improving the simulation speed
siml <- split(sim, sim$id)

# simulating the effect sizes for each study
resl <- lapply(siml, function(x) sim_study_m(x$theta_r_i, x$nt, x$nc, r))
res <- dplyr::bind_rows(resl) # combine everything
sim <- cbind(sim, res) # append to the sim dataframe

head(sim)
```

Now, we can use the `metafor::rma.mv()` function to fit the multivariate random-effects model. Compared to the multilevel model, the `metafor::rma.mv()` for a multivariate model uses the block variance-covariance matrix as the `vi` argument and the variable representing each outcome as a moderator estimating the average effect
size. We can create the block variance-covariance matrix. Given that we simulated the raw data, we can calculate and use the observed correlations (within the `sim_study_m()` function) and use a different correlation for each study within the `vcalc()` function (see [https://www.metafor-project.org/doku.php/analyses:berkey1998](https://www.metafor-project.org/doku.php/analyses:berkey1998https://www.metafor-project.org/doku.php/analyses:berkey1998) for an example). Often, these values are not reported; thus, we need to guess a plausible correlation to compute the block variance-covariance matrix. Here we use the $r$ value from our simulation and use the same for each study. We know the underlying model in this case, so our guessed value is correct.

```{r}
# create the block variance-covariance matrix
# cluster is the study (i.e. each "block")
# obs identify each outcome (i.e., each "block" number of rows/columns)
# r is the fixed sampling errors correlation
V <- vcalc(vi = vi, cluster = id, obs = p, rho = r, data = sim)

# first two studies (notice the off diagonal zeros for the correlation between different studies)
round(V, 3)[1:6, 1:6]

# outcome to character with a meaningful prefix
sim$p <- paste0("outcome", sim$p)

# fitting the model
res <- rma.mv(yi, V, mods = ~ 0 + p, random = ~p|id, data = sim, struct = "CS")
summary(res)
```

The output is more complicated compared to previous models. The `Variance Components` is the estimation of the $T^{2}$ matrix. In the two-level random-effects model, we have a single $\tau^{2}$ while in the multivariate model, we have a full variance-covariance matrix. Using the `struct = "CS"` argument, we are forcing the `rma.mv()` function to use a compound symmetry structure, thus estimating a single $\tau^{2}$ and a single $\rho$. Again, this is consistent with our simulation but could be a stringent assumption in a real-world analysis. The crucial part is the `random = ~ p|id` argument that specifies the random-effects structure. The syntax is clearly explained here [https://wviechtb.github.io/metafor/reference/rma.mv.html#specifying-random-effects](https://wviechtb.github.io/metafor/reference/rma.mv.html#specifying-random-effects) The basic idea is to specify the random effect structure as `~ inner|outer` where inner is the outcome variable and outer is the paper variable. The mods = `~ 0 + p` argument specify to estimate the average effect for each outcome (i.e., the $\theta_{r}$ vector of average effects). We are removing the intercept (`~ 0 + p` or equivalently `~ p - 1`) and the model will estimate the mean for each level of the `p` variable.

We could use also another structure for the $T^2$ matrix. We know that our data are generated using a CS structure thus using another structure will be less appropriate. In reality, is not known which is the real data generation model thus we can try different structure. Setting the `struct = "UN"` (unstructured) argument, `rma.mv` try to estimate a variance-covariance matrix where all correlations and variances need to be estimated. Compared to the previous model, we are know estimating 3 heterogeneity and three correlation parameters.

We could also use another structure for the $T^{2}$ matrix. We know that our data are generated using a CS structure; thus, using another structure will be less appropriate. In reality, it is not known which is the real data generation model; thus, we can try different structures. Setting the struct = "UN" (unstructured) argument, rma.mv tries to estimate a variance-covariance matrix where all correlations and variances need to be estimated. We are now estimating three
heterogeneity and three correlation parameters compared to the previous model.

```{r}
# fitting the model
res <- rma.mv(yi, V, mods = ~ 0 + p, random = ~p|id, data = sim, struct = "UN")
summary(res)
```

# Useful resources

- https://www.jepusto.com/simulating-correlated-smds/
- https://www.metafor-project.org/doku.php/metafor
- https://stat.ethz.ch/mailman/listinfo/r-sig-meta-analysis
- https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/

# References

<!-- CAPTIONS -->

(ref:fig-block-vcov) Example of a block variance-covariance matrix for three hypothetical studies. In the example, there are three outcomes ($p = 3$), but not all studies have all outcomes. For this reason, the matrix represents each study as a "block," and covariances between different studies are fixed to zero (because sampling errors are not correlated for different participants). The first study (pink) has three outcomes, thus a $3 \times 3$ matrix where the diagonal contains the sampling variances and off-diagonal elements are the sampling covariances. The second study has only one outcome; thus, the matrix is a single value for the sampling variance. The last study follows the same logic as the first study, with only two outcomes.

### 2023-03-28 R/sim-utils.R

@@ -1,0 +1,208 @@
+# Functions from the main paper -------------------------------------------
+
+#' sim_study
+#' @description Simulate a single study based on the input parameters and return a dataframe with effect size and variance or raw data.
+#' @param theta number indicating the population level effect size.
+#' @param nt number indicating the number of participants for the treated group.
+#' @param nc number indicating the number of participants for the control group. If \code{NULL} the value assigned to \code{nt} will be used.
+#' @param hedges logical. Whether the Hedges' correction should be applied. Default to \code{TRUE}.
+#' @param aggregate logical. Whether the participants data need to be aggregated calculating the effect size and variance. Default to \code{TRUE}.
+#'
+#' @return a dataframe
+#' @import metafor
+#' @export
+#' @examples
+#' sim_study(0.3, 30, 30, TRUE, TRUE) # return aggregated data
+#' sim_study(0.3, 30, 30, TRUE, TRUE) # return participants data
+
+sim_study <- function(theta, nt, nc = NULL, hedges = TRUE, aggregate = TRUE){
+  if(is.null(nc)) nc <- nt
+  # generate from normal distribution
+  yc <- rnorm(nc, 0, 1)
+  yt <- rnorm(nt, theta, 1)
+  # pooled variance
+  sp <- sqrt((var(yc)*(nc - 1) + var(yt)*(nt - 1)) / (nc + nt - 2))
+  # effect size
+  yi <- (mean(yt) - mean(yc)) / sp
+  
+  # transform to hedges' g
+  if(hedges){
+    yi <- yi * metafor:::.cmicalc(nc + nt - 2)
+  }
+  
+  # sampling variance
+  vi <- (nc + nt)/(nc * nt) + yi^2/(2 * (nc + nt - 2))
+  
+  if(!aggregate){
+    # return raw data
+    data.frame(id = 1:(nc + nt),
+               group = rep(c("c", "t"), c(nc, nt)),
+               y = c(yc, yt))
+  }else{
+    # compute effect size
+    data.frame(yi, vi)
+  }
+}
+
+#' sim_studies
+#' @description apply the \code{sim_study} function using a grid of parameters.
+#' @param ... arguments passed to \code{sim_study()}
+#' @param data a dataframe where the result will be attached (optional). Default to \code{NULL}
+#'
+#' @return a dataframe
[…]

### 2023-03-28 R/utils.R

@@ -1,0 +1,184 @@
+# Utils functions used in the paper and supplementary materials -----------
+
+#' svg_to_pdf
+#' @description convert all svg to pdf
+svg_to_pdf <- function(){
+  all_svg <- list.files(here::here("docs", "paper-draft", "img"), 
+                        pattern = ".svg", full.names = TRUE, recursive = TRUE)
+  out_names <- gsub(".svg", ".pdf", all_svg)
+  purrr::walk2(all_svg, out_names, rsvg::rsvg_pdf)
+}
+
+#' get_caption
+#' @description create text reference for captions
+get_caption <- function() {
+  chunk <- knitr::opts_current$get("label")
+  sprintf("(ref:%s)", chunk)
+}
+
+#' get_funs
+#' @description extract all functions as text from a \code{.R} file
+#' @param file path of the file
+#'
+#' @return named list
+#' @export
+#'
+get_funs <- function(file){
+  file <- suppressWarnings(readLines(file))
+  cutpoints <- grep("<- function", file)
+  cutpoints[length(cutpoints) + 1] <- length(file)
+  
+  out <- vector(mode = "list", length = length(cutpoints)-1)
+  fun_names <- vector(mode = "character", length = length(cutpoints)-1)
+  
+  for(i in 1:(length(cutpoints) - 1)){
+    if(i == length(cutpoints) - 1){
+      out[[i]] <- file[cutpoints[i]:(cutpoints[i + 1])]
+    }else{
+      out[[i]] <- file[cutpoints[i]:(cutpoints[i + 1] - 1)]
+    }
+    fun_names[i] <- stringr::str_extract(out[[i]][1], ".+?(?=<-)")
+  }
+  fun_names <- gsub(" ", "", fun_names)
+  names(out) <- fun_names
+  out <- lapply(out, function(x) x[!grepl("#'", x)]) # remove roxygen
+  return(out)
+}
+
+#' print_fun
+#'
+#' @param fun character vector with the function text
+#'
+#' @export
+#'
+print_fun <- function(fun){
+  cat("```r\n", fun, sep = "\n", "```\n")
+}
+
+#' rma_tidy
+#'
+#' @param res model fitted with \code{metafor::rma.uni()}
+#'
+#' @return a list
+#'
+rma_tidy <- function(res){
+  res_sum <- broom::tidy(res, conf.int = TRUE)
[…]

### 2023-03-28 scripts/r-img.R

@@ -1,0 +1,173 @@
+# Packages ----------------------------------------------------------------
+
+library(tidyverse)
+library(distributional)
+library(ggdist)
+library(here)
+library(latex2exp)
+library(ltxplot)
+
+ltxplot::load_theme_ltx()
+
+# Functions ---------------------------------------------------------------
+
+theme_rfig <- function(size = 15){
+  ltxplot::theme_latex(base_size = size)
+}
+
+# Fixed-vs-Random Effect --------------------------------------------------
+
+fixed_effect <- data.frame(
+  id = 1:4,
+  yi = rep(0.5, 4),
+  vi = c(0.01, 0.02, 0.1, 0.05),
+  model = "Fixed-Effects",
+  di = 0.5
+)
+
+fixed_effect$note <- sprintf("$d_i = \\theta_f + \\epsilon_i$")
+
+random_effect <- data.frame(
+  id = 1:4,
+  yi = c(-0.1, 0.4, 0.7, 0.25),
+  vi = c(0.01, 0.02, 0.08, 0.1),
+  model = "Random-Effects"
+)
+
+random_effect$note <- sprintf("$d_i = \\theta_r + theta_i + \\epsilon_i$")
+
+random_effect$di <- mean(random_effect$yi)
+
+models <- rbind(fixed_effect, random_effect)
+models$dist <- map2(models$yi, sqrt(models$vi), distributional::dist_normal)
+models$note <- latex2exp::TeX(models$note, output = "character")
+models$delta <- models$yi - models$di
+
+fixed_vs_random <- ggplot(models, aes(y = factor(id))) +
+  geom_vline(aes(xintercept = di), linetype = "dashed", alpha = 0.5) +
+  stat_dist_halfeye(aes(dist = dist), .width = 0.95) +
+  facet_wrap(~model, scales = "free_x") +
+  ylab("Study") +
+  xlab(TeX("$d_i$")) +
+  theme(axis.text.x = element_blank(),
+        axis.ticks = element_blank(),
+        axis.title.x = element_blank(),
+        strip.text = element_text(family = "lmroman")) +
+  theme_rfig() +
+  geom_label(aes(x = di, y = 0.6, label = note), parse = TRUE, size = 7, label.size = NA,
+             family = "lmroman") +
+  geom_segment(aes(x = di + delta, xend = di, y = id-0.08, yend = id-0.08),
[…]

### 2023-03-28 README.Rmd
updating README

@@
 
-- `R`: contains the simulation and utility functions used in the paper and supplementary materials
+- `R`: contains the simulation and utility functions used in the paper and the supplementary materials
 - `documents`: contains the `.Rmd` files to reproduce the paper and supplementary materials manuscripts

### 2023-06-27 documents/paper/paper.Rmd
fixing tab notation (rev)

@@
      ~`Equation`,     ~`Code`,                                               ~`Description`,
-       "$\\theta_f$",   "theta\\_f",                  "Fixed-effects model real effect size",
-       "$\\theta_r$",   "theta\\_r",                 "Random-effects model real effect size",
+       "$\\theta$",   "theta",                  "Fixed-effects model real effect size",
+       "$\\mu$",      "mu",                        "Random-effects model real effect size",
         "$\\tau^2$",    "tau2",                            "Effect size heterogeneity",
@@
   "$\\bar y_{T, C}$",        "",               "Mean of the experimental/control group",
-      "$s^2_{T, C}$",        "", "Standard deviation of the experimental/control group",
+      "$s_{T, C}$",        "", "Standard deviation of the experimental/control group",
         "$n_{T, C}$",  "nt, nc",        "Sample size of the experimental/control group",

### 2023-06-29 scripts/r-img.R
fixing random vs fixed figure

@@ -25,16 +25,16 @@ fixed_effect <- data.frame(
   di = 0.5
 )
 
-fixed_effect$note <- sprintf("$d_i = \\theta_f + \\epsilon_i$")
+fixed_effect$note <- sprintf("$y_i = \\theta + \\epsilon_i$")
 
 random_effect <- data.frame(
   id = 1:4,
   yi = c(-0.1, 0.4, 0.7, 0.25),
-  vi = c(0.01, 0.02, 0.08, 0.1),
+  vi = c(0.05, 0.08, 0.1, 0.2),
   model = "Random-Effects"
 )
 
-random_effect$note <- sprintf("$d_i = \\theta_r + theta_i + \\epsilon_i$")
+random_effect$note <- sprintf("$y_i = \\mu_{\\theta} + \\theta_i + \\epsilon_i$")
 
 random_effect$di <- mean(random_effect$yi)
 
@@ -43,7 +43,10 @@ models$dist <- map2(models$yi, sqrt(models$vi), distributional::dist_normal)
 models$note <- latex2exp::TeX(models$note, output = "character")
 models$delta <- models$yi - models$di
 
-fixed_vs_random <- ggplot(models, aes(y = factor(id))) +
+set.seed(2036)
+models$obs <- mapply(function(m, s) rnorm(1, m, s), models$yi, sqrt(models$vi))
+
+ggplot(models, aes(y = factor(id))) +
   geom_vline(aes(xintercept = di), linetype = "dashed", alpha = 0.5) +
   stat_dist_halfeye(aes(dist = dist), .width = 0.95) +
   facet_wrap(~model, scales = "free_x") +
@@ -58,10 +61,11 @@ fixed_vs_random <- ggplot(models, aes(y = factor(id))) +
              family = "lmroman") +
   geom_segment(aes(x = di + delta, xend = di, y = id-0.08, yend = id-0.08),
                position = position_dodge2(width = 1, padding = 1)) +
+  geom_point(aes(x = obs, y = factor(id)), color = "firebrick2", size = 2, shape = 15) +
   geom_point(aes(x = yi, y = id-0.08, alpha = model), show.legend = FALSE) +
   scale_alpha_manual(values = c(0, 1))
 
-# Metaregression with binary predictor ------------------------------------
+ # Metaregression with binary predictor ------------------------------------
 
 binary_metareg <- data.frame(
   id = 1:6,

### 2023-06-29 documents/paper/paper.Rmd
first notation fix and converting to UMD

@@
 The focus of the meta-analysis is collecting multiple studies with similar aims and methods and estimating the average effect of memory training.
-Despite differences in the type of cognitive task or experimental setup, each primary study collects an *experimental group* ($n_{C}$) and *control group* ($n_{C}$) and computes the average performance (*experimental group* ${\bar y}_{T}$ and *control group* ${\bar y}_{C}$) and standard deviations (*experimental group* $s_{T}$ and *control group*
+Despite differences in the type of cognitive task or experimental setup, each primary study collects an *experimental group* ($n_{C}$) and *control group* ($n_{C}$) and computes the average performance (*experimental group* $\overline T$ and *control group* $\overline C$) and standard deviations (*experimental group* $s_{T}$ and *control group*
 $s_{C}$).
@@
 
-The first step of a meta-analysis is to standardize information from each included study to convert each effect into a standard measure.
+The first step of a meta-analysis is to extract information from included studies.
 This common measure should give an immediate idea of the direction (i.e., the treatment improves or reduces performance) and the size of the effect.
 Standardized effect size measures [see @Lakens2013-eq for an overview], such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's $d$ [@Cohen1988-ad] or the Pearson correlation coefficient $\rho$ which could be calculated using the associated sample estimator *r* are commonly used to compare heterogeneous outcome variables.
-If all studies used the same raw measure, such as reaction times, it is possible to directly meta-analyze the studies without standardizing.
-In practice, studies often use heterogeneous measures, therefore requiring standardized measures.
+If all studies used the same raw measure, such as reaction times, it is possible to directly meta-analyze the studies without standardizing using (UMD, unstandardized mean difference).
 
@@
 We can divide meta-analysis models into two categories: the *fixed-effects* and the *random-effects model*.
-If we imagine combining multiple studies replicating the same experiment, we assume the real treatment effect $\theta_{f}$ is *fixed,* and each study is a more or less precise estimation of this parameter.
+If we imagine combining multiple studies replicating the same experiment, we assume the real treatment effect $\theta$ is *fixed,* and each study is a more or less precise estimation of this parameter.
 This is an essential and often inappropriate assumption because it ignores that study-level characteristics may impact the actual treatment effect.
@@
 In that case, the assumption of the *fixed-effects model* is no longer plausible.
-There is no unique real effect but a distribution of effects with mean $\theta_{r}$ and variance $\tau^{2}$.
+There is no unique real effect but a distribution of effects with mean $\mu_{\theta}$ and variance $\tau^{2}$.
 The mean is the *average* effect, and the variance is the heterogeneity of the phenomenon.
-The *random-effects model* will estimate the mean ($\theta_{r}$) and the heterogeneity $\tau^{2}$ will help generalize to the population of effects beyond the pool of included studies.
+The *random-effects model* will estimate the mean ($\mu_{\theta}$) and the heterogeneity $\tau^{2}$ will help generalize to the population of effects beyond the pool of included studies.
 
@@
 Before diving into the specific simulations, in this section, we define the common aspects of all simulations in the following sections.
-The current paper focuses on the *two-level*, *fixed,* and *random-effects model*s.
+The current paper focuses on the *two-level*, *fixed,* and *random-effects* models.
 All the examples refer to primary studies that assess the efficacy of a treatment by comparing a control and an experimental group.
-Equation \@ref(eq:datagen) formalizes the assumption about generating participants' data within a single study.
-The two groups (*experimental* $y_{T}$ and *control* $y_{C}$) are generated from normal distributions with $\mu = 0$ for the control group and $\mu = \theta$ (where $\theta$ is the effect size) for the experimental group.
-Both groups are assumed to be sampled from populations with homogeneous variances $\sigma^{2} = 1$.
+There are different simulation approaches to meta-analysis concerning how to generate study-level or participant-level data.
+In simulation studies where the purpose is not evaluating effect size estimators, it is convenient to simulate unstandardized effect size measures (e.g., UMD).
+Firstly the estimator is unbiased thus not requiring small sample correction [e.g., @Hedges1989-ip; @Hedges1981-za].
+Then the effect size and the sampling variance are independent.
+Similar to the simulation approach by Viechtbauer [@Viechtbauer2005-yh; @Viechtbauer2007-tq] the *experimental* group ($T$) and *control* groups $C$ are sampled from normal distributions respectively $T_i \sim \mathcal{N}(\Delta, 1)$ and $C_i \sim \mathcal{N}(0, 1)$ where $\Delta$ is the unstandardized mean difference (UMD).
+The UMD and the sampling variance are calculated using Equations \@ref(eq:effsize) and \@ref(eq:effsizevar) where $y$ is the sample estimator of $\Delta$.
 
@@
 \begin{gathered}
-y_{Ti} \sim \mathcal{N}(\theta, 1) \\
-y_{Ci} \sim \mathcal{N}(0, 1)
-(\#eq:datagen)
+y_i = \overline{T} - \overline{C}
+(\#eq:effsize)
 \end{gathered}
@@
 
-The true Standardized Mean Difference can be estimated using Cohen's $d$ as reported in Equation \@ref(eq:effsize).
-The difference between the two means is divided by the pooled standard deviation of the two groups.
-It is common to apply a small-sample correction version of this effect size called Hedges' $g$ [@Hedges1981-za]
-
 \begin{align}
 \begin{gathered}
-d = \frac{\bar y_T - \bar y_C}{S_p} \\
-S_p = \sqrt{\frac{S^2_{y_T}(n_{y_T} - 1) + S^2_{y_C}(n_{C} - 1)}{n_{T} + n_{C} - 2}} \\
-\sigma^2_d = \frac{n_{T} + n_{C}}{n_{T}n_{C}} + \frac{d^2}{2(n_{T} + n_{C} - 2)}
-(\#eq:effsize)
+\sigma_{\epsilon_i}^{2} = \frac{s^2_T}{n_T} + \frac{s^2_C}{n_C}
+(\#eq:effsizevar)
 \end{gathered}
@@
 
-1. Choose a $\theta$, $n_{T}$, and $n_{C}$ value.
+1. Choose a $\Delta$, $n_{T}$, and $n_{C}$ value.
 2.
-Simulate $n_{T}$ observations from a Gaussian distribution with $\mu = \theta$ and $\sigma^{2} = 1$ and $n_{T}$ observations from a Gaussian distribution with $\mu = 0$ and $\sigma^{2} = 1$.
-This way, the expected difference between groups will be $d$ and the variance is fixed at 1.
-3. Calculate the pooled standard deviation $S_{p}$ that is an estimation of the population standard deviation ($\sigma$) to standardize the mean difference.
-4. Calculate the observed effect size $d$ and the sampling variance $\sigma_{d}^{2}$.
+Simulate $n_{T}$ observations from a Gaussian distribution with $\mu = \Delta$ and $\sigma^{2} = 1$ and $n_{T}$ observations from a Gaussian distribution with $\mu = 0$ and $\sigma^{2} = 1$.
+This way, the expected difference between groups will be $\Delta$ and the variance is fixed at 1.
+3. Calculate the observed effect size $y$ and the sampling variance $\sigma_{\epsilon}^{2}$.
+
+The simulation approach can be easily extended calculating a standardized effect size measure [e.g., Cohens' $d$ @Cohen1988-ad] and the corresponding sampling variance.
+For example, after generating data for the two groups the mean difference can be standardized using the pooled standard deviation and applying the appropriate correction [e.g., @Hedges1989-ip; @Hedges1981-za].
 
@@
 
-The `theta` is the true effect size, `nc` and `nt` are the sample size for the control and experimental group, `aggregate` controls if returning the effect size and the corresponding variance or the participant-level data, and `hedges` specify whether applying the correction.
+The `es` is the true effect size ($\mu_{\theta}$ for the random-effects model and $\theta$ for the fixed-effect model), `nc` and `nt` are the sample size for the control and experimental group, `aggregate` controls if returning the effect size and the corresponding variance or the participant-level data.
 We can generate a single study with the desired parameters with this function.
@@
 set.seed(seed)
-sim_study(theta = 0.3, nc = 10000, nt = 10000, aggregate = FALSE)
+sim_study(es = 0.3, nc = 10000, nt = 10000, aggregate = FALSE)
 ```
@@
 set.seed(seed)
-sim <- sim_study(theta = 0.3, nc = 10000, nt = 10000, aggregate = FALSE) |> 
+sim <- sim_study(es = 0.3, nc = 10000, nt = 10000, aggregate = FALSE) |> 
   group_by(group) |> 
@@
 n <- 20 # number of participants per group, per study
-theta <- 0.3 # real effect size
-sim <- make_data(k = k, nt = n, nc = n, theta = theta)
+es <- 0.3 # real effect size
+sim <- make_data(k = k, nt = n, nc = n, es = es)
 ```
@@
              "$d_i$",      "yi",                                 "Observed effect size",
-      "$\\sigma^2_i$",      "vi",                          "Observed sampling variance",
+      "$$\\sigma_{\\epsilon_i}^{2}$$",      "vi",                          "Observed sampling variance",
        "$\\theta_i$", "theta\\_i",              "Random effect for the study i",
@@
 As reported in the previous sections, the *fixed-effects model* assumes the presence of a single true effect ($\theta_{f}$) and the observed variability is because each effect size is an imprecise estimation of the true effect.
-In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies
-(i.e., the sample size). Equation \@ref(eq:fixed-effect-model) depicts the *fixed-effect* equation.
+In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies (i.e., the sample size).
+Equations \@ref(eq:fixed-effect-model1) and \@ref(eq:fixed-effect-model2) depicts the *fixed-effects* model.
 
@@
 \begin{gathered}
-d_i = \theta_f + \epsilon_i \\
-\epsilon_i \sim \mathcal{N}(0,\sigma^2_i)
-(\#eq:fixed-effect-model)
+y_i = \theta + \epsilon_i
+(\#eq:fixed-effect-model1)
 \end{gathered}
@@
 
-Each observed effect size ($d_{i}$) is composed of the real effect size $\theta_{f}$ plus an error term ($\epsilon_{i}$) that is sampled from a normal distribution with $\mu = 0$ and $\sigma^{2} = \sigma_{i}^{2}$ (i.e., the known sampling variance of the study $i$).
-As demonstrated in
-Equation \@ref(eq:effsize), the increase in sample size will decrease the sampling variability.
-An exact study will essentially have $\theta_{f}$ as the observed effect size.
+\begin{align}
+\begin{gathered}
+\epsilon_i \sim \mathcal{N}(0,\sigma_{\epsilon_i}^{2})
+(\#eq:fixed-effect-model2)
+\end{gathered}
+\end{align}
+
+Each observed effect size ($y_{i}$) is composed of the real effect size $\theta$ plus an error term ($\epsilon_{i}$) that is sampled from a normal distribution with $\mu = 0$ and $\sigma^{2} = \sigma_{\epsilon_i}^{2}$ (i.e., the known sampling variance of the study $i$).
+As demonstrated in Equation \@ref(eq:effsize), the increase in sample size will decrease the sampling variability.
+An exact study will essentially have $\theta$ as the observed effect size.
 Since we are sampling participants' level data, the error component is already included in the `sim_study()` function.
@@
 n <- 20 # number of participants per group, per study
-theta_f <- 0.3 # real effect size
-
-sim <- make_data(k = k, nt = n, nc = n, theta_f = theta_f)
-sim <- sim_studies(sim$theta_f, sim$nc, sim$nt, data = sim) # simulated data
+theta <- 0.3 # real effect size
+
+sim <- make_data(k = k, nt = n, nc = n, es = theta)
+sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim) # simulated data
 res <- rma(yi = yi, vi = vi, method = "FE", data = sim)
@@
 
-We are using the $\theta_{f}$ parameter to generate random data using the `sim_study()` function that will introduce the random error component $\epsilon_{i}$.
+We are using the $\theta$ parameter to generate random data using the `sim_study()` function that will introduce the random error component $\epsilon_{i}$.
 Then we can fit the *fixed-effects* meta-analysis model with the `rma` function and `method = "FE"` of the `metafor` package.
-The parameter we are estimating is $\theta_{f}$ which is close to our simulation value (see Table \@ref(tab:res-fixed-effect)).
+The parameter we are estimating is $\theta$ which is close to our simulation value (see Table \@ref(tab:res-fixed-effect)).
 Increasing the number of studies ($k$) and/or the number of participants in each study ($n$) will improve estimation reducing the standard error.
@@
 k <- 10 # number of studies
-theta_f <- 0.3 # real effect size
+theta <- 0.3 # real effect size
 n <- c(30, 500)
 
-sim <- make_data(k = k, theta_f = theta_f)
+sim <- make_data(k = k, es = theta)
 
@@
 
-sim <- sim_studies(sim$theta_f, sim$nc, sim$nt, data = sim)
-
-sim <- escalc(yi = yi, vi = vi, data = sim)
+sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
+
+sim <- escalc(yi = yi, vi = vi, data = sim, measure = "MD")
 sim <- summary(sim)
@@
 The *fixed-effects model* assumes that the real effect is a single value.
-The random-effects model relaxes this assumption allowing $\theta_{f}$ (now $\theta_{r}$) to vary across studies.
+The random-effects model relaxes this assumption allowing the true effect size to vary across studies.
 For example, the difference between groups we are simulating could be influenced by the type of experiment or the participants' age.
-Now $\theta_{r}$ is no longer a single value but a distribution of values.
-Due to the effect size being a distribution, we need two parameters $\mu$ and $\sigma^{2}$ to describe the phenomenon.
-The parameter $\theta_{r}$ average true effect is now interpreted as the average effect size across different true effect sizes and $\sigma^{2}$ usually known as $\tau^{2}$ is the effect size variability or heterogeneity.
+Now $\theta$ is no longer a single value but a distribution of values.
+Due to the effect size being a distribution, we need to estimate both the mean and the variance.
+The parameter $\mu_{\theta}$ is mean of the distribution, interpreted as the average effect size across different true effect sizes and $\tau^{2}$ is the variance of the distribution interpreted as variability or heterogeneity of effect sizes.
 In practical terms, we now have two sources of variability: $\tau^{2}$, which express the real difference among effect sizes, and $\sigma_{i}^{2}$ which is the known sampling variance of each study as in the *fixed-effects model*.
-We can easily extend Equation \@ref(eq:fixed-effect-model) with Equation \@ref(eq:random-effect-model).
+We can easily extend Equation \@ref(eq:fixed-effect-model1) with Equations \@ref(eq:random-effect-model1), \@ref(eq:random-effect-model2) and \@ref(eq:random-effect-model3).
 
@@
 \begin{aligned}
-d_i = \theta_r + \theta_i + \epsilon_i \\
-\theta_i \sim \mathcal{N}(0, \tau^2) \\
-\epsilon_i \sim \mathcal{N}(0,\sigma^2_i)
-(\#eq:random-effect-model)
+y_i = \mu_{\theta} + \theta_i + \epsilon_i \\
+(\#eq:random-effect-model1)
 \end{aligned}
@@
 
-Compared to the *fixed-effects model* we need to generate another adjustment to the overall effect $\theta_{r}$ from a normal distribution with mean $\theta_{r}$ and variance $\tau^{2}$.
-The real effect size for the study $i$ will be $\theta_{r} + \theta_{i}$ where $\theta_{i}$ are the random-effects regulated by the $\tau^{2}$ parameter.
+\begin{align}
+\begin{aligned}
+\theta_i \sim \mathcal{N}(0,\tau^2)
+(\#eq:random-effect-model2)
+\end{aligned}
+\end{align}
+
+\begin{align}
+\begin{aligned}
+\epsilon_i \sim \mathcal{N}(0,\sigma_{\epsilon_i}^{2})
+(\#eq:random-effect-model3)
+\end{aligned}
+\end{align}
+
+Compared to the *fixed-effects model* we need to generate another adjustment to the overall effect $\mu_{\theta}$ from a normal distribution with mean 0 and variance $\tau^{2}$.
+The real effect size for the study $i$ will be $\mu_{\theta} + \theta_{i}$ where $\theta_{i}$ are the random-effects regulated by the $\tau^{2}$ parameter.
 
@@
 n <- 20 # number of participants per group, per study
-theta_r <- 0.3 # real effect size
+mu_theta <- 0.3 # real effect size
 tau2 <- 0.2 # the effect size heterogeneity
 
-sim <- make_data(k = k, nc = n, nt = n, theta_r = theta_r)
+sim <- make_data(k = k, nc = n, nt = n, mu_theta = mu_theta)
 
@@
 # adding the by-study adjustment to the average effect
-sim$theta_r_theta_i <- sim$theta_r + sim$theta_i
+sim$es <- sim$mu_theta + sim$theta_i
 
 # now we are using theta_r_theta_i and no longer theta_f
-sim <- sim_studies(sim$theta_r_theta_i, sim$nc, sim$nt, data = sim)
+sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
 res <- rma(yi = yi, vi = vi, method = "REML", data = sim)
@@
 See the `rma` documentation (https://wviechtb.github.io/metafor/reference/rma.uni.html#specifying-the-model) for an overview of other estimation methods] of the metafor package.
-We are now estimating two parameters: $\theta_{r}$ and $\tau^{2}$.
+We are now estimating two parameters: $\mu_{\theta}$ and $\tau^{2}$.
 As for the *fixed-effects model*, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
@@
 An important aspect of the *random-effects model* is the interplay between the heterogeneity ($\tau^{2}$) and the sampling variability ($\sigma_{i}^{2}$).
-As the number of studies $k$ increases, the estimation of $\theta_{r}$ and $\tau^{2}$ becomes more precise.
+As the number of studies $k$ increases, the estimation of $\mu_{\theta}$ and $\tau^{2}$ becomes more precise.
 Additionally, as each study\'s sample size decreases, each $\theta_{i}$ will be estimated with higher precision but as long as $\tau^{2} \neq 0$
@@
 This can be easily demonstrated using the previous simulation, increasing each study\'s sample size.
-Figure \@ref(fig:random-effect-high-low-precision) depicts the same meta-analysis but with different precision in estimating the study $d_{i}$.
+Figure \@ref(fig:random-effect-high-low-precision) depicts the same meta-analysis but with different precision in estimating the study $y_{i}$.
 Compared to Figure \@ref(fig:fixed-effect-high-low-precision), increasing the sample size of primary studies improves the estimation of each study without reducing the between-studies heterogeneity.
@@
 k <- 10 # number of studies
-theta_r <- 0.3 # real effect size
+mu_theta <- 0.3 # real effect size
 tau2 <- 0.2
@@
 
-sim <- make_data(k = k, theta_r = theta_r)
+sim <- make_data(k = k, mu_theta = mu_theta)
 
@@
 
-sim$theta_r_theta_i <- sim$theta_r + sim$theta_i
-sim <- sim_studies(sim$theta_r_theta_i, sim$nc, sim$nt, data = sim)
+sim$es <- sim$mu_theta + sim$theta_i
+sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
 
@@
 k <- 30 # number of studies
-theta_r <- 0.3 # real effect size
+mu_theta <- 0.3 # real effect size
 n <- 30 # sample size per group, per study
@@
 
-v <- (n + n)/(n * n) + theta_r^2/(2 * (n + n - 2)) # typical within study variance
+v <- (n + n)/(n * n) + mu_theta^2/(2 * (n + n - 2)) # typical within study variance
 tau2 <- -((I2*v)/(I2 - 1))
 
-sim <- make_data(k = k, nc = n, nt = n, theta_r = theta_r)
+sim <- make_data(k = k, nc = n, nt = n, mu_theta = mu_theta)
 
 sim$theta_i <- rnorm(k, 0, sqrt(tau2))
-sim$theta_r_theta_i <- sim$theta_r + sim$theta_i 
-sim <- sim_studies(sim$theta_r_theta_i, sim$nc, sim$nt, data = sim)
+sim$es <- sim$mu_theta + sim$theta_i 
+sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
 res <- rma(yi, vi, method = "REML", data = sim)
@@
 In our example, a group of studies uses an online memory task while others use a standard lab-based task.
-The Equation \@ref(eq:random-effect-model) can be easily extended for a meta-regression model by including a variable encoding the type of task (online vs.
+The Equation \@ref(eq:random-effect-model1) can be easily extended for a meta-regression model by including a variable encoding the type of task (online vs.
 lab-based) and the expected difference between the two levels of the moderator (i.e., the *lab vs.
@@
 To simulate this meta-regression, we need to fix the $\beta_{1}$ effect.
-Due to the fact that we are on a standardized scale, we can fix $\beta_{1}$ to be the expected difference in Cohen's $d$ scale between the two groups of studies.
+Due to the fact that we are on a standardized scale, we can fix $\beta_{1}$ to be the expected difference between the two groups of studies.
 Crucially, despite $\tau^{2}$ is still the heterogeneity between effect sizes, now we need to fix $\tau^{2}$ considering that we included a moderator.
-In other terms, $\tau_{r}^{2}$ is the residual heterogeneity after explaining the
-variance due to $X_{1}$. We can describe our model using two equations according to the value of $X_{1}$ as reported in Equation \@ref(eq:meta-regression-equations).
+In other terms, $\theta_i \sim \mathcal{N}(0, \tau_{r}^{2})$ where $\tau^2_r$ is the residual heterogeneity after including the moderator.
+We can describe our model using equation \@ref(eq:meta-regression-equations) according to the value of $X_{1i}$ (L = lab-based experiments, O = online experiments).
 
@@
 \begin{gathered}
-d_i = \beta_0 + \theta_i + \beta_1 X_1 + \epsilon_i \\
-\theta_i \sim \mathcal{N}(0, \tau^2) \\
-\epsilon_i \sim \mathcal{N}(0,\sigma^2_i)
+y_i = \beta_0 + \theta_i + \beta_1 X_{1i} + \epsilon_i
 (\#eq:meta-regression)
@@
 \begin{gathered}
-d_i = \beta_0 + \theta_i + \beta_1 X_1 + \epsilon_i \\
-d_{il} = \beta_0 + \theta_i + \beta_1 \times 0 + \epsilon_i \\
-d_{io} = \beta_0 + \theta_i + \beta_1 \times 1 + \epsilon_i \\
+y_{iL} = \beta_0 + \theta_i + \beta_1 \times 0 + \epsilon_i \\
+d_{iO} = \beta_0 + \theta_i + \beta_1 \times 1 + \epsilon_i \\
 (\#eq:meta-regression-equations)
@@
 
-We can simulate the same scenario of the random-effects model with $k_o = 15$ (*online tasks*) and $k_l = 15$ (*lab-based tasks*).
+We can simulate the same scenario of the random-effects model with $k_O = 15$ (*online tasks*) and $k_L = 15$ (*lab-based tasks*).
 Then we fix the $\beta_1 = 0.2$ and $\tau_r^2 = 0.1$ ($r$ for residual).
@@
 sim$theta_i <- rnorm(k, 0, sqrt(tau2r)) # the by-study residual adjustment
-sim$b0_theta_i_x <- b0 + sim$theta_i + b1*ifelse(sim$exp == "lab", 0, 1)
-
-sim <- sim_studies(sim$b0_theta_i_x, sim$nc, sim$nt, data = sim)
+sim$es <- b0 + sim$theta_i + b1*ifelse(sim$exp == "lab", 0, 1)
+
+sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
 res <- rma(yi, vi, mods = ~exp, method = "REML", data = sim)
@@
 
-```{r power-analysis-example, cache = TRUE}
+```{r power-analysis-example, cache = FALSE}
 set.seed(seed)

### 2023-06-29 documents/paper/paper.Rmd
fixing fixed vs random-effects figure caption

@@
 set.seed(123) # seed for simulation to ensure the reproducibility of results
-theta <- 0.5 # real mean difference in the populations
+es <- 0.5 # real mean difference in the populations
 sigma <- 1 # real standard deviations in the population
@@
 d_mc <- replicate(nsims, expr = {
-  g1 <- rnorm(n = n, mean = theta, sd = sigma)
+  g1 <- rnorm(n = n, mean = es, sd = sigma)
   g2 <- rnorm(n = n, mean = 0, sd = sigma)
@@
              "$d_i$",      "yi",                                 "Observed effect size",
-      "$$\\sigma_{\\epsilon_i}^{2}$$",      "vi",                          "Observed sampling variance",
+      "$\\sigma_{\\epsilon_i}^{2}$",      "vi",                          "Observed sampling variance",
        "$\\theta_i$", "theta\\_i",              "Random effect for the study i",
@@
 n <- 20 # number of participants per group, per study
-mu_theta <- 0.3 # real effect size
+mu <- 0.3 # real effect size
 tau2 <- 0.2 # the effect size heterogeneity
 
-sim <- make_data(k = k, nc = n, nt = n, mu_theta = mu_theta)
+sim <- make_data(k = k, nc = n, nt = n, mu = mu)
 
@@
 # adding the by-study adjustment to the average effect
-sim$es <- sim$mu_theta + sim$theta_i
+sim$es <- sim$mu + sim$theta_i
 
@@
 k <- 10 # number of studies
-mu_theta <- 0.3 # real effect size
+mu <- 0.3 # real effect size
 tau2 <- 0.2
@@
 
-sim <- make_data(k = k, mu_theta = mu_theta)
+sim <- make_data(k = k, mu = mu)
 
@@
 
-sim$es <- sim$mu_theta + sim$theta_i
+sim$es <- sim$mu + sim$theta_i
 sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
@@
 k <- 30 # number of studies
-mu_theta <- 0.3 # real effect size
+mu <- 0.3 # real effect size
 n <- 30 # sample size per group, per study
@@
 
-v <- (n + n)/(n * n) + mu_theta^2/(2 * (n + n - 2)) # typical within study variance
+v <- (n + n)/(n * n) + mu^2/(2 * (n + n - 2)) # typical within study variance
 tau2 <- -((I2*v)/(I2 - 1))
 
-sim <- make_data(k = k, nc = n, nt = n, mu_theta = mu_theta)
+sim <- make_data(k = k, nc = n, nt = n, mu = mu)
 
 sim$theta_i <- rnorm(k, 0, sqrt(tau2))
-sim$es <- sim$mu_theta + sim$theta_i 
+sim$es <- sim$mu + sim$theta_i 
 sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
@@
 online effect*).
-In regression terms, we could use a dummy variable ($X_{1}$) that takes the value of 0 for the *lab-based task* (*l*) and a value of 1 for an *online task* (*o*)^[The *dummy coding* (also known as treatment coding) is the default in R and `metafor` but other coding schemes could be used [see @Schad2020-ht for an overview of contrast coding schemes]].
+In regression terms (see equation \@ref(eq:meta-regression)), we could use a dummy variable ($X_{1}$) that takes the value of 0 for the *lab-based task* (*L*) and a value of 1 for an *online task* (*O*)^[The *dummy coding* (also known as treatment coding) is the default in R and `metafor` but other coding schemes could be used [see @Schad2020-ht for an overview of contrast coding schemes]].
 Now we fix $\beta_{1}$ to be the *lab vs.
-online effect* Equation \@ref(eq:random-effect-model) into Equation \@ref(eq:meta-regression).
-Now the product between $\beta_{1}X_{1}$ will consider the *lab vs.
+online effect* (i.e., the expected mean difference between the two groups of studies) and the product between $\beta_{1}X_{1}$ will consider the *lab vs.
 online effect*.
-To simulate this meta-regression, we need to fix the $\beta_{1}$ effect.
-Due to the fact that we are on a standardized scale, we can fix $\beta_{1}$ to be the expected difference between the two groups of studies.
 Crucially, despite $\tau^{2}$ is still the heterogeneity between effect sizes, now we need to fix $\tau^{2}$ considering that we included a moderator.
@@
 y_{iL} = \beta_0 + \theta_i + \beta_1 \times 0 + \epsilon_i \\
-d_{iO} = \beta_0 + \theta_i + \beta_1 \times 1 + \epsilon_i \\
+d_{iO} = \beta_0 + \theta_i + \beta_1 \times 1 + \epsilon_i
 (\#eq:meta-regression-equations)
@@
 
-sim$b0_theta_i_x <- b0 + sim$theta_i + sim$b1*sim$age0
-
-sim <- sim_studies(sim$b0_theta_i_x, sim$nc, sim$nt, data = sim)
+sim$es <- b0 + sim$theta_i + sim$b1*sim$age0
+
+sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
 ```
@@
 sim$age0 <- scale(sim$age, center = TRUE, scale = TRUE) # standardize the moderator
-sim$b0_theta_i_x <- b0 + sim$theta_i + b1*sim$age0
-
-sim <- sim_studies(sim$b0_theta_i_x, sim$nt, sim$nc, data = sim)
+sim$es <- b0 + sim$theta_i + b1*sim$age0
+
+sim <- sim_studies(sim$es, sim$nt, sim$nc, data = sim)
 res <- rma(yi, vi, mods = ~age0, method = "REML", data = sim)
@@
 
-```{r power-analysis-example, cache = FALSE}
+```{r power-analysis-example, cache = TRUE}
 set.seed(seed)
@@
 k <- c(5, 15, 25, 35, 50) # number of studies
-theta <- 0.3 # the average effect size
+delta <- 0.3 # the average effect size
 tau2 <- 0.3 # the heterogeneity
@@
 # creating all combinations
-sim_grid <- tidyr::expand_grid(k, theta, tau2, navg, nmin, nsim)
+sim_grid <- tidyr::expand_grid(k, mu, tau2, navg, nmin, nsim)
 
@@
 sim_grid_t <- sim_grid
-names(sim_grid_t) <- c("k", "$\\theta$", "$\\tau^2$", "$n_{avg}$", "$n_{min}$", "nsim", "Power")
+names(sim_grid_t) <- c("k", "$\\Delta$", "$\\tau^2$", "$n_{avg}$", "$n_{min}$", "nsim", "Power")
 
@@
 ```{r plot-power-analysis, echo = FALSE, fig.cap=get_caption()}
-title <- latex2exp::TeX(sprintf("$n_{avg} = %.1f$, $\\tau^2 = %s$, $\\theta_r = %s$, $\\alpha = 0.05$", sim_grid$navg[1], tau2, theta_r, 0.05))
+title <- latex2exp::TeX(sprintf("$n_{avg} = %.1f$, $\\tau^2 = %s$, $\\Delta = %s$, $\\alpha = 0.05$", sim_grid$navg[1], tau2, delta, 0.05))
 
@@
 (ref:img-fixed-vs-random) The difference between the assumptions of the *fixed* and *random-effects model*.
-Each distribution depicts the sampling distribution of $k = 4$ hypothetical studies ($i = 1$, $2$, ..., $4$) with a certain observed effect size $d_{i}$, sampling variability $\sigma_{i}^{2}$ and the 95% confidence interval (black segment).
-The *fixed-effects* plot on the left suggests that each observed effect size has the same underlying true effect $\theta_{f}$ (each distribution has the same mean) with a different degree of precision (e.g., Study 1 is more precise than Study 3).
-In practice, each study has a different observed effect size where studies with high precision (i.e., narrow sampling distributions) will be remarkably close to the real effect ($\theta_{f}$).
-The *random-effects model* on the right suggests that beyond the error term ($\epsilon_{i}$), each real effect size is composed of a fixed part (now $\theta_{r}$) and a random part ($\theta_{i}$, black segments below the sampling distributions).
+Each distribution depicts the sampling distribution of $k = 4$ hypothetical studies ($i = 1$, $2$, ..., $4$) with a certain observed effect size $y_{i}$ (red squares), sampling variability $\sigma_{i}^{2}$ and the 95% confidence interval (black segment).
+The *fixed-effects* plot on the left suggests that each observed effect size has the same underlying true effect $\theta$ (each distribution has the same mean) with a different degree of precision (e.g., Study 1 is more precise than Study 3).
+In practice, each study has a different observed effect size where studies with high precision (i.e., narrow sampling distributions) will be remarkably close to the real effect ($\theta$).
+The *random-effects model* on the right suggests that beyond the error term ($\epsilon_{i}$), each real effect size is composed of a fixed part (now $\mu_{\theta}$) and a random part ($\theta_{i}$, black segments below the sampling distributions).
 The random part is determined by $\tau^{2}$, which represents the variability of the effect size distribution at the population level.

### 2023-06-30 scripts/r-img.R
fixing random vs fixed plot

@@ -46,26 +46,110 @@ models$delta <- models$yi - models$di
 set.seed(2036)
 models$obs <- mapply(function(m, s) rnorm(1, m, s), models$yi, sqrt(models$vi))
 
-ggplot(models, aes(y = factor(id))) +
+p_fixed_down <- ggplot(filter(models, model == "Fixed-Effects"), aes(y = factor(id))) +
   geom_vline(aes(xintercept = di), linetype = "dashed", alpha = 0.5) +
   stat_dist_halfeye(aes(dist = dist), .width = 0.95) +
+  xlim(c(-1,2)) +
   facet_wrap(~model, scales = "free_x") +
   ylab("Study") +
-  xlab(TeX("$d_i$")) +
-  theme(axis.text.x = element_blank(),
-        axis.ticks = element_blank(),
-        axis.title.x = element_blank(),
-        strip.text = element_text(family = "lmroman")) +
+  xlab(TeX("$y_i$")) +
   theme_rfig() +
-  geom_label(aes(x = di, y = 0.6, label = note), parse = TRUE, size = 7, label.size = NA,
-             family = "lmroman") +
+  theme(strip.text = element_text(family = "lmroman"),
+        strip.text.x = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()) +
+  geom_label(aes(x = di, y = 0.6, label = note), parse = TRUE, size = 7, label.size = NA) +
   geom_segment(aes(x = di + delta, xend = di, y = id-0.08, yend = id-0.08),
                position = position_dodge2(width = 1, padding = 1)) +
   geom_point(aes(x = obs, y = factor(id)), color = "firebrick2", size = 2, shape = 15) +
   geom_point(aes(x = yi, y = id-0.08, alpha = model), show.legend = FALSE) +
   scale_alpha_manual(values = c(0, 1))
 
- # Metaregression with binary predictor ------------------------------------
+p_random_down <- ggplot(filter(models, model == "Random-Effects"), aes(y = factor(id))) +
+  geom_segment(aes(x = di + delta, xend = di + delta,
+                   y = factor(id), yend = Inf),
+               alpha = 0.3) +
+  geom_vline(aes(xintercept = di), linetype = "dashed", alpha = 0.5) +
[…]

### 2023-06-30 scripts/r-img.R
fixing theta to mu

@@ -134,7 +134,7 @@ p_up_right <- ggplot(up_right) +
         axis.text.y = element_blank(),
         axis.title.y = element_blank(),
         axis.ticks.y = element_blank()) +
-  geom_label(aes(x = theta, y = 3), 
+  geom_label(aes(x = mu, y = 3), 
              label = TeX("$N(\\mu_{\\theta}, \\tau^2)$"), 
              parse = TRUE,
              size = 7,

### 2023-07-03 documents/paper/paper.Rmd
adding equal-effects and fix equations

@@
 Firstly, it allows combining evidence from multiple studies assigning more weight to studies with lower estimation variability.
-Then using meta-regression, it is possible to assess the impact of study-level variables that could impact the phenomenon.
+Then using meta-regression, it is possible to include variables (i.e., moderators) to explain the observed heterogeneity [@Borenstein2009-mo, pp.187-203].
+More recently, location-scale models have been developed to inlcude predictors also on the residual heterogeneity [@Viechtbauer2022-rz].
 Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias.
@@
 
-### Fixed vs random-effects model
+### Equal vs fixed vs random-effects model
 
 The core of a meta-analysis is combining the results of multiple studies giving more weight to studies that provide a more precise effect estimation.
-We can divide meta-analysis models into two categories: the *fixed-effects* and the *random-effects model*.
-If we imagine combining multiple studies replicating the same experiment, we assume the real treatment effect $\theta$ is *fixed,* and each study is a more or less precise estimation of this parameter.
-This is an essential and often inappropriate assumption because it ignores that study-level characteristics may impact the actual treatment effect.
-
-Suppose some study-level characteristics (e.g., participants' age, sex, or socioeconomic status) or the experimental paradigm (e.g., type of memory task or difficulty) could impact the treatment effect.
-In that case, the assumption of the *fixed-effects model* is no longer plausible.
-There is no unique real effect but a distribution of effects with mean $\mu_{\theta}$ and variance $\tau^{2}$.
-The mean is the *average* effect, and the variance is the heterogeneity of the phenomenon.
-The *random-effects model* will estimate the mean ($\mu_{\theta}$) and the heterogeneity $\tau^{2}$ will help generalize to the population of effects beyond the pool of included studies.
-
-Most importantly, the parameters estimated with the *fixed* and *random-effects* do not represent the same quantity.
-The *random-effects model* estimates the *average* effect across a distribution of true effects, while the *fixed-effects model* estimates a single true effect.
-Figure 1 depicts the idea of the fixed vs.
-random-effects model.
-The *fixed-effects model* can be considered a particular type of *random-effects model* where $\tau^{2} = 0$ thus, the average effect is also the unique true effect.
+We can find essentially three meta-analysis models: the equal-effects, the fixed-effects and the random-effects model [@Laird1990-ro, @Hedges1998-uo].
+The *equal-effects* model assume that each study included in the meta-analysis is a more or less precise estimation of the true underlying effect ($\theta$).
+In other terms, we are assuming that there is no variability (i.e., heterogeneity) among true effect sizes.
+On the other side, suppose some study-level characteristics (e.g., participants' age, sex, or socioeconomic status) or the experimental paradigm (e.g., type of memory task or difficulty) could impact the treatment effect.
+The *random-effects* model assume a distribution of real effects with mean $\mu_{\theta}$ and variance $\tau^{2}$.
+Finally, the *fixed-effects* estimate the average effect of the included pool of studies ignoring the presence of heterogeneity.
+While the distinction between the equal and fixed-effects model is theoretical, the estimated model is the exactly the same^[see https://wviechtb.github.io/metafor/reference/misc-models.html for a detailed explanation].
+The *random-effects* model assume and estimate heterogeneity among effect sizes leading to different model parameters.
+From an inferential point of view, the *random-effects* model provide unconditional inference to the population of effect sizes while the *fixed-effects* model estimate the average effect of the selected studies, and not population-level parameters.
+The *equal-effects* model assume a single true population effect to be estimated.
+In this tutorial we will consider the equal and random-effects models.
 
@@
 
-We load relevant packages and set the `seed` for reproducibility to prepare the R environment for the simulations.
+This tutorial use several R packages for the simulations, meta-analysis fitting and figures/tables.
+For the data manipulation we used the *tidyverse* [@R-tidyverse] package.
+For the models fitting we used the *metafor* [@metafor2010] package.
+For figures and tables the *ggplot2* [@ggplot22016] or the `metafor::forest()` function, *kableExtra* [@R-kableExtra] and *papaja* [@R-papaja] packages.
+We set the `seed` for reproducibility to prepare the R environment for the simulations.
 
 ```{r sim-setup, eval = FALSE}
-library(dplyr) # for data manipulation
-library(tidyr) # for data manipulation
+library(dplyr) # for data manipulation (within tidyverse)
+library(tidyr) # for data manipulation (within tidyverse)
 library(metafor) # for meta-analysis models fitting
@@
 \begin{gathered}
-y_i = \overline{T} - \overline{C}
+D = \overline{T} - \overline{C}
 (\#eq:effsize)
@@
 \begin{gathered}
-\sigma_{\epsilon_i}^{2} = \frac{s^2_T}{n_T} + \frac{s^2_C}{n_C}
+\sigma_{\epsilon D}^{2} = \frac{s^2_T}{n_T} + \frac{s^2_C}{n_C}
 (\#eq:effsizevar)
@@
        "$\\tau^2_r$",   "tau2r",                   "Residual effect size heterogeneity",
-  "$\\bar y_{T, C}$",        "",               "Mean of the experimental/control group",
+  "$\\overline{T}, \\overline{C}$",        "",               "Mean of the experimental/control group",
       "$s_{T, C}$",        "", "Standard deviation of the experimental/control group",
@@
 Then the $\beta_1$ parameter represents the estimated difference in $y$ between the values of $X_1$ (i.e., lab-based vs online experiments).
-As said before, $\tau^2$ is now the residual heterogeneity that is interpreted as the variability between effect sizes after controlling for the moderator $X_1$.
+As said before, $\tau^2_r$ is now the residual heterogeneity that is interpreted as the variability between effect sizes after controlling for the moderator $X_1$.
 
@@
 online experiments).
-As mentioned above, $\tau^{2}$ the residual heterogeneity is now interpreted as the variability between effect sizes after controlling for the moderator $X_{1}$.
+As mentioned above, $\tau^{2}_r$ the residual heterogeneity is now interpreted as the variability between effect sizes after controlling for the moderator $X_{1}$.
 
@@
 In practice, each study has a different observed effect size where studies with high precision (i.e., narrow sampling distributions) will be remarkably close to the real effect ($\theta$).
-The *random-effects model* on the right suggests that beyond the error term ($\epsilon_{i}$), each real effect size is composed of a fixed part (now $\mu_{\theta}$) and a random part ($\theta_{i}$, black segments below the sampling distributions).
-The random part is determined by $\tau^{2}$, which represents the variability of the effect size distribution at the population level.
+The *random-effects model* on the right suggests that beyond the error term ($\epsilon_{i}$), each real effect size is composed of a fixed part (now $\mu_{\theta}$) and a random part ($\theta_{i}$) sampled from a normal distribution with mean $\mu_{\theta}$  and variance $\tau^{2}$.
 When $\tau^{2}$ is zero, the random-effects reduces to a fixed-effects model.

### 2023-07-03 scripts/r-img.R
adding observed effects sizes to all images

@@ -173,6 +173,9 @@ binary_metareg$gm <- mean(binary_metareg$yi)
 
 xlim <- c(mean(mm) - 1, mean(mm) + 1)
 
+set.seed(2027)
+binary_metareg$obs <- rnorm(nrow(binary_metareg), binary_metareg$yi, sqrt(binary_metareg$vi))
+
 plot_metareg_bin <- ggplot(binary_metareg,
                            aes(y = id)) +
   geom_segment(aes(x = gm, xend = yi, y = id-0.2, yend = id-0.2),
@@ -201,7 +204,8 @@ plot_metareg_bin <- ggplot(binary_metareg,
   annotate("text", x = 1, y = 2, 
            label = TeX("\\textbf{Explained $\\tau^2$}"), 
            parse = TRUE, color = "darkgreen", 
-           family = "bold", size = 8)
+           family = "bold", size = 8) +
+  geom_point(aes(x = obs, y = id), color = "firebrick", shape = 15, size = 2.5)
 
 # Metaregression with numerical predictor ---------------------------------
 
@@ -227,8 +231,10 @@ cont_metareg$pi <- predict(fit, newdata = data.frame(x = cont_metareg$x-0.1))
 cont_metareg$gm <- mean(cont_metareg$yi)
 cont_metareg$res <- cont_metareg$yi - cont_metareg$pi
 
-plot_metareg_cont <- ggplot(cont_metareg,
-                            aes(x = x, y = yi)) +
+set.seed(2027)
+cont_metareg$obs <- rnorm(nrow(cont_metareg), cont_metareg$yi, sqrt(cont_metareg$vi))
+
+plot_metareg_cont <- ggplot(cont_metareg, aes(x = x, y = yi)) +
   geom_segment(aes(x = x-0.1, xend = x-0.1, y = pi, yend = yi), color = "firebrick", linewidth = 1) +
   geom_segment(aes(x = x-0.1, xend = x-0.1, y = gm, yend = pi), color = "darkgreen", linewidth = 1) +
   geom_point(aes(x = x-0.1, y = yi), size = 3, color = "black") +
@@ -241,7 +247,7 @@ plot_metareg_cont <- ggplot(cont_metareg,
   theme(axis.text.x = element_blank(),
         axis.ticks.x = element_blank()) +
   theme_rfig() +
-  ylab(latex2exp::TeX("$d_i$")) +
+  ylab(latex2exp::TeX("$y_i$")) +
   ggtitle("Numerical Predictor") +
[…]

### 2023-07-03 documents/paper/paper.Rmd
fixing theta_i notation and equal-effects everywhere

@@
 Furthermore, data simulation is a powerful educational tool for understanding a statistical method.
-In this tutorial, we show how to simulate fixed-effects, random-effects, and meta-regression models and illustrate how to estimate statistical power.
+In this tutorial, we show how to simulate equal-effects, random-effects, and meta-regression models and illustrate how to estimate statistical power.
 In the supplementary materials, we extended these simulations also for multilevel and multivariate models.
@@
 This common measure should give an immediate idea of the direction (i.e., the treatment improves or reduces performance) and the size of the effect.
-Standardized effect size measures [see @Lakens2013-eq for an overview], such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's $d$ [@Cohen1988-ad] or the Pearson correlation coefficient $\rho$ which could be calculated using the associated sample estimator *r* are commonly used to compare heterogeneous outcome variables.
+Standardized effect size measures [see @Lakens2013-eq for an overview], such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's $d$ [@Cohen1988-ad] or the Pearson correlation coefficient $\rho$ which could be calculated using the associated sample estimator *r* are commonly used to compare heterogeneous outcome variables^[When using bounded effect size indexes such as the Pearson correlation, Fisher's *z* transformation is commonly used[@Borenstein2009-mo, pp.
+41-43]].
 If all studies used the same raw measure, such as reaction times, it is possible to directly meta-analyze the studies without standardizing using (UMD, unstandardized mean difference).
@@
 The core of a meta-analysis is combining the results of multiple studies giving more weight to studies that provide a more precise effect estimation.
-We can find essentially three meta-analysis models: the equal-effects, the fixed-effects and the random-effects model [@Laird1990-ro, @Hedges1998-uo].
+We can find essentially three meta-analysis models: the equal-effects, the fixed-effects and the random-effects model [@Laird1990-ro, @Hedges1998-uo, see also @Gronau2021-do and @Berkhout2023-fj for Bayesian a meta-analysis approach trying to combine evidence from multiple models e.g., equal and random-effects together].
 The *equal-effects* model assume that each study included in the meta-analysis is a more or less precise estimation of the true underlying effect ($\theta$).
@@
 The *equal-effects* model assume a single true population effect to be estimated.
-In this tutorial we will consider the equal and random-effects models.
+In the presence of heterogeneity, using meta-regression is it possible to include some predictors (e.g., the type of experimental setup) to explain effect sizes variability.
 
@@
 Still, the most common way is to use a *multilevel meta-analysis model,* considering the heterogeneity between and within studies.
-The standard *random-effects* (but also the *fixed-effects model*) model is also called a *two-level* model because it combines data after aggregating the participants-level (*first-level*) data.
+The standard *random-effects* (but also the *equal-effects model*) model is also called a *two-level* model because it combines data after aggregating the participants-level (*first-level*) data.
 The *three-level* model is commonly used to manage the multilevel data structure [@Cheung2014-fg; @Cheung2019-po].
@@
 Simulate $n_{T}$ observations from a Gaussian distribution with $\mu = \Delta$ and $\sigma^{2} = 1$ and $n_{T}$ observations from a Gaussian distribution with $\mu = 0$ and $\sigma^{2} = 1$.
-This way, the expected difference between groups will be $\Delta$ and the variance is fixed at 1.
+In this way, the expected difference between groups will be $\Delta$ and the expected variance for each group is 1.
 3. Calculate the observed effect size $y$ and the sampling variance $\sigma_{\epsilon}^{2}$.
@@
 
-The `es` is the true effect size ($\mu_{\theta}$ for the random-effects model and $\theta$ for the fixed-effect model), `nc` and `nt` are the sample size for the control and experimental group, `aggregate` controls if returning the effect size and the corresponding variance or the participant-level data.
+The `es` is the true effect size ($\mu_{\theta}$ for the random-effects model and $\theta$ for the equal-effects model), `nc` and `nt` are the sample size for the control and experimental group, `aggregate` controls if returning the effect size and the corresponding variance or the participant-level data.
 We can generate a single study with the desired parameters with this function.
@@
 Using the `sim_study()` function multiple times (with the appropriate adjustments) we can generate a series of studies simulating a data set for a meta-analysis (using the `sim_studies()` function).
-After each example, we will compute the appropriate model (e.g., *fixed* or *random-effects*) using the `metafor` package [@Viechtbauer2010-xz] to check the recovery of simulated parameters.
+After each example, we will compute the appropriate model (e.g., *fixed* or *random-effects*) using the `metafor` package [@Viechtbauer2010-xz] to check the recovery of simulated parameters^[By default `metafor` use the Wald *z* test on model parameters but other inference methods are available e.g.
+the Knapp and Hartung [-@Knapp2003-qx] method by setting `test="knha"` within the `rma()` function.
+This method is useful especially when the number of studies is low (see https://wviechtb.github.io/metafor/reference/misc-recs.html for more details)].
 Table \@ref(tab:notation-tab) summarizes the simulation parameters notation in equations and code.
@@
      ~`Equation`,     ~`Code`,                                               ~`Description`,
-       "$\\theta$",   "theta",                  "Fixed-effects model real effect size",
-       "$\\mu$",      "mu",                        "Random-effects model real effect size",
+       "$\\theta$",   "theta",                  "Equal-effects model true effect size",
+       "$\\mu$",      "mu",                        "Random-effects model true effect size",
         "$\\tau^2$",    "tau2",                            "Effect size heterogeneity",
@@
       "$\\sigma_{\\epsilon_i}^{2}$",      "vi",                          "Observed sampling variance",
-       "$\\theta_i$", "theta\\_i",              "Random effect for the study i",
+       "$\\delta_i$", "deltai",              "Random effect for the study i",
         "$\\beta_0$",      "b0",                            "Meta-regression intercept",
@@
 
-## Fixed-effects model
-
-The most basic model to simulate is the *fixed-effects model* (FE).
-As reported in the previous sections, the *fixed-effects model* assumes the presence of a single true effect ($\theta_{f}$) and the observed variability is because each effect size is an imprecise estimation of the true effect.
+## Equal-effects model
+
+The most basic model to simulate is the *equal-effects model* (FE).
+As reported in the previous sections, the *equal-effects model* assumes the presence of a single true effect ($\theta$) and the observed variability is because each effect size is an imprecise estimation of the true effect.
 In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies (i.e., the sample size).
-Equations \@ref(eq:fixed-effect-model1) and \@ref(eq:fixed-effect-model2) depicts the *fixed-effects* model.
+Equations \@ref(eq:equal-effects-model1) and \@ref(eq:equal-effects-model2) depicts the *equal-effects* model.
 
@@
 y_i = \theta + \epsilon_i
-(\#eq:fixed-effect-model1)
+(\#eq:equal-effects-model1)
 \end{gathered}
@@
 \epsilon_i \sim \mathcal{N}(0,\sigma_{\epsilon_i}^{2})
-(\#eq:fixed-effect-model2)
+(\#eq:equal-effects-model2)
 \end{gathered}
@@
 
-```{r sim-fixed-effect}
+```{r sim-equal-effects}
 set.seed(seed)
@@
 We are using the $\theta$ parameter to generate random data using the `sim_study()` function that will introduce the random error component $\epsilon_{i}$.
-Then we can fit the *fixed-effects* meta-analysis model with the `rma` function and `method = "FE"` of the `metafor` package.
-The parameter we are estimating is $\theta$ which is close to our simulation value (see Table \@ref(tab:res-fixed-effect)).
+Then we can fit the *equal-effects* meta-analysis model with the `rma` function and `method = "EE"` of the `metafor` package.
+The parameter we are estimating is $\theta$ which is close to our simulation value (see Table \@ref(tab:res-equal-effects)).
 Increasing the number of studies ($k$) and/or the number of participants in each study ($n$) will improve estimation reducing the standard error.
 
-```{r res-fixed-effect, echo = FALSE}
+```{r res-equal-effects, echo = FALSE}
 res_tidy <- rma_tidy(res)
@@
 Increasing the sample size for each study will increase the estimation precision thus the variability among studies will be reduced.
-This can be easily demonstrated by simulating studies with a high sample size as reported in Figure \@ref(fig:fixed-effect-high-low-precision).
+This can be easily demonstrated by simulating studies with a high sample size as reported in Figure \@ref(fig:equal-effects-high-low-precision).
 As the sample size increases the only source of variability (i.e., the error component $\epsilon_{i}$) is close to zero so each study is closer to the true simulated value.
 
-```{r fixed-effect-high-low-precision, echo = FALSE, fig.asp=0.8, fig.cap=get_caption()}
+```{r equal-effects-high-low-precision, echo = FALSE, fig.asp=0.8, fig.cap=get_caption()}
 set.seed(seed)
@@
 
-fit_lp <- rma(yi, vi, method = "FE", data = filter(sim, nt == 30 & nc == 30))
-fit_hp <- rma(yi, vi, method = "FE", data = filter(sim, nt == 500 & nc == 500))
+fit_lp <- rma(yi, vi, method = "EE", data = filter(sim, nt == 30 & nc == 30))
+fit_hp <- rma(yi, vi, method = "EE", data = filter(sim, nt == 500 & nc == 500))
 
@@
 
-The *random-effects model* (RE) can be considered an extension of the *fixed-effects model*.
-The *fixed-effects model* assumes that the real effect is a single value.
+The *random-effects model* (RE) can be considered an extension of the *equal-effects model*.
+The *equal-effects model* assumes that the real effect is a single value.
 The random-effects model relaxes this assumption allowing the true effect size to vary across studies.
@@
 The parameter $\mu_{\theta}$ is mean of the distribution, interpreted as the average effect size across different true effect sizes and $\tau^{2}$ is the variance of the distribution interpreted as variability or heterogeneity of effect sizes.
-In practical terms, we now have two sources of variability: $\tau^{2}$, which express the real difference among effect sizes, and $\sigma_{i}^{2}$ which is the known sampling variance of each study as in the *fixed-effects model*.
-We can easily extend Equation \@ref(eq:fixed-effect-model1) with Equations \@ref(eq:random-effect-model1), \@ref(eq:random-effect-model2) and \@ref(eq:random-effect-model3).
+In practical terms, we now have two sources of variability: $\tau^{2}$, which express the real difference among effect sizes, and $\sigma_{i}^{2}$ which is the known sampling variance of each study as in the *equal-effects model*.
+We can easily extend Equation \@ref(eq:equal-effects-model1) with Equations \@ref(eq:random-effect-model1), \@ref(eq:random-effect-model2) and \@ref(eq:random-effect-model3).
 
@@
 \begin{aligned}
-y_i = \mu_{\theta} + \theta_i + \epsilon_i \\
+y_i = \mu_{\theta} + \delta_i + \epsilon_i \\
 (\#eq:random-effect-model1)
@@
 \begin{aligned}
-\theta_i \sim \mathcal{N}(0,\tau^2)
+\delta_i \sim \mathcal{N}(0,\tau^2)
 (\#eq:random-effect-model2)
@@
 
-Compared to the *fixed-effects model* we need to generate another adjustment to the overall effect $\mu_{\theta}$ from a normal distribution with mean 0 and variance $\tau^{2}$.
+Compared to the *equal-effects model* we need to generate another adjustment to the overall effect $\mu_{\theta}$ from a normal distribution with mean 0 and variance $\tau^{2}$.
 The real effect size for the study $i$ will be $\mu_{\theta} + \theta_{i}$ where $\theta_{i}$ are the random-effects regulated by the $\tau^{2}$ parameter.
@@
 # simulate the random-effects adjustment
-sim$theta_i <- rnorm(k, 0, sqrt(tau2))
+sim$deltai <- rnorm(k, 0, sqrt(tau2))
 
 # adding the by-study adjustment to the average effect
-sim$es <- sim$mu + sim$theta_i
-
-# now we are using theta_r_theta_i and no longer theta_f
+sim$es <- sim$mu + sim$deltai
+
+# now we are using mu_deltai and no longer theta
 sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
@@
 We are now estimating two parameters: $\mu_{\theta}$ and $\tau^{2}$.
-As for the *fixed-effects model*, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
+As for the *equal-effects model*, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
 
@@
 As the number of studies $k$ increases, the estimation of $\mu_{\theta}$ and $\tau^{2}$ becomes more precise.
-Additionally, as each study\'s sample size decreases, each $\theta_{i}$ will be estimated with higher precision but as long as $\tau^{2} \neq 0$
-there will be variability among effect sizes.
+Additionally, as each study\'s sample size decreases, each $\theta_{i}$ will be estimated with higher precision but as long as $\tau^{2} \neq 0$ there will be variability among effect sizes.
 In other words, increasing the sample size of each study or the number of studies will not affect the value of $\tau^{2}$ but only the estimation precision [see @Borenstein2009-mo, Chapter 16, Figure 16.6 for a clear explanation].
@@
 Figure \@ref(fig:random-effect-high-low-precision) depicts the same meta-analysis but with different precision in estimating the study $y_{i}$.
-Compared to Figure \@ref(fig:fixed-effect-high-low-precision), increasing the sample size of primary studies improves the estimation of each study without reducing the between-studies heterogeneity.
+Compared to Figure \@ref(fig:equal-effects-high-low-precision), increasing the sample size of primary studies improves the estimation of each study without reducing the between-studies heterogeneity.
 
@@
 
-theta_i <- rnorm(k, 0, sqrt(tau2))
-
-sim$theta_i <- rep(theta_i, each = 2)
-
-sim$es <- sim$mu + sim$theta_i
+deltai <- rnorm(k, 0, sqrt(tau2))
+
+sim$deltai <- rep(deltai, each = 2)
+
+sim$es <- sim$mu + sim$deltai
 sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
@@
 From Equation \@ref(eq:isquared) and Figure \@ref(fig:random-effect-high-low-precision) is clear that if each included study has a considerable sample size the total variability will be reduced, mainly driven by real heterogeneity ($\tau^2$).
-This is the crucial difference between the fixed-effects and the random-effects model [see also @Borenstein2009-mo, pp.
+This is the crucial difference between the equal-effects and the random-effects model [see also @Borenstein2009-mo, pp.
 117–122].
@@
 Table \@ref(tab:res-random-effect-i2) depicts the results of the random-effects model fixing the $I^{2}$ value.
+Choosing a meaningful $\tau^2$ for the simulation can be sometimes difficult.
+Beyond fixing the $I^2$ value, one could use a plausible value from the literature.
+For example van Erp and colleagues [@Van_Erp2017-cf] estimate an empirical $\tau^2$ distribution across several published meta-analyses.
 
@@
 
-sim$theta_i <- rnorm(k, 0, sqrt(tau2))
-sim$es <- sim$mu + sim$theta_i 
+sim$deltai <- rnorm(k, 0, sqrt(tau2))
+sim$es <- sim$mu + sim$deltai 
 sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
@@
 Crucially, despite $\tau^{2}$ is still the heterogeneity between effect sizes, now we need to fix $\tau^{2}$ considering that we included a moderator.
-In other terms, $\theta_i \sim \mathcal{N}(0, \tau_{r}^{2})$ where $\tau^2_r$ is the residual heterogeneity after including the moderator.
+In other terms, $\delta_i \sim \mathcal{N}(0, \tau_{r}^{2})$ where $\tau^2_r$ is the residual heterogeneity after including the moderator.
 We can describe our model using equation \@ref(eq:meta-regression-equations) according to the value of $X_{1i}$ (L = lab-based experiments, O = online experiments).
@@
 \begin{gathered}
-y_i = \beta_0 + \theta_i + \beta_1 X_{1i} + \epsilon_i
+y_i = \beta_0 + \delta_i + \beta_1 X_{1i} + \epsilon_i
 (\#eq:meta-regression)
@@
 \begin{gathered}
-y_{iL} = \beta_0 + \theta_i + \beta_1 \times 0 + \epsilon_i \\
-d_{iO} = \beta_0 + \theta_i + \beta_1 \times 1 + \epsilon_i
+y_{iL} = \beta_0 + \delta_i + \beta_1 \times 0 + \epsilon_i \\
+d_{iO} = \beta_0 + \delta_i + \beta_1 \times 1 + \epsilon_i
 (\#eq:meta-regression-equations)
@@
 
-sim$theta_i <- rnorm(k, 0, sqrt(tau2r)) # the by-study residual adjustment
-sim$es <- b0 + sim$theta_i + b1*ifelse(sim$exp == "lab", 0, 1)
+sim$deltai <- rnorm(k, 0, sqrt(tau2r)) # the by-study residual adjustment
+sim$es <- b0 + sim$deltai + b1*ifelse(sim$exp == "lab", 0, 1)
 
@@
 
-theta_i <- rnorm(k, 0, sqrt(tau2r))
-sim$theta_i <- rep(theta_i, each = 2)
-
-sim$es <- b0 + sim$theta_i + sim$b1*sim$age0
+deltai <- rnorm(k, 0, sqrt(tau2r))
+sim$deltai <- rep(deltai, each = 2)
+
+sim$es <- b0 + sim$deltai + sim$b1*sim$age0
 
@@
 
-sim$theta_i <- rnorm(k, 0, sqrt(tau2r))
+sim$deltai <- rnorm(k, 0, sqrt(tau2r))
 sim$age0 <- scale(sim$age, center = TRUE, scale = TRUE) # standardize the moderator
-sim$es <- b0 + sim$theta_i + b1*sim$age0
+sim$es <- b0 + sim$deltai + b1*sim$age0
 
@@
 
-(ref:img-fixed-vs-random) The difference between the assumptions of the *fixed* and *random-effects model*.
+(ref:img-equal-vs-random) The difference between the assumptions of the *equal* and *random-effects model*.
 Each distribution depicts the sampling distribution of $k = 4$ hypothetical studies ($i = 1$, $2$, ..., $4$) with a certain observed effect size $y_{i}$ (red squares), sampling variability $\sigma_{i}^{2}$ and the 95% confidence interval (black segment).
-The *fixed-effects* plot on the left suggests that each observed effect size has the same underlying true effect $\theta$ (each distribution has the same mean) with a different degree of precision (e.g., Study 1 is more precise than Study 3).
+The *equal-effects* plot on the left suggests that each observed effect size has the same underlying true effect $\theta$ (each distribution has the same mean) with a different degree of precision (e.g., Study 1 is more precise than Study 3).
 In practice, each study has a different observed effect size where studies with high precision (i.e., narrow sampling distributions) will be remarkably close to the real effect ($\theta$).
 The *random-effects model* on the right suggests that beyond the error term ($\epsilon_{i}$), each real effect size is composed of a fixed part (now $\mu_{\theta}$) and a random part ($\theta_{i}$) sampled from a normal distribution with mean $\mu_{\theta}$  and variance $\tau^{2}$.
-When $\tau^{2}$ is zero, the random-effects reduces to a fixed-effects model.
+When $\tau^{2}$ is zero, the random-effects reduces to a equal-effects model.
 
@@
 
-(ref:res-fixed-effect) Summary of the simulated fixed-effects model. The only estimated parameter is the average effect* $\theta_{f}$ *with the standard error,
+(ref:res-equal-effects) Summary of the simulated equal-effects model. The only estimated parameter is the average effect* $\theta_{f}$ *with the standard error,
 95% confidence interval, and the Wald z-test. The z-test evaluates the null hypothesis that the real effect equals zero.
 
-(ref:fixed-effect-high-low-precision) Forest plots of two simulated *fixed-effects models*.
+(ref:equal-effects-high-low-precision) Forest plots of two simulated *equal-effects models*.
 On the left, the simulated model has $nt,nc = 30$ for each included study while on the right the sample size for each study is $nt,nc = 500$.
-Given that the data were generated under a *fixed-effects model*, when the sample size is high (on the right) each study is aligned on the real effect size because the error component ($\epsilon_{i}$) is almost entirely reduced.
+Given that the data were generated under a *equal-effects model*, when the sample size is high (on the right) each study is aligned on the real effect size because the error component ($\epsilon_{i}$) is almost entirely reduced.
 The average effect size is similar (depending on the random numbers generation) between the two scenarios while the estimation precision (the width of the black diamond) is narrower on the right.
 
-(ref:res-random-effect) Summary of the simulated random-effects model: Compared to the fixed-effects model, there are more parameters.
+(ref:res-random-effect) Summary of the simulated random-effects model: Compared to the equal-effects model, there are more parameters.
 The $\beta$ is the average effect ($\theta_{r}$) with the standard error, 95% confidence interval, and the Wald z-test.
@@
 On the left, the simulated model has $nt,nc = 30$ for each included study, while on the right, the sample size for each study is $nt,nc = 500$.
-Compared to Figure \@ref(fig:fixed-effect-high-low-precision), data were generated under a *random-effects model*.
+Compared to Figure \@ref(fig:equal-effects-high-low-precision), data were generated under a *random-effects model*.
 The estimated average effect is similar between the two scenarios regarding average effect and precision.
-However, compared to the *fixed-effect* simulation, increasing the sample size of primary studies only affects the precision without reducing the true heterogeneity (i.e., $\tau^{2} \neq 0$).
+However, compared to the *equal-effects* simulation, increasing the sample size of primary studies only affects the precision without reducing the true heterogeneity (i.e., $\tau^{2} \neq 0$).
 

### 2023-07-03 documents/paper/paper.Rmd
fixing s and equal_effects naming

@@
 Simulating data is also a powerful educational tool within this framework.
-For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for fixed and random-effects, and meta-regression models with applications also for statistical power calculation.
+For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for equal and random-effects, and meta-regression models with applications also for statistical power calculation.
 In the first section, we will introduce basic concepts of the meta-analysis that are useful for setting up the simulation.
@@
 Code and materials are available on the OSF repository ([https://osf.io/54djn/](https://osf.io/54djn/)).
-Multivariate and multilevel models’ simulation examples and more details about the coding approach are available in the supplementary materials.
+Simulation examples for multivariate and multilevel models and more details about the coding approach are available in the supplementary materials.
 
@@
 
-```{r img-fixed-vs-random, echo=FALSE, fig.asp=1, fig.cap=get_caption()}
-r_imgs$fixed_vs_random
+```{r img-equal-vs-random, echo=FALSE, fig.asp=1, fig.cap=get_caption()}
+r_imgs$equal_vs_random
 ```
@@
 In practical terms, we now have two sources of variability: $\tau^{2}$, which express the real difference among effect sizes, and $\sigma_{i}^{2}$ which is the known sampling variance of each study as in the *equal-effects model*.
-We can easily extend Equation \@ref(eq:equal-effects-model1) with Equations \@ref(eq:random-effect-model1), \@ref(eq:random-effect-model2) and \@ref(eq:random-effect-model3).
+We can easily extend Equation \@ref(eq:equal-effects-model1) with Equations \@ref(eq:random-effects-model1), \@ref(eq:random-effects-model2) and \@ref(eq:random-effects-model3).
 
@@
 y_i = \mu_{\theta} + \delta_i + \epsilon_i \\
-(\#eq:random-effect-model1)
+(\#eq:random-effects-model1)
 \end{aligned}
@@
 \delta_i \sim \mathcal{N}(0,\tau^2)
-(\#eq:random-effect-model2)
+(\#eq:random-effects-model2)
 \end{aligned}
@@
 \epsilon_i \sim \mathcal{N}(0,\sigma_{\epsilon_i}^{2})
-(\#eq:random-effect-model3)
+(\#eq:random-effects-model3)
 \end{aligned}
@@
 
-```{r sim-random-effect}
+```{r sim-random-effects}
 set.seed(seed)
@@
 Then we can fit the *random-effects* meta-analysis model with the `rma` function and `method = "REML"`.
-Table \@ref(tab:res-random-effect) depicts the model results^[The `method = "REML"` is not the only method to estimate a random-effects model.
+Table \@ref(tab:res-random-effects) depicts the model results^[The `method = "REML"` is not the only method to estimate a random-effects model.
 See the `rma` documentation (https://wviechtb.github.io/metafor/reference/rma.uni.html#specifying-the-model) for an overview of other estimation methods] of the metafor package.
@@
 
-```{r res-random-effect, echo = FALSE}
+```{r res-random-effects, echo = FALSE}
 res_tidy <- rma_tidy(res)
@@
 This can be easily demonstrated using the previous simulation, increasing each study\'s sample size.
-Figure \@ref(fig:random-effect-high-low-precision) depicts the same meta-analysis but with different precision in estimating the study $y_{i}$.
+Figure \@ref(fig:random-effects-high-low-precision) depicts the same meta-analysis but with different precision in estimating the study $y_{i}$.
 Compared to Figure \@ref(fig:equal-effects-high-low-precision), increasing the sample size of primary studies improves the estimation of each study without reducing the between-studies heterogeneity.
 
-```{r random-effect-high-low-precision, echo = FALSE, fig.asp=0.8, fig.cap=get_caption()}
+```{r random-effects-high-low-precision, echo = FALSE, fig.asp=0.8, fig.cap=get_caption()}
 set.seed(seed)
@@
 
-From Equation \@ref(eq:isquared) and Figure \@ref(fig:random-effect-high-low-precision) is clear that if each included study has a considerable sample size the total variability will be reduced, mainly driven by real heterogeneity ($\tau^2$).
+From Equation \@ref(eq:isquared) and Figure \@ref(fig:random-effects-high-low-precision) is clear that if each included study has a considerable sample size the total variability will be reduced, mainly driven by real heterogeneity ($\tau^2$).
 This is the crucial difference between the equal-effects and the random-effects model [see also @Borenstein2009-mo, pp.
@@
 However, with the assumption of homogeneous sample size across studies^[The homogeneous sample size of primary studies assumption is commonly used to calculate the statistical power [see @Borenstein2009-mo, Chapter 29]], we can solve Equation \@ref(eq:isquared) for $\tau^{2}$ obtaining the heterogeneity value associated with a certain $I^{2}$ as reported in Equation \@ref(eq:tau2-from-isquared).
-Table \@ref(tab:res-random-effect-i2) depicts the results of the random-effects model fixing the $I^{2}$ value.
+Table \@ref(tab:res-random-effects-i2) depicts the results of the random-effects model fixing the $I^{2}$ value.
 Choosing a meaningful $\tau^2$ for the simulation can be sometimes difficult.
@@
 
-```{r sim-random-effect-i2}
+```{r sim-random-effects-i2}
 set.seed(seed)
@@
 
-```{r res-random-effect-i2, echo = FALSE}
+```{r res-random-effects-i2, echo = FALSE}
 res_tidy <- rma_tidy(res)
@@
 In our example, a group of studies uses an online memory task while others use a standard lab-based task.
-The Equation \@ref(eq:random-effect-model1) can be easily extended for a meta-regression model by including a variable encoding the type of task (online vs.
+The Equation \@ref(eq:random-effects-model1) can be easily extended for a meta-regression model by including a variable encoding the type of task (online vs.
 lab-based) and the expected difference between the two levels of the moderator (i.e., the *lab vs.
@@
 
-(ref:res-random-effect) Summary of the simulated random-effects model: Compared to the equal-effects model, there are more parameters.
+(ref:res-random-effects) Summary of the simulated random-effects model: Compared to the equal-effects model, there are more parameters.
 The $\beta$ is the average effect ($\theta_{r}$) with the standard error, 95% confidence interval, and the Wald z-test.
@@
 
-(ref:random-effect-high-low-precision) Forest plots of two simulated *random-effects model*s.
+(ref:random-effects-high-low-precision) Forest plots of two simulated *random-effects model*s.
 On the left, the simulated model has $nt,nc = 30$ for each included study, while on the right, the sample size for each study is $nt,nc = 500$.
@@
 
-(ref:res-random-effect-i2) Summary of the *random-effect* model fixing the $I^2$ value. Parameters are the same as described in Table \@ref(tab:res-random-effect).
+(ref:res-random-effects-i2) Summary of the *random-effect* model fixing the $I^2$ value. Parameters are the same as described in Table \@ref(tab:res-random-effects).
 
@@
 The $R^2$ is the percentage of explained heterogeneity and $\tau^2_r$ is the estimated residual heterogeneity.
-Other parameters are the same as the standard *random-effect* model (see Table \@ref(tab:res-random-effect)).
+Other parameters are the same as the standard *random-effect* model (see Table \@ref(tab:res-random-effects)).
 
@@
 The `age0` parameter is the slope between `age` and the effect size interpreted as an increase in effect size for a unit increase in the average age.
-Other parameters are the same as described in Table \@ref(tab:res-random-effect) and \@ref(tab:res-meta-reg-dummy).
+Other parameters are the same as described in Table \@ref(tab:res-random-effects) and \@ref(tab:res-meta-reg-dummy).
 

### 2023-07-03 scripts/r-img.R

@@ -15,17 +15,17 @@ theme_rfig <- function(size = 15){
   ltxplot::theme_latex(base_size = size)
 }
 
-# Fixed-vs-Random Effect --------------------------------------------------
+# Equal-vs-Random Effect --------------------------------------------------
 
-fixed_effect <- data.frame(
+equal_effects <- data.frame(
   id = 1:4,
   yi = rep(0.5, 4),
   vi = c(0.01, 0.02, 0.1, 0.05),
-  model = "Fixed-Effects",
+  model = "Equal-Effects",
   di = 0.5
 )
 
-fixed_effect$note <- sprintf("$y_i = \\theta + \\epsilon_i$")
+equal_effects$note <- sprintf("$y_i = \\theta + \\epsilon_i$")
 
 random_effect <- data.frame(
   id = 1:4,
@@ -38,7 +38,7 @@ random_effect$note <- sprintf("$y_i = \\mu_{\\theta} + \\theta_i + \\epsilon_i$"
 
 random_effect$di <- mean(random_effect$yi)
 
-models <- rbind(fixed_effect, random_effect)
+models <- rbind(equal_effects, random_effect)
 models$dist <- map2(models$yi, sqrt(models$vi), distributional::dist_normal)
 models$note <- latex2exp::TeX(models$note, output = "character")
 models$delta <- models$yi - models$di
@@ -46,7 +46,7 @@ models$delta <- models$yi - models$di
 set.seed(2036)
 models$obs <- mapply(function(m, s) rnorm(1, m, s), models$yi, sqrt(models$vi))
 
-p_fixed_down <- ggplot(filter(models, model == "Fixed-Effects"), aes(y = factor(id))) +
+p_equal_down <- ggplot(filter(models, model == "Equal-Effects"), aes(y = factor(id))) +
   geom_vline(aes(xintercept = di), linetype = "dashed", alpha = 0.5) +
   stat_dist_halfeye(aes(dist = dist), .width = 0.95) +
   xlim(c(-1,2)) +
@@ -141,13 +141,13 @@ p_up_right <- ggplot(up_right) +
              label.size = NA)
 
 
-p_fixed <- cowplot::plot_grid(p_up_left, p_fixed_down, ncol = 1, align = "v",
+p_equal <- cowplot::plot_grid(p_up_left, p_equal_down, ncol = 1, align = "v",
                               rel_heights = c(0.33, 0.66))
 p_random <- cowplot::plot_grid(p_up_right, p_random_down, ncol = 1, align = "v",
[…]

### 2023-07-05 documents/paper/paper.Rmd
before adding \rev{}

@@
   - \usepackage{orcidlink}
+  - \newcommand{\rev}[1]{\textcolor{red}{#1}}
 
@@
 Then using meta-regression, it is possible to include variables (i.e., moderators) to explain the observed heterogeneity [@Borenstein2009-mo, pp.187-203].
-More recently, location-scale models have been developed to inlcude predictors also on the residual heterogeneity [@Viechtbauer2022-rz].
+More recently, location-scale models have been developed to include predictors also on the residual heterogeneity [@Viechtbauer2022-rz].
 Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias.
@@
 In the first section, we will introduce basic concepts of the meta-analysis that are useful for setting up the simulation.
-We will evaluate the effect size, variance calculation, and the fixed vs.
+We will evaluate the effect size, variance calculation, and the equal vs.
 random effects model distinction.
@@
 Code and materials are available on the OSF repository ([https://osf.io/54djn/](https://osf.io/54djn/)).
-Simulation examples for multivariate and multilevel models and more details about the coding approach are available in the supplementary materials.
+A theoretical introduction, simulation examples for multivariate and multilevel models, and more details about the coding approach are available in the supplementary materials.
 
@@
 
-### Multilevel and multivariate models
-
-The most straightforward situation for a meta-analysis is when each included study contributes with a single effect size.
-However, it is common to have multiple effect sizes belonging to the same study creating a situation of dependency between statistical units.
-There are numerous sources of dependency for a meta-analysis [@Cheung2014-fg; @Cheung2015-bw, pp.
-121-122].
-The first type of dependency (*multilevel*) consists of multiple effect sizes from independent participants.
-Despite being collected with different participants, effect sizes from the same study could be correlated.
-We obtain a multilevel data structure with independent effect sizes in the same study.
-There are multiple ways to model this data structure [see @Cheung2014-fg for an overview].
-Still, the most common way is to use a *multilevel meta-analysis model,* considering the heterogeneity between and within studies.
-The standard *random-effects* (but also the *equal-effects model*) model is also called a *two-level* model because it combines data after aggregating the participants-level (*first-level*) data.
-The *three-level* model is commonly used to manage the multilevel data structure [@Cheung2014-fg; @Cheung2019-po].
-Compared to the *two-level* model, the *three-level* model estimates two heterogeneity parameters ($\tau^{2}$ and $\omega^{2}$), one for the heterogeneity between studies and the other for the heterogeneity within studies.
-
-A second type of dependency (*multivariate*) consists of multiple effect sizes collected from the same pool of participants.
-For example, when using two cognitive tasks on the same pool of participants.
-This is commonly known as *multivariate meta-analysis* because the effect sizes sampling errors are correlated given that participants are assessed multiple times.
-A *multivariate meta-analysis model* could be used to
-consider the correlation between sampling errors.
-
-A third type of dependency arises when true effect sizes are correlated at the population level [@Cheung2015-bw, pp.
-121-122].
-For example, two different memory tasks could be correlated because the latent psychological constructs are intrinsically correlated.
-This correlation is present even when other groups of participants perform the two memory tasks (i.e., no *multivariate* dependency) or are administered by different researchers (i.e., no *multilevel* dependency).
-This type of dependency could be managed by estimating the correlation between outcomes using a *multivariate random-effects model* [@Cheung2015-bw, pp.
-127-130].
-Figures \@ref(fig:img-multilevel) and \@ref(fig:img-multivariate) depict the conceptual idea of *multilevel* and *multivariate* models.
-
-```{r img-multilevel, echo=FALSE, out.width="80%", fig.cap = get_caption()}
-knitr::include_graphics("img/multilevel.pdf")
-```
-
-```{r img-multivariate, echo=FALSE, out.width="80%", fig.cap = get_caption()}
-knitr::include_graphics("img/multivariate.pdf ")
-```
-
 # Simulation
@@
         "$n_{T, C}$",  "nt, nc",        "Sample size of the experimental/control group",
-             "$d_i$",      "yi",                                 "Observed effect size",
+             "$y_i$",      "yi",                                 "Observed effect size",
       "$\\sigma_{\\epsilon_i}^{2}$",      "vi",                          "Observed sampling variance",
@@
             placement = "H",
-            note = "The \\texttt{yi} and \\texttt{vi} notation for the observed effect size ($d_i$) and sampling variance ($\\sigma^2_i$) has been used to be consistent with the \\texttt{metafor} notation.")
+            note = "The \\texttt{yi} and \\texttt{vi} notation for the observed effect size ($y_i$) and sampling variance ($\\sigma^2_i$) has been used to be consistent with the \\texttt{metafor} notation.")
 ```
@@
 The relationship between the sampling error and the heterogeneity can be expressed using the $I^{2}$ statistics [e.g., @Higgins2002-fh] that is the percentage of the total variability $\tau^{2} + \tilde{v}$ that is attributable to real heterogeneity between studies ($\tau^{2}$).
-In Equation \@ref(eq:isquared), $\tilde{v}$ is the "typical" within-study sampling variability [e.g., @Higgins2002-fh]^[see also here https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate#fn__1 for an overview of the $I^2$ statistics also for complex models].
+In Equation \@ref(eq:isquared), $\tilde{v}$ is the "typical" within-study sampling variability as proposed by Higgins and Thompson [-@Higgins2002-fh]^[see also here https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate#fn__1 for an overview of the $I^2$ statistics also for complex models] and implemented in Equation  \@ref(eq:typical-v) with $w_i = 1/\sigma_{\epsilon_i}^{2}$ and $k$ is the number of studies.
 
@@
 (\#eq:isquared)
+\end{gathered}
+\end{align}
+
+\begin{align}
+\begin{gathered}
+\tilde{v} = \frac{(k-1) \sum w_i}{(\sum w_i)^2 - \sum w_i^2}
+(\#eq:typical-v)
 \end{gathered}
@@
 online effect*).
-In regression terms (see equation \@ref(eq:meta-regression)), we could use a dummy variable ($X_{1}$) that takes the value of 0 for the *lab-based task* (*L*) and a value of 1 for an *online task* (*O*)^[The *dummy coding* (also known as treatment coding) is the default in R and `metafor` but other coding schemes could be used [see @Schad2020-ht for an overview of contrast coding schemes]].
+In regression terms (see equation \@ref(eq:meta-regression)), we could use a dummy variable ($X_{1}$) that takes the value of 0 for the *lab-based task* (*L*) and a value of 1 for an *online task* (*O*)^[The *dummy coding* (also known as treatment coding) is the default in R and `metafor` but other coding schemes could be used [see @Schad2020-ht for an overview of contrast coding schemes].
+As in standard linear regression, the contrast coding of categorical factors or the presence/absence of the intercept influence the interpretation of estimated model parameters and omnibus meta-regression tests (see https://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept for a detailed discussion)].
 Now we fix $\beta_{1}$ to be the *lab vs.
@@
 A first strategy could be to use values estimated from the literature.
-Another approach consists in setting up the model and simulating several expected $d_i$ and calculating the range of simulated values.
+Another approach consists in setting up the model and simulating several expected $y_i$ and calculating the range of simulated values.
 A third possibility is fixing the proportion of explained heterogeneity calculating the $\beta_1$ value accordingly.
@@
   geom_point(alpha = 0.5) +
-  ylab(latex2exp::TeX("$d_i$")) +
+  ylab(latex2exp::TeX("$y_i$")) +
   xlab("Age") +
@@
              show.legend = FALSE) +
-  ylab(latex2exp::TeX("$d_i$")) +
+  ylab(latex2exp::TeX("$y_i$")) +
   xlab("Age") +
@@
 
-(ref:plot-meta-reg-plausible-implausible) Scatter plots with marginal histograms for the range of simulated $d_i$ with two $\beta_1$ values.
+(ref:plot-meta-reg-plausible-implausible) Scatter plots with marginal histograms for the range of simulated $y_i$ with two $\beta_1$ values.
 The x-axis depicts the average age of simulated studies and the y-axis depicts the simulated effect size.

### 2023-07-05 documents/paper/paper.Rmd
changes in red

@@
 Furthermore, data simulation is a powerful educational tool for understanding a statistical method.
-In this tutorial, we show how to simulate equal-effects, random-effects, and meta-regression models and illustrate how to estimate statistical power.
+In this tutorial, we show how to simulate \rev{equal-effects}, random-effects, and meta-regression models and illustrate how to estimate statistical power.
 In the supplementary materials, we extended these simulations also for multilevel and multivariate models.
@@
 Firstly, it allows combining evidence from multiple studies assigning more weight to studies with lower estimation variability.
-Then using meta-regression, it is possible to include variables (i.e., moderators) to explain the observed heterogeneity [@Borenstein2009-mo, pp.187-203].
+\color{red} Then using meta-regression, it is possible to include variables (i.e., moderators) to explain the observed heterogeneity [@Borenstein2009-mo, pp.187-203].
 More recently, location-scale models have been developed to include predictors also on the residual heterogeneity [@Viechtbauer2022-rz].
-Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias.
+\color{black} Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias.
 Despite the advantages, meta-analysis implementation is not always straightforward, especially for complex data structures.
@@
 Simulating data is also a powerful educational tool within this framework.
-For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for equal and random-effects, and meta-regression models with applications also for statistical power calculation.
+For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for \rev{equal} and random-effects, and meta-regression models with applications also for statistical power calculation.
 In the first section, we will introduce basic concepts of the meta-analysis that are useful for setting up the simulation.
-We will evaluate the effect size, variance calculation, and the equal vs.
+We will evaluate the effect size, variance calculation, and the \rev{equal} vs.
 random effects model distinction.
@@
 Code and materials are available on the OSF repository ([https://osf.io/54djn/](https://osf.io/54djn/)).
-A theoretical introduction, simulation examples for multivariate and multilevel models, and more details about the coding approach are available in the supplementary materials.
+\color{red} A theoretical introduction, simulation examples for multivariate and multilevel models, and more details about the coding approach are available in the supplementary materials.
+\color{black}
 
@@
 
-As an example that will be used throughout the paper, we consider the efficacy of memory training in improving memory performance during a cognitive task.
+\rev{As an example that will be used throughout the paper, we consider the efficacy of memory training in improving memory performance during a cognitive task.
 The typical primary study will collect data from a group of participants receiving the memory training (*experimental group*) and another group receiving a control treatment (*control group*).
@@
 Despite differences in the type of cognitive task or experimental setup, each primary study collects an *experimental group* ($n_{C}$) and *control group* ($n_{C}$) and computes the average performance (*experimental group* $\overline T$ and *control group* $\overline C$) and standard deviations (*experimental group* $s_{T}$ and *control group*
-$s_{C}$).
+$s_{C}$).}
 
@@
 This common measure should give an immediate idea of the direction (i.e., the treatment improves or reduces performance) and the size of the effect.
-Standardized effect size measures [see @Lakens2013-eq for an overview], such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's $d$ [@Cohen1988-ad] or the Pearson correlation coefficient $\rho$ which could be calculated using the associated sample estimator *r* are commonly used to compare heterogeneous outcome variables^[When using bounded effect size indexes such as the Pearson correlation, Fisher's *z* transformation is commonly used[@Borenstein2009-mo, pp.
-41-43]].
+Standardized effect size measures [see @Lakens2013-eq for an overview], such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's $d$ [@Cohen1988-ad] or the Pearson correlation coefficient $\rho$ which could be calculated using the associated sample estimator *r* are commonly used to compare heterogeneous outcome variables^[\color{red} When using bounded effect size indexes such as the Pearson correlation, Fisher's *z* transformation is commonly used[@Borenstein2009-mo, pp.
+41-43] \color{black}].
 If all studies used the same raw measure, such as reaction times, it is possible to directly meta-analyze the studies without standardizing using (UMD, unstandardized mean difference).
@@
 
+\color{red}
 The core of a meta-analysis is combining the results of multiple studies giving more weight to studies that provide a more precise effect estimation.
@@
 
+\color{black}
+
 ```{r img-equal-vs-random, echo=FALSE, fig.asp=1, fig.cap=get_caption()}
@@
 
+\color{red}
 This tutorial use several R packages for the simulations, meta-analysis fitting and figures/tables.
@@
 We set the `seed` for reproducibility to prepare the R environment for the simulations.
+\color{black}
 
@@
 All the examples refer to primary studies that assess the efficacy of a treatment by comparing a control and an experimental group.
-There are different simulation approaches to meta-analysis concerning how to generate study-level or participant-level data.
+\color{red} There are different simulation approaches to meta-analysis concerning how to generate study-level or participant-level data.
 In simulation studies where the purpose is not evaluating effect size estimators, it is convenient to simulate unstandardized effect size measures (e.g., UMD).
@@
 The UMD and the sampling variance are calculated using Equations \@ref(eq:effsize) and \@ref(eq:effsizevar) where $y$ is the sample estimator of $\Delta$.
+\color{black}
+
+\color{red}
 
@@
 
+\color{black}
+
 We can use the following algorithm implemented in the `sim_study()` function to simulate a single study. Before using the `sim_study()`
@@
 
+\color{red}
 1. Choose a $\Delta$, $n_{T}$, and $n_{C}$ value.
@@
 
+\color{black}
+
 ```{r fun-sim-study, results='asis', echo=FALSE}
@@
 
+\color{red}
 The `es` is the true effect size ($\mu_{\theta}$ for the random-effects model and $\theta$ for the equal-effects model), `nc` and `nt` are the sample size for the control and experimental group, `aggregate` controls if returning the effect size and the corresponding variance or the participant-level data.
@@
 The `mapply()` function is clearly explained in the supplementary materials.
+\color{black}
 
@@
 Using the `sim_study()` function multiple times (with the appropriate adjustments) we can generate a series of studies simulating a data set for a meta-analysis (using the `sim_studies()` function).
-After each example, we will compute the appropriate model (e.g., *fixed* or *random-effects*) using the `metafor` package [@Viechtbauer2010-xz] to check the recovery of simulated parameters^[By default `metafor` use the Wald *z* test on model parameters but other inference methods are available e.g.
+After each example, we will compute the appropriate model (e.g., *fixed* or *random-effects*) using the `metafor` package [@Viechtbauer2010-xz] to check the recovery of simulated parameters^[\color{red} By default `metafor` use the Wald *z* test on model parameters but other inference methods are available e.g.
 the Knapp and Hartung [-@Knapp2003-qx] method by setting `test="knha"` within the `rma()` function.
-This method is useful especially when the number of studies is low (see https://wviechtb.github.io/metafor/reference/misc-recs.html for more details)].
+This method is useful especially when the number of studies is low (see https://wviechtb.github.io/metafor/reference/misc-recs.html for more details) \color{black}].
 Table \@ref(tab:notation-tab) summarizes the simulation parameters notation in equations and code.
@@
 ```
+
+\color{red}
 
@@
 In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies (i.e., the sample size).
-Equations \@ref(eq:equal-effects-model1) and \@ref(eq:equal-effects-model2) depicts the *equal-effects* model.
+Equations \@ref(eq:equal-effects-model1) and \@ref(eq:equal-effects-model2) depicts the *equal-effects* model.}
 
@@
 
+\color{black}
+
 \begin{align}
@@
 
-Each observed effect size ($y_{i}$) is composed of the real effect size $\theta$ plus an error term ($\epsilon_{i}$) that is sampled from a normal distribution with $\mu = 0$ and $\sigma^{2} = \sigma_{\epsilon_i}^{2}$ (i.e., the known sampling variance of the study $i$).
+Each observed effect size \color{red} ($y_{i}$) \color{black} is composed of the real effect size \color{red} $\theta$ \color{black} plus an error term ($\epsilon_{i}$) that is sampled from a normal distribution with $\mu = 0$ and \color{red} $\sigma^{2} = \sigma_{\epsilon_i}^{2}$ \color{black} (i.e., the known sampling variance of the study $i$).
 As demonstrated in Equation \@ref(eq:effsize), the increase in sample size will decrease the sampling variability.
@@
 
+\color{red}
 We are using the $\theta$ parameter to generate random data using the `sim_study()` function that will introduce the random error component $\epsilon_{i}$.
@@
 Increasing the number of studies ($k$) and/or the number of participants in each study ($n$) will improve estimation reducing the standard error.
+\color{black}
 
@@
 
+\color{red}
 The *random-effects model* (RE) can be considered an extension of the *equal-effects model*.
@@
 
+\color{black}
+
 ```{r sim-random-effects}
@@
 We are now estimating two parameters: $\mu_{\theta}$ and $\tau^{2}$.
-As for the *equal-effects model*, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
+As for the \rev{*equal-effects model*}, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
 
@@
 In other words, increasing the sample size of each study or the number of studies will not affect the value of $\tau^{2}$ but only the estimation precision [see @Borenstein2009-mo, Chapter 16, Figure 16.6 for a clear explanation].
-This can be easily demonstrated using the previous simulation, increasing each study\'s sample size.
+This can be easily demonstrated using the previous simulation, increasing each study sample size.
 Figure \@ref(fig:random-effects-high-low-precision) depicts the same meta-analysis but with different precision in estimating the study $y_{i}$.
@@
 The relationship between the sampling error and the heterogeneity can be expressed using the $I^{2}$ statistics [e.g., @Higgins2002-fh] that is the percentage of the total variability $\tau^{2} + \tilde{v}$ that is attributable to real heterogeneity between studies ($\tau^{2}$).
-In Equation \@ref(eq:isquared), $\tilde{v}$ is the "typical" within-study sampling variability as proposed by Higgins and Thompson [-@Higgins2002-fh]^[see also here https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate#fn__1 for an overview of the $I^2$ statistics also for complex models] and implemented in Equation  \@ref(eq:typical-v) with $w_i = 1/\sigma_{\epsilon_i}^{2}$ and $k$ is the number of studies.
+\color{red} In Equation \@ref(eq:isquared), $\tilde{v}$ is the "typical" within-study sampling variability as proposed by Higgins and Thompson [-@Higgins2002-fh]^[see also here https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate#fn__1 for an overview of the $I^2$ statistics also for complex models] and implemented in Equation  \@ref(eq:typical-v) with $w_i = 1/\sigma_{\epsilon_i}^{2}$ and $k$ is the number of studies.
+
+\color{black}
 
@@
 
+\color{red}
 \begin{align}
@@
 \end{align}
+\color{black}
 
 From Equation \@ref(eq:isquared) and Figure \@ref(fig:random-effects-high-low-precision) is clear that if each included study has a considerable sample size the total variability will be reduced, mainly driven by real heterogeneity ($\tau^2$).
-This is the crucial difference between the equal-effects and the random-effects model [see also @Borenstein2009-mo, pp.
+This is the crucial difference between the \rev{equal-effects} and the random-effects model [see also @Borenstein2009-mo, pp.
 117–122].
@@
 When the sample size of each study is the same, $\tilde{v} = \sigma^2_i$.
-In the other case, $\tilde{v}$ needs to be calculated from sampling variances reported in Higgins and Thompson [-@Higgins2002-fh] and cannot be easily fixed a priori.
+In the other case, $\tilde{v}$ needs to be calculated from sampling variances as reported in Higgins and Thompson [-@Higgins2002-fh] and cannot be easily fixed a priori.
 However, with the assumption of homogeneous sample size across studies^[The homogeneous sample size of primary studies assumption is commonly used to calculate the statistical power [see @Borenstein2009-mo, Chapter 29]], we can solve Equation \@ref(eq:isquared) for $\tau^{2}$ obtaining the heterogeneity value associated with a certain $I^{2}$ as reported in Equation \@ref(eq:tau2-from-isquared).
@@
 Beyond fixing the $I^2$ value, one could use a plausible value from the literature.
-For example van Erp and colleagues [@Van_Erp2017-cf] estimate an empirical $\tau^2$ distribution across several published meta-analyses.
+\color{red} For example van Erp and colleagues [@Van_Erp2017-cf] estimate an empirical $\tau^2$ distribution across several published meta-analyses.
+\color{black}
+
+<!-- TODO add something on the simulation heterogeneous vs homogeneous sample sizes -->
 
@@
 In regression terms (see equation \@ref(eq:meta-regression)), we could use a dummy variable ($X_{1}$) that takes the value of 0 for the *lab-based task* (*L*) and a value of 1 for an *online task* (*O*)^[The *dummy coding* (also known as treatment coding) is the default in R and `metafor` but other coding schemes could be used [see @Schad2020-ht for an overview of contrast coding schemes].
-As in standard linear regression, the contrast coding of categorical factors or the presence/absence of the intercept influence the interpretation of estimated model parameters and omnibus meta-regression tests (see https://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept for a detailed discussion)].
+\color{red} As in standard linear regression, the contrast coding of categorical factors or the presence/absence of the intercept influence the interpretation of estimated model parameters and omnibus meta-regression tests (see https://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept for a detailed discussion)] \color{black}.
 Now we fix $\beta_{1}$ to be the *lab vs.
-online effect* (i.e., the expected mean difference between the two groups of studies) and the product between $\beta_{1}X_{1}$ will consider the *lab vs.
+online effect* (i.e., the expected mean difference between the two groups of studies) and the product between \rev{$\beta_{1}X_{1_i}$} will consider the *lab vs.
 online effect*.
 Crucially, despite $\tau^{2}$ is still the heterogeneity between effect sizes, now we need to fix $\tau^{2}$ considering that we included a moderator.
-In other terms, $\delta_i \sim \mathcal{N}(0, \tau_{r}^{2})$ where $\tau^2_r$ is the residual heterogeneity after including the moderator.
+In other terms, \color{red} $\delta_i \sim \mathcal{N}(0, \tau_{r}^{2})$ \color{black} where $\tau^2_r$ is the residual heterogeneity after including the moderator.
 We can describe our model using equation \@ref(eq:meta-regression-equations) according to the value of $X_{1i}$ (L = lab-based experiments, O = online experiments).
@@
 \begin{gathered}
-y_i = \beta_0 + \delta_i + \beta_1 X_{1i} + \epsilon_i
+y_i = \beta_0 + \delta_i + \beta_1 X_{1_i} + \epsilon_i
 (\#eq:meta-regression)
@@
 Then the $\beta_1$ parameter represents the estimated difference in $y$ between the values of $X_1$ (i.e., lab-based vs online experiments).
-As said before, $\tau^2_r$ is now the residual heterogeneity that is interpreted as the variability between effect sizes after controlling for the moderator $X_1$.
-
-Now we can fit the meta-regression model with the `rma` function as for the random-effects model with the addition of mods = `~ exp` that indicates which variable/s to consider as moderators.
-The results are presented in Table \@ref(tab:res-meta-reg-dummy).
-Next, the model will estimate an *intercept* parameter (i.e., $\beta_{0}$) that is the value of $y$ when $X_{1}$ is zero (i.e., for lab-based studies) or, in other terms, the expected value for lab-based studies.
-Then the $\beta_{1}$ parameter represents the estimated difference in $y$ between the values of $X_{1}$ (i.e., lab-based vs.
-online experiments).
-As mentioned above, $\tau^{2}_r$ the residual heterogeneity is now interpreted as the variability between effect sizes after controlling for the moderator $X_{1}$.
+As said before, \color{red} $\tau^2_r$ \color{black} is now the residual heterogeneity that is interpreted as the variability between effect sizes after controlling for the moderator $X_1$.
 
@@
 A first strategy could be to use values estimated from the literature.
-Another approach consists in setting up the model and simulating several expected $y_i$ and calculating the range of simulated values.
+Another approach consists in setting up the model and simulating several expected \color{red} $y_i$ \color{black} and calculating the range of simulated values.
 A third possibility is fixing the proportion of explained heterogeneity calculating the $\beta_1$ value accordingly.
@@
 In this specific example, if all studies target a specific population (e.g., adults below 50 years), the expected average age range can be easily simulated.
-In our case, we simulated $k$ average age values from a uniform distribution $\text{ag}e_{i} \sim U\left( 20,40 \right)$.
+In our case, we simulated $k$ average age values from a uniform distribution $\text{age}_{i} \sim U\left( 20,40 \right)$.
 Then we can plot the distribution of $d_{i}$ values to check the plausibility of simulated values.
@@
 
-(ref:img-equal-vs-random) The difference between the assumptions of the *equal* and *random-effects model*.
+(ref:img-equal-vs-random) \color{red} The difference between the assumptions of the *equal* and *random-effects model*.
 Each distribution depicts the sampling distribution of $k = 4$ hypothetical studies ($i = 1$, $2$, ..., $4$) with a certain observed effect size $y_{i}$ (red squares), sampling variability $\sigma_{i}^{2}$ and the 95% confidence interval (black segment).
@@
 When $\tau^{2}$ is zero, the random-effects reduces to a equal-effects model.
-
-(ref:img-multilevel) Graphical representation of a *multilevel* (three-level) model.
-On the top, the effect sizes are distributed at the population level.
-Then, each study is defined as the overall effect and the random effects determined by $\tau^{2}$, $\theta_{r} + \theta_{i}$.
-Finally, each experiment (effect size) within a study is composed of the study-specific effect $\theta_{r} + \theta_{i}$ and the experiment-specific random-effects $\theta_{r} + \theta_{i} + \theta_{\text{ij}}$ determined by $\omega^{2}$.
-
-(ref:img-multivariate) Graphical representation of a *multivariate* model.
-In this example, we have two outcomes (magenta and green) with an average effect, heterogeneity parameters, and a certain true correlation value ($\rho$, third type of dependency).
-Each study at level-2 (as in the standard random-effects model) has two effect sizes with corresponding sampling variances.
-The two effect sizes are correlated ($\rho_{s}$, with $s$ for sampling errors) because each effect is collected on the same participants.
+\color{black}
 
@@
 
-(ref:res-equal-effects) Summary of the simulated equal-effects model. The only estimated parameter is the average effect* $\theta_{f}$ *with the standard error,
-95% confidence interval, and the Wald z-test. The z-test evaluates the null hypothesis that the real effect equals zero.
-
-(ref:equal-effects-high-low-precision) Forest plots of two simulated *equal-effects models*.
+(ref:res-equal-effects) \color{red} Summary of the simulated equal-effects model. The only estimated parameter is the average effect* $\theta_{f}$ *with the standard error,
+95% confidence interval, and the Wald z-test. The z-test evaluates the null hypothesis that the real effect equals zero.\color{black}
+
+(ref:equal-effects-high-low-precision) \color{red} Forest plots of two simulated *equal-effects models*.
 On the left, the simulated model has $nt,nc = 30$ for each included study while on the right the sample size for each study is $nt,nc = 500$.
@@
 The average effect size is similar (depending on the random numbers generation) between the two scenarios while the estimation precision (the width of the black diamond) is narrower on the right.
-
-(ref:res-random-effects) Summary of the simulated random-effects model: Compared to the equal-effects model, there are more parameters.
+\color{black}
+
+(ref:res-random-effects) \color{red} Summary of the simulated random-effects model: Compared to the equal-effects model, there are more parameters.
 The $\beta$ is the average effect ($\theta_{r}$) with the standard error, 95% confidence interval, and the Wald z-test.
 The $\tau^{2}$ is the estimated heterogeneity and the $I^{2}$ (explained in Section \@ref(re-model)) represents the percentage of total variability due to between-study heterogeneity.
-
-(ref:random-effects-high-low-precision) Forest plots of two simulated *random-effects model*s.
+\color{black}
+
+(ref:random-effects-high-low-precision) \color{red} Forest plots of two simulated *random-effects model*s.
 On the left, the simulated model has $nt,nc = 30$ for each included study, while on the right, the sample size for each study is $nt,nc = 500$.
@@
 However, compared to the *equal-effects* simulation, increasing the sample size of primary studies only affects the precision without reducing the true heterogeneity (i.e., $\tau^{2} \neq 0$).
-
-(ref:res-random-effects-i2) Summary of the *random-effect* model fixing the $I^2$ value. Parameters are the same as described in Table \@ref(tab:res-random-effects).
+\color{black}
+
+(ref:res-random-effects-i2) Summary of the *random-effects* model fixing the $I^2$ value. Parameters are the same as described in Table \@ref(tab:res-random-effects).
 
@@
 
-(ref:res-meta-reg-dummy) Summary of the *random-effect* model with a categorical predictor (lab vs online experiments).
+(ref:res-meta-reg-dummy) \color{red} Summary of the *random-effect* model with a categorical predictor (lab vs online experiments).
 The `intercept` is the average effect for lab-based experiments and `exponline` is the difference between lab-based and online experiments.
@@
 Other parameters are the same as the standard *random-effect* model (see Table \@ref(tab:res-random-effects)).
-
-(ref:res-meta-reg-numerical-r2) Summary of the *random-effects* model fixing the $R^{2}$ value.
+\color{black}
+
+(ref:res-meta-reg-numerical-r2) \color{red} Summary of the *random-effects* model fixing the $R^{2}$ value.
 The `intercept` is the effect size for the average `age` (given that age is mean-centered).
@@
 Other parameters are the same as described in Table \@ref(tab:res-random-effects) and \@ref(tab:res-meta-reg-dummy).
-
-(ref:plot-meta-reg-plausible-implausible) Scatter plots with marginal histograms for the range of simulated $y_i$ with two $\beta_1$ values.
+\color{black}
+
+(ref:plot-meta-reg-plausible-implausible) \color{red} Scatter plots with marginal histograms for the range of simulated $y_i$ with two $\beta_1$ values.
 The x-axis depicts the average age of simulated studies and the y-axis depicts the simulated effect size.
@@
 On the right, ($\beta = 0.7$) values range between -10 to 10 that despite theoretically possible is highly implausible values in real meta-analysis for psychological data.
+\color{black}
 

### 2023-07-07 documents/paper/paper.Rmd
changes in red (addign oberved effect sizes into captions)
fixing typos

@@
 within-subjects) and available information and find the proper formula to compute the sampling variability.
-In addition, there are formulas and approaches to convert from one effect size measure to another [@Borenstein2009-mo; @Lipsey2001-vp; @Lakens2013-eq].^[See also the blog post by James Pustejovsky ([link](https://www.jepusto.com/alternative-formulas-for-the-smd/)) for a general method to compute the sampling variability for standardized mean difference measures.] Usually, the effect size sampling variability depends mainly on the sample size that determines the weight assigned during the meta-analytic estimation.
+In addition, there are formulas and approaches to convert from one effect size measure to another [@Borenstein2009-mo; @Lipsey2001-vp; @Lakens2013-eq].^[See also the blog post by James Pustejovsky ([www.jepusto.com/alternative-formulas-for-the-smd](https://www.jepusto.com/alternative-formulas-for-the-smd/)) for a general method to compute the sampling variability for standardized mean difference measures.] Usually, the effect size sampling variability depends mainly on the sample size that determines the weight assigned during the meta-analytic estimation.
 
@@
 Before diving into the specific simulations, in this section, we define the common aspects of all simulations in the following sections.
-The current paper focuses on the *two-level*, *fixed,* and *random-effects* models.
+The current paper focuses on the *two-level*, *equal,* and *random-effects* models.
 All the examples refer to primary studies that assess the efficacy of a treatment by comparing a control and an experimental group.
 \color{red} There are different simulation approaches to meta-analysis concerning how to generate study-level or participant-level data.
-In simulation studies where the purpose is not evaluating effect size estimators, it is convenient to simulate unstandardized effect size measures (e.g., UMD).
+In simulation studies where the purpose is not evaluating effect size estimators, it is convenient to simulate unstandardized effect size measures (e.g., unstandardized mean difference, UMD).
 Firstly the estimator is unbiased thus not requiring small sample correction [e.g., @Hedges1989-ip; @Hedges1981-za].
@@
 
-The most basic model to simulate is the *equal-effects model* (FE).
+The most basic model to simulate is the *equal-effects model* (EE).
 As reported in the previous sections, the *equal-effects model* assumes the presence of a single true effect ($\theta$) and the observed variability is because each effect size is an imprecise estimation of the true effect.
@@
 sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim) # simulated data
-res <- rma(yi = yi, vi = vi, method = "FE", data = sim)
+res <- rma(yi = yi, vi = vi, method = "EE", data = sim)
 ```
@@
 
-From a linear regression perspective, both the FE and the RE models can be seen as *intercept-only* models where only the mean (i.e., the linear regression intercept) is estimated Borenstein et al.
+From a linear regression perspective, both the EE and the RE models can be seen as *intercept-only* models where only the mean (i.e., the linear regression intercept) is estimated Borenstein et al.
 (2009).
@@
 y_{iL} = \beta_0 + \delta_i + \beta_1 \times 0 + \epsilon_i \\
-d_{iO} = \beta_0 + \delta_i + \beta_1 \times 1 + \epsilon_i
+y_{iO} = \beta_0 + \delta_i + \beta_1 \times 1 + \epsilon_i
 (\#eq:meta-regression-equations)
@@
 
-As explained in the introduction, there are several approaches and tools to estimate the power of *fixed* and *random-effects models* [see @Borenstein2009-mo, Chapter 29; @Harrer2019-rl, Chapter 14].
+As explained in the introduction, there are several approaches and tools to estimate the power of *equal* and *random-effects models* [see @Borenstein2009-mo, Chapter 29; @Harrer2019-rl, Chapter 14].
 These methods are easy to implement but made strong assumptions, such as the homogeneity of sample size, and did not consider the uncertainty in estimating $\tau^{2}$.
@@
 
+The supplementary materials can be also found on the Open Science Framework repository (https://osf.io/54djn/)
+
 ### Prior versions {-}
@@
 
-(ref:res-equal-effects) \color{red} Summary of the simulated equal-effects model. The only estimated parameter is the average effect* $\theta_{f}$ *with the standard error,
-95% confidence interval, and the Wald z-test. The z-test evaluates the null hypothesis that the real effect equals zero.\color{black}
+(ref:res-equal-effects) \color{red} Summary of the simulated equal-effects model.
+The only estimated parameter is the average effect $\theta$ ($\beta$) with the standard error, 95% confidence interval, and the Wald z-test.
+The z-test evaluates the null hypothesis that the real effect equals zero.\color{black}
 
@@
 (ref:res-random-effects) \color{red} Summary of the simulated random-effects model: Compared to the equal-effects model, there are more parameters.
-The $\beta$ is the average effect ($\theta_{r}$) with the standard error, 95% confidence interval, and the Wald z-test.
+The $\beta$ is the average effect ($\mu_{\theta}$) with the standard error, 95% confidence interval, and the Wald z-test.
 The $\tau^{2}$ is the estimated heterogeneity and the $I^{2}$ (explained in Section \@ref(re-model)) represents the percentage of total variability due to between-study heterogeneity.
@@
 The green segments depict the explained heterogeneity, and the red segments the residual (unexplained) heterogeneity.
+\color{red} The red squares are simulated observed effect sizes from the sampling distribution.
+\color{black}
 

### 2023-07-18 README.Rmd
add tests and fix packages

@@
   collapse = TRUE,
+  message = FALSE,
+  warning = FALSE,
   comment = "#>",
@@
 devtools::load_all()
+pkgs <- simulatingMeta:::.get_packages()
 ```
@@
 
+## Testing
+
+The custom functions are tested within the `scripts/test.R` script.
+The idea is to use the same simulation codes of the paper increasing the $k$ and $n$ values and check the recovery of simulated parameters using the `check_sim()` function.
+
 # Session
@@
 ```{r, echo = FALSE, message=FALSE, warning=FALSE, results='asis'}
-pkgs <- get_packages()
 sprintf("- `%s` (%s)", pkgs$pkg, pkgs$version) |> 

### 2023-07-18 R/sim-utils.R
unstandardized effect sizes implementation

@@ -2,36 +2,29 @@
 
 #' sim_study
 #' @description Simulate a single study based on the input parameters and return a dataframe with effect size and variance or raw data.
-#' @param theta number indicating the population level effect size.
-#' @param nt number indicating the number of participants for the treated group.
+#' @param es number indicating the population level effect size.
+#' @param nt number indicating the number of participants for the experimental group.
 #' @param nc number indicating the number of participants for the control group. If \code{NULL} the value assigned to \code{nt} will be used.
-#' @param hedges logical. Whether the Hedges' correction should be applied. Default to \code{TRUE}.
-#' @param aggregate logical. Whether the participants data need to be aggregated calculating the effect size and variance. Default to \code{TRUE}.
+#' @param aggregate logical. Whether the participants data need to be aggregated calculating the effect size and variance or not. Default to \code{TRUE}.
 #'
-#' @return a dataframe
+#' @return dataframe
 #' @import metafor
 #' @export
 #' @examples
-#' sim_study(0.3, 30, 30, TRUE, TRUE) # return aggregated data
-#' sim_study(0.3, 30, 30, TRUE, TRUE) # return participants data
+#' sim_study(0.3, 30, 30, aggregate = TRUE) # return aggregated data
+#' sim_study(0.3, 30, 30, aggregate = FALSE) # return participants data
 
-sim_study <- function(theta, nt, nc = NULL, hedges = TRUE, aggregate = TRUE){
+sim_study <- function(es, nt, nc = NULL, aggregate = TRUE){
   if(is.null(nc)) nc <- nt
   # generate from normal distribution
   yc <- rnorm(nc, 0, 1)
-  yt <- rnorm(nt, theta, 1)
-  # pooled variance
-  sp <- sqrt((var(yc)*(nc - 1) + var(yt)*(nt - 1)) / (nc + nt - 2))
-  # effect size
-  yi <- (mean(yt) - mean(yc)) / sp
+  yt <- rnorm(nt, es, 1)
   
-  # transform to hedges' g
-  if(hedges){
[…]

### 2023-07-18 R/utils.R
cleaning

@@ -1,14 +1,5 @@
 # Utils functions used in the paper and supplementary materials -----------
 
-#' svg_to_pdf
-#' @description convert all svg to pdf
-svg_to_pdf <- function(){
-  all_svg <- list.files(here::here("docs", "paper-draft", "img"), 
-                        pattern = ".svg", full.names = TRUE, recursive = TRUE)
-  out_names <- gsub(".svg", ".pdf", all_svg)
-  purrr::walk2(all_svg, out_names, rsvg::rsvg_pdf)
-}
-
 #' get_caption
 #' @description create text reference for captions
 get_caption <- function() {
@@ -153,23 +144,40 @@ as_tex <- function(x){
 
 
 #' check_sim
-#'
+#' @description
+#' The function compare a (named) list of simulation parameters with a (named) list of simulations results
+#' computing the absolute value of the difference
+#' 
 #' @param fixed named list of population level parameters
 #' @param results named list of simulation results
-#'
+#' @param name optional name for the simulation to be displayed
 #' @return NULL
 #' @import cli
 #' @export
 #'
-check_sim <- function(fixed, results){
+check_sim <- function(fixed, results, name = NULL){
   results <- results[names(fixed)] # arrange
   diff <- abs(unlist(results) - unlist(fixed))
   out <- sprintf("%s: Fixed = %.3f, Simulated = %.3f --> %s = %.3f",
                  cli::col_blue(names(results)), fixed, results, cli::col_magenta("|Fixed - Simulated|"), diff)
-  cat(out, sep = "\n\n")
+  
+  cli::cli_rule()
+  if(!is.null(name)){
+    name <- sprintf("Simulation: %s", cli::col_green(name))
+    cat(name, out, sep = "\n\n")
+  }else{
+    cat(out, sep = "\n\n")
+  }
+  cli::cli_rule()
 }
 
-get_packages <- function(){
+#' .get_packages
+#' @description
+#' Get all packages with the corresponding version used in the current project 
+#' @importFrom renv dependencies
+#' @return dataframe
+#' 
+.get_packages <- function(){
   options(renv.verbose = FALSE)
   pkgs <- renv::dependencies()
   pkgs <- unique(pkgs$Package)

### 2023-07-18 scripts/r-img.R
fix equations

@@ -30,11 +30,11 @@ equal_effects$note <- sprintf("$y_i = \\theta + \\epsilon_i$")
 random_effect <- data.frame(
   id = 1:4,
   yi = c(-0.1, 0.4, 0.7, 0.25),
-  vi = c(0.05, 0.08, 0.1, 0.2),
+  vi = c(0.01, 0.02, 0.1, 0.05),
   model = "Random-Effects"
 )
 
-random_effect$note <- sprintf("$y_i = \\mu_{\\theta} + \\theta_i + \\epsilon_i$")
+random_effect$note <- sprintf("$y_i = \\mu_{\\theta} + \\delta_i + \\epsilon_i$")
 
 random_effect$di <- mean(random_effect$yi)
 
@@ -194,7 +194,7 @@ plot_metareg_bin <- ggplot(binary_metareg,
              size = 6, label.size = NA) +
   theme(axis.text.x = element_blank()) +
   theme_rfig() +
-  xlab(latex2exp::TeX("$d_i$")) +
+  xlab(latex2exp::TeX("$y_i$")) +
   ggtitle("Categorical Predictor") +
   xlim(xlim) +
   annotate("text", x = 1, y = 1.5, 

### 2023-07-18 scripts/test.R
adding simulation tests

@@ -1,0 +1,318 @@
+# Test Simulations
+
+library(metafor)
+devtools::load_all()
+seed <- 2023
+
+# Equal-Effects Model -----------------------------------------------------
+
+set.seed(seed)
+k <- 1e3 # number of studies
+n <- 1e3 # number of participants per group, per study
+theta <- 0.3 # real effect size
+
+sim <- make_data(k = k, nt = n, nc = n, es = theta)
+sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim) # simulated data
+res <- rma(yi = yi, vi = vi, method = "EE", data = sim)
+
+check_sim(fixed = list(theta = theta), 
+          results = list(theta = res$b[[1]]),
+          name = "Equal-Effects Model")
+
+# Random-effects Model ----------------------------------------------------
+
+set.seed(seed)
+k <- 1e3 # number of studies
+n <- 1e3 # number of participants per group, per study
+mu <- 0.3 # real effect size
+tau2 <- 0.2 # the effect size heterogeneity
+
+sim <- make_data(k = k, nc = n, nt = n, mu = mu)
+
+# simulate the random-effects adjustment
+sim$deltai <- rnorm(k, 0, sqrt(tau2))
+
+# adding the by-study adjustment to the average effect
+sim$es <- sim$mu + sim$deltai 
+
+# now we are using mu_deltai and no longer theta
+sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
+res <- rma(yi = yi, vi = vi, method = "REML", data = sim)
+
+check_sim(fixed = list(theta = theta,
+                       tau2 = tau2), 
+          results = list(theta = res$b[[1]],
+                         tau2 = res$tau2),
+          name = "Random-effects Model")
+
+# Random-effects model fixing I2 ------------------------------------------
+
+set.seed(seed)
+theta <- 0.3
+k <- 1e3 # number of studies
+mu <- 0.3 # real effect size
+n <- 1e3 # sample size per group, per study
+I2 <- 0.6 # desired I2 value
+
+v <- (n + n)/(n * n) + mu^2/(2 * (n + n - 2)) # typical within study variance
+tau2 <- -((I2*v)/(I2 - 1))
+
+sim <- make_data(k = k, nc = n, nt = n, mu = mu)
+
+sim$deltai <- rnorm(k, 0, sqrt(tau2))
+sim$es <- sim$mu + sim$deltai 
[…]

### 2023-07-18 documents/supplementary/supplementary.Rmd
fixing typos and revision

@@
     toc: true
+    toc_depth: 1
     latex_engine: xelatex
@@
   - \usepackage{xcolor}
+  - \newcommand{\rev}[1]{\textcolor{red}{#1}}
 ---
@@
 devtools::load_all() # loading all custom functions
+
+# theme for plots
+theme_paper <- function(base_size = 15){
+  ltxplot::theme_latex(base_size = base_size)
+}
 ```
@@
 
-In the main paper, we presented the basic meta-analysis models and simulations.
+In the main paper, we presented the most straightforward situation for a meta-analysis when each included study contributes with a single effect size.
 However, in real-world meta-analyses, it is common to find complex data structures such as multilevel and multivariate.
-As reported in the paper, there could be a different source of dependencies when each study reports multiple effect sizes.
-Rephrasing the [@Cheung2015-bw, pp.
-121--122] taxonomy, the first type of dependency, which we call *multilevel* dependency, arises when multiple independent experiments are associated with the same study.
-Level 1 (i.e., participants' level) is different, but the effects are likely correlated because the same research group conducted the experiments.
-The second type of dependency, *multivariate* dependency, arises when multiple effects are collected on the same pool of participants within each study.
-Now sampling errors within each study are correlated because the same participant is tested multiple times.
-Finally, we have a dependency on multiple effects at the population level.
-Two outcomes could be correlated (e.g., the type of
-memory tasks) even if the pool of participants is different.
-The first two types of dependency are the reasons to use a multilevel and multivariate model.
+There are numerous sources of dependency for a meta-analysis [@Cheung2014-fg; @Cheung2015-bw, pp.
+121-122].
+
+The first type of dependency (*multilevel*) consists of multiple effect sizes from independent participants.
+Despite being collected with different participants, effect sizes from the same study could be correlated.
+We obtain a multilevel data structure with independent effect sizes in the same study.
+Figure \@ref(fig:img-multilevel) depict the general idea of a *multivariate* model.
+There are multiple ways to model this data structure [see @Cheung2014-fg for an overview].
+Still, the most common way is to use a *multilevel meta-analysis model,* considering the heterogeneity between and within studies.
+The standard *random-effects* (but also the *equal-effects model*) model is also called a *two-level* model because it combines data after aggregating the participants-level (*first-level*) data.
+The *three-level* model is commonly used to manage the multilevel data structure [@Cheung2014-fg; @Cheung2019-po].
+Compared to the *two-level* model, the *three-level* model estimates two heterogeneity parameters ($\tau^{2}$ and $\omega^{2}$), one for the heterogeneity between studies and the other for the heterogeneity within studies.
+
+A second type of dependency (*multivariate*) consists of multiple effect sizes collected from the same pool of participants.
+Figure \@ref(fig:img-multivariate) depict the general idea of a *multivariate* model.
+For example, when using two cognitive tasks on the same pool of participants.
+This is commonly known as *multivariate meta-analysis* because the effect sizes sampling errors are correlated given that participants are assessed multiple times.
+A *multivariate meta-analysis model* could be used to
+consider the correlation between sampling errors.
+
+A third type of dependency arises when true effect sizes are correlated at the population level [@Cheung2015-bw, pp.
+121-122].
+For example, two different memory tasks could be correlated because the latent psychological constructs are intrinsically correlated.
+This correlation is present even when other groups of participants perform the two memory tasks (i.e., no *multivariate* dependency) or are administered by different researchers (i.e., no *multilevel* dependency).
+This type of dependency could be managed by estimating the correlation between outcomes using a *multivariate random-effects model* [@Cheung2015-bw, pp.
+127-130].
+Figures \@ref(fig:img-multilevel) and \@ref(fig:img-multivariate) depict the conceptual idea of *multilevel* and *multivariate* models.
 The following sections will show two examples to simulate these complex data structures.
+
+```{r img-multilevel, echo=FALSE, out.width="80%", fig.cap = get_caption()}
+knitr::include_graphics("img/multilevel.pdf")
+```
+
+```{r img-multivariate, echo=FALSE, out.width="80%", fig.cap = get_caption()}
+knitr::include_graphics("img/multivariate.pdf ")
+```
 
@@
 In the paper, we used the `mapply()` function, a particularly useful tool in R, to apply the same function (in parallel) with multiple parameters using multiple vectors.
-In our case, the `sim_study()` function has several parameters (`theta`, `nt`, among others), and we can apply the function multiple times with a combination of parameters.
+In our case, the `sim_study()` function has several parameters (`es`, `nt`, among others), and we can apply the function multiple times with a combination of parameters.
 It can be considered essentially a `for` loop but more compact and less verbose.
@@
 for(i in 1:length(res)){
-  res[[i]] <- sim_study(theta = sim$theta[i], nc = sim$nc[i], nt = sim$nt[i])
+  res[[i]] <- sim_study(theta = sim$es[i], nc = sim$nc[i], nt = sim$nt[i])
 }
@@
 # using mapply
-sim <- mapply(sim_study, sim$theta, sim$nc, sim$nt)
+sim <- mapply(sim_study, sim$es, sim$nc, sim$nt)
 
 # using sim_studies (with mapply inside)
-sim <- sim_studies(sim$theta, sim$nc, sim$nt, data = sim)
+sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
 ```
@@
 \begin{gathered}
-d_{ij} = \theta_r + \theta_i + \theta_{ij} + \epsilon_{ij} \\
-\theta_i \sim N(0, \tau^2) \\
-\theta_{ij} \sim N(0, \omega^2) \\
-\epsilon_{ij} \sim N(0,\sigma^2_i)
-(\#eq:three-level-model)
+y_{ij} = \mu_{\theta} + \delta_i + \zeta_{ij} + \epsilon_{ij}
+(\#eq:three-level-model1)
+\end{gathered}
+\end{align}
+
+\begin{align}
+\begin{gathered}
+\delta_i\sim N(0, \tau^2)
+(\#eq:three-level-model2)
+\end{gathered}
+\end{align}
+
+\begin{align}
+\begin{gathered}
+\zeta_{ij} \sim N(0, \omega^2)
+(\#eq:three-level-model3)
+\end{gathered}
+\end{align}
+
+\begin{align}
+\begin{gathered}
+\epsilon_{ij} \sim N(0,\sigma_{\epsilon_i}^{2})
+(\#eq:three-level-model4)
 \end{gathered}
@@
 The notation suggests that each effect size belongs to an experiment (effect size) ($j$) nested within a study ($i$).
-The variability between studies (level 3) is handled by the $\tau^{2}$ parameter and the variability within studies (level 2) is handled by the $\omega^{2}$ parameter.
-Practically, each study is composed of the average effect size $\theta_{r}$ plus a study-specific adjustment determined by $\tau^{2}$.
-Then each effect size is composed of the study-specific
-effect ($\theta_{r} + \theta_{i}$) plus the adjustment determined by $\omega^{2}$ thus $\theta_{r} + \theta_{i} + \theta_{\text{ij}}$.
+The variability between studies (level 3) is handled by the $\tau^{2}$ parameter (see Equation \@ref(eq:three-level-model2) and the variability within studies (level 2) is handled by the $\omega^{2}$ parameter (see Equation \@ref(eq:three-level-model3)).
+Practically, each study is composed of the average effect size $\mu_{\theta}$ plus a study-specific adjustment determined by $\tau^{2}$.
+Then each effect size is composed of the study-specific effect ($\mu_{\theta} + \delta_{i}$) plus the adjustment determined by $\omega^{2}$ thus $\mu_{\theta} + \delta_{i} + \zeta_{ij}$.
+As in the two-level meta-analysis model there is also the random error component $\sigma_{\epsilon_i}^{2}$ (see Equation \@ref(eq:three-level-model4)).
 The number of effect sizes within each study will determine the estimation precision of the $\omega^{2}$ parameter and the number of studies will determine the estimation precision of the $\tau^{2}$ parameter.
-The
-relationship between $\tau^{2}$ and $\omega^{2}$ can be expressed using the intraclass correlation coefficient (ICC) expressed as $ICC = \frac{\tau^{2}}{\tau^{2} + \omega^{2}}$ representing the proportion of total heterogeneity associated with level 2 or level 3.
+The relationship between $\tau^{2}$ and $\omega^{2}$ can be expressed using the intraclass correlation coefficient (ICC) expressed as $ICC = \frac{\tau^{2}}{\tau^{2} + \omega^{2}}$ representing the proportion of total heterogeneity associated with level 2 or level 3 [see @Cheung2014-fg].
 For example, an ICC of 0.5 suggests that 50% of heterogeneity is caused by between-studies variability.
-On the extreme, an ICC close to 1 suggests that the between-studies variability causes most heterogeneity
-(i.e., the effect sizes within each study are very similar).
+On the extreme, an ICC close to 1 suggests that the between-studies variability causes most heterogeneity (i.e., the effect sizes within each study are very similar).
 
@@
 n <- 30 # sample size for each group, within each study
-theta_r <- 0.3 # real average effect size
+mu <- 0.3 # real average effect size
 
-theta_i <- rnorm(i, 0, sqrt(tau2))
-theta_ij <- rnorm(i*j, 0, sqrt(omega2))
+delta_i <- rnorm(i, 0, sqrt(tau2))
+zeta_ij <- rnorm(i*j, 0, sqrt(omega2))
 
@@
   nc = n,
-  theta_r = theta_r
+  mu = mu
 )
 
-sim$theta_i <- theta_i[sim$study]
-sim$theta_ij <- theta_ij
+sim$delta_i <- delta_i[sim$study]
+sim$zeta_ij <- zeta_ij
 
@@
 
-Each study $i$ is repeated $j$ times^[To create a more realistic simulation, the number of effect sizes for each paper could be heterogeneous (e.g., sampled from a Poisson distribution)] along the $\theta_{i}$ while each effect has a unique $\theta_{\text{ij}}$.
+Each study $i$ is repeated $j$ times^[To create a more realistic simulation, the number of effect sizes for each paper could be heterogeneous (e.g., sampled from a Poisson distribution)] along the $\delta_{i}$ while each effect has a unique $\zeta_{\text{ij}}$.
 We can use the same approach as the main paper using the `sim_studies()` function.
@@
 set.seed(seed)
-sim$theta_r_i_ij <- sim$theta_r + sim$theta_i + sim$theta_ij
-sim <- sim_studies(theta = sim$theta_r_i_ij, nt = sim$nt, nc = sim$nc, data = sim)
+sim$es <- sim$mu + sim$delta_i + sim$zeta_ij
+sim <- sim_studies(es = sim$es, nt = sim$nt, nc = sim$nc, data = sim)
 ```
@@
 
-As explained above, the model estimates the average effect $\theta_{r}$ and the two heterogeneity parameters.
+As explained above, the model estimates the average effect $\mu_{\theta}$ and the two heterogeneity parameters.
 To create a more realistic simulation, we can create some variability in the number of experiments within the same paper, similar to what we did for the number of participants in the main paper.
@@
 Following the example of the main paper, we have three memory tests collected on the same pool of participants.
-For this reason we have three real effects ($\theta_{r_{1}}$, $\theta_{r_{2}}$ and $\theta_{r_{3}}$), three heterogeneity ($\tau^2_1$, $\tau^2_2$ and $\tau^2_3$) and the population-level correlations between these outcomes ($\rho_{12}$, $\rho_{13}$ and $\rho_{23}$).
-Moreover, the sampling errors ($\epsilon_i$) are now correlated; thus each study will have three variances for each outcome ($\sigma^2_1$, $\sigma^2_2$, and $\sigma^2_3$), and the correlations between sampling errors ($\rho_{s12}$, $\rho_{s13}$ and $\rho_{s23}$)^[We used the subscript $s$ for sampling errors to distinguish between between-outcomes correlations and correlations between sampling errors].
-Equation \@ref(eq:multivariate-model) formalize the multivariate model for a single study $i$ with three outcomes.
+For this reason we have three real effects ($\mu_{\theta_{1}}$, $\mu_{\theta_{2}}$ and $\mu_{\theta_{3}}$), three heterogeneity ($\tau^2_1$, $\tau^2_2$ and $\tau^2_3$) and the population-level correlations between these outcomes ($\rho_{12}$, $\rho_{13}$ and $\rho_{23}$).
+Moreover, the sampling errors ($\epsilon_i$) are now correlated; thus each study will have three sampling variances for each outcome ($\sigma_{\epsilon_1}^{2}$, $\sigma_{\epsilon_1}^{2}$, and $\sigma_{\epsilon_1}^{2}$), and the correlations between sampling errors ($\rho_{s12}$, $\rho_{s13}$ and $\rho_{s23}$)^[We used the subscript $s$ for sampling errors to distinguish between between-outcomes correlations and correlations between sampling errors].
+Equations \@ref(eq:multivariate-model1), \@ref(eq:multivariate-model2), \@ref(eq:multivariate-model3), \@ref(eq:multivariate-model4) \@ref(eq:multivariate-model5) formalize the multivariate model for a single study $i$ with three outcomes.
 The random-effects adjustments ($\theta_i$) to the overall effect and sampling errors are now sampled from a multivariate normal distribution, respectively $\theta_i \sim \mathcal{MVN}(0, \mathrm{T}^2)$ and $\theta_i \sim \mathcal{MVN}(0, \mathrm{V}^2)$.
@@
 \begin{bmatrix}
-d_{i1} \\
-d_{i2} \\
-d_{i3}
+y_{1_i} \\
+y_{2_i} \\
+y_{3_i}
 \end{bmatrix} 
@@
 \begin{bmatrix}
-\theta_{1} \\
-\theta_{2} \\
-\theta_{3}
+\mu_{{\theta}_{1}} \\
+\mu_{{\theta}_{2}} \\
+\mu_{{\theta}_{3}}
 \end{bmatrix}
@@
 \begin{bmatrix}
-\theta_{i1} \\
-\theta_{i2} \\
-\theta_{i3}
+\delta_{1_i} \\
+\delta_{2_i} \\
+\delta_{3_i}
 \end{bmatrix}
@@
 \begin{bmatrix}
-\epsilon_{i1} \\
-\epsilon_{i2} \\
-\epsilon_{i3}
-\end{bmatrix} \\
+\epsilon_{1_i} \\
+\epsilon_{2_i} \\
+\epsilon_{3_i}
+\end{bmatrix}
+(\#eq:multivariate-model1)
+\end{gathered}
+\end{align}
+
+\begin{align}
+\begin{gathered}
 \begin{bmatrix}
-\theta_{i1} \\
-\theta_{i2} \\
-\theta_{i3}
+\delta_{1_i} \\
+\delta_{2_i} \\
+\delta_{3_i}
 \end{bmatrix}
-\sim \mathcal{MVN}(0, \mathrm{T}^2) \\
+\sim \mathcal{MVN}(0, \mathrm{T})
+(\#eq:multivariate-model2)
+\end{gathered}
+\end{align}
+
+\begin{align}
+\begin{gathered}
 \begin{bmatrix}
-\epsilon_{i1} \\
-\epsilon_{i2} \\
-\epsilon_{i3}
+\epsilon_{1_i} \\
+\epsilon_{2_i} \\
+\epsilon_{3_i}
 \end{bmatrix}
-\sim \mathcal{MVN}(0, \mathrm{V}^2) \\
-\mathrm{T}^2 = \begin{bmatrix} 
+\sim \mathcal{MVN}(0, \mathrm{V})
+(\#eq:multivariate-model3)
+\end{gathered}
+\end{align}
+
+\begin{align}
+\begin{gathered}
+\mathrm{T} = \begin{bmatrix}
 \tau_1^2 & & & \\
@@
 \rho_{31}\tau_3\tau_1 & \rho_{32}\tau_3\tau_2 & \tau_3^2
-\end{bmatrix} \\
-\mathrm{V}^2 = \begin{bmatrix} 
-\sigma_1^2 & & & \\ 
-\rho_{s21}\sigma_2\sigma_1 & \sigma_2^2 & & \\ 
-\rho_{s31}\sigma_3\sigma_1 & \rho_{s32}\sigma_3\sigma_2 & \sigma_3^2
-\end{bmatrix} \\
-(\#eq:multivariate-model)
+\end{bmatrix}
+(\#eq:multivariate-model4)
+\end{gathered}
+\end{align}
+
+\begin{align}
+\begin{gathered}
+\mathrm{V} = \begin{bmatrix}
+\sigma^2_{\epsilon_1} & & & \\
+\rho_{s_{21}}\sigma_{\epsilon_2}\sigma_{\epsilon_1} & \sigma^2_{\epsilon_2} & & \\
+\rho_{s_{31}}\sigma_{\epsilon_3}\sigma_{\epsilon_1} & \rho_{s_{32}}\sigma_{\epsilon_3}\sigma_{\epsilon_2} & \sigma^2_{\epsilon_3}
+\end{bmatrix}
+(\#eq:multivariate-model5)
 \end{gathered}
@@
 The previous code simulated a *compound symmetry* (CS) structure where all variances ($\tau^2$) and covariances ($\rho\tau_1\tau_2$) are the same.
-There could be different structures of the $T^2$ matrix that are clearly explained in the `metafor::rma.mv()` documentation (see https://wviechtb.github.io/metafor/reference/rma.mv.html#specifying-random-effects).
+There could be different structures of the $T$ matrix that are clearly explained in the `metafor::rma.mv()` documentation (see https://wviechtb.github.io/metafor/reference/rma.mv.html#specifying-random-effects).
 For the sake of simplicity we will use the CS structure.
@@
 We can use the `sim_study_m()` that follows the same idea of the `sim_study()` function but extends to multiple outcomes.
-We generate a single study with $p$ outcomes where sampling errors are correlated according to a compound symmetry structure^[Also for the $V^2$ matrix we can assume different variance-covariance structures.
+We generate a single study with $p$ outcomes where sampling errors are correlated according to a compound symmetry structure^[Also for the $V$ matrix we can assume different variance-covariance structures.
 In this example we are using a CS structure].
@@
 set.seed(seed)
-thetas <- c(0.1, 0, 0.7) # three real average outcomes
+mus <- c(0.1, 0, 0.7) # three real average outcomes
 n <- 30 # number of participant per study
 r <- 0.6 # correlation between sampling errors
-sim_study_m(thetas, nt = n, nc = n, r = r)
+sim_study_m(mus, nt = n, nc = n, r = r)
 ```
@@
 
-thetas_r <- c(0.1, 0, 1) # population level real effect sizes
+mus <- c(0.1, 0, 1) # population level real effect sizes
 taus2 <- c(0.3, 0.3, 0.3) # population level real tau2s
@@
 # adding the real effects and the random-effect adjustments
-sim$theta_r <- rep(thetas_r, k)
-sim$theta_i <- sim_T(k, p, taus2, rho)
+sim$mu <- rep(mus, k)
+sim$delta_i <- sim_T(k, p, taus2, rho)
 
 # average effect + random-effect adjustment
-sim$theta_r_i <- sim$theta_r + sim$theta_i
+sim$es <- sim$mu + sim$delta_i
 
@@
 # simulating the effect sizes for each study
-resl <- lapply(siml, function(x) sim_study_m(x$theta_r_i, x$nt, x$nc, r))
+resl <- lapply(siml, function(x) sim_study_m(x$es, x$nt, x$nc, r))
 res <- dplyr::bind_rows(resl) # combine everything
@@
 The syntax is clearly explained here [https://wviechtb.github.io/metafor/reference/rma.mv.html#specifying-random-effects](https://wviechtb.github.io/metafor/reference/rma.mv.html#specifying-random-effects) The basic idea is to specify the random effect structure as `~ inner|outer` where inner is the outcome variable and outer is the paper variable.
-The mods = `~ 0 + p` argument specify to estimate the average effect for each outcome (i.e., the $\theta_{r}$ vector of average effects).
+The mods = `~ 0 + p` argument specify to estimate the average effect for each outcome (i.e., the $\mathbf{\mu_{\theta}}$ vector of average effects).
 We are removing the intercept (`~ 0 + p` or equivalently `~ p - 1`) and the model will estimate the mean for each level of the `p` variable.
 
-We could use also another structure for the $T^2$ matrix.
+We could use also another structure for the $T$ matrix.
 We know that our data are generated using a CS structure thus using another structure will be less appropriate.
@@
 
+# Extra
+
+## $I^2$ with homogeneous vs heterogeneous sample sizes
+
+In the paper we suggested that is not straightforward to simulate a meta-analysis fixing the $I^2$ as a simulation parameter with heterogeneous sample sizes.
+Here we show that when sample sizes are generated from a probability distribution (e.g., Gaussian) we can fix the $I^2$ value using the $\tilde{v}$ calculated using the expected value of the sample size distribution.
+For example, if the sample sizes of our $k$ studies is generate from a Poisson distribution $n_i \sim \mathcal{P}(\lambda)$, then $\tilde{v}$ can be calculated using $\lambda$ as *typical* sample size (i.e., as if the sample sizes were homogeneous).
+The following code simulate this scenario showing the distribution of $I^2$ with heterogeneous sample sizes^[Compared to the simulations of the paper here we are simulating the effect size and the sampling variances from the respective sampling distributions [e.g., supplementary materials by @Van_Aert2019-xt]].
+
+```{r}
+set.seed(seed)
+tau2 <- 0.2 # real tau2 value
+navg <- 40 # average sample size
+nmin <- 10 # minimum sample size
+k <- 30 # number of studies
+k <- 30 # number of studies
+nsim <- 1000 # number of simulations
+mu <- 0.3 # real average effect size
+
+# here we simulate some effect sizes using the aggregated
+# data approach, thus from the summary statistics distributions
+
+I2 <- rep(0, nsim)
+
+for(i in 1:nsim){
+  ni <- nmin + rpois(k, navg - nmin)
+
+  # effect sizes
+  yi <- rnorm(k, rnorm(k, mu, sqrt(tau2)), sqrt(1/ni + 1/ni))
+
+  # sampling variances
+  vi <- (rchisq(k, ni + ni - 2) / (ni + ni - 2)) * (1/ni + 1/ni)
+
+  # model
+  fit <- rma(yi, vi)
+
+  # extracting I2
+  I2[i] <- fit$I2
+}
+```
+
+```{r, echo = FALSE}
+title <- sprintf("Simulated $I^2$ values (Mean = %.3f)", mean(I2))
+
+hist(I2, 
+     breaks = seq(min(I2), 100, 3),
+     xlab = latex2exp::TeX("$I^2$"),
+     main = latex2exp::TeX(title),
+     col = "salmon",
+     cex.lab = 1.1,
+     cex.main = 1.3)
+```
+
+The $I^2$ calculated using the expected value of the sample sizes distribution ($\lambda = 40$) and assuming sample size homogeneity is:
+
+```{r}
+v <- 1/navg + 1/navg
+tau2 / (v + tau2)
+```
+
+The value is very close to the mean of the simulated $I^2$ distribution.
+In conclusion, by generating the sample sizes from a probability distribution (e.g., Poisson) and using the Higgins and Thompson [-@Higgins2002-fh] formula to calculate $\tilde{v}$ we can somehow fix the $I^2$ with heterogeneous sample sizes using the expected value of the distribution.
+
 # Useful resources
@@
 
-<!-- CAPTIONS -->
-
 (ref:fig-block-vcov) Example of a block variance-covariance matrix for three hypothetical studies.
@@
 The last study follows the same logic as the first study, with only two outcomes.
+
+(ref:img-multilevel) Graphical representation of a *multilevel* (three-level) model.
+Each observed effect size is defined as the overall effect average effect $\mu_{\theta}$, a study-specific random-effect $\delta_{i} \sim N(0, \tau^2)$ and a experiment-specific random-effects $\zeta_{ij} N(0, \omega^2)$.
+
+(ref:img-multivariate) Graphical representation of a *multivariate* model.
+In this example, we have two outcomes ($p$) (magenta and green) with an average effect ($\mu_{\theta_p}$), heterogeneity parameters ($\tau^2_p$), and a certain true correlation value ($\rho_{p_1p_2}$, third type of dependency).
+Each study at level-2 (as in the standard random-effects model) has two effect sizes with corresponding sampling variances.
+The two effect sizes are correlated ($\rho_{s}$, with $s$ for sampling errors) because each effect is collected on the same participants.

### 2023-07-18 documents/paper/paper.Rmd
fixing after revision

@@
 \color{red} Then using meta-regression, it is possible to include variables (i.e., moderators) to explain the observed heterogeneity [@Borenstein2009-mo, pp.187-203].
-More recently, location-scale models have been developed to include predictors also on the residual heterogeneity [@Viechtbauer2022-rz].
+More recently, location-scale models have been developed to include predictors also on the residual heterogeneity [i.e., @Viechtbauer2022-rz].
 \color{black} Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias.
@@
 
-\rev{As an example that will be used throughout the paper, we consider the efficacy of memory training in improving memory performance during a cognitive task.
+\color{red} As an example that will be used throughout the paper, we consider the efficacy of memory training in improving memory performance during a cognitive task.
 The typical primary study will collect data from a group of participants receiving the memory training (*experimental group*) and another group receiving a control treatment (*control group*).
 The focus of the meta-analysis is collecting multiple studies with similar aims and methods and estimating the average effect of memory training.
-Despite differences in the type of cognitive task or experimental setup, each primary study collects an *experimental group* ($n_{C}$) and *control group* ($n_{C}$) and computes the average performance (*experimental group* $\overline T$ and *control group* $\overline C$) and standard deviations (*experimental group* $s_{T}$ and *control group*
-$s_{C}$).}
+Despite differences in the type of cognitive task or experimental setup, each primary study collects an *experimental group* ($n_{T}$) and *control group* ($n_{C}$) and computes the average performance (*experimental group* $\overline T$ and *control group* $\overline C$) and standard deviations (*experimental group* $s_{T}$ and *control group*
+$s_{C}$). \color{black}
 
@@
 This common measure should give an immediate idea of the direction (i.e., the treatment improves or reduces performance) and the size of the effect.
-Standardized effect size measures [see @Lakens2013-eq for an overview], such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's $d$ [@Cohen1988-ad] or the Pearson correlation coefficient $\rho$ which could be calculated using the associated sample estimator *r* are commonly used to compare heterogeneous outcome variables^[\color{red} When using bounded effect size indexes such as the Pearson correlation, Fisher's *z* transformation is commonly used[@Borenstein2009-mo, pp.
+Standardized effect size measures [see @Lakens2013-eq for an overview], such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's $d$ [@Cohen1988-ad] or the Pearson correlation coefficient $\rho$ which could be calculated using the associated sample estimator *r* are commonly used to compare heterogeneous outcome variables^[\color{red} When using bounded effect size indexes such as the Pearson correlation, Fisher's *z* transformation is commonly used [@Borenstein2009-mo, pp.
 41-43] \color{black}].
-If all studies used the same raw measure, such as reaction times, it is possible to directly meta-analyze the studies without standardizing using (UMD, unstandardized mean difference).
+If all studies used the same raw measure, such as reaction times, it is possible to directly meta-analyze the studies without standardizing e.g.
+using the unstandardized mean difference [UMD, @Borenstein2009-mo, pp.
+21-24].
 
@@
 The core of a meta-analysis is combining the results of multiple studies giving more weight to studies that provide a more precise effect estimation.
-We can find essentially three meta-analysis models: the equal-effects, the fixed-effects and the random-effects model [@Laird1990-ro, @Hedges1998-uo, see also @Gronau2021-do and @Berkhout2023-fj for Bayesian a meta-analysis approach trying to combine evidence from multiple models e.g., equal and random-effects together].
+We can find essentially three meta-analysis models: the equal-effects, the fixed-effects and the random-effects model [@Laird1990-ro; @Hedges1998-uo].
 The *equal-effects* model assume that each study included in the meta-analysis is a more or less precise estimation of the true underlying effect ($\theta$).
@@
 On the other side, suppose some study-level characteristics (e.g., participants' age, sex, or socioeconomic status) or the experimental paradigm (e.g., type of memory task or difficulty) could impact the treatment effect.
-The *random-effects* model assume a distribution of real effects with mean $\mu_{\theta}$ and variance $\tau^{2}$.
+In this case we have true variability (i.e., heterogeneity) among studies.
+The *random-effects* model assume a distribution of real effects with mean $\mu_{\theta}$ and variance $\tau^{2}$ thus estimating the heterogeneity.
 Finally, the *fixed-effects* estimate the average effect of the included pool of studies ignoring the presence of heterogeneity.
-While the distinction between the equal and fixed-effects model is theoretical, the estimated model is the exactly the same^[see https://wviechtb.github.io/metafor/reference/misc-models.html for a detailed explanation].
+While the distinction between the equal and fixed-effects model is theoretical, the estimated model is the exactly the same^[See [wviechtb.github.io/metafor/reference/misc-models.html](https://wviechtb.github.io/metafor/reference/misc-models.html) for a detailed explanation].
 The *random-effects* model assume and estimate heterogeneity among effect sizes leading to different model parameters.
@@
 The *equal-effects* model assume a single true population effect to be estimated.
-In the presence of heterogeneity, using meta-regression is it possible to include some predictors (e.g., the type of experimental setup) to explain effect sizes variability.
+The Figure \@ref(fig:img-equal-vs-random) depicts the theoretical distinction between the equal and random-effects model.
+In the presence of heterogeneity, we can include moderators (e.g., the type of experimental setup) into a meta-regression model to explain the effect sizes heterogeneity.
+\color{red} An interesting proposal to combine evidence from these different models into a single analysis is called Bayesian model-averaged meta-analysis [@Gronau2021-do; @Berkhout2023-fj].
+\color{black}
 
@@
 For the data manipulation we used the *tidyverse* [@R-tidyverse] package.
-For the models fitting we used the *metafor* [@metafor2010] package.
-For figures and tables the *ggplot2* [@ggplot22016] or the `metafor::forest()` function, *kableExtra* [@R-kableExtra] and *papaja* [@R-papaja] packages.
-We set the `seed` for reproducibility to prepare the R environment for the simulations.
+For the models fitting we used the *metafor* [@Viechtbauer2010-xz] package.
+For figures and tables the *ggplot2* [@R-ggplot2] or the `metafor::forest()` function, *kableExtra* [@R-kableExtra] and *papaja* [@R-papaja] packages.
+We set the `seed` for reproducibility of the simulation environment.
 \color{black}
@@
 library(metafor) # for meta-analysis models fitting
+library(purrr) # for *apply like functions
 seed <- 2023 # general seed for all simulations
@@
 
+\color{red}
 Before diving into the specific simulations, in this section, we define the common aspects of all simulations in the following sections.
@@
 All the examples refer to primary studies that assess the efficacy of a treatment by comparing a control and an experimental group.
-\color{red} There are different simulation approaches to meta-analysis concerning how to generate study-level or participant-level data.
-In simulation studies where the purpose is not evaluating effect size estimators, it is convenient to simulate unstandardized effect size measures (e.g., unstandardized mean difference, UMD).
-Firstly the estimator is unbiased thus not requiring small sample correction [e.g., @Hedges1989-ip; @Hedges1981-za].
-Then the effect size and the sampling variance are independent.
-Similar to the simulation approach by Viechtbauer [@Viechtbauer2005-yh; @Viechtbauer2007-tq] the *experimental* group ($T$) and *control* groups $C$ are sampled from normal distributions respectively $T_i \sim \mathcal{N}(\Delta, 1)$ and $C_i \sim \mathcal{N}(0, 1)$ where $\Delta$ is the unstandardized mean difference (UMD).
-The UMD and the sampling variance are calculated using Equations \@ref(eq:effsize) and \@ref(eq:effsizevar) where $y$ is the sample estimator of $\Delta$.
+In simulation studies where the purpose is not evaluating effect size estimators, it is convenient to simulate unstandardized effect size measures [see @Viechtbauer2005-yh; @Viechtbauer2007-tq].
+The estimator is unbiased thus not requiring small sample correction [e.g., @Hedges1989-ip; @Hedges1981-za].
+Furthermore, the effect size and the sampling variance are independent.
+Similar to the simulation approach by Viechtbauer [-@Viechtbauer2005-yh; -@Viechtbauer2007-tq] the *experimental* group ($T$) and *control* groups $C$ are sampled from normal distributions respectively $T_i \sim \mathcal{N}(\Delta, 1)$ and $C_i \sim \mathcal{N}(0, 1)$ where $\Delta$ is the unstandardized mean difference (UMD).
+The UMD and the sampling variance are calculated using Equations \@ref(eq:effsize) and \@ref(eq:effsizevar) where $y$ is the estimated value of $\Delta$.
 \color{black}
@@
 
-The simulation approach can be easily extended calculating a standardized effect size measure [e.g., Cohens' $d$ @Cohen1988-ad] and the corresponding sampling variance.
+The simulation approach can be easily extended calculating a standardized effect size measures [e.g., Cohens' $d$ @Cohen1988-ad] and the corresponding sampling variance.
 For example, after generating data for the two groups the mean difference can be standardized using the pooled standard deviation and applying the appropriate correction [e.g., @Hedges1989-ip; @Hedges1981-za].
@@
 \color{red}
-The `es` is the true effect size ($\mu_{\theta}$ for the random-effects model and $\theta$ for the equal-effects model), `nc` and `nt` are the sample size for the control and experimental group, `aggregate` controls if returning the effect size and the corresponding variance or the participant-level data.
+The `es` is the true effect size ($\mu_{\theta}$ for the random-effects model and $\theta$ for the equal-effects model), `nc` and `nt` are the sample size for the control and experimental group, `aggregate` controls if returning the effect size and the corresponding sampling variance or the participant-level data.
 We can generate a single study with the desired parameters with this function.
@@
 Using the `sim_study()` function multiple times (with the appropriate adjustments) we can generate a series of studies simulating a data set for a meta-analysis (using the `sim_studies()` function).
-After each example, we will compute the appropriate model (e.g., *fixed* or *random-effects*) using the `metafor` package [@Viechtbauer2010-xz] to check the recovery of simulated parameters^[\color{red} By default `metafor` use the Wald *z* test on model parameters but other inference methods are available e.g.
+After each example, we will compute the appropriate model (e.g., *equal* or *random-effects*) using the `metafor` package [@Viechtbauer2010-xz] to check the recovery of simulated parameters^[\color{red} By default `metafor` use the Wald *z* test on model parameters but other inference methods are available e.g.
 the Knapp and Hartung [-@Knapp2003-qx] method by setting `test="knha"` within the `rma()` function.
@@
      "$\\theta$",                        "theta",      "Equal-effects model true effect size",
-       "$\\mu$",      "mu",                        "Random-effects model true effect size",
-        "$\\tau^2$",    "tau2",                            "Effect size heterogeneity",
-       "$\\tau^2_r$",   "tau2r",                   "Residual effect size heterogeneity",
+     "$\\mu_{\\theta}$",                 "mu\\_theta", "Random-effects model true effect size",
+     "$\\tau^2$",                        "tau2",       "Effect sizes heterogeneity",
+     "$\\tau^2_r$",                      "tau2r",      "Residual effect sizes heterogeneity",
      "$\\overline{T}, \\overline{C}$",   "",           "Mean of the experimental/control group",
@@
             placement = "H",
-            note = "The \\texttt{yi} and \\texttt{vi} notation for the observed effect size ($y_i$) and sampling variance ($\\sigma^2_i$) has been used to be consistent with the \\texttt{metafor} notation.")
+            note = "The \\texttt{yi} and \\texttt{vi} notation for the observed effect size ($y_i$) and sampling variance ($\\sigma_{\\epsilon_i}^{2}$) has been used to be consistent with the \\texttt{metafor} notation.")
 ```
@@
 The most basic model to simulate is the *equal-effects model* (EE).
-As reported in the previous sections, the *equal-effects model* assumes the presence of a single true effect ($\theta$) and the observed variability is because each effect size is an imprecise estimation of the true effect.
-In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies (i.e., the sample size).
-Equations \@ref(eq:equal-effects-model1) and \@ref(eq:equal-effects-model2) depicts the *equal-effects* model.}
+As reported in the previous sections, the *equal-effects model* assumes the presence of a single true effect ($\theta$) and the \color{red} observed variability is caused by each effect size being a more or less imprecise estimation of the true effect.
+\color{black} In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies (i.e., the sample size).
+Equations \@ref(eq:equal-effects-model1) and \@ref(eq:equal-effects-model2) \rev{formalize} the *equal-effects* model.
 
@@
 As demonstrated in Equation \@ref(eq:effsize), the increase in sample size will decrease the sampling variability.
-An exact study will essentially have $\theta$ as the observed effect size.
+\rev{A study with a extremely large sample size} will essentially have $\theta$ as the observed effect size.
 Since we are sampling participants' level data, the error component is already included in the `sim_study()` function.
@@
 The parameter $\mu_{\theta}$ is mean of the distribution, interpreted as the average effect size across different true effect sizes and $\tau^{2}$ is the variance of the distribution interpreted as variability or heterogeneity of effect sizes.
-In practical terms, we now have two sources of variability: $\tau^{2}$, which express the real difference among effect sizes, and $\sigma_{i}^{2}$ which is the known sampling variance of each study as in the *equal-effects model*.
+In practical terms, we now have two sources of variability: $\tau^{2}$, which express the real difference among effect sizes, and $\sigma_{\epsilon_i}^{2}$ which is the known sampling variance of each study as in the *equal-effects model*.
 We can easily extend Equation \@ref(eq:equal-effects-model1) with Equations \@ref(eq:random-effects-model1), \@ref(eq:random-effects-model2) and \@ref(eq:random-effects-model3).
@@
 n <- 20 # number of participants per group, per study
-mu <- 0.3 # real effect size
+mu_theta <- 0.3 # real effect size
 tau2 <- 0.2 # the effect size heterogeneity
 
-sim <- make_data(k = k, nc = n, nt = n, mu = mu)
+sim <- make_data(k = k, nc = n, nt = n, mu_theta = mu_theta)
 
@@
 # adding the by-study adjustment to the average effect
-sim$es <- sim$mu + sim$deltai
+sim$es <- sim$mu_theta + sim$deltai
 
@@
 Table \@ref(tab:res-random-effects) depicts the model results^[The `method = "REML"` is not the only method to estimate a random-effects model.
-See the `rma` documentation (https://wviechtb.github.io/metafor/reference/rma.uni.html#specifying-the-model) for an overview of other estimation methods] of the metafor package.
+See the `rma` documentation (https://wviechtb.github.io/metafor/reference/rma.uni.html#specifying-the-model) for an overview of other estimation methods] of the `metafor` package.
 We are now estimating two parameters: $\mu_{\theta}$ and $\tau^{2}$.
-As for the \rev{*equal-effects model*}, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
+As for the \color{red} *equal-effects model* \color{black}, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
 
@@
 An important aspect of the *random-effects model* is the interplay between the heterogeneity ($\tau^{2}$) and the sampling variability ($\sigma_{i}^{2}$).
-As the number of studies $k$ increases, the estimation of $\mu_{\theta}$ and $\tau^{2}$ becomes more precise.
-Additionally, as each study\'s sample size decreases, each $\theta_{i}$ will be estimated with higher precision but as long as $\tau^{2} \neq 0$ there will be variability among effect sizes.
+As the number of studies $k$ increases, the estimation of $\mu_{\theta}$ and $\tau^{2}$ becomes more precise \color{red} [@Blazquez-Rincon2023-kp; @Rubio-Aparicio2018-pd] \color{black}.
+Additionally, as the sample size of each study decreases, each $\theta_{i}$ will be estimated with higher precision but as long as $\tau^{2} \neq 0$ there will be variability among effect sizes.
 In other words, increasing the sample size of each study or the number of studies will not affect the value of $\tau^{2}$ but only the estimation precision [see @Borenstein2009-mo, Chapter 16, Figure 16.6 for a clear explanation].
@@
 k <- 10 # number of studies
-mu <- 0.3 # real effect size
+mu_theta <- 0.3 # real effect size
 tau2 <- 0.2
@@
 
-sim <- make_data(k = k, mu = mu)
+sim <- make_data(k = k, mu_theta = mu_theta)
 
@@
 
-sim$es <- sim$mu + sim$deltai
+sim$es <- sim$mu_theta + sim$deltai
 sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
@@
 The relationship between the sampling error and the heterogeneity can be expressed using the $I^{2}$ statistics [e.g., @Higgins2002-fh] that is the percentage of the total variability $\tau^{2} + \tilde{v}$ that is attributable to real heterogeneity between studies ($\tau^{2}$).
-\color{red} In Equation \@ref(eq:isquared), $\tilde{v}$ is the "typical" within-study sampling variability as proposed by Higgins and Thompson [-@Higgins2002-fh]^[see also here https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate#fn__1 for an overview of the $I^2$ statistics also for complex models] and implemented in Equation  \@ref(eq:typical-v) with $w_i = 1/\sigma_{\epsilon_i}^{2}$ and $k$ is the number of studies.
-
+\color{red} In Equation \@ref(eq:isquared), $\tilde{v}$ is the *typical* within-study sampling variability (i.e., a summary statistics of sampling variances of included studies) as proposed by Higgins and Thompson [-@Higgins2002-fh]^[See https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate#fn__1 for an overview of the $I^2$ statistics also for complex models.
+There are also other proposals to compute the $\tilde{v}$ statistics [see @Takkouche1999-sv; @Takkouche2013-nd]] and implemented in Equation \@ref(eq:typical-v) with $w_i = 1/\sigma_{\epsilon_i}^{2}$ and $k$ is the number of studies.
 \color{black}
@@
 \begin{align}
-\begin{gathered}
+\begin{gathered}Random-effects model true effect
 \tilde{v} = \frac{(k-1) \sum w_i}{(\sum w_i)^2 - \sum w_i^2}
@@
 
+\color{red}
 Given the interpretation of $I^{2}$ it is possible to simulate a meta-analysis fixing a certain $I^{2}$ value.
 The only caveat is fixing the $\tilde{v}$.
-When the sample size of each study is the same, $\tilde{v} = \sigma^2_i$.
+When the sample size of each study is the same^[The homogeneous sample size of primary studies assumption is commonly used to calculate the statistical power [see @Borenstein2009-mo, Chapter 29]], $\tilde{v} = \sigma_{\epsilon_i}^{2}$.
 In the other case, $\tilde{v}$ needs to be calculated from sampling variances as reported in Higgins and Thompson [-@Higgins2002-fh] and cannot be easily fixed a priori.
-However, with the assumption of homogeneous sample size across studies^[The homogeneous sample size of primary studies assumption is commonly used to calculate the statistical power [see @Borenstein2009-mo, Chapter 29]], we can solve Equation \@ref(eq:isquared) for $\tau^{2}$ obtaining the heterogeneity value associated with a certain $I^{2}$ as reported in Equation \@ref(eq:tau2-from-isquared).
+With the assumption of homogeneous sample size across studies^[In case of heterogeneous sample sizes generated from a probability distribution (e.g., Gaussian) the *typical* sampling variance calculated using the expected value of the distribution can be used to somehow fix the (average) $I^2$ (see the supplementary materials for a simulated example)], we can solve Equation \@ref(eq:isquared) for $\tau^{2}$ obtaining the heterogeneity value associated with a certain $I^{2}$ as reported in Equation \@ref(eq:tau2-from-isquared).
+\color{black}
+
+\color{red}
 Table \@ref(tab:res-random-effects-i2) depicts the results of the random-effects model fixing the $I^{2}$ value.
 Choosing a meaningful $\tau^2$ for the simulation can be sometimes difficult.
-Beyond fixing the $I^2$ value, one could use a plausible value from the literature.
-\color{red} For example van Erp and colleagues [@Van_Erp2017-cf] estimate an empirical $\tau^2$ distribution across several published meta-analyses.
-\color{black}
-
-<!-- TODO add something on the simulation heterogeneous vs homogeneous sample sizes -->
+Beyond fixing the $I^2$ value, a possibility is choosing plausible $\tau^2$ value from the literature.
+For example van Erp and colleagues [-@Van_Erp2017-cf] estimate an empirical $\tau^2$ distribution across several published meta-analyses that can be used to simulate a plausible scenario.
+\color{black}
 
@@
 k <- 30 # number of studies
-mu <- 0.3 # real effect size
+mu_theta <- 0.3 # real effect size
 n <- 30 # sample size per group, per study
@@
 
-v <- (n + n)/(n * n) + mu^2/(2 * (n + n - 2)) # typical within study variance
+v <- (n + n)/(n * n) + mu_theta^2/(2 * (n + n - 2)) # typical within study variance
 tau2 <- -((I2*v)/(I2 - 1))
 
-sim <- make_data(k = k, nc = n, nt = n, mu = mu)
+sim <- make_data(k = k, nc = n, nt = n, mu_theta = mu_theta)
 
 sim$deltai <- rnorm(k, 0, sqrt(tau2))
-sim$es <- sim$mu + sim$deltai 
+sim$es <- sim$mu_theta + sim$deltai 
 sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
@@
 
-From a linear regression perspective, both the EE and the RE models can be seen as *intercept-only* models where only the mean (i.e., the linear regression intercept) is estimated Borenstein et al.
-(2009).
+From a linear regression perspective, both the EE and the RE models can be seen as *intercept-only* models where only the mean (i.e., the linear regression intercept) is estimated.
 As reported in the meta-analysis introduction, the between-study heterogeneity usually represents the true variability of the effect due to differences among primary studies.
@@
 For example, a group of studies could use a particular memory task where the expected effect is higher than another.
-In this way, considering the
-type of task will explain part of the observed heterogeneity as in standard regression models.
+In this way, considering the type of task will explain part of the observed heterogeneity as in standard regression models.
 Figures \@ref(fig:img-metaregression-bin) and \@ref(fig:img-metaregression-num) depict a random-effects meta-regression model for a categorical and numerical predictor.
@@
 Now we fix $\beta_{1}$ to be the *lab vs.
-online effect* (i.e., the expected mean difference between the two groups of studies) and the product between \rev{$\beta_{1}X_{1_i}$} will consider the *lab vs.
+online effect* (i.e., the expected mean difference between the two groups of studies) and the product between \color{red}$\beta_{1}X_{1_i}$ \color{black} will consider the *lab vs.
 online effect*.
@@
 In other terms, \color{red} $\delta_i \sim \mathcal{N}(0, \tau_{r}^{2})$ \color{black} where $\tau^2_r$ is the residual heterogeneity after including the moderator.
-We can describe our model using equation \@ref(eq:meta-regression-equations) according to the value of $X_{1i}$ (L = lab-based experiments, O = online experiments).
+We can describe our model using equation \@ref(eq:meta-regression-equations) according to the value of $X_{1_i}$ (L = lab-based experiments, O = online experiments).
 
@@
 \begin{gathered}
-y_{iL} = \beta_0 + \delta_i + \beta_1 \times 0 + \epsilon_i \\
-y_{iO} = \beta_0 + \delta_i + \beta_1 \times 1 + \epsilon_i
+y_{L_i} = \beta_0 + \delta_i + \beta_1 \times 0 + \epsilon_i \\
+y_{O_i} = \beta_0 + \delta_i + \beta_1 \times 1 + \epsilon_i
 (\#eq:meta-regression-equations)
@@
 Now the intercept is still the average effect size when the age is zero but now zero is the average age.
-Importantly, the contrast coding for categorical predictors or centering numerical variables does not affect the overall model estimation but only parameters values and interpretation.
+Importantly, the contrast coding for categorical predictors or centering numerical variables does not affect the overall model but only parameters values and interpretation.
 
@@
 In our case, we simulated $k$ average age values from a uniform distribution $\text{age}_{i} \sim U\left( 20,40 \right)$.
-Then we can plot the distribution of $d_{i}$ values to check the plausibility of simulated values.
+Then we can plot the distribution of $y_{i}$ values to check the plausibility of simulated values.
 As shown in Figure \@ref(fig:plot-meta-reg-plausible-implausible), with the same range for the moderator, a $\beta_{1} = 0.1$ gives a plausible effect sizes range while a $\beta_{1} = 0.7$ predicts very extreme values.
@@
 This approach has been implemented by Lopez-Lopez and colleagues [-@Lopez-Lopez2014-it].
-We can use Equation \@ref(eq:beta-from-r2) to find the $\beta_{1}$ value that is associated with a certain $R^{2}$.
+We can use Equation \@ref(eq:beta-from-r21) and \@ref(eq:beta-from-r22) to find the $\beta_{1}$ value that is associated with a certain $R^{2}$.
 
@@
 \begin{aligned}
-\beta^2_1 = \tau^2R^2 \\
+\beta^2_1 = \tau^2R^2
+(\#eq:beta-from-r21)
+\end{aligned}
+\end{align}
+
+\begin{align}
+\begin{aligned}
 \tau^2_{r} = \tau^2 - \beta^2_1
-(\#eq:beta-from-r2)
+(\#eq:beta-from-r22)
 \end{aligned}
@@
 Lopez-Lopez and colleagues [-@Lopez-Lopez2014-it] generated the moderator ($X_{1}$) values from a standard normal distribution.
-In the following example, we standardized the moderator (`scale()`) after simulating values on the *age* scale
-(e.g., `runif(k, 20, 40)`).
+In the following example, we standardized the moderator (`scale()`) after simulating values on the *age* scale (e.g., `runif(k, 20, 40)`).
 
@@
 However, one of the most critical applications is estimating the power of a specific statistical model.
+\color{red} One of the purpose of power analysis by simulations is estimating the required number of studies to detect an hypothetical effect size when planning a meta-analysis.
+At the same time, the meta-analysis that we presented (*two-level* equal or random-effects model) can be considered as a multi-lab [e.g., @Klein2018-he] study where experiments are planned and not collected from the literature.
+In fact, we can use the same simulation approach to optimize the number of participants and studies when planning a multi-lab project.
+\color{black}
 
@@
 These methods are easy to implement but made strong assumptions, such as the homogeneity of sample size, and did not consider the uncertainty in estimating $\tau^{2}$.
-Jackson and colleagues [-@Jackson2017-dv]
-partially solve the issue by developing an interesting method that takes into account the uncertainty in estimating $\tau^{2}$ without using simulations.
-However, only Monte Carlo simulations can consider complex scenarios and extra parameters.
+Jackson and colleagues [-@Jackson2017-dv] partially solve the issue by developing an interesting method that takes into account the uncertainty in estimating $\tau^{2}$ without using simulations.
+However, only Monte Carlo simulations can \rev{handle} complex scenarios and extra parameters.
 
@@
 
-1.  Choose the model that generates the data (e.g., fixed, or random-effects model).
+1.  Choose the model that generates the data (e.g., equal or random-effects model).
 2.  Fix the relevant parameters (e.g., $\tau^{2}$ and $\theta$).
@@
 
-For example, we can estimate the power of a *random-effects model* by repeating the simulation presented in Section \@ref(re-model) many times.
+For example, we can estimate the power of a *random-effects* model by repeating the simulation presented in Section \@ref(re-model) many times.
 We simulated heterogeneity of sample sizes sampling $n_{T}$ and $n_{C}$ values from a Poisson distribution with $\lambda = 20$.
@@
 Then the `summary_sim()` function analyzes each simulation returning the relevant values.
-We repeat the simulation of the *random-effects model* several times with different parameters.
+We repeat the simulation of the *random-effects* model several times with different parameters.
 
@@
 # creating all combinations
-sim_grid <- tidyr::expand_grid(k, mu, tau2, navg, nmin, nsim)
+sim_grid <- tidyr::expand_grid(k, mu = delta, tau2, navg, nmin, nsim)
 
@@
 
-The present work introduced the basic concepts of the meta-analysis regarding fixed, random-effects models and meta-regression with a simulation-based approach.
+The present work introduced the basic concepts of the meta-analysis regarding equal, random-effects models and meta-regression with a simulation-based approach.
 We believe the presented examples can help implement alternative or more complex models.
-For example, the `sim_study()` function can be easily modified to simulate another effect sizes index, such as a paired sample Cohen's $d$ or a Pearson's
-correlation.
-In addition, more complex models, such as *multivariate* or *multilevel* models, can be implemented following the same approach (see supplementary materials).
-For example, the *three-level* model estimates another heterogeneity component representing the variability of multiple effect sizes in the same study.
+For example, the `sim_study()` function can be easily modified to simulate another effect sizes index, \color{red} such as correlations or odds ratios.
+\color{black} In addition, more complex models, such as *multivariate* or *multilevel* models, can be implemented following the same approach (see \rev{the} supplementary materials).
+For example, the \color{red} *multilevel* (e.g., *three-level*) \color{black} model estimates another heterogeneity component representing the variability of multiple \rev{independent} effect sizes \rev{within} the same study.
 Similarly, the *multivariate* model could include the correlation between multiple outcomes and the correlation between sampling errors.
@@
 We decided to give the foundations to understand meta-analyses with a simulation approach because more complex models are still based on the same principles.
-Second, there are limitations
-concerning simulating participant-level data.
+Second, there are limitations concerning simulating participant-level data.
 We decided to simulate the meta-analysis dataset starting from the participants' level to maximize the flexibility and clearness of each step.
 The downside concerns the efficiency and scalability of the simulation setup.
-For large-scale simulations (e.g., many conditions, iterations, or complex models), simulating from aggregated statistics is probably more efficient (see
-Heuvel et al., 2020 for an example) to improve the computation speed^[see https://www.jepusto.com/simulating-correlated-smds/ for a very clear example of the participant-level vs aggregated data simulation].
+For large-scale simulations (e.g., many conditions, iterations, or complex models), simulating from aggregated statistics is probably more efficient [see @van-den-Heuvel2020-rv for an example] to improve the \rev{simulation efficency}^[See https://www.jepusto.com/simulating-correlated-smds/ for a very clear example of the participant-level vs aggregated data simulation approaches].
 
@@
 (ref:img-equal-vs-random) \color{red} The difference between the assumptions of the *equal* and *random-effects model*.
-Each distribution depicts the sampling distribution of $k = 4$ hypothetical studies ($i = 1$, $2$, ..., $4$) with a certain observed effect size $y_{i}$ (red squares), sampling variability $\sigma_{i}^{2}$ and the 95% confidence interval (black segment).
+Each distribution depicts the sampling distribution of $k = 4$ hypothetical studies ($i = 1$, $2$, ..., $4$) with a certain observed effect size $y_{i}$ (red squares), sampling variability $\sigma_{\epsilon_i}^{2}$ and the 95% confidence interval (black segment).
 The *equal-effects* plot on the left suggests that each observed effect size has the same underlying true effect $\theta$ (each distribution has the same mean) with a different degree of precision (e.g., Study 1 is more precise than Study 3).
-In practice, each study has a different observed effect size where studies with high precision (i.e., narrow sampling distributions) will be remarkably close to the real effect ($\theta$).
-The *random-effects model* on the right suggests that beyond the error term ($\epsilon_{i}$), each real effect size is composed of a fixed part (now $\mu_{\theta}$) and a random part ($\theta_{i}$) sampled from a normal distribution with mean $\mu_{\theta}$  and variance $\tau^{2}$.
+In practice, each study has a different observed effect size where studies with high precision (i.e., narrow sampling distributions) will be close to the real effect ($\theta$).
+The *random-effects model* on the right suggests that beyond the error term ($\epsilon_{i}$), each real effect size is composed of a fixed part (now $\mu_{\theta}$) and a random part ($\theta_{i}$) sampled from a normal distribution with mean zero and variance $\tau^{2}$.
 When $\tau^{2}$ is zero, the random-effects reduces to a equal-effects model.
@@
 
-(ref:res-random-effects-i2) Summary of the *random-effects* model fixing the $I^2$ value. Parameters are the same as described in Table \@ref(tab:res-random-effects).
+(ref:res-random-effects-i2) Summary of the random-effects model fixing the $I^2$ value. Parameters are the same as described in Table \@ref(tab:res-random-effects).
 
@@
 The green segments depict the explained heterogeneity, and the red segments the residual (unexplained) heterogeneity.
-\color{red} The red squares are simulated observed effect sizes from the sampling distribution.
+\color{red} The red squares are simulated observed effect sizes from the sampling distributions.
 \color{black}
@@
 On the left, the majority of simulated values range between -1.5 and 1.5 thus $\beta_1 = 0.1$ can be considered a plausible value.
-On the right, ($\beta = 0.7$) values range between -10 to 10 that despite theoretically possible is highly implausible values in real meta-analysis for psychological data.
+On the right, ($\beta = 0.7$) values range between -10 to 10 that despite theoretically possible can be considered highly implausible values in real meta-analyses for psychological data.
 \color{black}

### 2023-07-18 documents/paper/paper.tex
fixing after revision

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
 \fi
@@
 \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
 \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
@@
 \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
 \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
@@
 \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
 \newcommand{\ImportTok}[1]{#1}
@@
 \newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
 \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
@@
 \usepackage{orcidlink}
+\newcommand{\rev}[1]{\textcolor{red}{#1}}
 \ifLuaTeX
@@
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
@@
 Furthermore, data simulation is a powerful educational tool for understanding a statistical method.
-In this tutorial, we show how to simulate fixed-effects, random-effects, and meta-regression models and illustrate how to estimate statistical power.
+In this tutorial, we show how to simulate \rev{equal-effects}, random-effects, and meta-regression models and illustrate how to estimate statistical power.
 In the supplementary materials, we extended these simulations also for multilevel and multivariate models.
@@
 Firstly, it allows combining evidence from multiple studies assigning more weight to studies with lower estimation variability.
-Then using meta-regression, it is possible to assess the impact of study-level variables that could impact the phenomenon.
-Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias.
+\color{red} Then using meta-regression, it is possible to include variables (i.e., moderators) to explain the observed heterogeneity (Borenstein, Hedges, Higgins, \& Rothstein, 2009, pp.
+187--203).
+More recently, location-scale models have been developed to include predictors also on the residual heterogeneity (i.e., Viechtbauer \& López-López, 2022).
+\color{black} Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias.
 Despite the advantages, meta-analysis implementation is not always straightforward, especially for complex data structures.
-Additionally, the fact that there are several introductory and advanced resources to understand meta-analysis (Borenstein, Hedges, Higgins, \& Rothstein, 2009; Harrer, Cuijpers, Furukawa, \& Ebert, 2021; Schmid, Stijnen, \& White, 2022), to our knowledge, there are no introductory resources about how to simulate realistic meta-analytic data.
+Additionally, the fact that there are several introductory and advanced resources to understand meta-analysis (Borenstein et al., 2009; Harrer, Cuijpers, Furukawa, \& Ebert, 2021; Schmid, Stijnen, \& White, 2022), to our knowledge, there are no introductory resources about how to simulate realistic meta-analytic data.
 Simulating data has several advantages because it requires understanding the statistical method and the data-generation process.
@@
 Simulating data is also a powerful educational tool within this framework.
-For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for fixed and random-effects, and meta-regression models with applications also for statistical power calculation.
+For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for \textcolor{red}{equal} and random-effects, and meta-regression models with applications also for statistical power calculation.
 In the first section, we will introduce basic concepts of the meta-analysis that are useful for setting up the simulation.
-We will evaluate the effect size, variance calculation, and the fixed vs.~random effects model distinction.
+We will evaluate the effect size, variance calculation, and the \textcolor{red}{equal} vs.~random effects model distinction.
 Then we will describe how to simulate data for these models and simulate a meta-regression model with a categorical and numerical predictor.
@@
 Code and materials are available on the OSF repository (\url{https://osf.io/54djn/}).
-Multivariate and multilevel models' simulation examples and more details about the coding approach are available in the supplementary materials.
+\color{red} A theoretical introduction, simulation examples for multivariate and multilevel models, and more details about the coding approach are available in the supplementary materials.
+\color{black}
 
@@
 
-As an example that will be used throughout the paper, we consider the efficacy of memory training in improving memory performance during a cognitive task.
+\color{red} As an example that will be used throughout the paper, we consider the efficacy of memory training in improving memory performance during a cognitive task.
 The typical primary study will collect data from a group of participants receiving the memory training (\emph{experimental group}) and another group receiving a control treatment (\emph{control group}).
 The focus of the meta-analysis is collecting multiple studies with similar aims and methods and estimating the average effect of memory training.
-Despite differences in the type of cognitive task or experimental setup, each primary study collects an \emph{experimental group} (\(n_{C}\)) and \emph{control group} (\(n_{C}\)) and computes the average performance (\emph{experimental group} \({\bar y}_{T}\) and \emph{control group} \({\bar y}_{C}\)) and standard deviations (\emph{experimental group} \(s_{T}\) and \emph{control group}
-\(s_{C}\)).
+Despite differences in the type of cognitive task or experimental setup, each primary study collects an \emph{experimental group} (\(n_{T}\)) and \emph{control group} (\(n_{C}\)) and computes the average performance (\emph{experimental group} \(\overline T\) and \emph{control group} \(\overline C\)) and standard deviations (\emph{experimental group} \(s_{T}\) and \emph{control group}
+\(s_{C}\)). \color{black}
 
@@
 
-The first step of a meta-analysis is to standardize information from each included study to convert each effect into a standard measure.
+The first step of a meta-analysis is to extract information from included studies.
 This common measure should give an immediate idea of the direction (i.e., the treatment improves or reduces performance) and the size of the effect.
-Standardized effect size measures (see Lakens, 2013 for an overview), such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's \(d\) (Cohen, 1988) or the Pearson correlation coefficient \(\rho\) which could be calculated using the associated sample estimator \emph{r} are commonly used to compare heterogeneous outcome variables.
-If all studies used the same raw measure, such as reaction times, it is possible to directly meta-analyze the studies without standardizing.
-In practice, studies often use heterogeneous measures, therefore requiring standardized measures.
+Standardized effect size measures (see Lakens, 2013 for an overview), such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's \(d\) (Cohen, 1988) or the Pearson correlation coefficient \(\rho\) which could be calculated using the associated sample estimator \emph{r} are commonly used to compare heterogeneous outcome variables\footnote{\color{red} When using bounded effect size indexes such as the Pearson correlation, Fisher's \emph{z} transformation is commonly used (Borenstein et al., 2009, pp.
+41--43) \color{black}}.
+If all studies used the same raw measure, such as reaction times, it is possible to directly meta-analyze the studies without standardizing e.g.~using the unstandardized mean difference (UMD, Borenstein et al., 2009, pp.
+21--24).
 
@@
 The idea is to choose the appropriate measure considering the study design (e.g., between vs.~within-subjects) and available information and find the proper formula to compute the sampling variability.
-In addition, there are formulas and approaches to convert from one effect size measure to another (Borenstein et al., 2009; Lakens, 2013; Lipsey \& Wilson, 2001).\footnote{See also the blog post by James Pustejovsky (\href{https://www.jepusto.com/alternative-formulas-for-the-smd/}{link}) for a general method to compute the sampling variability for standardized mean difference measures.} Usually, the effect size sampling variability depends mainly on the sample size that determines the weight assigned during the meta-analytic estimation.
-
-\hypertarget{fixed-vs-random-effects-model}{%
-\subsubsection{Fixed vs random-effects model}\label{fixed-vs-random-effects-model}}
+In addition, there are formulas and approaches to convert from one effect size measure to another (Borenstein et al., 2009; Lakens, 2013; Lipsey \& Wilson, 2001).\footnote{See also the blog post by James Pustejovsky (\href{https://www.jepusto.com/alternative-formulas-for-the-smd/}{www.jepusto.com/alternative-formulas-for-the-smd}) for a general method to compute the sampling variability for standardized mean difference measures.} Usually, the effect size sampling variability depends mainly on the sample size that determines the weight assigned during the meta-analytic estimation.
+
+\hypertarget{equal-vs-fixed-vs-random-effects-model}{%
+\subsubsection{Equal vs fixed vs random-effects model}\label{equal-vs-fixed-vs-random-effects-model}}
+
+\color{red}
 
 The core of a meta-analysis is combining the results of multiple studies giving more weight to studies that provide a more precise effect estimation.
-We can divide meta-analysis models into two categories: the \emph{fixed-effects} and the \emph{random-effects model}.
-If we imagine combining multiple studies replicating the same experiment, we assume the real treatment effect \(\theta_{f}\) is \emph{fixed,} and each study is a more or less precise estimation of this parameter.
-This is an essential and often inappropriate assumption because it ignores that study-level characteristics may impact the actual treatment effect.
-
-Suppose some study-level characteristics (e.g., participants' age, sex, or socioeconomic status) or the experimental paradigm (e.g., type of memory task or difficulty) could impact the treatment effect.
-In that case, the assumption of the \emph{fixed-effects model} is no longer plausible.
-There is no unique real effect but a distribution of effects with mean \(\theta_{r}\) and variance \(\tau^{2}\).
-The mean is the \emph{average} effect, and the variance is the heterogeneity of the phenomenon.
-The \emph{random-effects model} will estimate the mean (\(\theta_{r}\)) and the heterogeneity \(\tau^{2}\) will help generalize to the population of effects beyond the pool of included studies.
-
-Most importantly, the parameters estimated with the \emph{fixed} and \emph{random-effects} do not represent the same quantity.
-The \emph{random-effects model} estimates the \emph{average} effect across a distribution of true effects, while the \emph{fixed-effects model} estimates a single true effect.
-Figure 1 depicts the idea of the fixed vs.~random-effects model.
-The \emph{fixed-effects model} can be considered a particular type of \emph{random-effects model} where \(\tau^{2} = 0\) thus, the average effect is also the unique true effect.
+We can find essentially three meta-analysis models: the equal-effects, the fixed-effects and the random-effects model (Hedges \& Vevea, 1998; Laird \& Mosteller, 1990).
+The \emph{equal-effects} model assume that each study included in the meta-analysis is a more or less precise estimation of the true underlying effect (\(\theta\)).
+In other terms, we are assuming that there is no variability (i.e., heterogeneity) among true effect sizes.
+On the other side, suppose some study-level characteristics (e.g., participants' age, sex, or socioeconomic status) or the experimental paradigm (e.g., type of memory task or difficulty) could impact the treatment effect.
+In this case we have true variability (i.e., heterogeneity) among studies.
+The \emph{random-effects} model assume a distribution of real effects with mean \(\mu_{\theta}\) and variance \(\tau^{2}\) thus estimating the heterogeneity.
+Finally, the \emph{fixed-effects} estimate the average effect of the included pool of studies ignoring the presence of heterogeneity.
+While the distinction between the equal and fixed-effects model is theoretical, the estimated model is the exactly the same\footnote{See \href{https://wviechtb.github.io/metafor/reference/misc-models.html}{wviechtb.github.io/metafor/reference/misc-models.html} for a detailed explanation}.
+The \emph{random-effects} model assume and estimate heterogeneity among effect sizes leading to different model parameters.
+From an inferential point of view, the \emph{random-effects} model provide unconditional inference to the population of effect sizes while the \emph{fixed-effects} model estimate the average effect of the selected studies, and not population-level parameters.
+The \emph{equal-effects} model assume a single true population effect to be estimated.
+The Figure \ref{fig:img-equal-vs-random} depicts the theoretical distinction between the equal and random-effects model.
+In the presence of heterogeneity, we can include moderators (e.g., the type of experimental setup) into a meta-regression model to explain the effect sizes heterogeneity.
+\color{red} An interesting proposal to combine evidence from these different models into a single analysis is called Bayesian model-averaged meta-analysis (Berkhout, Haaf, Gronau, Heck, \& Wagenmakers, 2023; Gronau, Heck, Berkhout, Haaf, \& Wagenmakers, 2021).
+\color{black}
+
+\color{black}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/img-fixed-vs-random-1}
+{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/img-equal-vs-random-1}
 
@@
 
-\caption{The difference between the assumptions of the \emph{fixed} and \emph{random-effects model}.
-Each distribution depicts the sampling distribution of \(k = 4\) hypothetical studies (\(i = 1\), \(2\), \ldots, \(4\)) with a certain observed effect size \(d_{i}\), sampling variability \(\sigma_{i}^{2}\) and the 95\% confidence interval (black segment).
-The \emph{fixed-effects} plot on the left suggests that each observed effect size has the same underlying true effect \(\theta_{f}\) (each distribution has the same mean) with a different degree of precision (e.g., Study 1 is more precise than Study 3).
-In practice, each study has a different observed effect size where studies with high precision (i.e., narrow sampling distributions) will be remarkably close to the real effect (\(\theta_{f}\)).
-The \emph{random-effects model} on the right suggests that beyond the error term (\(\epsilon_{i}\)), each real effect size is composed of a fixed part (now \(\theta_{r}\)) and a random part (\(\theta_{i}\), black segments below the sampling distributions).
-The random part is determined by \(\tau^{2}\), which represents the variability of the effect size distribution at the population level.
-When \(\tau^{2}\) is zero, the random-effects reduces to a fixed-effects model.}\label{fig:img-fixed-vs-random}
-\end{figure}
-
-\normalsize
-
-\hypertarget{multilevel-and-multivariate-models}{%
-\subsubsection{Multilevel and multivariate models}\label{multilevel-and-multivariate-models}}
-
-The most straightforward situation for a meta-analysis is when each included study contributes with a single effect size.
-However, it is common to have multiple effect sizes belonging to the same study creating a situation of dependency between statistical units.
-There are numerous sources of dependency for a meta-analysis (Cheung, 2014; 2015, pp.
-121--122).
-The first type of dependency (\emph{multilevel}) consists of multiple effect sizes from independent participants.
-Despite being collected with different participants, effect sizes from the same study could be correlated.
-We obtain a multilevel data structure with independent effect sizes in the same study.
-There are multiple ways to model this data structure (see Cheung, 2014 for an overview).
-Still, the most common way is to use a \emph{multilevel meta-analysis model,} considering the heterogeneity between and within studies.
-The standard \emph{random-effects} (but also the \emph{fixed-effects model}) model is also called a \emph{two-level} model because it combines data after aggregating the participants-level (\emph{first-level}) data.
-The \emph{three-level} model is commonly used to manage the multilevel data structure (Cheung, 2014, 2019).
-Compared to the \emph{two-level} model, the \emph{three-level} model estimates two heterogeneity parameters (\(\tau^{2}\) and \(\omega^{2}\)), one for the heterogeneity between studies and the other for the heterogeneity within studies.
-
-A second type of dependency (\emph{multivariate}) consists of multiple effect sizes collected from the same pool of participants.
-For example, when using two cognitive tasks on the same pool of participants.
-This is commonly known as \emph{multivariate meta-analysis} because the effect sizes sampling errors are correlated given that participants are assessed multiple times.
-A \emph{multivariate meta-analysis model} could be used to
-consider the correlation between sampling errors.
-
-A third type of dependency arises when true effect sizes are correlated at the population level (Cheung, 2015, pp.
-121--122).
-For example, two different memory tasks could be correlated because the latent psychological constructs are intrinsically correlated.
-This correlation is present even when other groups of participants perform the two memory tasks (i.e., no \emph{multivariate} dependency) or are administered by different researchers (i.e., no \emph{multilevel} dependency).
-This type of dependency could be managed by estimating the correlation between outcomes using a \emph{multivariate random-effects model} (Cheung, 2015, pp.
-127--130).
-Figures \ref{fig:img-multilevel} and \ref{fig:img-multivariate} depict the conceptual idea of \emph{multilevel} and \emph{multivariate} models.
-
-\scriptsize
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth]{img/multilevel}
-
-}
-
-\caption{Graphical representation of a \emph{multilevel} (three-level) model.
-On the top, the effect sizes are distributed at the population level.
-Then, each study is defined as the overall effect and the random effects determined by \(\tau^{2}\), \(\theta_{r} + \theta_{i}\).
-Finally, each experiment (effect size) within a study is composed of the study-specific effect \(\theta_{r} + \theta_{i}\) and the experiment-specific random-effects \(\theta_{r} + \theta_{i} + \theta_{\text{ij}}\) determined by \(\omega^{2}\).}\label{fig:img-multilevel}
-\end{figure}
-
-\normalsize
-
-\scriptsize
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth]{img/multivariate.pdf }
-
-}
-
-\caption{Graphical representation of a \emph{multivariate} model.
-In this example, we have two outcomes (magenta and green) with an average effect, heterogeneity parameters, and a certain true correlation value (\(\rho\), third type of dependency).
-Each study at level-2 (as in the standard random-effects model) has two effect sizes with corresponding sampling variances.
-The two effect sizes are correlated (\(\rho_{s}\), with \(s\) for sampling errors) because each effect is collected on the same participants.}\label{fig:img-multivariate}
+\caption{\color{red} The difference between the assumptions of the \emph{equal} and \emph{random-effects model}.
+Each distribution depicts the sampling distribution of \(k = 4\) hypothetical studies (\(i = 1\), \(2\), \ldots, \(4\)) with a certain observed effect size \(y_{i}\) (red squares), sampling variability \(\sigma_{\epsilon_i}^{2}\) and the 95\% confidence interval (black segment).
+The \emph{equal-effects} plot on the left suggests that each observed effect size has the same underlying true effect \(\theta\) (each distribution has the same mean) with a different degree of precision (e.g., Study 1 is more precise than Study 3).
+In practice, each study has a different observed effect size where studies with high precision (i.e., narrow sampling distributions) will be close to the real effect (\(\theta\)).
+The \emph{random-effects model} on the right suggests that beyond the error term (\(\epsilon_{i}\)), each real effect size is composed of a fixed part (now \(\mu_{\theta}\)) and a random part (\(\theta_{i}\)) sampled from a normal distribution with mean zero and variance \(\tau^{2}\).
+When \(\tau^{2}\) is zero, the random-effects reduces to a equal-effects model.
+\color{black}}\label{fig:img-equal-vs-random}
 \end{figure}
@@
 \FunctionTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{) }\CommentTok{\# seed for simulation to ensure the reproducibility of results}
-\NormalTok{theta }\OtherTok{\textless{}{-}} \FloatTok{0.5} \CommentTok{\# real mean difference in the populations}
+\NormalTok{es }\OtherTok{\textless{}{-}} \FloatTok{0.5} \CommentTok{\# real mean difference in the populations}
 \NormalTok{sigma }\OtherTok{\textless{}{-}} \DecValTok{1} \CommentTok{\# real standard deviations in the population}
@@
 \NormalTok{d\_mc }\OtherTok{\textless{}{-}} \FunctionTok{replicate}\NormalTok{(nsims, }\AttributeTok{expr =}\NormalTok{ \{}
-\NormalTok{  g1 }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\AttributeTok{n =}\NormalTok{ n, }\AttributeTok{mean =}\NormalTok{ theta, }\AttributeTok{sd =}\NormalTok{ sigma)}
+\NormalTok{  g1 }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\AttributeTok{n =}\NormalTok{ n, }\AttributeTok{mean =}\NormalTok{ es, }\AttributeTok{sd =}\NormalTok{ sigma)}
 \NormalTok{  g2 }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\AttributeTok{n =}\NormalTok{ n, }\AttributeTok{mean =} \DecValTok{0}\NormalTok{, }\AttributeTok{sd =}\NormalTok{ sigma)}
@@
 
-We load relevant packages and set the \texttt{seed} for reproducibility to prepare the R environment for the simulations.
+\color{red}
+
+This tutorial use several R packages for the simulations, meta-analysis fitting and figures/tables.
+For the data manipulation we used the \emph{tidyverse} (Wickham, 2023) package.
+For the models fitting we used the \emph{metafor} (Viechtbauer, 2010) package.
+For figures and tables the \emph{ggplot2} (Wickham et al., 2023) or the \texttt{metafor::forest()} function, \emph{kableExtra} (Zhu, 2021) and \emph{papaja} (Aust \& Barth, 2022) packages.
+We set the \texttt{seed} for reproducibility of the simulation environment.
+\color{black}
 
@@
 \begin{Highlighting}[]
-\FunctionTok{library}\NormalTok{(dplyr) }\CommentTok{\# for data manipulation}
-\FunctionTok{library}\NormalTok{(tidyr) }\CommentTok{\# for data manipulation}
+\FunctionTok{library}\NormalTok{(dplyr) }\CommentTok{\# for data manipulation (within tidyverse)}
+\FunctionTok{library}\NormalTok{(tidyr) }\CommentTok{\# for data manipulation (within tidyverse)}
 \FunctionTok{library}\NormalTok{(metafor) }\CommentTok{\# for meta{-}analysis models fitting}
+\FunctionTok{library}\NormalTok{(purrr) }\CommentTok{\# for *apply like functions}
 \NormalTok{seed }\OtherTok{\textless{}{-}} \DecValTok{2023} \CommentTok{\# general seed for all simulations}
@@
 
+\color{red}
+
 Before diving into the specific simulations, in this section, we define the common aspects of all simulations in the following sections.
-The current paper focuses on the \emph{two-level}, \emph{fixed,} and \emph{random-effects model}s.
+The current paper focuses on the \emph{two-level}, \emph{equal,} and \emph{random-effects} models.
 All the examples refer to primary studies that assess the efficacy of a treatment by comparing a control and an experimental group.
-Equation \eqref{eq:datagen} formalizes the assumption about generating participants' data within a single study.
-The two groups (\emph{experimental} \(y_{T}\) and \emph{control} \(y_{C}\)) are generated from normal distributions with \(\mu = 0\) for the control group and \(\mu = \theta\) (where \(\theta\) is the effect size) for the experimental group.
-Both groups are assumed to be sampled from populations with homogeneous variances \(\sigma^{2} = 1\).
+In simulation studies where the purpose is not evaluating effect size estimators, it is convenient to simulate unstandardized effect size measures (see Viechtbauer, 2005, 2007).
+The estimator is unbiased thus not requiring small sample correction (Hedges, 1981; e.g., Hedges, 1989).
+Furthermore, the effect size and the sampling variance are independent.
+Similar to the simulation approach by Viechtbauer (2005, 2007) the \emph{experimental} group (\(T\)) and \emph{control} groups \(C\) are sampled from normal distributions respectively \(T_i \sim \mathcal{N}(\Delta, 1)\) and \(C_i \sim \mathcal{N}(0, 1)\) where \(\Delta\) is the unstandardized mean difference (UMD).
+The UMD and the sampling variance are calculated using Equations \eqref{eq:effsize} and \eqref{eq:effsizevar} where \(y\) is the estimated value of \(\Delta\).
+\color{black}
+
+\color{red}
 
@@
 \begin{gathered}
-y_{Ti} \sim \mathcal{N}(\theta, 1) \\
-y_{Ci} \sim \mathcal{N}(0, 1)
-\label{eq:datagen}
-\end{gathered}
-\end{align}
-
-The true Standardized Mean Difference can be estimated using Cohen's \(d\) as reported in Equation \eqref{eq:effsize}.
-The difference between the two means is divided by the pooled standard deviation of the two groups.
-It is common to apply a small-sample correction version of this effect size called Hedges' \(g\) (Hedges, 1981)
-
-\begin{align}
-\begin{gathered}
-d = \frac{\bar y_T - \bar y_C}{S_p} \\
-S_p = \sqrt{\frac{S^2_{y_T}(n_{y_T} - 1) + S^2_{y_C}(n_{C} - 1)}{n_{T} + n_{C} - 2}} \\
-\sigma^2_d = \frac{n_{T} + n_{C}}{n_{T}n_{C}} + \frac{d^2}{2(n_{T} + n_{C} - 2)}
+D = \overline{T} - \overline{C}
 \label{eq:effsize}
@@
 
+\begin{align}
+\begin{gathered}
+\sigma_{\epsilon D}^{2} = \frac{s^2_T}{n_T} + \frac{s^2_C}{n_C}
+\label{eq:effsizevar}
+\end{gathered}
+\end{align}
+
+\color{black}
+
 We can use the following algorithm implemented in the \texttt{sim\_study()} function to simulate a single study. Before using the \texttt{sim\_study()}
@@
 Table \ref{tab:sim-data-example-tab} depicts an example of the \texttt{make\_data()} output.
+
+\color{red}
 
@@
 \item
-  Choose a \(\theta\), \(n_{T}\), and \(n_{C}\) value.
+  Choose a \(\Delta\), \(n_{T}\), and \(n_{C}\) value.
 \item
-  Simulate \(n_{T}\) observations from a Gaussian distribution with \(\mu = \theta\) and \(\sigma^{2} = 1\) and \(n_{T}\) observations from a Gaussian distribution with \(\mu = 0\) and \(\sigma^{2} = 1\).
-This way, the expected difference between groups will be \(d\) and the variance is fixed at 1.
+  Simulate \(n_{T}\) observations from a Gaussian distribution with \(\mu = \Delta\) and \(\sigma^{2} = 1\) and \(n_{T}\) observations from a Gaussian distribution with \(\mu = 0\) and \(\sigma^{2} = 1\).
+In this way, the expected difference between groups will be \(\Delta\) and the expected variance for each group is 1.
 \item
-  Calculate the pooled standard deviation \(S_{p}\) that is an estimation of the population standard deviation (\(\sigma\)) to standardize the mean difference.
-\item
-  Calculate the observed effect size \(d\) and the sampling variance \(\sigma_{d}^{2}\).
+  Calculate the observed effect size \(y\) and the sampling variance \(\sigma_{\epsilon}^{2}\).
 \end{enumerate}
+
+The simulation approach can be easily extended calculating a standardized effect size measures (e.g., Cohens' \(d\) Cohen, 1988) and the corresponding sampling variance.
+For example, after generating data for the two groups the mean difference can be standardized using the pooled standard deviation and applying the appropriate correction (Hedges, 1981; e.g., Hedges, 1989).
+
+\color{black}
 
@@
 
-\NormalTok{sim\_study }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(theta, nt, }\AttributeTok{nc =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{hedges =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{aggregate =} \ConstantTok{TRUE}\NormalTok{)\{}
+\NormalTok{sim\_study }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(es, nt, }\AttributeTok{nc =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{aggregate =} \ConstantTok{TRUE}\NormalTok{)\{}
   \ControlFlowTok{if}\NormalTok{(}\FunctionTok{is.null}\NormalTok{(nc)) nc }\OtherTok{\textless{}{-}}\NormalTok{ nt}
@@
 \NormalTok{  yc }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(nc, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)}
-\NormalTok{  yt }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(nt, theta, }\DecValTok{1}\NormalTok{)}
-  \CommentTok{\# pooled variance}
-\NormalTok{  sp }\OtherTok{\textless{}{-}} \FunctionTok{sqrt}\NormalTok{((}\FunctionTok{var}\NormalTok{(yc)}\SpecialCharTok{*}\NormalTok{(nc }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{var}\NormalTok{(yt)}\SpecialCharTok{*}\NormalTok{(nt }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{)) }\SpecialCharTok{/}\NormalTok{ (nc }\SpecialCharTok{+}\NormalTok{ nt }\SpecialCharTok{{-}} \DecValTok{2}\NormalTok{))}
+\NormalTok{  yt }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(nt, es, }\DecValTok{1}\NormalTok{)}
+
   \CommentTok{\# effect size}
-\NormalTok{  yi }\OtherTok{\textless{}{-}}\NormalTok{ (}\FunctionTok{mean}\NormalTok{(yt) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(yc)) }\SpecialCharTok{/}\NormalTok{ sp}
-
-  \CommentTok{\# transform to hedges\textquotesingle{} g}
-  \ControlFlowTok{if}\NormalTok{(hedges)\{}
-\NormalTok{    yi }\OtherTok{\textless{}{-}}\NormalTok{ yi }\SpecialCharTok{*}\NormalTok{ metafor}\SpecialCharTok{:::}\FunctionTok{.cmicalc}\NormalTok{(nc }\SpecialCharTok{+}\NormalTok{ nt }\SpecialCharTok{{-}} \DecValTok{2}\NormalTok{)}
-\NormalTok{  \}}
+\NormalTok{  yi }\OtherTok{\textless{}{-}}\NormalTok{ (}\FunctionTok{mean}\NormalTok{(yt) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(yc))}
 
   \CommentTok{\# sampling variance}
-\NormalTok{  vi }\OtherTok{\textless{}{-}}\NormalTok{ (nc }\SpecialCharTok{+}\NormalTok{ nt)}\SpecialCharTok{/}\NormalTok{(nc }\SpecialCharTok{*}\NormalTok{ nt) }\SpecialCharTok{+}\NormalTok{ yi}\SpecialCharTok{\^{}}\DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\DecValTok{2} \SpecialCharTok{*}\NormalTok{ (nc }\SpecialCharTok{+}\NormalTok{ nt }\SpecialCharTok{{-}} \DecValTok{2}\NormalTok{))}
+\NormalTok{  vi }\OtherTok{\textless{}{-}} \FunctionTok{var}\NormalTok{(yt)}\SpecialCharTok{/}\NormalTok{nt }\SpecialCharTok{+} \FunctionTok{var}\NormalTok{(yt)}\SpecialCharTok{/}\NormalTok{nc}
 
@@
 
-The \texttt{theta} is the true effect size, \texttt{nc} and \texttt{nt} are the sample size for the control and experimental group, \texttt{aggregate} controls if returning the effect size and the corresponding variance or the participant-level data, and \texttt{hedges} specify whether applying the correction.
+\color{red}
+
+The \texttt{es} is the true effect size (\(\mu_{\theta}\) for the random-effects model and \(\theta\) for the equal-effects model), \texttt{nc} and \texttt{nt} are the sample size for the control and experimental group, \texttt{aggregate} controls if returning the effect size and the corresponding sampling variance or the participant-level data.
 We can generate a single study with the desired parameters with this function.
@@
 The \texttt{mapply()} function is clearly explained in the supplementary materials.
+\color{black}
 
@@
 \FunctionTok{set.seed}\NormalTok{(seed)}
-\FunctionTok{sim\_study}\NormalTok{(}\AttributeTok{theta =} \FloatTok{0.3}\NormalTok{, }\AttributeTok{nc =} \DecValTok{10000}\NormalTok{, }\AttributeTok{nt =} \DecValTok{10000}\NormalTok{, }\AttributeTok{aggregate =} \ConstantTok{FALSE}\NormalTok{)}
+\FunctionTok{sim\_study}\NormalTok{(}\AttributeTok{es =} \FloatTok{0.3}\NormalTok{, }\AttributeTok{nc =} \DecValTok{10000}\NormalTok{, }\AttributeTok{nt =} \DecValTok{10000}\NormalTok{, }\AttributeTok{aggregate =} \ConstantTok{FALSE}\NormalTok{)}
 \end{Highlighting}
@@
 Using the \texttt{sim\_study()} function multiple times (with the appropriate adjustments) we can generate a series of studies simulating a data set for a meta-analysis (using the \texttt{sim\_studies()} function).
-After each example, we will compute the appropriate model (e.g., \emph{fixed} or \emph{random-effects}) using the \texttt{metafor} package (Viechtbauer, 2010) to check the recovery of simulated parameters.
+After each example, we will compute the appropriate model (e.g., \emph{equal} or \emph{random-effects}) using the \texttt{metafor} package (Viechtbauer, 2010) to check the recovery of simulated parameters\footnote{\color{red} By default \texttt{metafor} use the Wald \emph{z} test on model parameters but other inference methods are available e.g.~the Knapp and Hartung (2003) method by setting \texttt{test="knha"} within the \texttt{rma()} function.
+This method is useful especially when the number of studies is low (see \url{https://wviechtb.github.io/metafor/reference/misc-recs.html} for more details) \color{black}}.
 Table \ref{tab:notation-tab} summarizes the simulation parameters notation in equations and code.
@@
 \NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{20} \CommentTok{\# number of participants per group, per study}
-\NormalTok{theta }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# real effect size}
-\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{theta =}\NormalTok{ theta)}
+\NormalTok{es }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# real effect size}
+\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{es =}\NormalTok{ es)}
 \end{Highlighting}
@@
 \toprule
-id & \multicolumn{1}{c}{nt} & \multicolumn{1}{c}{nc} & \multicolumn{1}{c}{theta}\\
+id & \multicolumn{1}{c}{nt} & \multicolumn{1}{c}{nc} & \multicolumn{1}{c}{es}\\
 \midrule
@@
 
+\color{red}
+
 \scriptsize
@@
 \midrule
-$\theta_f$ & \texttt{theta\_f} & Fixed-effects model real effect size\\
-$\theta_r$ & \texttt{theta\_r} & Random-effects model real effect size\\
-$\tau^2$ & \texttt{tau2} & Effect size heterogeneity\\
-$\tau^2_r$ & \texttt{tau2r} & Residual effect size heterogeneity\\
-$\bar y_{T, C}$ & \texttt{} & Mean of the experimental/control group\\
-$s^2_{T, C}$ & \texttt{} & Standard deviation of the experimental/control group\\
+$\theta$ & \texttt{theta} & Equal-effects model true effect size\\
+$\mu_{\theta}$ & \texttt{mu\_theta} & Random-effects model true effect size\\
+$\tau^2$ & \texttt{tau2} & Effect sizes heterogeneity\\
+$\tau^2_r$ & \texttt{tau2r} & Residual effect sizes heterogeneity\\
+$\overline{T}, \overline{C}$ & \texttt{} & Mean of the experimental/control group\\
+$s_{T, C}$ & \texttt{} & Standard deviation of the experimental/control group\\
 $n_{T, C}$ & \texttt{nt, nc} & Sample size of the experimental/control group\\
-$d_i$ & \texttt{yi} & Observed effect size\\
-$\sigma^2_i$ & \texttt{vi} & Observed sampling variance\\
-$\theta_i$ & \texttt{theta\_i} & Random effect for the study i\\
+$y_i$ & \texttt{yi} & Observed effect size\\
+$\sigma_{\epsilon_i}^{2}$ & \texttt{vi} & Observed sampling variance\\
+$\delta_i$ & \texttt{deltai} & Random effect for the study i\\
 $\beta_0$ & \texttt{b0} & Meta-regression intercept\\
@@
 \begin{tablenotes}[para]
-\normalsize{\textit{Note.} The \texttt{yi} and \texttt{vi} notation for the observed effect size ($d_i$) and sampling variance ($\sigma^2_i$) has been used to be consistent with the \texttt{metafor} notation.}
+\normalsize{\textit{Note.} The \texttt{yi} and \texttt{vi} notation for the observed effect size ($y_i$) and sampling variance ($\sigma_{\epsilon_i}^{2}$) has been used to be consistent with the \texttt{metafor} notation.}
 \end{tablenotes}
@@
 
-\hypertarget{fixed-effects-model}{%
-\subsection{Fixed-effects model}\label{fixed-effects-model}}
-
-The most basic model to simulate is the \emph{fixed-effects model} (FE).
-As reported in the previous sections, the \emph{fixed-effects model} assumes the presence of a single true effect (\(\theta_{f}\)) and the observed variability is because each effect size is an imprecise estimation of the true effect.
-In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies
-(i.e., the sample size). Equation \eqref{eq:fixed-effect-model} depicts the \emph{fixed-effect} equation.
+\hypertarget{equal-effects-model}{%
+\subsection{Equal-effects model}\label{equal-effects-model}}
+
+The most basic model to simulate is the \emph{equal-effects model} (EE).
+As reported in the previous sections, the \emph{equal-effects model} assumes the presence of a single true effect (\(\theta\)) and the \color{red} observed variability is caused by each effect size being a more or less imprecise estimation of the true effect.
+\color{black} In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies (i.e., the sample size).
+Equations \eqref{eq:equal-effects-model1} and \eqref{eq:equal-effects-model2} \textcolor{red}{formalize} the \emph{equal-effects} model.
 
@@
 \begin{gathered}
-d_i = \theta_f + \epsilon_i \\
-\epsilon_i \sim \mathcal{N}(0,\sigma^2_i)
-\label{eq:fixed-effect-model}
+y_i = \theta + \epsilon_i
+\label{eq:equal-effects-model1}
 \end{gathered}
@@
 
-Each observed effect size (\(d_{i}\)) is composed of the real effect size \(\theta_{f}\) plus an error term (\(\epsilon_{i}\)) that is sampled from a normal distribution with \(\mu = 0\) and \(\sigma^{2} = \sigma_{i}^{2}\) (i.e., the known sampling variance of the study \(i\)).
-As demonstrated in
-Equation \eqref{eq:effsize}, the increase in sample size will decrease the sampling variability.
-An exact study will essentially have \(\theta_{f}\) as the observed effect size.
+\color{black}
+
+\begin{align}
+\begin{gathered}
+\epsilon_i \sim \mathcal{N}(0,\sigma_{\epsilon_i}^{2})
+\label{eq:equal-effects-model2}
+\end{gathered}
+\end{align}
+
+Each observed effect size \color{red} (\(y_{i}\)) \color{black} is composed of the real effect size \color{red} \(\theta\) \color{black} plus an error term (\(\epsilon_{i}\)) that is sampled from a normal distribution with \(\mu = 0\) and \color{red} \(\sigma^{2} = \sigma_{\epsilon_i}^{2}\) \color{black} (i.e., the known sampling variance of the study \(i\)).
+As demonstrated in Equation \eqref{eq:effsize}, the increase in sample size will decrease the sampling variability.
+\textcolor{red}{A study with a extremely large sample size} will essentially have \(\theta\) as the observed effect size.
 Since we are sampling participants' level data, the error component is already included in the \texttt{sim\_study()} function.
@@
 \NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{20} \CommentTok{\# number of participants per group, per study}
-\NormalTok{theta\_f }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# real effect size}
-
-\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{theta\_f =}\NormalTok{ theta\_f)}
-\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{theta\_f, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim) }\CommentTok{\# simulated data}
-\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(}\AttributeTok{yi =}\NormalTok{ yi, }\AttributeTok{vi =}\NormalTok{ vi, }\AttributeTok{method =} \StringTok{"FE"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
+\NormalTok{theta }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# real effect size}
+
+\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{es =}\NormalTok{ theta)}
+\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{es, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim) }\CommentTok{\# simulated data}
+\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(}\AttributeTok{yi =}\NormalTok{ yi, }\AttributeTok{vi =}\NormalTok{ vi, }\AttributeTok{method =} \StringTok{"EE"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
 \end{Highlighting}
@@
 
-We are using the \(\theta_{f}\) parameter to generate random data using the \texttt{sim\_study()} function that will introduce the random error component \(\epsilon_{i}\).
-Then we can fit the \emph{fixed-effects} meta-analysis model with the \texttt{rma} function and \texttt{method\ =\ "FE"} of the \texttt{metafor} package.
-The parameter we are estimating is \(\theta_{f}\) which is close to our simulation value (see Table \ref{tab:res-fixed-effect}).
+\color{red}
+
+We are using the \(\theta\) parameter to generate random data using the \texttt{sim\_study()} function that will introduce the random error component \(\epsilon_{i}\).
+Then we can fit the \emph{equal-effects} meta-analysis model with the \texttt{rma} function and \texttt{method\ =\ "EE"} of the \texttt{metafor} package.
+The parameter we are estimating is \(\theta\) which is close to our simulation value (see Table \ref{tab:res-equal-effects}).
 Increasing the number of studies (\(k\)) and/or the number of participants in each study (\(n\)) will improve estimation reducing the standard error.
+\color{black}
 
@@
 
-\caption{\label{tab:res-fixed-effect}Summary of the simulated fixed-effects model. The only estimated parameter is the average effect* \(\theta_{f}\) *with the standard error,
-95\% confidence interval, and the Wald z-test. The z-test evaluates the null hypothesis that the real effect equals zero.}
+\caption{\label{tab:res-equal-effects}\color{red} Summary of the simulated equal-effects model.
+The only estimated parameter is the average effect \(\theta\) (\(\beta\)) with the standard error, 95\% confidence interval, and the Wald z-test.
+The z-test evaluates the null hypothesis that the real effect equals zero.\color{black}}
 \centering
@@
 \midrule
-overall & 0.334 (SE = 0.059) & $[0.219, 0.449]$ & 5.701 & < 0.001\\
+overall & 0.347 (SE = 0.056) & $[0.237, 0.456]$ & 6.223 & < 0.001\\
 \bottomrule
@@
 Increasing the sample size for each study will increase the estimation precision thus the variability among studies will be reduced.
-This can be easily demonstrated by simulating studies with a high sample size as reported in Figure \ref{fig:fixed-effect-high-low-precision}.
+This can be easily demonstrated by simulating studies with a high sample size as reported in Figure \ref{fig:equal-effects-high-low-precision}.
 As the sample size increases the only source of variability (i.e., the error component \(\epsilon_{i}\)) is close to zero so each study is closer to the true simulated value.
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/fixed-effect-high-low-precision-1}
+{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/equal-effects-high-low-precision-1}
 
@@
 
-\caption{Forest plots of two simulated \emph{fixed-effects models}.
+\caption{\color{red} Forest plots of two simulated \emph{equal-effects models}.
 On the left, the simulated model has \(nt,nc = 30\) for each included study while on the right the sample size for each study is \(nt,nc = 500\).
-Given that the data were generated under a \emph{fixed-effects model}, when the sample size is high (on the right) each study is aligned on the real effect size because the error component (\(\epsilon_{i}\)) is almost entirely reduced.
-The average effect size is similar (depending on the random numbers generation) between the two scenarios while the estimation precision (the width of the black diamond) is narrower on the right.}\label{fig:fixed-effect-high-low-precision}
+Given that the data were generated under a \emph{equal-effects model}, when the sample size is high (on the right) each study is aligned on the real effect size because the error component (\(\epsilon_{i}\)) is almost entirely reduced.
+The average effect size is similar (depending on the random numbers generation) between the two scenarios while the estimation precision (the width of the black diamond) is narrower on the right.
+\color{black}}\label{fig:equal-effects-high-low-precision}
 \end{figure}
@@
 
-The \emph{random-effects model} (RE) can be considered an extension of the \emph{fixed-effects model}.
-The \emph{fixed-effects model} assumes that the real effect is a single value.
-The random-effects model relaxes this assumption allowing \(\theta_{f}\) (now \(\theta_{r}\)) to vary across studies.
+\color{red}
+
+The \emph{random-effects model} (RE) can be considered an extension of the \emph{equal-effects model}.
+The \emph{equal-effects model} assumes that the real effect is a single value.
+The random-effects model relaxes this assumption allowing the true effect size to vary across studies.
 For example, the difference between groups we are simulating could be influenced by the type of experiment or the participants' age.
-Now \(\theta_{r}\) is no longer a single value but a distribution of values.
-Due to the effect size being a distribution, we need two parameters \(\mu\) and \(\sigma^{2}\) to describe the phenomenon.
-The parameter \(\theta_{r}\) average true effect is now interpreted as the average effect size across different true effect sizes and \(\sigma^{2}\) usually known as \(\tau^{2}\) is the effect size variability or heterogeneity.
-In practical terms, we now have two sources of variability: \(\tau^{2}\), which express the real difference among effect sizes, and \(\sigma_{i}^{2}\) which is the known sampling variance of each study as in the \emph{fixed-effects model}.
-We can easily extend Equation \eqref{eq:fixed-effect-model} with Equation \eqref{eq:random-effect-model}.
+Now \(\theta\) is no longer a single value but a distribution of values.
+Due to the effect size being a distribution, we need to estimate both the mean and the variance.
+The parameter \(\mu_{\theta}\) is mean of the distribution, interpreted as the average effect size across different true effect sizes and \(\tau^{2}\) is the variance of the distribution interpreted as variability or heterogeneity of effect sizes.
+In practical terms, we now have two sources of variability: \(\tau^{2}\), which express the real difference among effect sizes, and \(\sigma_{\epsilon_i}^{2}\) which is the known sampling variance of each study as in the \emph{equal-effects model}.
+We can easily extend Equation \eqref{eq:equal-effects-model1} with Equations \eqref{eq:random-effects-model1}, \eqref{eq:random-effects-model2} and \eqref{eq:random-effects-model3}.
 
@@
 \begin{aligned}
-d_i = \theta_r + \theta_i + \epsilon_i \\
-\theta_i \sim \mathcal{N}(0, \tau^2) \\
-\epsilon_i \sim \mathcal{N}(0,\sigma^2_i)
-\label{eq:random-effect-model}
+y_i = \mu_{\theta} + \delta_i + \epsilon_i \\
+\label{eq:random-effects-model1}
 \end{aligned}
@@
 
-Compared to the \emph{fixed-effects model} we need to generate another adjustment to the overall effect \(\theta_{r}\) from a normal distribution with mean \(\theta_{r}\) and variance \(\tau^{2}\).
-The real effect size for the study \(i\) will be \(\theta_{r} + \theta_{i}\) where \(\theta_{i}\) are the random-effects regulated by the \(\tau^{2}\) parameter.
+\begin{align}
+\begin{aligned}
+\delta_i \sim \mathcal{N}(0,\tau^2)
+\label{eq:random-effects-model2}
+\end{aligned}
+\end{align}
+
+\begin{align}
+\begin{aligned}
+\epsilon_i \sim \mathcal{N}(0,\sigma_{\epsilon_i}^{2})
+\label{eq:random-effects-model3}
+\end{aligned}
+\end{align}
+
+Compared to the \emph{equal-effects model} we need to generate another adjustment to the overall effect \(\mu_{\theta}\) from a normal distribution with mean 0 and variance \(\tau^{2}\).
+The real effect size for the study \(i\) will be \(\mu_{\theta} + \theta_{i}\) where \(\theta_{i}\) are the random-effects regulated by the \(\tau^{2}\) parameter.
+
+\color{black}
 
@@
 \NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{20} \CommentTok{\# number of participants per group, per study}
-\NormalTok{theta\_r }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# real effect size}
+\NormalTok{mu\_theta }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# real effect size}
 \NormalTok{tau2 }\OtherTok{\textless{}{-}} \FloatTok{0.2} \CommentTok{\# the effect size heterogeneity}
 
-\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{theta\_r =}\NormalTok{ theta\_r)}
+\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{mu\_theta =}\NormalTok{ mu\_theta)}
 
 \CommentTok{\# simulate the random{-}effects adjustment}
-\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2))}
+\NormalTok{sim}\SpecialCharTok{$}\NormalTok{deltai }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2))}
 
 \CommentTok{\# adding the by{-}study adjustment to the average effect}
-\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_r\_theta\_i }\OtherTok{\textless{}{-}}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_r }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_i }
-
-\CommentTok{\# now we are using theta\_r\_theta\_i and no longer theta\_f}
-\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{theta\_r\_theta\_i, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim)}
+\NormalTok{sim}\SpecialCharTok{$}\NormalTok{es }\OtherTok{\textless{}{-}}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{mu\_theta }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{deltai }
+
+\CommentTok{\# now we are using mu\_deltai and no longer theta}
+\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{es, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim)}
 \NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(}\AttributeTok{yi =}\NormalTok{ yi, }\AttributeTok{vi =}\NormalTok{ vi, }\AttributeTok{method =} \StringTok{"REML"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
@@
 Then we can fit the \emph{random-effects} meta-analysis model with the \texttt{rma} function and \texttt{method\ =\ "REML"}.
-Table \ref{tab:res-random-effect} depicts the model results\footnote{The \texttt{method\ =\ "REML"} is not the only method to estimate a random-effects model.
-See the \texttt{rma} documentation (\url{https://wviechtb.github.io/metafor/reference/rma.uni.html\#specifying-the-model}) for an overview of other estimation methods} of the metafor package.
-We are now estimating two parameters: \(\theta_{r}\) and \(\tau^{2}\).
-As for the \emph{fixed-effects model}, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
+Table \ref{tab:res-random-effects} depicts the model results\footnote{The \texttt{method\ =\ "REML"} is not the only method to estimate a random-effects model.
+See the \texttt{rma} documentation (\url{https://wviechtb.github.io/metafor/reference/rma.uni.html\#specifying-the-model}) for an overview of other estimation methods} of the \texttt{metafor} package.
+We are now estimating two parameters: \(\mu_{\theta}\) and \(\tau^{2}\).
+As for the \color{red} \emph{equal-effects model} \color{black}, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
 
@@
 
-\caption{\label{tab:res-random-effect}Summary of the simulated random-effects model: Compared to the fixed-effects model, there are more parameters.
-The \(\beta\) is the average effect (\(\theta_{r}\)) with the standard error, 95\% confidence interval, and the Wald z-test.
-The \(\tau^{2}\) is the estimated heterogeneity and the \(I^{2}\) (explained in Section \ref{re-model}) represents the percentage of total variability due to between-study heterogeneity.}
+\caption{\label{tab:res-random-effects}\color{red} Summary of the simulated random-effects model: Compared to the equal-effects model, there are more parameters.
+The \(\beta\) is the average effect (\(\mu_{\theta}\)) with the standard error, 95\% confidence interval, and the Wald z-test.
+The \(\tau^{2}\) is the estimated heterogeneity and the \(I^{2}\) (explained in Section \ref{re-model}) represents the percentage of total variability due to between-study heterogeneity.
+\color{black}}
 \centering
@@
 \midrule
-overall & 0.381 (SE = 0.107) & $[0.172, 0.590]$ & 3.578 & < 0.001\\
+overall & 0.394 (SE = 0.110) & $[0.179, 0.609]$ & 3.596 & < 0.001\\
 \bottomrule
 \multicolumn{5}{l}{\textsuperscript{} $k = 30$}\\
-\multicolumn{5}{l}{\textsuperscript{} $\tau^2 = 0.234$ ($SE = 0.089$)}\\
-\multicolumn{5}{l}{\textsuperscript{} $I^2 = 68.842\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $\tau^2 = 0.262$ ($SE = 0.095$)}\\
+\multicolumn{5}{l}{\textsuperscript{} $I^2 = 73.480\%$}\\
 \end{tabular}
@@
 An important aspect of the \emph{random-effects model} is the interplay between the heterogeneity (\(\tau^{2}\)) and the sampling variability (\(\sigma_{i}^{2}\)).
-As the number of studies \(k\) increases, the estimation of \(\theta_{r}\) and \(\tau^{2}\) becomes more precise.
-Additionally, as each study\textquotesingle s sample size decreases, each \(\theta_{i}\) will be estimated with higher precision but as long as \(\tau^{2} \neq 0\)
-there will be variability among effect sizes.
+As the number of studies \(k\) increases, the estimation of \(\mu_{\theta}\) and \(\tau^{2}\) becomes more precise \color{red} (Blázquez-Rincón, Sánchez-Meca, Botella, \& Suero, 2023; Rubio-Aparicio et al., 2018) \color{black}.
+Additionally, as the sample size of each study decreases, each \(\theta_{i}\) will be estimated with higher precision but as long as \(\tau^{2} \neq 0\) there will be variability among effect sizes.
 In other words, increasing the sample size of each study or the number of studies will not affect the value of \(\tau^{2}\) but only the estimation precision (see Borenstein et al., 2009, Chapter 16, Figure 16.6 for a clear explanation).
-This can be easily demonstrated using the previous simulation, increasing each study\textquotesingle s sample size.
-Figure \ref{fig:random-effect-high-low-precision} depicts the same meta-analysis but with different precision in estimating the study \(d_{i}\).
-Compared to Figure \ref{fig:fixed-effect-high-low-precision}, increasing the sample size of primary studies improves the estimation of each study without reducing the between-studies heterogeneity.
+This can be easily demonstrated using the previous simulation, increasing each study sample size.
+Figure \ref{fig:random-effects-high-low-precision} depicts the same meta-analysis but with different precision in estimating the study \(y_{i}\).
+Compared to Figure \ref{fig:equal-effects-high-low-precision}, increasing the sample size of primary studies improves the estimation of each study without reducing the between-studies heterogeneity.
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/random-effect-high-low-precision-1}
+{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/random-effects-high-low-precision-1}
 
@@
 
-\caption{Forest plots of two simulated \emph{random-effects model}s.
+\caption{\color{red} Forest plots of two simulated \emph{random-effects model}s.
 On the left, the simulated model has \(nt,nc = 30\) for each included study, while on the right, the sample size for each study is \(nt,nc = 500\).
-Compared to Figure \ref{fig:fixed-effect-high-low-precision}, data were generated under a \emph{random-effects model}.
+Compared to Figure \ref{fig:equal-effects-high-low-precision}, data were generated under a \emph{random-effects model}.
 The estimated average effect is similar between the two scenarios regarding average effect and precision.
-However, compared to the \emph{fixed-effect} simulation, increasing the sample size of primary studies only affects the precision without reducing the true heterogeneity (i.e., \(\tau^{2} \neq 0\)).}\label{fig:random-effect-high-low-precision}
+However, compared to the \emph{equal-effects} simulation, increasing the sample size of primary studies only affects the precision without reducing the true heterogeneity (i.e., \(\tau^{2} \neq 0\)).
+\color{black}}\label{fig:random-effects-high-low-precision}
 \end{figure}
@@
 The relationship between the sampling error and the heterogeneity can be expressed using the \(I^{2}\) statistics (e.g., Higgins \& Thompson, 2002) that is the percentage of the total variability \(\tau^{2} + \tilde{v}\) that is attributable to real heterogeneity between studies (\(\tau^{2}\)).
-In Equation \eqref{eq:isquared}, \(\tilde{v}\) is the ``typical'' within-study sampling variability (e.g., Higgins \& Thompson, 2002)\footnote{see also here \url{https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate\#fn__1} for an overview of the \(I^2\) statistics also for complex models}.
+\color{red} In Equation \eqref{eq:isquared}, \(\tilde{v}\) is the \emph{typical} within-study sampling variability (i.e., a summary statistics of sampling variances of included studies) as proposed by Higgins and Thompson (2002)\footnote{See \url{https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate\#fn__1} for an overview of the \(I^2\) statistics also for complex models.
+There are also other proposals to compute the \(\tilde{v}\) statistics (see Takkouche, Cadarso-Suárez, \& Spiegelman, 1999; Takkouche, Khudyakov, Costa-Bouzas, \& Spiegelman, 2013)} and implemented in Equation \eqref{eq:typical-v} with \(w_i = 1/\sigma_{\epsilon_i}^{2}\) and \(k\) is the number of studies.
+\color{black}
 
@@
 
-From Equation \eqref{eq:isquared} and Figure \ref{fig:random-effect-high-low-precision} is clear that if each included study has a considerable sample size the total variability will be reduced, mainly driven by real heterogeneity (\(\tau^2\)).
-This is the crucial difference between the fixed-effects and the random-effects model (see also Borenstein et al., 2009, pp.
+\color{red}
+
+\begin{align}
+\begin{gathered}Random-effects model true effect
+\tilde{v} = \frac{(k-1) \sum w_i}{(\sum w_i)^2 - \sum w_i^2}
+\label{eq:typical-v}
+\end{gathered}
+\end{align}
+\color{black}
+
+From Equation \eqref{eq:isquared} and Figure \ref{fig:random-effects-high-low-precision} is clear that if each included study has a considerable sample size the total variability will be reduced, mainly driven by real heterogeneity (\(\tau^2\)).
+This is the crucial difference between the \textcolor{red}{equal-effects} and the random-effects model (see also Borenstein et al., 2009, pp.
 117--122).
+
+\color{red}
 
@@
 The only caveat is fixing the \(\tilde{v}\).
-When the sample size of each study is the same, \(\tilde{v} = \sigma^2_i\).
-In the other case, \(\tilde{v}\) needs to be calculated from sampling variances reported in Higgins and Thompson (2002) and cannot be easily fixed a priori.
-However, with the assumption of homogeneous sample size across studies\footnote{The homogeneous sample size of primary studies assumption is commonly used to calculate the statistical power (see Borenstein et al., 2009, Chapter 29)}, we can solve Equation \eqref{eq:isquared} for \(\tau^{2}\) obtaining the heterogeneity value associated with a certain \(I^{2}\) as reported in Equation \eqref{eq:tau2-from-isquared}.
-Table \ref{tab:res-random-effect-i2} depicts the results of the random-effects model fixing the \(I^{2}\) value.
+When the sample size of each study is the same\footnote{The homogeneous sample size of primary studies assumption is commonly used to calculate the statistical power (see Borenstein et al., 2009, Chapter 29)}, \(\tilde{v} = \sigma_{\epsilon_i}^{2}\).
+In the other case, \(\tilde{v}\) needs to be calculated from sampling variances as reported in Higgins and Thompson (2002) and cannot be easily fixed a priori.
+With the assumption of homogeneous sample size across studies\footnote{In case of heterogeneous sample sizes generated from a probability distribution (e.g., Gaussian) the \emph{typical} sampling variance calculated using the expected value of the distribution can be used to somehow fix the (average) \(I^2\) (see the supplementary materials for a simulated example)}, we can solve Equation \eqref{eq:isquared} for \(\tau^{2}\) obtaining the heterogeneity value associated with a certain \(I^{2}\) as reported in Equation \eqref{eq:tau2-from-isquared}.
+\color{black}
+
+\color{red}
+
+Table \ref{tab:res-random-effects-i2} depicts the results of the random-effects model fixing the \(I^{2}\) value.
+Choosing a meaningful \(\tau^2\) for the simulation can be sometimes difficult.
+Beyond fixing the \(I^2\) value, a possibility is choosing plausible \(\tau^2\) value from the literature.
+For example van Erp and colleagues (2017) estimate an empirical \(\tau^2\) distribution across several published meta-analyses that can be used to simulate a plausible scenario.
+\color{black}
 
@@
 \NormalTok{k }\OtherTok{\textless{}{-}} \DecValTok{30} \CommentTok{\# number of studies}
-\NormalTok{theta\_r }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# real effect size}
+\NormalTok{mu\_theta }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# real effect size}
 \NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{30} \CommentTok{\# sample size per group, per study}
@@
 
-\NormalTok{v }\OtherTok{\textless{}{-}}\NormalTok{ (n }\SpecialCharTok{+}\NormalTok{ n)}\SpecialCharTok{/}\NormalTok{(n }\SpecialCharTok{*}\NormalTok{ n) }\SpecialCharTok{+}\NormalTok{ theta\_r}\SpecialCharTok{\^{}}\DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\DecValTok{2} \SpecialCharTok{*}\NormalTok{ (n }\SpecialCharTok{+}\NormalTok{ n }\SpecialCharTok{{-}} \DecValTok{2}\NormalTok{)) }\CommentTok{\# typical within study variance}
+\NormalTok{v }\OtherTok{\textless{}{-}}\NormalTok{ (n }\SpecialCharTok{+}\NormalTok{ n)}\SpecialCharTok{/}\NormalTok{(n }\SpecialCharTok{*}\NormalTok{ n) }\SpecialCharTok{+}\NormalTok{ mu\_theta}\SpecialCharTok{\^{}}\DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\DecValTok{2} \SpecialCharTok{*}\NormalTok{ (n }\SpecialCharTok{+}\NormalTok{ n }\SpecialCharTok{{-}} \DecValTok{2}\NormalTok{)) }\CommentTok{\# typical within study variance}
 \NormalTok{tau2 }\OtherTok{\textless{}{-}} \SpecialCharTok{{-}}\NormalTok{((I2}\SpecialCharTok{*}\NormalTok{v)}\SpecialCharTok{/}\NormalTok{(I2 }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{))}
 
-\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{theta\_r =}\NormalTok{ theta\_r)}
-
-\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2))}
-\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_r\_theta\_i }\OtherTok{\textless{}{-}}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_r }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_i }
-\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{theta\_r\_theta\_i, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim)}
+\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{make\_data}\NormalTok{(}\AttributeTok{k =}\NormalTok{ k, }\AttributeTok{nc =}\NormalTok{ n, }\AttributeTok{nt =}\NormalTok{ n, }\AttributeTok{mu\_theta =}\NormalTok{ mu\_theta)}
+
+\NormalTok{sim}\SpecialCharTok{$}\NormalTok{deltai }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2))}
+\NormalTok{sim}\SpecialCharTok{$}\NormalTok{es }\OtherTok{\textless{}{-}}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{mu\_theta }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{deltai }
+\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{es, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim)}
 \NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(yi, vi, }\AttributeTok{method =} \StringTok{"REML"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
@@
 
-\caption{\label{tab:res-random-effect-i2}Summary of the \emph{random-effect} model fixing the \(I^2\) value. Parameters are the same as described in Table \ref{tab:res-random-effect}.}
+\caption{\label{tab:res-random-effects-i2}Summary of the random-effects model fixing the \(I^2\) value. Parameters are the same as described in Table \ref{tab:res-random-effects}.}
 \centering
@@
 \midrule
-overall & 0.298 (SE = 0.071) & $[0.159, 0.438]$ & 4.191 & < 0.001\\
+overall & 0.300 (SE = 0.071) & $[0.160, 0.440]$ & 4.201 & < 0.001\\
 \bottomrule
 \multicolumn{5}{l}{\textsuperscript{} $k = 30$}\\
-\multicolumn{5}{l}{\textsuperscript{} $\tau^2 = 0.083$ ($SE = 0.040$)}\\
-\multicolumn{5}{l}{\textsuperscript{} $I^2 = 54.830\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $\tau^2 = 0.087$ ($SE = 0.040$)}\\
+\multicolumn{5}{l}{\textsuperscript{} $I^2 = 57.904\%$}\\
 \end{tabular}
@@
 
-From a linear regression perspective, both the FE and the RE models can be seen as \emph{intercept-only} models where only the mean (i.e., the linear regression intercept) is estimated Borenstein et al.~(2009).
+From a linear regression perspective, both the EE and the RE models can be seen as \emph{intercept-only} models where only the mean (i.e., the linear regression intercept) is estimated.
 As reported in the meta-analysis introduction, the between-study heterogeneity usually represents the true variability of the effect due to differences among primary studies.
@@
 For example, a group of studies could use a particular memory task where the expected effect is higher than another.
-In this way, considering the
-type of task will explain part of the observed heterogeneity as in standard regression models.
+In this way, considering the type of task will explain part of the observed heterogeneity as in standard regression models.
 Figures \ref{fig:img-metaregression-bin} and \ref{fig:img-metaregression-num} depict a random-effects meta-regression model for a categorical and numerical predictor.
@@
 The effect size differs between conditions A and B, including the \emph{condition} moderator explaining part of the total heterogeneity (red plus green segments).
-The green segments depict the explained heterogeneity, and the red segments the residual (unexplained) heterogeneity.}\label{fig:img-metaregression-bin}
+The green segments depict the explained heterogeneity, and the red segments the residual (unexplained) heterogeneity.
+\color{red} The red squares are simulated observed effect sizes from the sampling distributions.
+\color{black}}\label{fig:img-metaregression-bin}
 \end{figure}
@@
 In our example, a group of studies uses an online memory task while others use a standard lab-based task.
-The Equation \eqref{eq:random-effect-model} can be easily extended for a meta-regression model by including a variable encoding the type of task (online vs.~lab-based) and the expected difference between the two levels of the moderator (i.e., the \emph{lab vs.~online effect}).
-In regression terms, we could use a dummy variable (\(X_{1}\)) that takes the value of 0 for the \emph{lab-based task} (\emph{l}) and a value of 1 for an \emph{online task} (\emph{o})\footnote{The \emph{dummy coding} (also known as treatment coding) is the default in R and \texttt{metafor} but other coding schemes could be used (see Schad, Vasishth, Hohenstein, \& Kliegl, 2020 for an overview of contrast coding schemes)}.
-Now we fix \(\beta_{1}\) to be the \emph{lab vs.~online effect} Equation \eqref{eq:random-effect-model} into Equation \eqref{eq:meta-regression}.
-Now the product between \(\beta_{1}X_{1}\) will consider the \emph{lab vs.~online effect}.
-To simulate this meta-regression, we need to fix the \(\beta_{1}\) effect.
-Due to the fact that we are on a standardized scale, we can fix \(\beta_{1}\) to be the expected difference in Cohen's \(d\) scale between the two groups of studies.
+The Equation \eqref{eq:random-effects-model1} can be easily extended for a meta-regression model by including a variable encoding the type of task (online vs.~lab-based) and the expected difference between the two levels of the moderator (i.e., the \emph{lab vs.~online effect}).
+In regression terms (see equation \eqref{eq:meta-regression}), we could use a dummy variable (\(X_{1}\)) that takes the value of 0 for the \emph{lab-based task} (\emph{L}) and a value of 1 for an \emph{online task} (\emph{O})\footnote{The \emph{dummy coding} (also known as treatment coding) is the default in R and \texttt{metafor} but other coding schemes could be used (see Schad, Vasishth, Hohenstein, \& Kliegl, 2020 for an overview of contrast coding schemes).
+\color{red} As in standard linear regression, the contrast coding of categorical factors or the presence/absence of the intercept influence the interpretation of estimated model parameters and omnibus meta-regression tests (see \url{https://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept} for a detailed discussion)} \color{black}.
+Now we fix \(\beta_{1}\) to be the \emph{lab vs.~online effect} (i.e., the expected mean difference between the two groups of studies) and the product between \color{red}\(\beta_{1}X_{1_i}\) \color{black} will consider the \emph{lab vs.~online effect}.
 Crucially, despite \(\tau^{2}\) is still the heterogeneity between effect sizes, now we need to fix \(\tau^{2}\) considering that we included a moderator.
-In other terms, \(\tau_{r}^{2}\) is the residual heterogeneity after explaining the
-variance due to \(X_{1}\). We can describe our model using two equations according to the value of \(X_{1}\) as reported in Equation \eqref{eq:meta-regression-equations}.
+In other terms, \color{red} \(\delta_i \sim \mathcal{N}(0, \tau_{r}^{2})\) \color{black} where \(\tau^2_r\) is the residual heterogeneity after including the moderator.
+We can describe our model using equation \eqref{eq:meta-regression-equations} according to the value of \(X_{1_i}\) (L = lab-based experiments, O = online experiments).
 
@@
 \begin{gathered}
-d_i = \beta_0 + \theta_i + \beta_1 X_1 + \epsilon_i \\
-\theta_i \sim \mathcal{N}(0, \tau^2) \\
-\epsilon_i \sim \mathcal{N}(0,\sigma^2_i)
+y_i = \beta_0 + \delta_i + \beta_1 X_{1_i} + \epsilon_i
 \label{eq:meta-regression}
@@
 \begin{gathered}
-d_i = \beta_0 + \theta_i + \beta_1 X_1 + \epsilon_i \\
-d_{il} = \beta_0 + \theta_i + \beta_1 \times 0 + \epsilon_i \\
-d_{io} = \beta_0 + \theta_i + \beta_1 \times 1 + \epsilon_i \\
+y_{L_i} = \beta_0 + \delta_i + \beta_1 \times 0 + \epsilon_i \\
+y_{O_i} = \beta_0 + \delta_i + \beta_1 \times 1 + \epsilon_i
 \label{eq:meta-regression-equations}
@@
 
-We can simulate the same scenario of the random-effects model with \(k_o = 15\) (\emph{online tasks}) and \(k_l = 15\) (\emph{lab-based tasks}).
+We can simulate the same scenario of the random-effects model with \(k_O = 15\) (\emph{online tasks}) and \(k_L = 15\) (\emph{lab-based tasks}).
 Then we fix the \(\beta_1 = 0.2\) and \(\tau_r^2 = 0.1\) (\(r\) for residual).
@@
 
-\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2r)) }\CommentTok{\# the by{-}study residual adjustment}
-\NormalTok{sim}\SpecialCharTok{$}\NormalTok{b0\_theta\_i\_x }\OtherTok{\textless{}{-}}\NormalTok{ b0 }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_i }\SpecialCharTok{+}\NormalTok{ b1}\SpecialCharTok{*}\FunctionTok{ifelse}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{exp }\SpecialCharTok{==} \StringTok{"lab"}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)}
-
-\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{b0\_theta\_i\_x, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim)}
+\NormalTok{sim}\SpecialCharTok{$}\NormalTok{deltai }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2r)) }\CommentTok{\# the by{-}study residual adjustment}
+\NormalTok{sim}\SpecialCharTok{$}\NormalTok{es }\OtherTok{\textless{}{-}}\NormalTok{ b0 }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{deltai }\SpecialCharTok{+}\NormalTok{ b1}\SpecialCharTok{*}\FunctionTok{ifelse}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{exp }\SpecialCharTok{==} \StringTok{"lab"}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)}
+
+\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{es, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim)}
 \NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(yi, vi, }\AttributeTok{mods =} \SpecialCharTok{\textasciitilde{}}\NormalTok{exp, }\AttributeTok{method =} \StringTok{"REML"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
@@
 Then the \(\beta_1\) parameter represents the estimated difference in \(y\) between the values of \(X_1\) (i.e., lab-based vs online experiments).
-As said before, \(\tau^2\) is now the residual heterogeneity that is interpreted as the variability between effect sizes after controlling for the moderator \(X_1\).
-
-Now we can fit the meta-regression model with the \texttt{rma} function as for the random-effects model with the addition of mods = \texttt{\textasciitilde{}\ exp} that indicates which variable/s to consider as moderators.
-The results are presented in Table \ref{tab:res-meta-reg-dummy}.
-Next, the model will estimate an \emph{intercept} parameter (i.e., \(\beta_{0}\)) that is the value of \(y\) when \(X_{1}\) is zero (i.e., for lab-based studies) or, in other terms, the expected value for lab-based studies.
-Then the \(\beta_{1}\) parameter represents the estimated difference in \(y\) between the values of \(X_{1}\) (i.e., lab-based vs.~online experiments).
-As mentioned above, \(\tau^{2}\) the residual heterogeneity is now interpreted as the variability between effect sizes after controlling for the moderator \(X_{1}\).
+As said before, \color{red} \(\tau^2_r\) \color{black} is now the residual heterogeneity that is interpreted as the variability between effect sizes after controlling for the moderator \(X_1\).
 
@@
 
-\caption{\label{tab:res-meta-reg-dummy}Summary of the \emph{random-effect} model with a categorical predictor (lab vs online experiments).
+\caption{\label{tab:res-meta-reg-dummy}\color{red} Summary of the \emph{random-effect} model with a categorical predictor (lab vs online experiments).
 The \texttt{intercept} is the average effect for lab-based experiments and \texttt{exponline} is the difference between lab-based and online experiments.
 The \(R^2\) is the percentage of explained heterogeneity and \(\tau^2_r\) is the estimated residual heterogeneity.
-Other parameters are the same as the standard \emph{random-effect} model (see Table \ref{tab:res-random-effect}).}
+Other parameters are the same as the standard \emph{random-effect} model (see Table \ref{tab:res-random-effects}).
+\color{black}}
 \centering
@@
 \midrule
-intercept & 0.041 (SE = 0.101) & $[-0.157, 0.238]$ & 0.403 & 0.687\\
-exponline & 0.321 (SE = 0.143) & $[0.040, 0.601]$ & 2.243 & 0.025\\
+intercept & 0.038 (SE = 0.101) & $[-0.161, 0.236]$ & 0.372 & 0.710\\
+exponline & 0.323 (SE = 0.143) & $[0.043, 0.602]$ & 2.263 & 0.024\\
 \bottomrule
 \multicolumn{5}{l}{\textsuperscript{} $k = 30$}\\
-\multicolumn{5}{l}{\textsuperscript{} $\tau^2_r = 0.085$ ($SE = 0.041$)}\\
-\multicolumn{5}{l}{\textsuperscript{} $R^2 = 19.902\%$}\\
-\multicolumn{5}{l}{\textsuperscript{} $I^2 = 55.435\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $\tau^2_r = 0.087$ ($SE = 0.041$)}\\
+\multicolumn{5}{l}{\textsuperscript{} $R^2 = 20.659\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $I^2 = 57.741\%$}\\
 \end{tabular}
@@
 A first strategy could be to use values estimated from the literature.
-Another approach consists in setting up the model and simulating several expected \(d_i\) and calculating the range of simulated values.
+Another approach consists in setting up the model and simulating several expected \color{red} \(y_i\) \color{black} and calculating the range of simulated values.
 A third possibility is fixing the proportion of explained heterogeneity calculating the \(\beta_1\) value accordingly.
@@
 In this specific example, if all studies target a specific population (e.g., adults below 50 years), the expected average age range can be easily simulated.
-In our case, we simulated \(k\) average age values from a uniform distribution \(\text{ag}e_{i} \sim U\left( 20,40 \right)\).
-Then we can plot the distribution of \(d_{i}\) values to check the plausibility of simulated values.
+In our case, we simulated \(k\) average age values from a uniform distribution \(\text{age}_{i} \sim U\left( 20,40 \right)\).
+Then we can plot the distribution of \(y_{i}\) values to check the plausibility of simulated values.
 As shown in Figure \ref{fig:plot-meta-reg-plausible-implausible}, with the same range for the moderator, a \(\beta_{1} = 0.1\) gives a plausible effect sizes range while a \(\beta_{1} = 0.7\) predicts very extreme values.
@@
 
-\NormalTok{theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2r))}
-\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(theta\_i, }\AttributeTok{each =} \DecValTok{2}\NormalTok{)}
-
-\NormalTok{sim}\SpecialCharTok{$}\NormalTok{b0\_theta\_i\_x }\OtherTok{\textless{}{-}}\NormalTok{ b0 }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_i }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{b1}\SpecialCharTok{*}\NormalTok{sim}\SpecialCharTok{$}\NormalTok{age0}
-
-\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{b0\_theta\_i\_x, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim)}
+\NormalTok{deltai }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2r))}
+\NormalTok{sim}\SpecialCharTok{$}\NormalTok{deltai }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(deltai, }\AttributeTok{each =} \DecValTok{2}\NormalTok{)}
+
+\NormalTok{sim}\SpecialCharTok{$}\NormalTok{es }\OtherTok{\textless{}{-}}\NormalTok{ b0 }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{deltai }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{b1}\SpecialCharTok{*}\NormalTok{sim}\SpecialCharTok{$}\NormalTok{age0}
+
+\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{es, sim}\SpecialCharTok{$}\NormalTok{nc, sim}\SpecialCharTok{$}\NormalTok{nt, }\AttributeTok{data =}\NormalTok{ sim)}
 \end{Highlighting}
@@
 
-\caption{Scatter plots with marginal histograms for the range of simulated \(d_i\) with two \(\beta_1\) values.
+\caption{\color{red} Scatter plots with marginal histograms for the range of simulated \(y_i\) with two \(\beta_1\) values.
 The x-axis depicts the average age of simulated studies and the y-axis depicts the simulated effect size.
 On the left, the majority of simulated values range between -1.5 and 1.5 thus \(\beta_1 = 0.1\) can be considered a plausible value.
-On the right, (\(\beta = 0.7\)) values range between -10 to 10 that despite theoretically possible is highly implausible values in real meta-analysis for psychological data.}\label{fig:plot-meta-reg-plausible-implausible}
+On the right, (\(\beta = 0.7\)) values range between -10 to 10 that despite theoretically possible can be considered highly implausible values in real meta-analyses for psychological data.
+\color{black}}\label{fig:plot-meta-reg-plausible-implausible}
 \end{figure}
@@
 This approach has been implemented by Lopez-Lopez and colleagues (2014).
-We can use Equation \eqref{eq:beta-from-r2} to find the \(\beta_{1}\) value that is associated with a certain \(R^{2}\).
+We can use Equation \eqref{eq:beta-from-r21} and \eqref{eq:beta-from-r22} to find the \(\beta_{1}\) value that is associated with a certain \(R^{2}\).
 
@@
 \begin{aligned}
-\beta^2_1 = \tau^2R^2 \\
+\beta^2_1 = \tau^2R^2
+\label{eq:beta-from-r21}
+\end{aligned}
+\end{align}
+
+\begin{align}
+\begin{aligned}
 \tau^2_{r} = \tau^2 - \beta^2_1
-\label{eq:beta-from-r2}
+\label{eq:beta-from-r22}
 \end{aligned}
@@
 Lopez-Lopez and colleagues (2014) generated the moderator (\(X_{1}\)) values from a standard normal distribution.
-In the following example, we standardized the moderator (\texttt{scale()}) after simulating values on the \emph{age} scale
-(e.g., \texttt{runif(k,\ 20,\ 40)}).
+In the following example, we standardized the moderator (\texttt{scale()}) after simulating values on the \emph{age} scale (e.g., \texttt{runif(k,\ 20,\ 40)}).
 
@@
 
-\NormalTok{sim}\SpecialCharTok{$}\NormalTok{theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2r))}
+\NormalTok{sim}\SpecialCharTok{$}\NormalTok{deltai }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2r))}
 \NormalTok{sim}\SpecialCharTok{$}\NormalTok{age0 }\OtherTok{\textless{}{-}} \FunctionTok{scale}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{age, }\AttributeTok{center =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{scale =} \ConstantTok{TRUE}\NormalTok{) }\CommentTok{\# standardize the moderator}
-\NormalTok{sim}\SpecialCharTok{$}\NormalTok{b0\_theta\_i\_x }\OtherTok{\textless{}{-}}\NormalTok{ b0 }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{theta\_i }\SpecialCharTok{+}\NormalTok{ b1}\SpecialCharTok{*}\NormalTok{sim}\SpecialCharTok{$}\NormalTok{age0}
-
-\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{b0\_theta\_i\_x, sim}\SpecialCharTok{$}\NormalTok{nt, sim}\SpecialCharTok{$}\NormalTok{nc, }\AttributeTok{data =}\NormalTok{ sim)}
+\NormalTok{sim}\SpecialCharTok{$}\NormalTok{es }\OtherTok{\textless{}{-}}\NormalTok{ b0 }\SpecialCharTok{+}\NormalTok{ sim}\SpecialCharTok{$}\NormalTok{deltai }\SpecialCharTok{+}\NormalTok{ b1}\SpecialCharTok{*}\NormalTok{sim}\SpecialCharTok{$}\NormalTok{age0}
+
+\NormalTok{sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(sim}\SpecialCharTok{$}\NormalTok{es, sim}\SpecialCharTok{$}\NormalTok{nt, sim}\SpecialCharTok{$}\NormalTok{nc, }\AttributeTok{data =}\NormalTok{ sim)}
 \NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(yi, vi, }\AttributeTok{mods =} \SpecialCharTok{\textasciitilde{}}\NormalTok{age0, }\AttributeTok{method =} \StringTok{"REML"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
@@
 
-\caption{\label{tab:res-meta-reg-numerical-r2}Summary of the \emph{random-effects} model fixing the \(R^{2}\) value.
+\caption{\label{tab:res-meta-reg-numerical-r2}\color{red} Summary of the \emph{random-effects} model fixing the \(R^{2}\) value.
 The \texttt{intercept} is the effect size for the average \texttt{age} (given that age is mean-centered).
 The \texttt{age0} parameter is the slope between \texttt{age} and the effect size interpreted as an increase in effect size for a unit increase in the average age.
-Other parameters are the same as described in Table \ref{tab:res-random-effect} and \ref{tab:res-meta-reg-dummy}.}
+Other parameters are the same as described in Table \ref{tab:res-random-effects} and \ref{tab:res-meta-reg-dummy}.
+\color{black}}
 \centering
@@
 \midrule
-intercept & 0.335 (SE = 0.051) & $[0.236, 0.435]$ & 6.586 & < 0.001\\
-age0 & 0.163 (SE = 0.051) & $[0.063, 0.263]$ & 3.184 & 0.001\\
+intercept & 0.339 (SE = 0.051) & $[0.239, 0.439]$ & 6.643 & < 0.001\\
+age0 & 0.168 (SE = 0.051) & $[0.067, 0.268]$ & 3.272 & 0.001\\
 \bottomrule
 \multicolumn{5}{l}{\textsuperscript{} $k = 100$}\\
-\multicolumn{5}{l}{\textsuperscript{} $\tau^2_r = 0.189$ ($SE = 0.037$)}\\
-\multicolumn{5}{l}{\textsuperscript{} $R^2 = 11.283\%$}\\
-\multicolumn{5}{l}{\textsuperscript{} $I^2 = 73.038\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $\tau^2_r = 0.197$ ($SE = 0.037$)}\\
+\multicolumn{5}{l}{\textsuperscript{} $R^2 = 11.029\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $I^2 = 76.611\%$}\\
 \end{tabular}
@@
 However, one of the most critical applications is estimating the power of a specific statistical model.
-
-As explained in the introduction, there are several approaches and tools to estimate the power of \emph{fixed} and \emph{random-effects models} (see Borenstein et al., 2009, Chapter 29; Harrer, Cuijpers, \& Ebert, 2019, Chapter 14).
+\color{red} One of the purpose of power analysis by simulations is estimating the required number of studies to detect an hypothetical effect size when planning a meta-analysis.
+At the same time, the meta-analysis that we presented (\emph{two-level} equal or random-effects model) can be considered as a multi-lab (e.g., Klein et al., 2018) study where experiments are planned and not collected from the literature.
+In fact, we can use the same simulation approach to optimize the number of participants and studies when planning a multi-lab project.
+\color{black}
+
+As explained in the introduction, there are several approaches and tools to estimate the power of \emph{equal} and \emph{random-effects models} (see Borenstein et al., 2009, Chapter 29; Harrer, Cuijpers, \& Ebert, 2019, Chapter 14).
 These methods are easy to implement but made strong assumptions, such as the homogeneity of sample size, and did not consider the uncertainty in estimating \(\tau^{2}\).
-Jackson and colleagues (2017)
-partially solve the issue by developing an interesting method that takes into account the uncertainty in estimating \(\tau^{2}\) without using simulations.
-However, only Monte Carlo simulations can consider complex scenarios and extra parameters.
+Jackson and colleagues (2017) partially solve the issue by developing an interesting method that takes into account the uncertainty in estimating \(\tau^{2}\) without using simulations.
+However, only Monte Carlo simulations can \textcolor{red}{handle} complex scenarios and extra parameters.
 
@@
 \item
-  Choose the model that generates the data (e.g., fixed, or random-effects model).
+  Choose the model that generates the data (e.g., equal or random-effects model).
 \item
@@
 
-For example, we can estimate the power of a \emph{random-effects model} by repeating the simulation presented in Section \ref{re-model} many times.
+For example, we can estimate the power of a \emph{random-effects} model by repeating the simulation presented in Section \ref{re-model} many times.
 We simulated heterogeneity of sample sizes sampling \(n_{T}\) and \(n_{C}\) values from a Poisson distribution with \(\lambda = 20\).
@@
 Then the \texttt{summary\_sim()} function analyzes each simulation returning the relevant values.
-We repeat the simulation of the \emph{random-effects model} several times with different parameters.
+We repeat the simulation of the \emph{random-effects} model several times with different parameters.
 
@@
 
-\NormalTok{do\_sim }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(k, theta, tau2, navg, nmin, nsim, }\AttributeTok{alpha =} \FloatTok{0.05}\NormalTok{, }\AttributeTok{summary =} \ConstantTok{TRUE}\NormalTok{)\{}
+\NormalTok{do\_sim }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(k, mu, tau2, navg, nmin, nsim, }\AttributeTok{alpha =} \FloatTok{0.05}\NormalTok{, }\AttributeTok{summary =} \ConstantTok{TRUE}\NormalTok{)\{}
   \CommentTok{\# preallocate for computation speed}
@@
   \ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{nsim)\{}
-\NormalTok{    theta\_i }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2))}
-\NormalTok{    theta\_theta\_i }\OtherTok{\textless{}{-}}\NormalTok{ theta }\SpecialCharTok{+}\NormalTok{ theta\_i}
-    \CommentTok{\# simulate sample size}
+\NormalTok{    deltai }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(k, }\DecValTok{0}\NormalTok{, }\FunctionTok{sqrt}\NormalTok{(tau2))}
+\NormalTok{    es }\OtherTok{\textless{}{-}}\NormalTok{ mu }\SpecialCharTok{+}\NormalTok{ deltai}
+    \CommentTok{\# simulate sample size, it is possible to use other distributions e.g., Gaussian}
 \NormalTok{    n }\OtherTok{\textless{}{-}}\NormalTok{ nmin }\SpecialCharTok{+} \FunctionTok{rpois}\NormalTok{(k, navg }\SpecialCharTok{{-}}\NormalTok{ nmin)}
     \CommentTok{\# simulate the studies}
-\NormalTok{    sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(theta\_theta\_i, n, n, }\AttributeTok{data =} \ConstantTok{NULL}\NormalTok{)}
+\NormalTok{    sim }\OtherTok{\textless{}{-}} \FunctionTok{sim\_studies}\NormalTok{(es, n, n, }\AttributeTok{data =} \ConstantTok{NULL}\NormalTok{)}
 \NormalTok{    res }\OtherTok{\textless{}{-}} \FunctionTok{rma}\NormalTok{(yi, vi, }\AttributeTok{method =} \StringTok{"REML"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sim)}
@@
 \NormalTok{k }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{15}\NormalTok{, }\DecValTok{25}\NormalTok{, }\DecValTok{35}\NormalTok{, }\DecValTok{50}\NormalTok{) }\CommentTok{\# number of studies}
-\NormalTok{theta }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# the average effect size}
+\NormalTok{delta }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# the average effect size}
 \NormalTok{tau2 }\OtherTok{\textless{}{-}} \FloatTok{0.3} \CommentTok{\# the heterogeneity}
@@
 \CommentTok{\# creating all combinations}
-\NormalTok{sim\_grid }\OtherTok{\textless{}{-}}\NormalTok{ tidyr}\SpecialCharTok{::}\FunctionTok{expand\_grid}\NormalTok{(k, theta, tau2, navg, nmin, nsim)}
+\NormalTok{sim\_grid }\OtherTok{\textless{}{-}}\NormalTok{ tidyr}\SpecialCharTok{::}\FunctionTok{expand\_grid}\NormalTok{(k, }\AttributeTok{mu =}\NormalTok{ delta, tau2, navg, nmin, nsim)}
 
@@
 \toprule
-k & \multicolumn{1}{c}{$\theta$} & \multicolumn{1}{c}{$\tau^2$} & \multicolumn{1}{c}{$n_{avg}$} & \multicolumn{1}{c}{$n_{min}$} & \multicolumn{1}{c}{nsim} & \multicolumn{1}{c}{Power}\\
+k & \multicolumn{1}{c}{$\Delta$} & \multicolumn{1}{c}{$\tau^2$} & \multicolumn{1}{c}{$n_{avg}$} & \multicolumn{1}{c}{$n_{min}$} & \multicolumn{1}{c}{nsim} & \multicolumn{1}{c}{Power}\\
 \midrule
-5 & 0.30 & 0.30 & 20 & 10 & 5000 & 0.25\\
-15 & 0.30 & 0.30 & 20 & 10 & 5000 & 0.46\\
+5 & 0.30 & 0.30 & 20 & 10 & 5000 & 0.26\\
+15 & 0.30 & 0.30 & 20 & 10 & 5000 & 0.47\\
 25 & 0.30 & 0.30 & 20 & 10 & 5000 & 0.65\\
@@
 
-The present work introduced the basic concepts of the meta-analysis regarding fixed, random-effects models and meta-regression with a simulation-based approach.
+The present work introduced the basic concepts of the meta-analysis regarding equal, random-effects models and meta-regression with a simulation-based approach.
 We believe the presented examples can help implement alternative or more complex models.
-For example, the \texttt{sim\_study()} function can be easily modified to simulate another effect sizes index, such as a paired sample Cohen's \(d\) or a Pearson's
-correlation.
-In addition, more complex models, such as \emph{multivariate} or \emph{multilevel} models, can be implemented following the same approach (see supplementary materials).
-For example, the \emph{three-level} model estimates another heterogeneity component representing the variability of multiple effect sizes in the same study.
+For example, the \texttt{sim\_study()} function can be easily modified to simulate another effect sizes index, \color{red} such as correlations or odds ratios.
+\color{black} In addition, more complex models, such as \emph{multivariate} or \emph{multilevel} models, can be implemented following the same approach (see \textcolor{red}{the} supplementary materials).
+For example, the \color{red} \emph{multilevel} (e.g., \emph{three-level}) \color{black} model estimates another heterogeneity component representing the variability of multiple \textcolor{red}{independent} effect sizes \textcolor{red}{within} the same study.
 Similarly, the \emph{multivariate} model could include the correlation between multiple outcomes and the correlation between sampling errors.
@@
 We decided to give the foundations to understand meta-analyses with a simulation approach because more complex models are still based on the same principles.
-Second, there are limitations
-concerning simulating participant-level data.
+Second, there are limitations concerning simulating participant-level data.
 We decided to simulate the meta-analysis dataset starting from the participants' level to maximize the flexibility and clearness of each step.
 The downside concerns the efficiency and scalability of the simulation setup.
-For large-scale simulations (e.g., many conditions, iterations, or complex models), simulating from aggregated statistics is probably more efficient (see
-Heuvel et al., 2020 for an example) to improve the computation speed\footnote{see \url{https://www.jepusto.com/simulating-correlated-smds/} for a very clear example of the participant-level vs aggregated data simulation}.
+For large-scale simulations (e.g., many conditions, iterations, or complex models), simulating from aggregated statistics is probably more efficient (see Heuvel, Almalik, \& Zhan, 2020 for an example) to improve the \textcolor{red}{simulation efficency}\footnote{See \url{https://www.jepusto.com/simulating-correlated-smds/} for a very clear example of the participant-level vs aggregated data simulation approaches}.
 
@@
 
+The supplementary materials can be also found on the Open Science Framework repository (\url{https://osf.io/54djn/})
+
 \hypertarget{prior-versions}{%
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-R-papaja}{}}%
+Aust, F., \& Barth, M. (2022). \emph{Papaja: Prepare american psychological association journal articles with r markdown}. Retrieved from \url{https://github.com/crsh/papaja}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Berkhout2023-fj}{}}%
+Berkhout, S.
+W., Haaf, J.
+M., Gronau, Q.
+F., Heck, D.
+W., \& Wagenmakers, E.-J.
+(2023).
+A tutorial on bayesian model-averaged meta-analysis in {JASP}.
+\emph{Behav.
+Res.
+Methods}.
+\url{https://doi.org/10.3758/s13428-023-02093-6}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Blazquez-Rincon2023-kp}{}}%
+Blázquez-Rincón, D., Sánchez-Meca, J., Botella, J., \& Suero, M.
+(2023).
+Heterogeneity estimation in meta-analysis of standardized mean differences when the distribution of random effects departs from normal: A monte carlo simulation study.
+\emph{BMC Med.
+Res.
+Methodol.}, \emph{23}(1), 19.
+\url{https://doi.org/10.1186/s12874-022-01809-0}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Borenstein2009-mo}{}}%
 Borenstein, M., Hedges, L. V., Higgins, J. P. T., \& Rothstein, H. R. (2009). \emph{Introduction to {Meta-Analysis}}. \url{https://doi.org/10.1002/9780470743386}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Cheung2014-fg}{}}%
-Cheung, M.
-W.-L.
-(2014).
-Modeling dependent effect sizes with three-level meta-analyses: A structural equation modeling approach.
-\emph{Psychol.
-Methods}, \emph{19}(2), 211--229.
-\url{https://doi.org/10.1037/a0032968}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Cheung2015-bw}{}}%
-Cheung, M. W.-L. (2015). \emph{{Meta-Analysis}: A structural equation modeling approach}. John Wiley \& Sons.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Cheung2019-po}{}}%
-Cheung, M.
-W.-L.
-(2019).
-A guide to conducting a {Meta-Analysis} with {Non-Independent} effect sizes.
-\emph{Neuropsychol.
-Rev.}, \emph{29}(4), 387--396.
-\url{https://doi.org/10.1007/s11065-019-09415-6}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Gronau2021-do}{}}%
+Gronau, Q.
+F., Heck, D.
+W., Berkhout, S.
+W., Haaf, J.
+M., \& Wagenmakers, E.-J.
+(2021).
+A primer on bayesian model-averaged meta-analysis.
+\emph{Adv.
+Methods Pract.
+Psychol.
+Sci.}, \emph{4}(3), 251524592110312.
+\url{https://doi.org/10.1177/25152459211031256}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Harrer2019-rl}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Hedges1989-ip}{}}%
+Hedges, L. V. (1989). An unbiased correction for sampling error in validity generalization studies. \emph{J. Appl. Psychol.}, \emph{74}(3), 469--477. \url{https://doi.org/10.1037/0021-9010.74.3.469}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hedges1998-uo}{}}%
+Hedges, L. V., \& Vevea, J. L. (1998). Fixed- and random-effects models in meta-analysis. \emph{Psychol. Methods}, \emph{3}(4), 486--504. \url{https://doi.org/10.1037/1082-989X.3.4.486}
+
+\leavevmode\vadjust pre{\hypertarget{ref-van-den-Heuvel2020-rv}{}}%
+Heuvel, E.
+R.
+van den, Almalik, O., \& Zhan, Z.
+(2020).
+\emph{Simulation models for aggregated data meta-analysis: Evaluation of pooling effect sizes and publication biases}.
+Retrieved from \url{https://arxiv.org/abs/2009.06305}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Higgins2002-fh}{}}%
@@
 Jackson, D., \& Turner, R. (2017). Power analysis for random-effects meta-analysis. \emph{Res. Synth. Methods}, \emph{8}(3), 290--302. \url{https://doi.org/10.1002/jrsm.1240}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Klein2018-he}{}}%
+Klein, R.
+A., Vianello, M., Hasselman, F., Adams, B.
+G., Adams, R.
+B., Jr, Alper, S., \ldots{} Nosek, B.
+A.
+(2018).
+Many labs 2: Investigating variation in replicability across samples and settings.
+\emph{Adv.
+Methods Pract.
+Psychol.
+Sci.}, \emph{1}(4), 443--490.
+\url{https://doi.org/10.1177/2515245918810225}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Knapp2003-qx}{}}%
+Knapp, G., \& Hartung, J. (2003). Improved tests for a random effects meta-regression with a single covariate. \emph{Stat. Med.}, \emph{22}(17), 2693--2710. \url{https://doi.org/10.1002/sim.1482}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Laird1990-ro}{}}%
+Laird, N.
+M., \& Mosteller, F.
+(1990).
+Some statistical methods for combining experimental results.
+\emph{Int.
+J.
+Technol.
+Assess.
+Health Care}, \emph{6}(1), 5--30.
+\url{https://doi.org/10.1017/s0266462300008916}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Rubio-Aparicio2018-pd}{}}%
+Rubio-Aparicio, M., López-López, J.
+A., Sánchez-Meca, J., Marín-Martínez, F., Viechtbauer, W., \& Van den Noortgate, W.
+(2018).
+Estimation of an overall standardized mean difference in random‐effects meta‐analysis if the distribution of random effects departs from normal.
+\emph{Res.
+Synth.
+Methods}, \emph{9}(3), 489--503.
+\url{https://doi.org/10.1002/jrsm.1312}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Schad2020-ht}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Takkouche1999-sv}{}}%
+Takkouche, B., Cadarso-Suárez, C., \& Spiegelman, D.
+(1999).
+Evaluation of old and new tests of heterogeneity in epidemiologic meta-analysis.
+\emph{Am.
+J.
+Epidemiol.}, \emph{150}(2), 206--215.
+\url{https://doi.org/10.1093/oxfordjournals.aje.a009981}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Takkouche2013-nd}{}}%
+Takkouche, B., Khudyakov, P., Costa-Bouzas, J., \& Spiegelman, D.
+(2013).
+Confidence intervals for heterogeneity measures in meta-analysis.
+\emph{Am.
+J.
+Epidemiol.}, \emph{178}(6), 993--1004.
+\url{https://doi.org/10.1093/aje/kwt060}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Van_Erp2017-cf}{}}%
+Van Erp, S., Verhagen, J., Grasman, R.
+P.
+P.
+P., \& Wagenmakers, E.-J.
+(2017).
+Estimates of between-study heterogeneity for 705 meta-analyses reported in \emph{psychological bulletin} from 1990--2013.
+\emph{Journal of Open Psychology Data}, \emph{5}(1), 4.
+\url{https://doi.org/10.5334/jopd.33}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2005-yh}{}}%
+Viechtbauer, W.
+(2005).
+Bias and efficiency of meta-analytic variance estimators in the random-effects model.
+\emph{J.
+Educ.
+Behav.
+Stat.}, \emph{30}(3), 261--293.
+\url{https://doi.org/10.3102/10769986030003261}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2007-tq}{}}%
+Viechtbauer, W.
+(2007).
+Approximate confidence intervals for standardized effect sizes in the two-independent and two-dependent samples design.
+\emph{J.
+Educ.
+Behav.
+Stat.}, \emph{32}(1), 39--60.
+\url{https://doi.org/10.3102/1076998606298034}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2010-xz}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2022-rz}{}}%
+Viechtbauer, W., \& López-López, J. A. (2022). Location-scale models for meta-analysis. \emph{Res. Synth. Methods}, \emph{13}(6), 697--715. \url{https://doi.org/10.1002/jrsm.1562}
+
+\leavevmode\vadjust pre{\hypertarget{ref-R-tidyverse}{}}%
+Wickham, H. (2023). \emph{Tidyverse: Easily install and load the tidyverse}. Retrieved from \url{https://CRAN.R-project.org/package=tidyverse}
+
+\leavevmode\vadjust pre{\hypertarget{ref-R-ggplot2}{}}%
+Wickham, H., Chang, W., Henry, L., Pedersen, T.
+L., Takahashi, K., Wilke, C., \ldots{} Dunnington, D.
+(2023).
+\emph{ggplot2: Create elegant data visualisations using the grammar of graphics}.
+Retrieved from \url{https://CRAN.R-project.org/package=ggplot2}
+
+\leavevmode\vadjust pre{\hypertarget{ref-R-kableExtra}{}}%
+Zhu, H. (2021). \emph{kableExtra: Construct complex table with kable and pipe syntax}. Retrieved from \url{https://CRAN.R-project.org/package=kableExtra}
+
 \end{CSLReferences}

### 2023-07-20 scripts/r-img.R
cleaning fig and res to obs

@@ -154,7 +154,7 @@ equal_vs_random <- cowplot::plot_grid(p_equal, p_random, ncol = 2)
 binary_metareg <- data.frame(
   id = 1:6,
   yi = c(0.1, -0.1, -0.1, 0.5, 0.7, 0.8),
-  vi = c(0.01, 0.02, 0.01, 0.01, 0.02, 0.03),
+  vi = c(0.001, 0.002, 0.001, 0.001, 0.002, 0.003),
   cond = rep(c("a", "b"), each = 3)
 )
 
@@ -178,16 +178,16 @@ binary_metareg$obs <- rnorm(nrow(binary_metareg), binary_metareg$yi, sqrt(binary
 
 plot_metareg_bin <- ggplot(binary_metareg,
                            aes(y = id)) +
-  geom_segment(aes(x = gm, xend = yi, y = id-0.2, yend = id-0.2),
+  geom_segment(aes(x = gm, xend = obs, y = id-0.2, yend = id-0.2),
                color = "darkgreen", size = 1) +
-  geom_segment(aes(x = cond_mean, xend = yi, y = id-0.2, yend = id-0.2),
+  geom_segment(aes(x = cond_mean, xend = obs, y = id-0.2, yend = id-0.2),
                color = "firebrick", size = 1) +
-  geom_point(aes(x = yi, y = id-0.2), size = 3) +
   geom_vline(xintercept = mm) +
   geom_vline(xintercept = mean(binary_metareg$cond_mean), linetype = "dashed") +
   stat_dist_halfeye(aes(dist = dist),
                     show.legend = FALSE,
-                    .width = 0.95) +
+                    .width = 0.95,
+                    scale = 1) +
   theme(axis.ticks.x = element_blank()) +
   ylab("Study") +
   geom_label(aes(x = cond_mean, y = 7, label = note), 
@@ -213,36 +213,42 @@ x <- seq(1, 6, 1)
 
 cont_metareg <- data.frame(
   id = 1:6,
-  #yi = 0.2 + x*0.1 + rnorm(6, 0, 0.15),
-  yi = c(0.2656016, 0.5375599, 0.6202332, 0.3823055, 0.5343109, 0.9233668),
+  yi = c(0.2656016, 0.5375599, 0.4202332, 0.3823055, 0.5343109, 0.9233668),
   vi = c(0.001, 0.002, 0.001, 0.001, 0.002, 0.003),
   x
 )
+
[…]

### 2023-07-24 documents/supplementary/supplementary.Rmd
update

@@
   - \usepackage{xcolor}
+  - \usepackage{amsmath}
   - \newcommand{\rev}[1]{\textcolor{red}{#1}}
@@
 In the main paper, we presented the most straightforward situation for a meta-analysis when each included study contributes with a single effect size.
-However, in real-world meta-analyses, it is common to find complex data structures such as multilevel and multivariate.
-There are numerous sources of dependency for a meta-analysis [@Cheung2014-fg; @Cheung2015-bw, pp.
+However, in real-world meta-analyses, it is common to find complex data structures such as multilevel and multivariate creating dependency between effect sizes [@Cheung2014-fg; @Cheung2015-bw, pp.
 121-122].
 
-The first type of dependency (*multilevel*) consists of multiple effect sizes from independent participants.
+A first type of dependency (*multilevel*) consists of multiple effect sizes from independent participants.
 Despite being collected with different participants, effect sizes from the same study could be correlated.
@@
 Figure \@ref(fig:img-multilevel) depict the general idea of a *multivariate* model.
-There are multiple ways to model this data structure [see @Cheung2014-fg for an overview].
-Still, the most common way is to use a *multilevel meta-analysis model,* considering the heterogeneity between and within studies.
+There are multiple ways to model this situation [see @Cheung2014-fg for an overview].
+Still, the most common way is to use a *multilevel meta-analysis model,* estimating the heterogeneity between and within studies.
 The standard *random-effects* (but also the *equal-effects model*) model is also called a *two-level* model because it combines data after aggregating the participants-level (*first-level*) data.
@@
 127-130].
-Figures \@ref(fig:img-multilevel) and \@ref(fig:img-multivariate) depict the conceptual idea of *multilevel* and *multivariate* models.
-The following sections will show two examples to simulate these complex data structures.
+The following sections will show two examples of data simulation with these complex data structures.
 
@@
 
-# Simulations
+As for the main paper, we started by loading relevant packages and setting the `seed`:
 
@@
 
-In the paper, we used the `mapply()` function, a particularly useful tool in R, to apply the same function (in parallel) with multiple parameters using multiple vectors.
-In our case, the `sim_study()` function has several parameters (`es`, `nt`, among others), and we can apply the function multiple times with a combination of parameters.
-It can be considered essentially a `for` loop but more compact and less verbose.
-
-The following code illustrates the idea of `mapply()` compared to a standard `for` loop.
-The sim data set contains our pool of $k$ studies, and we want to apply the `sim_study()` function to each dataset row.
-We can use a `for` loop that is verbose but more explicit or the `mapply()` function that is more compact but less explicit.
-The computation time and results `for` the two approaches are equivalent, given the appropriate setup.
-
-```{r mapply-vs-for, eval = FALSE}
-# using a for loop
-
-# preallocate for speed (not necessary with mapply)
-res <- vector(mode = "list", length = nrow(sim))
-for(i in 1:length(res)){
-  res[[i]] <- sim_study(theta = sim$es[i], nc = sim$nc[i], nt = sim$nt[i])
-}
-
-# using mapply
-sim <- mapply(sim_study, sim$es, sim$nc, sim$nt)
-
-# using sim_studies (with mapply inside)
-sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
-```
+\newpage
 
@@
 Compared to the standard two-level model, the three-level model estimates another source of heterogeneity ($\omega^{2}$) representing the variability within studies.
-We can easily extend the two-level *random-effects* equation into equation \@ref(eq:three-level-model) from the main paper adding another "adjustment" to the overall effect.
+We can easily extend the two-level *random-effects* equation into equation \@ref(eq:three-level-model1) from the main paper adding another "adjustment" to the overall effect.
 
@@
 
-The notation suggests that each effect size belongs to an experiment (effect size) ($j$) nested within a study ($i$).
+The notation suggests that each effect size belongs to an experiment ($j$) nested within a study ($i$).
 The variability between studies (level 3) is handled by the $\tau^{2}$ parameter (see Equation \@ref(eq:three-level-model2) and the variability within studies (level 2) is handled by the $\omega^{2}$ parameter (see Equation \@ref(eq:three-level-model3)).
@@
 n <- 30 # sample size for each group, within each study
-mu <- 0.3 # real average effect size
+mu_theta <- 0.3 # real average effect size
 
@@
   nc = n,
-  mu = mu
+  mu_theta = mu_theta
 )
@@
 set.seed(seed)
-sim$es <- sim$mu + sim$delta_i + sim$zeta_ij
+sim$es <- sim$mu_theta + sim$delta_i + sim$zeta_ij
 sim <- sim_studies(es = sim$es, nt = sim$nt, nc = sim$nc, data = sim)
@@
 Finally we fit a *three-level models* using the `rma.mv` from the `metafor` package.
-The syntax differs slightly from the rma function and is clearly explained here https://www.metafor-project.org/doku.php/analyses:konstantopoulos2011.
-The essential part is adding the nested random effect using the `random = ` argument specifying that the `study` variable is nested within the `study` variable with `~ 1|study/experiment`.
+The syntax differs slightly from the `rma` function and is clearly explained on the `metafor` documentation ( [https://www.metafor-project.org/doku.php/analyses:konstantopoulos2011](https://www.metafor-project.org/doku.php/analyses:konstantopoulos2011)).
+The essential part is adding the nested random effect using the `random = ` argument specifying that the `study` variable is nested within the `study` variable with `~ 1|study/effect`.
 
@@
 For this reason we have three real effects ($\mu_{\theta_{1}}$, $\mu_{\theta_{2}}$ and $\mu_{\theta_{3}}$), three heterogeneity ($\tau^2_1$, $\tau^2_2$ and $\tau^2_3$) and the population-level correlations between these outcomes ($\rho_{12}$, $\rho_{13}$ and $\rho_{23}$).
-Moreover, the sampling errors ($\epsilon_i$) are now correlated; thus each study will have three sampling variances for each outcome ($\sigma_{\epsilon_1}^{2}$, $\sigma_{\epsilon_1}^{2}$, and $\sigma_{\epsilon_1}^{2}$), and the correlations between sampling errors ($\rho_{s12}$, $\rho_{s13}$ and $\rho_{s23}$)^[We used the subscript $s$ for sampling errors to distinguish between between-outcomes correlations and correlations between sampling errors].
+Moreover, the sampling errors ($\epsilon_i$) are now correlated; thus each study will have three sampling variances for each outcome ($\sigma_{\epsilon_1}^{2}$, $\sigma_{\epsilon_2}^{2}$, and $\sigma_{\epsilon_3}^{2}$), and the correlations between sampling errors ($\rho_{s12}$, $\rho_{s13}$ and $\rho_{s23}$)^[We used the subscript $s$ for sampling errors to distinguish between between-outcomes correlations and correlations between sampling errors].
 Equations \@ref(eq:multivariate-model1), \@ref(eq:multivariate-model2), \@ref(eq:multivariate-model3), \@ref(eq:multivariate-model4) \@ref(eq:multivariate-model5) formalize the multivariate model for a single study $i$ with three outcomes.
-The random-effects adjustments ($\theta_i$) to the overall effect and sampling errors are now sampled from a multivariate normal distribution, respectively $\theta_i \sim \mathcal{MVN}(0, \mathrm{T}^2)$ and $\theta_i \sim \mathcal{MVN}(0, \mathrm{V}^2)$.
+The random-effects adjustments ($\delta_i$) to the overall effect and sampling errors ($\epsilon_i$) are now sampled from multivariate normal distributions (see Equations \@ref(eq:multivariate-model2), and \@ref(eq:multivariate-model3))
 
@@
 
-The previous code simulated a *compound symmetry* (CS) structure where all variances ($\tau^2$) and covariances ($\rho\tau_1\tau_2$) are the same.
+The previous code simulated a *compound symmetry* (CS) structure where all variances ($\tau^2$) and covariances ($\rho$) are the same.
 There could be different structures of the $T$ matrix that are clearly explained in the `metafor::rma.mv()` documentation (see https://wviechtb.github.io/metafor/reference/rma.mv.html#specifying-random-effects).
@@
 
-For the sampling errors, the matrix $V^{2}$ for the study, $i$ is essentially a $p_{i} \times p_{i}$ matrix, where the diagonal contains the sampling variances, and off-diagonal elements are the covariances.
-Crucially, sampling errors are correlated within a study but uncorrelated between studies.
-For this reason, stacking all matrices
-from $k$ studies will create a block variance-covariance matrix where the correlations between sampling errors of different studies are fixed to zero.
+For the sampling errors, the matrix $V$ is block variance-covariance matrix, where each study is represented by a $p_{i} \times p_{i}$ matrix.
+Diagonal elements are the sampling variances of different outcomes and off-diagonal elements are the covariances.
+The full $V$ matrix has the covariance fixed to zero for effects coming from different studies because sampling errors are not correlated.
 Figure \@ref(fig:fig-block-vcov) depicts the idea of the block-variance covariance matrix for three studies with a different number of outcomes.
@@
 We have all pieces to simulate the multivariate model.
-We can use the `sim_study_m()` that follows the same idea of the `sim_study()` function but extends to multiple outcomes.
+We can use the `sim_study_m()` that follows the same idea of the `sim_study()` function including multiple outcomes.
 We generate a single study with $p$ outcomes where sampling errors are correlated according to a compound symmetry structure^[Also for the $V$ matrix we can assume different variance-covariance structures.
@@
 In practical terms, we assume that the
-correlation at the participant level between multiple outcomes is $r$ and fixed between studies.
+correlation at the participant level between multiple outcomes is $\rho_s$ and is the same across different studies.
 
@@
 
-As in the main paper, we can create a dataframe with simulation parameters and iterating the `sim_study_m()` producing a meta-analysis dataframe.
+As in the main paper, we can create a dataframe with simulation parameters and we use the `sim_study_m()` producing the meta-analysis dataframe.
 
@@
 rho <- 0.7 # population level correlation between outcomes
-r <- 0.5 # correlation between sampling errors
+rho_s <- 0.5 # correlation between sampling errors
 n <- 30 # number of participants per group per study
@@
 # adding the real effects and the random-effect adjustments
-sim$mu <- rep(mus, k)
+sim$mu_theta <- rep(mus, k)
 sim$delta_i <- sim_T(k, p, taus2, rho)
@@
 # average effect + random-effect adjustment
-sim$es <- sim$mu + sim$delta_i
+sim$es <- sim$mu_theta + sim$delta_i
 
@@
 # simulating the effect sizes for each study
-resl <- lapply(siml, function(x) sim_study_m(x$es, x$nt, x$nc, r))
+resl <- lapply(siml, function(x) sim_study_m(x$es, x$nt, x$nc, rho_s))
 res <- dplyr::bind_rows(resl) # combine everything
@@
 Given that we simulated the raw data, we can calculate and use the observed correlations (within the `sim_study_m()` function) and use a different correlation for each study within the `vcalc()` function (see [https://www.metafor-project.org/doku.php/analyses:berkey1998](https://www.metafor-project.org/doku.php/analyses:berkey1998https://www.metafor-project.org/doku.php/analyses:berkey1998) for an example).
-Often, these values are not reported; thus, we need to guess a plausible correlation to compute the block variance-covariance matrix.
-Here we use the $r$ value from our simulation and use the same for each study.
-We know the underlying model in this case, so our guessed value is correct.
+Often, these values are not reported; thus, we need to guess a plausible correlation to approximate the block variance-covariance matrix.
+Here we use the $\rho_s$ value from our simulation and use the same for each study.
+We know the underlying model in this case, so our guessed value is appropriate.
 
@@
 # r is the fixed sampling errors correlation
-V <- vcalc(vi = vi, cluster = id, obs = p, rho = r, data = sim)
+V <- vcalc(vi = vi, cluster = id, obs = p, rho = rho_s, data = sim)
 
@@
 The output is more complicated compared to previous models.
-The `Variance Components` is the estimation of the $T^{2}$ matrix.
+The `Variance Components` is the estimation of the $T$ matrix.
 In the two-level random-effects model, we have a single $\tau^{2}$ while in the multivariate model, we have a full variance-covariance matrix.
@@
 The crucial part is the `random = ~ p|id` argument that specifies the random-effects structure.
-The syntax is clearly explained here [https://wviechtb.github.io/metafor/reference/rma.mv.html#specifying-random-effects](https://wviechtb.github.io/metafor/reference/rma.mv.html#specifying-random-effects) The basic idea is to specify the random effect structure as `~ inner|outer` where inner is the outcome variable and outer is the paper variable.
-The mods = `~ 0 + p` argument specify to estimate the average effect for each outcome (i.e., the $\mathbf{\mu_{\theta}}$ vector of average effects).
+The syntax is clearly explained in the `metafor` documentation [https://wviechtb.github.io/metafor/reference/rma.mv.html#specifying-random-effects](https://wviechtb.github.io/metafor/reference/rma.mv.html#specifying-random-effects).
+The basic idea is to specify the random effects structure as `~ inner|outer` where inner is the outcome variable and outer is the paper variable.
+The mods = `~ 0 + p` argument specify to estimate the average effect for each outcome (i.e., $\mu_{\theta}$ vector of average effects).
 We are removing the intercept (`~ 0 + p` or equivalently `~ p - 1`) and the model will estimate the mean for each level of the `p` variable.
@@
 Setting the `struct = "UN"` (unstructured) argument, `rma.mv` try to estimate a variance-covariance matrix where all correlations and variances need to be estimated.
-Compared to the previous model, we are know estimating 3 heterogeneity and three correlation parameters.
-
-We could also use another structure for the $T^{2}$ matrix.
-We know that our data are generated using a CS structure; thus, using another structure will be less appropriate.
-In reality, it is not known which is the real data generation model; thus, we can try different structures.
-Setting the struct = "UN" (unstructured) argument, rma.mv tries to estimate a variance-covariance matrix where all correlations and variances need to be estimated.
-We are now estimating three
-heterogeneity and three correlation parameters compared to the previous model.
+Compared to the previous model, we are know estimating three heterogeneity parameters and three correlation parameters.
 
@@
 
+## Simulating using `mapply()`
+
+In the paper, we used the `mapply()` function, a particularly useful tool in R, to apply the same function (in parallel) with multiple parameters using multiple vectors.
+In our case, the `sim_study()` function has several parameters (`es`, `nt`, among others), and we can apply the function multiple times with a combination of parameters.
+It can be considered essentially a `for` loop but more compact and less verbose.
+
+The following code illustrates the idea of `mapply()` compared to a standard `for` loop.
+The sim data set contains our pool of $k$ studies, and we want to apply the `sim_study()` function to each dataset row.
+We can use a `for` loop that is verbose but more explicit or the `mapply()` function that is more compact but less explicit.
+The computation time and results `for` the two approaches are equivalent, given the appropriate setup.
+
+```{r mapply-vs-for, eval = FALSE}
+# using a for loop
+
+# preallocate for speed (not necessary with mapply)
+res <- vector(mode = "list", length = nrow(sim))
+for(i in 1:length(res)){
+  res[[i]] <- sim_study(theta = sim$es[i], nc = sim$nc[i], nt = sim$nt[i])
+}
+
+# using mapply
+sim <- mapply(sim_study, sim$es, sim$nc, sim$nt)
+
+# using sim_studies (with mapply inside)
+sim <- sim_studies(sim$es, sim$nc, sim$nt, data = sim)
+```
+
 ## $I^2$ with homogeneous vs heterogeneous sample sizes
@@
 For example, if the sample sizes of our $k$ studies is generate from a Poisson distribution $n_i \sim \mathcal{P}(\lambda)$, then $\tilde{v}$ can be calculated using $\lambda$ as *typical* sample size (i.e., as if the sample sizes were homogeneous).
-The following code simulate this scenario showing the distribution of $I^2$ with heterogeneous sample sizes^[Compared to the simulations of the paper here we are simulating the effect size and the sampling variances from the respective sampling distributions [e.g., supplementary materials by @Van_Aert2019-xt]].
+The following code simulate this scenario showing the distribution of $I^2$ with heterogeneous sample sizes^[Compared to the simulations of the paper here we are simulating the effect size and the sampling variances from the respective sampling distributions.
+In other terms we are not simulating participant-level data but aggregated data [e.g., supplementary materials by @Van_Aert2019-xt]].
 
@@
 nsim <- 1000 # number of simulations
-mu <- 0.3 # real average effect size
+mu_theta <- 0.3 # real average effect size
 
@@
   # effect sizes
-  yi <- rnorm(k, rnorm(k, mu, sqrt(tau2)), sqrt(1/ni + 1/ni))
+  yi <- rnorm(k, rnorm(k, mu_theta, sqrt(tau2)), sqrt(1/ni + 1/ni))
 
@@
 (ref:img-multilevel) Graphical representation of a *multilevel* (three-level) model.
-Each observed effect size is defined as the overall effect average effect $\mu_{\theta}$, a study-specific random-effect $\delta_{i} \sim N(0, \tau^2)$ and a experiment-specific random-effects $\zeta_{ij} N(0, \omega^2)$.
+Each observed effect size is defined as the overall effect average effect $\mu_{\theta}$, a study-specific random-effect $\delta_{i} \sim N(0, \tau^2)$ and a experiment-specific random-effects $\zeta_{ij} \sim N(0, \omega^2)$.
 

### 2023-07-24 documents/paper/paper.Rmd
update

@@
 keywords          : "meta-analysis, monte carlo simulations, power analysis"
-wordcount         : "7877"
+wordcount         : "7782"
 bibliography      : "../files/references.bib"
@@
 155-176].
-A recent paper by DeBruine and Barr [-@DeBruine2021-id], which deeply inspired the current work, proposed a stimulating way to understand linear mixed effects via data simulation.
+A recent paper by DeBruine and Barr [-@DeBruine2021-id], which deeply inspired the current work, proposed a stimulating way to understand linear \rev{mixed-effects models} via data simulation.
 Simulating data is also a powerful educational tool within this framework.
-For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for \rev{equal} and random-effects, and meta-regression models with applications also for statistical power calculation.
+For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for \rev{equal-effects}, random-effects, and meta-regression models with applications also to statistical power calculation.
 In the first section, we will introduce basic concepts of the meta-analysis that are useful for setting up the simulation.
@@
 random effects model distinction.
-Then we will describe how to simulate data for these models and simulate a meta-regression model with a categorical and numerical predictor.
+Then we will describe how to simulate data for these models and simulate a meta-regression with categorical and numerical predictors.
 Finally, we will introduce the power analysis extending the previous examples to estimate the statistical power.
@@
 This common measure should give an immediate idea of the direction (i.e., the treatment improves or reduces performance) and the size of the effect.
-Standardized effect size measures [see @Lakens2013-eq for an overview], such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's $d$ [@Cohen1988-ad] or the Pearson correlation coefficient $\rho$ which could be calculated using the associated sample estimator *r* are commonly used to compare heterogeneous outcome variables^[\color{red} When using bounded effect size indexes such as the Pearson correlation, Fisher's *z* transformation is commonly used [@Borenstein2009-mo, pp.
+Standardized effect size measures [see @Lakens2013-eq for an overview], such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's $d$ [@Cohen1988-ad] or the Pearson correlation coefficient $\rho$ which could be calculated using the associated sample estimator *r* are commonly used to compare heterogeneous outcome variables^[\color{red} When using bounded effect sizes such as the Pearson correlation, Fisher Z-transformation is commonly used [@Borenstein2009-mo, pp.
 41-43] \color{black}].
@@
 The core of a meta-analysis is combining the results of multiple studies giving more weight to studies that provide a more precise effect estimation.
-We can find essentially three meta-analysis models: the equal-effects, the fixed-effects and the random-effects model [@Laird1990-ro; @Hedges1998-uo].
-The *equal-effects* model assume that each study included in the meta-analysis is a more or less precise estimation of the true underlying effect ($\theta$).
+We can find essentially three meta-analysis models: the equal-effects, the fixed-effects, and the random-effects model [@Laird1990-ro; @Hedges1998-uo].
+The *equal-effects* model assumes that each study included in the meta-analysis is a more or less precise estimation of the true underlying effect ($\theta$).
 In other terms, we are assuming that there is no variability (i.e., heterogeneity) among true effect sizes.
 On the other side, suppose some study-level characteristics (e.g., participants' age, sex, or socioeconomic status) or the experimental paradigm (e.g., type of memory task or difficulty) could impact the treatment effect.
-In this case we have true variability (i.e., heterogeneity) among studies.
-The *random-effects* model assume a distribution of real effects with mean $\mu_{\theta}$ and variance $\tau^{2}$ thus estimating the heterogeneity.
-Finally, the *fixed-effects* estimate the average effect of the included pool of studies ignoring the presence of heterogeneity.
+In this case, we have true variability (i.e., heterogeneity) among studies.
+The *random-effects* model assumes a distribution of real effects with mean $\mu_{\theta}$ and variance $\tau^{2}$ thus estimating the heterogeneity.
+Finally, the *fixed-effects* model estimate the average effect of the included pool of studies ignoring the presence of heterogeneity.
 While the distinction between the equal and fixed-effects model is theoretical, the estimated model is the exactly the same^[See [wviechtb.github.io/metafor/reference/misc-models.html](https://wviechtb.github.io/metafor/reference/misc-models.html) for a detailed explanation].
-The *random-effects* model assume and estimate heterogeneity among effect sizes leading to different model parameters.
-From an inferential point of view, the *random-effects* model provide unconditional inference to the population of effect sizes while the *fixed-effects* model estimate the average effect of the selected studies, and not population-level parameters.
-The *equal-effects* model assume a single true population effect to be estimated.
-The Figure \@ref(fig:img-equal-vs-random) depicts the theoretical distinction between the equal and random-effects model.
-In the presence of heterogeneity, we can include moderators (e.g., the type of experimental setup) into a meta-regression model to explain the effect sizes heterogeneity.
+The *random-effects* model assumes and estimates heterogeneity among effect sizes leading to different model parameters.
+From an inferential point of view, the *random-effects* model provides unconditional inference to the population of effect sizes while the *fixed-effects* model estimates the average effect of the selected studies and not population-level parameters.
+The *equal-effects* model assumes a single true population effect to be estimated.
+Figure \@ref(fig:img-equal-vs-random) depicts the theoretical distinction between the equal and random-effects model.
+In the presence of heterogeneity, we can include moderators (e.g., the type of experimental setup) into a meta-regression model to explain the effect size heterogeneity.
 \color{red} An interesting proposal to combine evidence from these different models into a single analysis is called Bayesian model-averaged meta-analysis [@Gronau2021-do; @Berkhout2023-fj].
+However, this model is beyond the scope of the tutorial.
 \color{black}
@@
 \color{red}
-This tutorial use several R packages for the simulations, meta-analysis fitting and figures/tables.
-For the data manipulation we used the *tidyverse* [@R-tidyverse] package.
+This tutorial uses several R packages for the simulations, meta-analysis fitting, and figures/tables.
+For the data manipulation, we used the *tidyverse* [@R-tidyverse] package.
 For the models fitting we used the *metafor* [@Viechtbauer2010-xz] package.
@@
 library(metafor) # for meta-analysis models fitting
-library(purrr) # for *apply like functions
+library(purrr) # for *apply like functions (within tidyverse)
 seed <- 2023 # general seed for all simulations
@@
 the Knapp and Hartung [-@Knapp2003-qx] method by setting `test="knha"` within the `rma()` function.
-This method is useful especially when the number of studies is low (see https://wviechtb.github.io/metafor/reference/misc-recs.html for more details) \color{black}].
-Table \@ref(tab:notation-tab) summarizes the simulation parameters notation in equations and code.
+This method is useful especially when the number of studies is low (see [wviechtb.github.io/metafor/reference/misc-recs.html](https://wviechtb.github.io/metafor/reference/misc-recs.html) for more details) \color{black}].
+Table \@ref(tab:notation-tab) \rev{summarizes the notation used in simulations and code in equations and code.}
 
@@
 Compared to the *equal-effects model* we need to generate another adjustment to the overall effect $\mu_{\theta}$ from a normal distribution with mean 0 and variance $\tau^{2}$.
-The real effect size for the study $i$ will be $\mu_{\theta} + \theta_{i}$ where $\theta_{i}$ are the random-effects regulated by the $\tau^{2}$ parameter.
+The real effect size for the study $i$ will be $\mu_{\theta} + \delta_{i}$ where $\delta_{i}$ are the random-effects regulated by the $\tau^{2}$ parameter.
 
@@
 As the number of studies $k$ increases, the estimation of $\mu_{\theta}$ and $\tau^{2}$ becomes more precise \color{red} [@Blazquez-Rincon2023-kp; @Rubio-Aparicio2018-pd] \color{black}.
-Additionally, as the sample size of each study decreases, each $\theta_{i}$ will be estimated with higher precision but as long as $\tau^{2} \neq 0$ there will be variability among effect sizes.
+Additionally, as the sample size of each study decreases, each $\delta_{i}$ will be estimated with higher precision but as long as $\tau^{2} \neq 0$ there will be variability among effect sizes.
 In other words, increasing the sample size of each study or the number of studies will not affect the value of $\tau^{2}$ but only the estimation precision [see @Borenstein2009-mo, Chapter 16, Figure 16.6 for a clear explanation].
@@
 \begin{align}
-\begin{gathered}Random-effects model true effect
+\begin{gathered}
 \tilde{v} = \frac{(k-1) \sum w_i}{(\sum w_i)^2 - \sum w_i^2}
@@
 
-From Equation \@ref(eq:isquared) and Figure \@ref(fig:random-effects-high-low-precision) is clear that if each included study has a considerable sample size the total variability will be reduced, mainly driven by real heterogeneity ($\tau^2$).
+From Equation \@ref(eq:isquared) and Figure \@ref(fig:random-effects-high-low-precision) is clear that if each included study has a considerable sample size \color{red} the sampling variability ($\tilde{v}$) will be reduced, and the total variability will be mainly driven by real heterogeneity ($\tau^2$) \color{black}.
 This is the crucial difference between the \rev{equal-effects} and the random-effects model [see also @Borenstein2009-mo, pp.
@@
 The only caveat is fixing the $\tilde{v}$.
-When the sample size of each study is the same^[The homogeneous sample size of primary studies assumption is commonly used to calculate the statistical power [see @Borenstein2009-mo, Chapter 29]], $\tilde{v} = \sigma_{\epsilon_i}^{2}$.
+When the sample size of each study is the same^[\color{red} The assumption of studies with homogeneous sample sizes is commonly used to calculate the statistical power \color{black}[see @Borenstein2009-mo, Chapter 29]], $\tilde{v} = \sigma_{\epsilon_i}^{2}$.
 In the other case, $\tilde{v}$ needs to be calculated from sampling variances as reported in Higgins and Thompson [-@Higgins2002-fh] and cannot be easily fixed a priori.
@@
 In regression terms (see equation \@ref(eq:meta-regression)), we could use a dummy variable ($X_{1}$) that takes the value of 0 for the *lab-based task* (*L*) and a value of 1 for an *online task* (*O*)^[The *dummy coding* (also known as treatment coding) is the default in R and `metafor` but other coding schemes could be used [see @Schad2020-ht for an overview of contrast coding schemes].
-\color{red} As in standard linear regression, the contrast coding of categorical factors or the presence/absence of the intercept influence the interpretation of estimated model parameters and omnibus meta-regression tests (see https://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept for a detailed discussion)] \color{black}.
+\color{red} As in standard linear regression, the contrast coding of categorical factors or the presence/absence of the intercept influences the interpretation of estimated model parameters and omnibus meta-regression tests (see https://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept for a detailed discussion)] \color{black}.
 Now we fix $\beta_{1}$ to be the *lab vs.
@@
 
-Now we can fit the meta-regression model with the `rma` function as for the random-effects model with the addition of `mods = ~ exp` that indicates which variable/s to consider as moderators.
+Now we can fit the meta-regression model with the `rma` function as for the random-effects model with the addition of `mods = ~ exp` that indicates which variables to consider as moderators.
 The results are presented in Table \@ref(tab:res-meta-reg-dummy).
@@
 However, one of the most critical applications is estimating the power of a specific statistical model.
-\color{red} One of the purpose of power analysis by simulations is estimating the required number of studies to detect an hypothetical effect size when planning a meta-analysis.
+\color{red} One of the purposes of power analysis by simulations is to estimate the required number of studies to detect an hypothetical effect size when planning a meta-analysis.
 At the same time, the meta-analysis that we presented (*two-level* equal or random-effects model) can be considered as a multi-lab [e.g., @Klein2018-he] study where experiments are planned and not collected from the literature.
-In fact, we can use the same simulation approach to optimize the number of participants and studies when planning a multi-lab project.
+We can use the same simulation approach to optimize the number of participants and studies when planning a multi-lab project.
 \color{black}
@@
 Jackson and colleagues [-@Jackson2017-dv] partially solve the issue by developing an interesting method that takes into account the uncertainty in estimating $\tau^{2}$ without using simulations.
-However, only Monte Carlo simulations can \rev{handle} complex scenarios and extra parameters.
+\rev{However, complex simulation scenarios can be handled only using Monte Carlo methods.}
 
@@
 
-The present work introduced the basic concepts of the meta-analysis regarding equal, random-effects models and meta-regression with a simulation-based approach.
-We believe the presented examples can help implement alternative or more complex models.
+The present work introduced the basic concepts of the meta-analysis regarding equal-effects, random-effects and meta-regression models with a simulation-based approach.
+\rev{We believe the presented examples are useful to implement alternative or more complex models}.
 For example, the `sim_study()` function can be easily modified to simulate another effect sizes index, \color{red} such as correlations or odds ratios.
-\color{black} In addition, more complex models, such as *multivariate* or *multilevel* models, can be implemented following the same approach (see \rev{the} supplementary materials).
-For example, the \color{red} *multilevel* (e.g., *three-level*) \color{black} model estimates another heterogeneity component representing the variability of multiple \rev{independent} effect sizes \rev{within} the same study.
-Similarly, the *multivariate* model could include the correlation between multiple outcomes and the correlation between sampling errors.
+\color{black} In addition, more complex models, such as *multivariate* or *multilevel* models, can be simulated following a similar approach (see \rev{the} supplementary materials).
+The \color{red} *multilevel* (e.g., *three-level*) \color{black} model estimates another heterogeneity component representing the variability of multiple \rev{independent} effect sizes \rev{within} the same study.
+Similarly, the *multivariate* model \rev{includes} the correlation between multiple outcomes and the correlation between sampling errors.
 
 The present work did have a few limitations.
-First, we only introduced basic concepts about meta-analysis and Monte Carlo simulations, while setting up complex simulations requires more knowledge and complexity of the simulation setup.
+\rev{Firstly}, we only introduced basic concepts about meta-analysis and Monte Carlo simulations, while setting up complex simulations requires more knowledge and complexity of the simulation setup.
 We decided to give the foundations to understand meta-analyses with a simulation approach because more complex models are still based on the same principles.
 Second, there are limitations concerning simulating participant-level data.
-We decided to simulate the meta-analysis dataset starting from the participants' level to maximize the flexibility and clearness of each step.
+We decided to simulate the meta-analysis data starting from the participants' level to maximize the flexibility and clearness of each step.
 The downside concerns the efficiency and scalability of the simulation setup.
@@
 
-We thank the [R-sig-meta-analysis](https://stat.ethz.ch/mailman/listinfo/r-sig-meta-analysis) mailing-list where suggestions and clarifications significantly improved our simulations approach and the R code.
+We thank the *[R-sig-meta-analysis](https://stat.ethz.ch/mailman/listinfo/r-sig-meta-analysis)* mailing-list.
+\rev{Their} suggestions and clarifications significantly improved \rev{the} simulations approach and the R code.
 
@@
 
-The code to reproduce simulations, figures and tables can be found on Open Science Framework (https://osf.io/54djn/)
+The code to reproduce simulations, figures and tables can be found on Open Science Framework (https://osf.io/54djn/) \color{red} and Github (https://github.com/shared-research/simulating-meta-analysis).
+\color{black}
 
@@
 
-The supplementary materials can be also found on the Open Science Framework repository (https://osf.io/54djn/)
+The supplementary materials can be also found on the Open Science Framework repository (https://osf.io/54djn/) \color{red} and Github (https://github.com/shared-research/simulating-meta-analysis).
+\color{black}
 
@@
 In practice, each study has a different observed effect size where studies with high precision (i.e., narrow sampling distributions) will be close to the real effect ($\theta$).
-The *random-effects model* on the right suggests that beyond the error term ($\epsilon_{i}$), each real effect size is composed of a fixed part (now $\mu_{\theta}$) and a random part ($\theta_{i}$) sampled from a normal distribution with mean zero and variance $\tau^{2}$.
+The *random-effects model* on the right suggests that beyond the error term ($\epsilon_{i}$), each real effect size is composed of a fixed part (now $\mu_{\theta}$) and a random part ($\delta_{i}$) sampled from a normal distribution with mean zero and variance $\tau^{2}$.
 When $\tau^{2}$ is zero, the random-effects reduces to a equal-effects model.
@@
 On the left, the simulated model has $nt,nc = 30$ for each included study while on the right the sample size for each study is $nt,nc = 500$.
-Given that the data were generated under a *equal-effects model*, when the sample size is high (on the right) each study is aligned on the real effect size because the error component ($\epsilon_{i}$) is almost entirely reduced.
+Given that the data were generated under a *equal-effects model*, when the sample size is high (on the right) each study is aligned on the real effect size because the error component ($\epsilon_{i}$) \rev{is close to zero}.
 The average effect size is similar (depending on the random numbers generation) between the two scenarios while the estimation precision (the width of the black diamond) is narrower on the right.
@@
 (ref:plot-meta-reg-numerical-r2) Meta-regression results for the *random-effects model* with a numerical moderator.
-Each effect size is represented with a black dot where the dimension represents the weight according to the inverse of the precision.
+Each effect size is represented with a black dot where the dimension represents the weight according to the inverse of the variance.
 The line represents the estimated meta-regression slope with the 95% confidence interval (grey bands).

### 2023-07-24 documents/paper/paper.tex
update

@@
 {}{\PatchFailed}
-\keywords{meta-analysis, monte carlo simulations, power analysis\newline\indent Word count: 7877}
+\keywords{meta-analysis, monte carlo simulations, power analysis\newline\indent Word count: 7782}
 \usepackage{csquotes}
@@
 69--76; Ingalls, 2011).
-A recent paper by DeBruine and Barr (2021), which deeply inspired the current work, proposed a stimulating way to understand linear mixed effects via data simulation.
+A recent paper by DeBruine and Barr (2021), which deeply inspired the current work, proposed a stimulating way to understand linear \textcolor{red}{mixed-effects models} via data simulation.
 Simulating data is also a powerful educational tool within this framework.
-For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for \textcolor{red}{equal} and random-effects, and meta-regression models with applications also for statistical power calculation.
+For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for \textcolor{red}{equal-effects}, random-effects, and meta-regression models with applications also to statistical power calculation.
 In the first section, we will introduce basic concepts of the meta-analysis that are useful for setting up the simulation.
 We will evaluate the effect size, variance calculation, and the \textcolor{red}{equal} vs.~random effects model distinction.
-Then we will describe how to simulate data for these models and simulate a meta-regression model with a categorical and numerical predictor.
+Then we will describe how to simulate data for these models and simulate a meta-regression with categorical and numerical predictors.
 Finally, we will introduce the power analysis extending the previous examples to estimate the statistical power.
@@
 This common measure should give an immediate idea of the direction (i.e., the treatment improves or reduces performance) and the size of the effect.
-Standardized effect size measures (see Lakens, 2013 for an overview), such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's \(d\) (Cohen, 1988) or the Pearson correlation coefficient \(\rho\) which could be calculated using the associated sample estimator \emph{r} are commonly used to compare heterogeneous outcome variables\footnote{\color{red} When using bounded effect size indexes such as the Pearson correlation, Fisher's \emph{z} transformation is commonly used (Borenstein et al., 2009, pp.
+Standardized effect size measures (see Lakens, 2013 for an overview), such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's \(d\) (Cohen, 1988) or the Pearson correlation coefficient \(\rho\) which could be calculated using the associated sample estimator \emph{r} are commonly used to compare heterogeneous outcome variables\footnote{\color{red} When using bounded effect sizes such as the Pearson correlation, Fisher Z-transformation is commonly used (Borenstein et al., 2009, pp.
 41--43) \color{black}}.
@@
 The core of a meta-analysis is combining the results of multiple studies giving more weight to studies that provide a more precise effect estimation.
-We can find essentially three meta-analysis models: the equal-effects, the fixed-effects and the random-effects model (Hedges \& Vevea, 1998; Laird \& Mosteller, 1990).
-The \emph{equal-effects} model assume that each study included in the meta-analysis is a more or less precise estimation of the true underlying effect (\(\theta\)).
+We can find essentially three meta-analysis models: the equal-effects, the fixed-effects, and the random-effects model (Hedges \& Vevea, 1998; Laird \& Mosteller, 1990).
+The \emph{equal-effects} model assumes that each study included in the meta-analysis is a more or less precise estimation of the true underlying effect (\(\theta\)).
 In other terms, we are assuming that there is no variability (i.e., heterogeneity) among true effect sizes.
 On the other side, suppose some study-level characteristics (e.g., participants' age, sex, or socioeconomic status) or the experimental paradigm (e.g., type of memory task or difficulty) could impact the treatment effect.
-In this case we have true variability (i.e., heterogeneity) among studies.
-The \emph{random-effects} model assume a distribution of real effects with mean \(\mu_{\theta}\) and variance \(\tau^{2}\) thus estimating the heterogeneity.
-Finally, the \emph{fixed-effects} estimate the average effect of the included pool of studies ignoring the presence of heterogeneity.
+In this case, we have true variability (i.e., heterogeneity) among studies.
+The \emph{random-effects} model assumes a distribution of real effects with mean \(\mu_{\theta}\) and variance \(\tau^{2}\) thus estimating the heterogeneity.
+Finally, the \emph{fixed-effects} model estimate the average effect of the included pool of studies ignoring the presence of heterogeneity.
 While the distinction between the equal and fixed-effects model is theoretical, the estimated model is the exactly the same\footnote{See \href{https://wviechtb.github.io/metafor/reference/misc-models.html}{wviechtb.github.io/metafor/reference/misc-models.html} for a detailed explanation}.
-The \emph{random-effects} model assume and estimate heterogeneity among effect sizes leading to different model parameters.
-From an inferential point of view, the \emph{random-effects} model provide unconditional inference to the population of effect sizes while the \emph{fixed-effects} model estimate the average effect of the selected studies, and not population-level parameters.
-The \emph{equal-effects} model assume a single true population effect to be estimated.
-The Figure \ref{fig:img-equal-vs-random} depicts the theoretical distinction between the equal and random-effects model.
-In the presence of heterogeneity, we can include moderators (e.g., the type of experimental setup) into a meta-regression model to explain the effect sizes heterogeneity.
+The \emph{random-effects} model assumes and estimates heterogeneity among effect sizes leading to different model parameters.
+From an inferential point of view, the \emph{random-effects} model provides unconditional inference to the population of effect sizes while the \emph{fixed-effects} model estimates the average effect of the selected studies and not population-level parameters.
+The \emph{equal-effects} model assumes a single true population effect to be estimated.
+Figure \ref{fig:img-equal-vs-random} depicts the theoretical distinction between the equal and random-effects model.
+In the presence of heterogeneity, we can include moderators (e.g., the type of experimental setup) into a meta-regression model to explain the effect size heterogeneity.
 \color{red} An interesting proposal to combine evidence from these different models into a single analysis is called Bayesian model-averaged meta-analysis (Berkhout, Haaf, Gronau, Heck, \& Wagenmakers, 2023; Gronau, Heck, Berkhout, Haaf, \& Wagenmakers, 2021).
+However, this model is beyond the scope of the tutorial.
 \color{black}
@@
 In practice, each study has a different observed effect size where studies with high precision (i.e., narrow sampling distributions) will be close to the real effect (\(\theta\)).
-The \emph{random-effects model} on the right suggests that beyond the error term (\(\epsilon_{i}\)), each real effect size is composed of a fixed part (now \(\mu_{\theta}\)) and a random part (\(\theta_{i}\)) sampled from a normal distribution with mean zero and variance \(\tau^{2}\).
+The \emph{random-effects model} on the right suggests that beyond the error term (\(\epsilon_{i}\)), each real effect size is composed of a fixed part (now \(\mu_{\theta}\)) and a random part (\(\delta_{i}\)) sampled from a normal distribution with mean zero and variance \(\tau^{2}\).
 When \(\tau^{2}\) is zero, the random-effects reduces to a equal-effects model.
@@
 
-This tutorial use several R packages for the simulations, meta-analysis fitting and figures/tables.
-For the data manipulation we used the \emph{tidyverse} (Wickham, 2023) package.
+This tutorial uses several R packages for the simulations, meta-analysis fitting, and figures/tables.
+For the data manipulation, we used the \emph{tidyverse} (Wickham, 2023) package.
 For the models fitting we used the \emph{metafor} (Viechtbauer, 2010) package.
@@
 \FunctionTok{library}\NormalTok{(metafor) }\CommentTok{\# for meta{-}analysis models fitting}
-\FunctionTok{library}\NormalTok{(purrr) }\CommentTok{\# for *apply like functions}
+\FunctionTok{library}\NormalTok{(purrr) }\CommentTok{\# for *apply like functions (within tidyverse)}
 \NormalTok{seed }\OtherTok{\textless{}{-}} \DecValTok{2023} \CommentTok{\# general seed for all simulations}
@@
 After each example, we will compute the appropriate model (e.g., \emph{equal} or \emph{random-effects}) using the \texttt{metafor} package (Viechtbauer, 2010) to check the recovery of simulated parameters\footnote{\color{red} By default \texttt{metafor} use the Wald \emph{z} test on model parameters but other inference methods are available e.g.~the Knapp and Hartung (2003) method by setting \texttt{test="knha"} within the \texttt{rma()} function.
-This method is useful especially when the number of studies is low (see \url{https://wviechtb.github.io/metafor/reference/misc-recs.html} for more details) \color{black}}.
-Table \ref{tab:notation-tab} summarizes the simulation parameters notation in equations and code.
+This method is useful especially when the number of studies is low (see \href{https://wviechtb.github.io/metafor/reference/misc-recs.html}{wviechtb.github.io/metafor/reference/misc-recs.html} for more details) \color{black}}.
+Table \ref{tab:notation-tab} \textcolor{red}{summarizes the notation used in simulations and code in equations and code.}
 
@@
 On the left, the simulated model has \(nt,nc = 30\) for each included study while on the right the sample size for each study is \(nt,nc = 500\).
-Given that the data were generated under a \emph{equal-effects model}, when the sample size is high (on the right) each study is aligned on the real effect size because the error component (\(\epsilon_{i}\)) is almost entirely reduced.
+Given that the data were generated under a \emph{equal-effects model}, when the sample size is high (on the right) each study is aligned on the real effect size because the error component (\(\epsilon_{i}\)) \textcolor{red}{is close to zero}.
 The average effect size is similar (depending on the random numbers generation) between the two scenarios while the estimation precision (the width of the black diamond) is narrower on the right.
@@
 Compared to the \emph{equal-effects model} we need to generate another adjustment to the overall effect \(\mu_{\theta}\) from a normal distribution with mean 0 and variance \(\tau^{2}\).
-The real effect size for the study \(i\) will be \(\mu_{\theta} + \theta_{i}\) where \(\theta_{i}\) are the random-effects regulated by the \(\tau^{2}\) parameter.
+The real effect size for the study \(i\) will be \(\mu_{\theta} + \delta_{i}\) where \(\delta_{i}\) are the random-effects regulated by the \(\tau^{2}\) parameter.
 
@@
 As the number of studies \(k\) increases, the estimation of \(\mu_{\theta}\) and \(\tau^{2}\) becomes more precise \color{red} (Blázquez-Rincón, Sánchez-Meca, Botella, \& Suero, 2023; Rubio-Aparicio et al., 2018) \color{black}.
-Additionally, as the sample size of each study decreases, each \(\theta_{i}\) will be estimated with higher precision but as long as \(\tau^{2} \neq 0\) there will be variability among effect sizes.
+Additionally, as the sample size of each study decreases, each \(\delta_{i}\) will be estimated with higher precision but as long as \(\tau^{2} \neq 0\) there will be variability among effect sizes.
 In other words, increasing the sample size of each study or the number of studies will not affect the value of \(\tau^{2}\) but only the estimation precision (see Borenstein et al., 2009, Chapter 16, Figure 16.6 for a clear explanation).
@@
 
-From Equation \eqref{eq:isquared} and Figure \ref{fig:random-effects-high-low-precision} is clear that if each included study has a considerable sample size the total variability will be reduced, mainly driven by real heterogeneity (\(\tau^2\)).
+From Equation \eqref{eq:isquared} and Figure \ref{fig:random-effects-high-low-precision} is clear that if each included study has a considerable sample size \color{red} the sampling variability (\(\tilde{v}\)) will be reduced, and the total variability will be mainly driven by real heterogeneity (\(\tau^2\)) \color{black}.
 This is the crucial difference between the \textcolor{red}{equal-effects} and the random-effects model (see also Borenstein et al., 2009, pp.
@@
 The only caveat is fixing the \(\tilde{v}\).
-When the sample size of each study is the same\footnote{The homogeneous sample size of primary studies assumption is commonly used to calculate the statistical power (see Borenstein et al., 2009, Chapter 29)}, \(\tilde{v} = \sigma_{\epsilon_i}^{2}\).
+When the sample size of each study is the same\footnote{\color{red} The assumption of studies with homogeneous sample sizes is commonly used to calculate the statistical power \color{black}(see Borenstein et al., 2009, Chapter 29)}, \(\tilde{v} = \sigma_{\epsilon_i}^{2}\).
 In the other case, \(\tilde{v}\) needs to be calculated from sampling variances as reported in Higgins and Thompson (2002) and cannot be easily fixed a priori.
@@
 In regression terms (see equation \eqref{eq:meta-regression}), we could use a dummy variable (\(X_{1}\)) that takes the value of 0 for the \emph{lab-based task} (\emph{L}) and a value of 1 for an \emph{online task} (\emph{O})\footnote{The \emph{dummy coding} (also known as treatment coding) is the default in R and \texttt{metafor} but other coding schemes could be used (see Schad, Vasishth, Hohenstein, \& Kliegl, 2020 for an overview of contrast coding schemes).
-\color{red} As in standard linear regression, the contrast coding of categorical factors or the presence/absence of the intercept influence the interpretation of estimated model parameters and omnibus meta-regression tests (see \url{https://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept} for a detailed discussion)} \color{black}.
+\color{red} As in standard linear regression, the contrast coding of categorical factors or the presence/absence of the intercept influences the interpretation of estimated model parameters and omnibus meta-regression tests (see \url{https://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept} for a detailed discussion)} \color{black}.
 Now we fix \(\beta_{1}\) to be the \emph{lab vs.~online effect} (i.e., the expected mean difference between the two groups of studies) and the product between \color{red}\(\beta_{1}X_{1_i}\) \color{black} will consider the \emph{lab vs.~online effect}.
@@
 
-Now we can fit the meta-regression model with the \texttt{rma} function as for the random-effects model with the addition of \texttt{mods\ =\ \textasciitilde{}\ exp} that indicates which variable/s to consider as moderators.
+Now we can fit the meta-regression model with the \texttt{rma} function as for the random-effects model with the addition of \texttt{mods\ =\ \textasciitilde{}\ exp} that indicates which variables to consider as moderators.
 The results are presented in Table \ref{tab:res-meta-reg-dummy}.
@@
 \caption{Meta-regression results for the \emph{random-effects model} with a numerical moderator.
-Each effect size is represented with a black dot where the dimension represents the weight according to the inverse of the precision.
+Each effect size is represented with a black dot where the dimension represents the weight according to the inverse of the variance.
 The line represents the estimated meta-regression slope with the 95\% confidence interval (grey bands).}\label{fig:plot-meta-reg-numerical-r2}
@@
 However, one of the most critical applications is estimating the power of a specific statistical model.
-\color{red} One of the purpose of power analysis by simulations is estimating the required number of studies to detect an hypothetical effect size when planning a meta-analysis.
+\color{red} One of the purposes of power analysis by simulations is to estimate the required number of studies to detect an hypothetical effect size when planning a meta-analysis.
 At the same time, the meta-analysis that we presented (\emph{two-level} equal or random-effects model) can be considered as a multi-lab (e.g., Klein et al., 2018) study where experiments are planned and not collected from the literature.
-In fact, we can use the same simulation approach to optimize the number of participants and studies when planning a multi-lab project.
+We can use the same simulation approach to optimize the number of participants and studies when planning a multi-lab project.
 \color{black}
@@
 Jackson and colleagues (2017) partially solve the issue by developing an interesting method that takes into account the uncertainty in estimating \(\tau^{2}\) without using simulations.
-However, only Monte Carlo simulations can \textcolor{red}{handle} complex scenarios and extra parameters.
+\textcolor{red}{However, complex simulation scenarios can be handled only using Monte Carlo methods.}
 
@@
 
-The present work introduced the basic concepts of the meta-analysis regarding equal, random-effects models and meta-regression with a simulation-based approach.
-We believe the presented examples can help implement alternative or more complex models.
+The present work introduced the basic concepts of the meta-analysis regarding equal-effects, random-effects and meta-regression models with a simulation-based approach.
+\textcolor{red}{We believe the presented examples are useful to implement alternative or more complex models}.
 For example, the \texttt{sim\_study()} function can be easily modified to simulate another effect sizes index, \color{red} such as correlations or odds ratios.
-\color{black} In addition, more complex models, such as \emph{multivariate} or \emph{multilevel} models, can be implemented following the same approach (see \textcolor{red}{the} supplementary materials).
-For example, the \color{red} \emph{multilevel} (e.g., \emph{three-level}) \color{black} model estimates another heterogeneity component representing the variability of multiple \textcolor{red}{independent} effect sizes \textcolor{red}{within} the same study.
-Similarly, the \emph{multivariate} model could include the correlation between multiple outcomes and the correlation between sampling errors.
+\color{black} In addition, more complex models, such as \emph{multivariate} or \emph{multilevel} models, can be simulated following a similar approach (see \textcolor{red}{the} supplementary materials).
+The \color{red} \emph{multilevel} (e.g., \emph{three-level}) \color{black} model estimates another heterogeneity component representing the variability of multiple \textcolor{red}{independent} effect sizes \textcolor{red}{within} the same study.
+Similarly, the \emph{multivariate} model \textcolor{red}{includes} the correlation between multiple outcomes and the correlation between sampling errors.
 
 The present work did have a few limitations.
-First, we only introduced basic concepts about meta-analysis and Monte Carlo simulations, while setting up complex simulations requires more knowledge and complexity of the simulation setup.
+\textcolor{red}{Firstly}, we only introduced basic concepts about meta-analysis and Monte Carlo simulations, while setting up complex simulations requires more knowledge and complexity of the simulation setup.
 We decided to give the foundations to understand meta-analyses with a simulation approach because more complex models are still based on the same principles.
 Second, there are limitations concerning simulating participant-level data.
-We decided to simulate the meta-analysis dataset starting from the participants' level to maximize the flexibility and clearness of each step.
+We decided to simulate the meta-analysis data starting from the participants' level to maximize the flexibility and clearness of each step.
 The downside concerns the efficiency and scalability of the simulation setup.
@@
 
-We thank the \href{https://stat.ethz.ch/mailman/listinfo/r-sig-meta-analysis}{R-sig-meta-analysis} mailing-list where suggestions and clarifications significantly improved our simulations approach and the R code.
+We thank the \emph{\href{https://stat.ethz.ch/mailman/listinfo/r-sig-meta-analysis}{R-sig-meta-analysis}} mailing-list.
+\textcolor{red}{Their} suggestions and clarifications significantly improved \textcolor{red}{the} simulations approach and the R code.
 
@@
 
-The code to reproduce simulations, figures and tables can be found on Open Science Framework (\url{https://osf.io/54djn/})
+The code to reproduce simulations, figures and tables can be found on Open Science Framework (\url{https://osf.io/54djn/}) \color{red} and Github (\url{https://github.com/shared-research/simulating-meta-analysis}).
+\color{black}
 
@@
 
-The supplementary materials can be also found on the Open Science Framework repository (\url{https://osf.io/54djn/})
+The supplementary materials can be also found on the Open Science Framework repository (\url{https://osf.io/54djn/}) \color{red} and Github (\url{https://github.com/shared-research/simulating-meta-analysis}).
+\color{black}
 

### 2023-09-06 documents/paper/paper.Rmd
splitting intro into para

@@
 Furthermore, data simulation is a powerful educational tool for understanding a statistical method.
-In this tutorial, we show how to simulate \rev{equal-effects}, random-effects, and meta-regression models and illustrate how to estimate statistical power.
+In this tutorial, we show how to simulate equal-effects, random-effects, and meta-regression models and illustrate how to estimate statistical power.
 In the supplementary materials, we extended these simulations also for multilevel and multivariate models.
@@
 mask              : no
-#toc               : yes
 figurelist        : no
@@
 classoption       : "man"
-numbersections    : true
+numbersections    : false
 nocite: |
@@
   - \newcommand{\rev}[1]{\textcolor{red}{#1}}
-
-#csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
 documentclass     : "apa6"
@@
   linenumbers: "yes"
-  draft: "no"
   finalsection: TRUE
@@
 
+\setcounter{secnumdepth}{5}
+
 # Introduction
@@
 Firstly, it allows combining evidence from multiple studies assigning more weight to studies with lower estimation variability.
-\color{red} Then using meta-regression, it is possible to include variables (i.e., moderators) to explain the observed heterogeneity [@Borenstein2009-mo, pp.187-203].
+Then using meta-regression, it is possible to include variables (i.e., moderators) to explain the observed heterogeneity [@Borenstein2009-mo, pp.187-203].
 More recently, location-scale models have been developed to include predictors also on the residual heterogeneity [i.e., @Viechtbauer2022-rz].
-\color{black} Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias.
+Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias.
 Despite the advantages, meta-analysis implementation is not always straightforward, especially for complex data structures.
 Additionally, the fact that there are several introductory and advanced resources to understand meta-analysis [@Borenstein2009-mo; @Schmid2022-xe; @Harrer2021-bz], to our knowledge, there are no introductory resources about how to simulate realistic meta-analytic data.
+
 Simulating data has several advantages because it requires understanding the statistical method and the data-generation process.
@@
 155-176].
-A recent paper by DeBruine and Barr [-@DeBruine2021-id], which deeply inspired the current work, proposed a stimulating way to understand linear \rev{mixed-effects models} via data simulation.
+A recent paper by DeBruine and Barr [-@DeBruine2021-id], which deeply inspired the current work, proposed a stimulating way to understand linear mixed-effects models via data simulation.
 Simulating data is also a powerful educational tool within this framework.
-For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for \rev{equal-effects}, random-effects, and meta-regression models with applications also to statistical power calculation.
+
+For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for equal-effects, random-effects, and meta-regression models with applications also to statistical power calculation.
 In the first section, we will introduce basic concepts of the meta-analysis that are useful for setting up the simulation.
-We will evaluate the effect size, variance calculation, and the \rev{equal} vs.
+We will evaluate the effect size, variance calculation, and the equal vs.
 random effects model distinction.
@@
 We used the R statistical programming language `r rcit`.
+
 We assume the reader is familiar with the basic concepts of R, but core functions will be explained.
 Code and materials are available on the OSF repository ([https://osf.io/54djn/](https://osf.io/54djn/)).
-\color{red} A theoretical introduction, simulation examples for multivariate and multilevel models, and more details about the coding approach are available in the supplementary materials.
-\color{black}
+A theoretical introduction, simulation examples for multivariate and multilevel models, and more details about the coding approach are available in the supplementary materials.
 
@@
 
-\color{red} As an example that will be used throughout the paper, we consider the efficacy of memory training in improving memory performance during a cognitive task.
+ As an example that will be used throughout the paper, we consider the efficacy of memory training in improving memory performance during a cognitive task.
 The typical primary study will collect data from a group of participants receiving the memory training (*experimental group*) and another group receiving a control treatment (*control group*).
@@
 Despite differences in the type of cognitive task or experimental setup, each primary study collects an *experimental group* ($n_{T}$) and *control group* ($n_{C}$) and computes the average performance (*experimental group* $\overline T$ and *control group* $\overline C$) and standard deviations (*experimental group* $s_{T}$ and *control group*
-$s_{C}$). \color{black}
+$s_{C}$).
 
@@
 This common measure should give an immediate idea of the direction (i.e., the treatment improves or reduces performance) and the size of the effect.
-Standardized effect size measures [see @Lakens2013-eq for an overview], such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's $d$ [@Cohen1988-ad] or the Pearson correlation coefficient $\rho$ which could be calculated using the associated sample estimator *r* are commonly used to compare heterogeneous outcome variables^[\color{red} When using bounded effect sizes such as the Pearson correlation, Fisher Z-transformation is commonly used [@Borenstein2009-mo, pp.
-41-43] \color{black}].
+Standardized effect size measures [see @Lakens2013-eq for an overview], such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's $d$ [@Cohen1988-ad] or the Pearson correlation coefficient $\rho$ which could be calculated using the associated sample estimator *r* are commonly used to compare heterogeneous outcome variables^[ When using bounded effect sizes such as the Pearson correlation, Fisher Z-transformation is commonly used [@Borenstein2009-mo, pp.
+41-43] ].
 If all studies used the same raw measure, such as reaction times, it is possible to directly meta-analyze the studies without standardizing e.g.
@@
 
-\color{red}
 The core of a meta-analysis is combining the results of multiple studies giving more weight to studies that provide a more precise effect estimation.
@@
 In the presence of heterogeneity, we can include moderators (e.g., the type of experimental setup) into a meta-regression model to explain the effect size heterogeneity.
-\color{red} An interesting proposal to combine evidence from these different models into a single analysis is called Bayesian model-averaged meta-analysis [@Gronau2021-do; @Berkhout2023-fj].
+An interesting proposal to combine evidence from these different models into a single analysis is called Bayesian model-averaged meta-analysis [@Gronau2021-do; @Berkhout2023-fj].
 However, this model is beyond the scope of the tutorial.
-\color{black}
-
-\color{black}
 
@@
 
-\color{red}
 This tutorial uses several R packages for the simulations, meta-analysis fitting, and figures/tables.
@@
 We set the `seed` for reproducibility of the simulation environment.
-\color{black}
 
@@
 
-\color{red}
 Before diving into the specific simulations, in this section, we define the common aspects of all simulations in the following sections.
@@
 The UMD and the sampling variance are calculated using Equations \@ref(eq:effsize) and \@ref(eq:effsizevar) where $y$ is the estimated value of $\Delta$.
-\color{black}
-
-\color{red}
 
@@
 
-\color{black}
-
 We can use the following algorithm implemented in the `sim_study()` function to simulate a single study. Before using the `sim_study()`
@@
 
-\color{red}
 1. Choose a $\Delta$, $n_{T}$, and $n_{C}$ value.
@@
 
-\color{black}
-
 ```{r fun-sim-study, results='asis', echo=FALSE}
@@
 
-\color{red}
 The `es` is the true effect size ($\mu_{\theta}$ for the random-effects model and $\theta$ for the equal-effects model), `nc` and `nt` are the sample size for the control and experimental group, `aggregate` controls if returning the effect size and the corresponding sampling variance or the participant-level data.
@@
 The `mapply()` function is clearly explained in the supplementary materials.
-\color{black}
 
@@
 Using the `sim_study()` function multiple times (with the appropriate adjustments) we can generate a series of studies simulating a data set for a meta-analysis (using the `sim_studies()` function).
-After each example, we will compute the appropriate model (e.g., *equal* or *random-effects*) using the `metafor` package [@Viechtbauer2010-xz] to check the recovery of simulated parameters^[\color{red} By default `metafor` use the Wald *z* test on model parameters but other inference methods are available e.g.
+After each example, we will compute the appropriate model (e.g., *equal* or *random-effects*) using the `metafor` package [@Viechtbauer2010-xz] to check the recovery of simulated parameters^[ By default `metafor` use the Wald *z* test on model parameters but other inference methods are available e.g.
 the Knapp and Hartung [-@Knapp2003-qx] method by setting `test="knha"` within the `rma()` function.
-This method is useful especially when the number of studies is low (see [wviechtb.github.io/metafor/reference/misc-recs.html](https://wviechtb.github.io/metafor/reference/misc-recs.html) for more details) \color{black}].
-Table \@ref(tab:notation-tab) \rev{summarizes the notation used in simulations and code in equations and code.}
+This method is useful especially when the number of studies is low (see [wviechtb.github.io/metafor/reference/misc-recs.html](https://wviechtb.github.io/metafor/reference/misc-recs.html) for more details) ].
+Table \@ref(tab:notation-tab) summarizes the notation used in simulations and code in equations and code.
 
@@
 ```
-
-\color{red}
 
@@
 The most basic model to simulate is the *equal-effects model* (EE).
-As reported in the previous sections, the *equal-effects model* assumes the presence of a single true effect ($\theta$) and the \color{red} observed variability is caused by each effect size being a more or less imprecise estimation of the true effect.
-\color{black} In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies (i.e., the sample size).
-Equations \@ref(eq:equal-effects-model1) and \@ref(eq:equal-effects-model2) \rev{formalize} the *equal-effects* model.
+As reported in the previous sections, the *equal-effects model* assumes the presence of a single true effect ($\theta$) and the  observed variability is caused by each effect size being a more or less imprecise estimation of the true effect.
+In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies (i.e., the sample size).
+Equations \@ref(eq:equal-effects-model1) and \@ref(eq:equal-effects-model2) formalize the *equal-effects* model.
 
@@
 
-\color{black}
-
 \begin{align}
@@
 
-Each observed effect size \color{red} ($y_{i}$) \color{black} is composed of the real effect size \color{red} $\theta$ \color{black} plus an error term ($\epsilon_{i}$) that is sampled from a normal distribution with $\mu = 0$ and \color{red} $\sigma^{2} = \sigma_{\epsilon_i}^{2}$ \color{black} (i.e., the known sampling variance of the study $i$).
+Each observed effect size  ($y_{i}$)  is composed of the real effect size  $\theta$  plus an error term ($\epsilon_{i}$) that is sampled from a normal distribution with $\mu = 0$ and  $\sigma^{2} = \sigma_{\epsilon_i}^{2}$  (i.e., the known sampling variance of the study $i$).
 As demonstrated in Equation \@ref(eq:effsize), the increase in sample size will decrease the sampling variability.
-\rev{A study with a extremely large sample size} will essentially have $\theta$ as the observed effect size.
+A study with a extremely large sample size will essentially have $\theta$ as the observed effect size.
 Since we are sampling participants' level data, the error component is already included in the `sim_study()` function.
@@
 
-\color{red}
 We are using the $\theta$ parameter to generate random data using the `sim_study()` function that will introduce the random error component $\epsilon_{i}$.
@@
 Increasing the number of studies ($k$) and/or the number of participants in each study ($n$) will improve estimation reducing the standard error.
-\color{black}
 
@@
 
-\color{red}
 The *random-effects model* (RE) can be considered an extension of the *equal-effects model*.
@@
 
-\color{black}
-
 ```{r sim-random-effects}
@@
 We are now estimating two parameters: $\mu_{\theta}$ and $\tau^{2}$.
-As for the \color{red} *equal-effects model* \color{black}, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
+As for the  *equal-effects model* , increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
 
@@
 An important aspect of the *random-effects model* is the interplay between the heterogeneity ($\tau^{2}$) and the sampling variability ($\sigma_{i}^{2}$).
-As the number of studies $k$ increases, the estimation of $\mu_{\theta}$ and $\tau^{2}$ becomes more precise \color{red} [@Blazquez-Rincon2023-kp; @Rubio-Aparicio2018-pd] \color{black}.
+As the number of studies $k$ increases, the estimation of $\mu_{\theta}$ and $\tau^{2}$ becomes more precise  [@Blazquez-Rincon2023-kp; @Rubio-Aparicio2018-pd] .
 Additionally, as the sample size of each study decreases, each $\delta_{i}$ will be estimated with higher precision but as long as $\tau^{2} \neq 0$ there will be variability among effect sizes.
@@
 The relationship between the sampling error and the heterogeneity can be expressed using the $I^{2}$ statistics [e.g., @Higgins2002-fh] that is the percentage of the total variability $\tau^{2} + \tilde{v}$ that is attributable to real heterogeneity between studies ($\tau^{2}$).
-\color{red} In Equation \@ref(eq:isquared), $\tilde{v}$ is the *typical* within-study sampling variability (i.e., a summary statistics of sampling variances of included studies) as proposed by Higgins and Thompson [-@Higgins2002-fh]^[See https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate#fn__1 for an overview of the $I^2$ statistics also for complex models.
+In Equation \@ref(eq:isquared), $\tilde{v}$ is the *typical* within-study sampling variability (i.e., a summary statistics of sampling variances of included studies) as proposed by Higgins and Thompson [-@Higgins2002-fh]^[See https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate#fn__1 for an overview of the $I^2$ statistics also for complex models.
 There are also other proposals to compute the $\tilde{v}$ statistics [see @Takkouche1999-sv; @Takkouche2013-nd]] and implemented in Equation \@ref(eq:typical-v) with $w_i = 1/\sigma_{\epsilon_i}^{2}$ and $k$ is the number of studies.
-\color{black}
 
@@
 
-\color{red}
 \begin{align}
@@
 \end{align}
-\color{black}
-
-From Equation \@ref(eq:isquared) and Figure \@ref(fig:random-effects-high-low-precision) is clear that if each included study has a considerable sample size \color{red} the sampling variability ($\tilde{v}$) will be reduced, and the total variability will be mainly driven by real heterogeneity ($\tau^2$) \color{black}.
-This is the crucial difference between the \rev{equal-effects} and the random-effects model [see also @Borenstein2009-mo, pp.
+
+From Equation \@ref(eq:isquared) and Figure \@ref(fig:random-effects-high-low-precision) is clear that if each included study has a considerable sample size  the sampling variability ($\tilde{v}$) will be reduced, and the total variability will be mainly driven by real heterogeneity ($\tau^2$) .
+This is the crucial difference between the equal-effects and the random-effects model [see also @Borenstein2009-mo, pp.
 117–122].
 
-\color{red}
 Given the interpretation of $I^{2}$ it is possible to simulate a meta-analysis fixing a certain $I^{2}$ value.
 The only caveat is fixing the $\tilde{v}$.
-When the sample size of each study is the same^[\color{red} The assumption of studies with homogeneous sample sizes is commonly used to calculate the statistical power \color{black}[see @Borenstein2009-mo, Chapter 29]], $\tilde{v} = \sigma_{\epsilon_i}^{2}$.
+When the sample size of each study is the same^[ The assumption of studies with homogeneous sample sizes is commonly used to calculate the statistical power [see @Borenstein2009-mo, Chapter 29]], $\tilde{v} = \sigma_{\epsilon_i}^{2}$.
 In the other case, $\tilde{v}$ needs to be calculated from sampling variances as reported in Higgins and Thompson [-@Higgins2002-fh] and cannot be easily fixed a priori.
 With the assumption of homogeneous sample size across studies^[In case of heterogeneous sample sizes generated from a probability distribution (e.g., Gaussian) the *typical* sampling variance calculated using the expected value of the distribution can be used to somehow fix the (average) $I^2$ (see the supplementary materials for a simulated example)], we can solve Equation \@ref(eq:isquared) for $\tau^{2}$ obtaining the heterogeneity value associated with a certain $I^{2}$ as reported in Equation \@ref(eq:tau2-from-isquared).
-\color{black}
-
-\color{red}
+
 Table \@ref(tab:res-random-effects-i2) depicts the results of the random-effects model fixing the $I^{2}$ value.
@@
 For example van Erp and colleagues [-@Van_Erp2017-cf] estimate an empirical $\tau^2$ distribution across several published meta-analyses that can be used to simulate a plausible scenario.
-\color{black}
 
@@
 In regression terms (see equation \@ref(eq:meta-regression)), we could use a dummy variable ($X_{1}$) that takes the value of 0 for the *lab-based task* (*L*) and a value of 1 for an *online task* (*O*)^[The *dummy coding* (also known as treatment coding) is the default in R and `metafor` but other coding schemes could be used [see @Schad2020-ht for an overview of contrast coding schemes].
-\color{red} As in standard linear regression, the contrast coding of categorical factors or the presence/absence of the intercept influences the interpretation of estimated model parameters and omnibus meta-regression tests (see https://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept for a detailed discussion)] \color{black}.
+As in standard linear regression, the contrast coding of categorical factors or the presence/absence of the intercept influences the interpretation of estimated model parameters and omnibus meta-regression tests (see https://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept for a detailed discussion)] .
 Now we fix $\beta_{1}$ to be the *lab vs.
-online effect* (i.e., the expected mean difference between the two groups of studies) and the product between \color{red}$\beta_{1}X_{1_i}$ \color{black} will consider the *lab vs.
+online effect* (i.e., the expected mean difference between the two groups of studies) and the product between $\beta_{1}X_{1_i}$  will consider the *lab vs.
 online effect*.
 Crucially, despite $\tau^{2}$ is still the heterogeneity between effect sizes, now we need to fix $\tau^{2}$ considering that we included a moderator.
-In other terms, \color{red} $\delta_i \sim \mathcal{N}(0, \tau_{r}^{2})$ \color{black} where $\tau^2_r$ is the residual heterogeneity after including the moderator.
+In other terms,  $\delta_i \sim \mathcal{N}(0, \tau_{r}^{2})$  where $\tau^2_r$ is the residual heterogeneity after including the moderator.
 We can describe our model using equation \@ref(eq:meta-regression-equations) according to the value of $X_{1_i}$ (L = lab-based experiments, O = online experiments).
@@
 Then the $\beta_1$ parameter represents the estimated difference in $y$ between the values of $X_1$ (i.e., lab-based vs online experiments).
-As said before, \color{red} $\tau^2_r$ \color{black} is now the residual heterogeneity that is interpreted as the variability between effect sizes after controlling for the moderator $X_1$.
+As said before,  $\tau^2_r$  is now the residual heterogeneity that is interpreted as the variability between effect sizes after controlling for the moderator $X_1$.
 
@@
 A first strategy could be to use values estimated from the literature.
-Another approach consists in setting up the model and simulating several expected \color{red} $y_i$ \color{black} and calculating the range of simulated values.
+Another approach consists in setting up the model and simulating several expected  $y_i$  and calculating the range of simulated values.
 A third possibility is fixing the proportion of explained heterogeneity calculating the $\beta_1$ value accordingly.
@@
 However, one of the most critical applications is estimating the power of a specific statistical model.
-\color{red} One of the purposes of power analysis by simulations is to estimate the required number of studies to detect an hypothetical effect size when planning a meta-analysis.
+One of the purposes of power analysis by simulations is to estimate the required number of studies to detect an hypothetical effect size when planning a meta-analysis.
 At the same time, the meta-analysis that we presented (*two-level* equal or random-effects model) can be considered as a multi-lab [e.g., @Klein2018-he] study where experiments are planned and not collected from the literature.
 We can use the same simulation approach to optimize the number of participants and studies when planning a multi-lab project.
-\color{black}
 
@@
 Jackson and colleagues [-@Jackson2017-dv] partially solve the issue by developing an interesting method that takes into account the uncertainty in estimating $\tau^{2}$ without using simulations.
-\rev{However, complex simulation scenarios can be handled only using Monte Carlo methods.}
+However, complex simulation scenarios can be handled only using Monte Carlo methods.
 
@@
 The present work introduced the basic concepts of the meta-analysis regarding equal-effects, random-effects and meta-regression models with a simulation-based approach.
-\rev{We believe the presented examples are useful to implement alternative or more complex models}.
-For example, the `sim_study()` function can be easily modified to simulate another effect sizes index, \color{red} such as correlations or odds ratios.
-\color{black} In addition, more complex models, such as *multivariate* or *multilevel* models, can be simulated following a similar approach (see \rev{the} supplementary materials).
-The \color{red} *multilevel* (e.g., *three-level*) \color{black} model estimates another heterogeneity component representing the variability of multiple \rev{independent} effect sizes \rev{within} the same study.
-Similarly, the *multivariate* model \rev{includes} the correlation between multiple outcomes and the correlation between sampling errors.
+We believe the presented examples are useful to implement alternative or more complex models.
+For example, the `sim_study()` function can be easily modified to simulate another effect sizes index,  such as correlations or odds ratios.
+In addition, more complex models, such as *multivariate* or *multilevel* models, can be simulated following a similar approach (see the supplementary materials).
+The  *multilevel* (e.g., *three-level*)  model estimates another heterogeneity component representing the variability of multiple independent effect sizes within the same study.
+Similarly, the *multivariate* model includes the correlation between multiple outcomes and the correlation between sampling errors.
 
 The present work did have a few limitations.
-\rev{Firstly}, we only introduced basic concepts about meta-analysis and Monte Carlo simulations, while setting up complex simulations requires more knowledge and complexity of the simulation setup.
+Firstly, we only introduced basic concepts about meta-analysis and Monte Carlo simulations, while setting up complex simulations requires more knowledge and complexity of the simulation setup.
 We decided to give the foundations to understand meta-analyses with a simulation approach because more complex models are still based on the same principles.
@@
 The downside concerns the efficiency and scalability of the simulation setup.
-For large-scale simulations (e.g., many conditions, iterations, or complex models), simulating from aggregated statistics is probably more efficient [see @van-den-Heuvel2020-rv for an example] to improve the \rev{simulation efficency}^[See https://www.jepusto.com/simulating-correlated-smds/ for a very clear example of the participant-level vs aggregated data simulation approaches].
+For large-scale simulations (e.g., many conditions, iterations, or complex models), simulating from aggregated statistics is probably more efficient [see @van-den-Heuvel2020-rv for an example] to improve the simulation efficency^[See https://www.jepusto.com/simulating-correlated-smds/ for a very clear example of the participant-level vs aggregated data simulation approaches].
 
@@
 We thank the *[R-sig-meta-analysis](https://stat.ethz.ch/mailman/listinfo/r-sig-meta-analysis)* mailing-list.
-\rev{Their} suggestions and clarifications significantly improved \rev{the} simulations approach and the R code.
+Their suggestions and clarifications significantly improved the simulations approach and the R code.
 
@@
 
-The code to reproduce simulations, figures and tables can be found on Open Science Framework (https://osf.io/54djn/) \color{red} and Github (https://github.com/shared-research/simulating-meta-analysis).
-\color{black}
+The code to reproduce simulations, figures and tables can be found on Open Science Framework (https://osf.io/54djn/) and Github (https://github.com/shared-research/simulating-meta-analysis).
 
@@
 
-The supplementary materials can be also found on the Open Science Framework repository (https://osf.io/54djn/) \color{red} and Github (https://github.com/shared-research/simulating-meta-analysis).
-\color{black}
+The supplementary materials can be also found on the Open Science Framework repository (https://osf.io/54djn/) and Github (https://github.com/shared-research/simulating-meta-analysis).
 
@@
 
-(ref:img-equal-vs-random) \color{red} The difference between the assumptions of the *equal* and *random-effects model*.
+(ref:img-equal-vs-random)  The difference between the assumptions of the *equal* and *random-effects model*.
 Each distribution depicts the sampling distribution of $k = 4$ hypothetical studies ($i = 1$, $2$, ..., $4$) with a certain observed effect size $y_{i}$ (red squares), sampling variability $\sigma_{\epsilon_i}^{2}$ and the 95% confidence interval (black segment).
@@
 When $\tau^{2}$ is zero, the random-effects reduces to a equal-effects model.
-\color{black}
 
@@
 
-(ref:res-equal-effects) \color{red} Summary of the simulated equal-effects model.
+(ref:res-equal-effects)  Summary of the simulated equal-effects model.
 The only estimated parameter is the average effect $\theta$ ($\beta$) with the standard error, 95% confidence interval, and the Wald z-test.
-The z-test evaluates the null hypothesis that the real effect equals zero.\color{black}
-
-(ref:equal-effects-high-low-precision) \color{red} Forest plots of two simulated *equal-effects models*.
+The z-test evaluates the null hypothesis that the real effect equals zero.
+
+(ref:equal-effects-high-low-precision)  Forest plots of two simulated *equal-effects models*.
 On the left, the simulated model has $nt,nc = 30$ for each included study while on the right the sample size for each study is $nt,nc = 500$.
-Given that the data were generated under a *equal-effects model*, when the sample size is high (on the right) each study is aligned on the real effect size because the error component ($\epsilon_{i}$) \rev{is close to zero}.
+Given that the data were generated under a *equal-effects model*, when the sample size is high (on the right) each study is aligned on the real effect size because the error component ($\epsilon_{i}$) is close to zero.
 The average effect size is similar (depending on the random numbers generation) between the two scenarios while the estimation precision (the width of the black diamond) is narrower on the right.
-\color{black}
-
-(ref:res-random-effects) \color{red} Summary of the simulated random-effects model: Compared to the equal-effects model, there are more parameters.
+
+(ref:res-random-effects)  Summary of the simulated random-effects model: Compared to the equal-effects model, there are more parameters.
 The $\beta$ is the average effect ($\mu_{\theta}$) with the standard error, 95% confidence interval, and the Wald z-test.
 The $\tau^{2}$ is the estimated heterogeneity and the $I^{2}$ (explained in Section \@ref(re-model)) represents the percentage of total variability due to between-study heterogeneity.
-\color{black}
-
-(ref:random-effects-high-low-precision) \color{red} Forest plots of two simulated *random-effects model*s.
+
+(ref:random-effects-high-low-precision)  Forest plots of two simulated *random-effects model*s.
 On the left, the simulated model has $nt,nc = 30$ for each included study, while on the right, the sample size for each study is $nt,nc = 500$.
@@
 However, compared to the *equal-effects* simulation, increasing the sample size of primary studies only affects the precision without reducing the true heterogeneity (i.e., $\tau^{2} \neq 0$).
-\color{black}
 
@@
 The green segments depict the explained heterogeneity, and the red segments the residual (unexplained) heterogeneity.
-\color{red} The red squares are simulated observed effect sizes from the sampling distributions.
-\color{black}
+The red squares are simulated observed effect sizes from the sampling distributions.
 
@@
 
-(ref:res-meta-reg-dummy) \color{red} Summary of the *random-effect* model with a categorical predictor (lab vs online experiments).
+(ref:res-meta-reg-dummy)  Summary of the *random-effect* model with a categorical predictor (lab vs online experiments).
 The `intercept` is the average effect for lab-based experiments and `exponline` is the difference between lab-based and online experiments.
@@
 Other parameters are the same as the standard *random-effect* model (see Table \@ref(tab:res-random-effects)).
-\color{black}
-
-(ref:res-meta-reg-numerical-r2) \color{red} Summary of the *random-effects* model fixing the $R^{2}$ value.
+
+(ref:res-meta-reg-numerical-r2)  Summary of the *random-effects* model fixing the $R^{2}$ value.
 The `intercept` is the effect size for the average `age` (given that age is mean-centered).
@@
 Other parameters are the same as described in Table \@ref(tab:res-random-effects) and \@ref(tab:res-meta-reg-dummy).
-\color{black}
-
-(ref:plot-meta-reg-plausible-implausible) \color{red} Scatter plots with marginal histograms for the range of simulated $y_i$ with two $\beta_1$ values.
+
+(ref:plot-meta-reg-plausible-implausible)  Scatter plots with marginal histograms for the range of simulated $y_i$ with two $\beta_1$ values.
 The x-axis depicts the average age of simulated studies and the y-axis depicts the simulated effect size.
@@
 On the right, ($\beta = 0.7$) values range between -10 to 10 that despite theoretically possible can be considered highly implausible values in real meta-analyses for psychological data.
-\color{black}
 

### 2023-09-06 scripts/r-img.R
adding color-blind-friendly palettes

@@ -15,6 +15,12 @@ theme_rfig <- function(size = 15){
   ltxplot::theme_latex(base_size = size)
 }
 
+# color-blind-friendly palettes see http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
+
+cb_blue <- "#0072B2"
+cb_red <- "#CC79A7"
+cb_green <- "#009E73"
+
 # Equal-vs-Random Effect --------------------------------------------------
 
 equal_effects <- data.frame(
@@ -63,7 +69,7 @@ p_equal_down <- ggplot(filter(models, model == "Equal-Effects"), aes(y = factor(
   geom_label(aes(x = di, y = 0.6, label = note), parse = TRUE, size = 7, label.size = NA) +
   geom_segment(aes(x = di + delta, xend = di, y = id-0.08, yend = id-0.08),
                position = position_dodge2(width = 1, padding = 1)) +
-  geom_point(aes(x = obs, y = factor(id)), color = "firebrick2", size = 2, shape = 15) +
+  geom_point(aes(x = obs, y = factor(id)), color = cb_red, size = 2, shape = 15) +
   geom_point(aes(x = yi, y = id-0.08, alpha = model), show.legend = FALSE) +
   scale_alpha_manual(values = c(0, 1))
 
@@ -87,7 +93,7 @@ p_random_down <- ggplot(filter(models, model == "Random-Effects"), aes(y = facto
   geom_label(aes(x = di, y = 0.6, label = note), parse = TRUE, size = 7, label.size = NA) +
   # geom_segment(aes(x = di + delta, xend = di, y = id-0.08, yend = id-0.08),
   #              position = position_dodge2(width = 1, padding = 1)) +
-  geom_point(aes(x = obs, y = factor(id)), color = "firebrick2", size = 2, shape = 15) +
+  geom_point(aes(x = obs, y = factor(id)), color = cb_red, size = 2, shape = 15) +
   geom_point(aes(x = yi, y = id-0.08, alpha = model), show.legend = FALSE) +
   scale_alpha_manual(values = c(0, 1)) +
   geom_segment(aes(x = di + delta, xend = di + delta,
@@ -124,7 +130,7 @@ p_up_left <- ggplot(up_left) +
         axis.ticks.y = element_blank())
 
 p_up_right <- ggplot(up_right) +
-  stat_halfeye(aes(y = id, dist = dist), .width = 0.95, fill = "#7FB3D5") +
[…]

### 2023-09-06 documents/paper/paper.Rmd
adding section on tutorial usage

@@
 keywords          : "meta-analysis, monte carlo simulations, power analysis"
-wordcount         : "7782"
+wordcount         : "7866"
 bibliography      : "../files/references.bib"
@@
 
+The aim of the tutorial is not providing a complete theoretical introduction to meta-analysis but presenting the core topics using a simulation-based approach.
+Readers expert in conducting meta-analyses could benefit more from the proposed approach in terms of simulation setup and coding strategies.
+Readers without experience in meta-analysis could benefit both from the theoretical introduction and the simulation approach.
+However, for a better and extensive overview of meta-analysis topics the reader could refer to meta-analysis textbooks [e.g., @Harrer2019-rl; @Borenstein2009-mo].
+
 We assume the reader is familiar with the basic concepts of R, but core functions will be explained.
@@
   ggplot(aes(x = k, y = power)) +
-  geom_hline(yintercept = 0.8, linetype = "dashed", color = "firebrick4") +
+  geom_hline(yintercept = 0.8, linetype = "dashed", color = "#CC79A7") +
   geom_line() +
@@
 (ref:img-equal-vs-random)  The difference between the assumptions of the *equal* and *random-effects model*.
-Each distribution depicts the sampling distribution of $k = 4$ hypothetical studies ($i = 1$, $2$, ..., $4$) with a certain observed effect size $y_{i}$ (red squares), sampling variability $\sigma_{\epsilon_i}^{2}$ and the 95% confidence interval (black segment).
+Each distribution depicts the sampling distribution of $k = 4$ hypothetical studies ($i = 1$, $2$, ..., $4$) with a certain observed effect size $y_{i}$ (pink squares), sampling variability $\sigma_{\epsilon_i}^{2}$ and the 95% confidence interval (black segment).
 The *equal-effects* plot on the left suggests that each observed effect size has the same underlying true effect $\theta$ (each distribution has the same mean) with a different degree of precision (e.g., Study 1 is more precise than Study 3).
@@
 The dotted line is the average effect (i.e., random-effects model without moderators).
-The effect size differs between conditions A and B, including the *condition* moderator explaining part of the total heterogeneity (red plus green segments).
-The green segments depict the explained heterogeneity, and the red segments the residual (unexplained) heterogeneity.
-The red squares are simulated observed effect sizes from the sampling distributions.
+The effect size differs between conditions A and B, including the *condition* moderator explaining part of the total heterogeneity (pink plus green segments).
+The green segments depict the explained heterogeneity, and the pink segments the residual (unexplained) heterogeneity.
+The pink squares are simulated observed effect sizes from the sampling distributions.
 
@@
 The x-axis depicts the number of studies ($k$), and the y-axis the estimated power.
-The red dotted line is the 80% power level, usually considered a good value for power analysis.
+The pink dotted line is the 80% power level, usually considered a good value for power analysis.
 

### 2023-09-06 documents/paper/paper.tex
updating after revision

@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 % Make \paragraph and \subparagraph free-standing
@@
 \maketitle
+
+\setcounter{secnumdepth}{5}
 
@@
 Firstly, it allows combining evidence from multiple studies assigning more weight to studies with lower estimation variability.
-\color{red} Then using meta-regression, it is possible to include variables (i.e., moderators) to explain the observed heterogeneity (Borenstein, Hedges, Higgins, \& Rothstein, 2009, pp.
+Then using meta-regression, it is possible to include variables (i.e., moderators) to explain the observed heterogeneity (Borenstein, Hedges, Higgins, \& Rothstein, 2009, pp.
 187--203).
 More recently, location-scale models have been developed to include predictors also on the residual heterogeneity (i.e., Viechtbauer \& López-López, 2022).
-\color{black} Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias.
+Finally, considering the replication crisis, there are statistical methods to determine the presence and extent of the publication bias.
 Despite the advantages, meta-analysis implementation is not always straightforward, especially for complex data structures.
 Additionally, the fact that there are several introductory and advanced resources to understand meta-analysis (Borenstein et al., 2009; Harrer, Cuijpers, Furukawa, \& Ebert, 2021; Schmid, Stijnen, \& White, 2022), to our knowledge, there are no introductory resources about how to simulate realistic meta-analytic data.
+
 Simulating data has several advantages because it requires understanding the statistical method and the data-generation process.
@@
 69--76; Ingalls, 2011).
-A recent paper by DeBruine and Barr (2021), which deeply inspired the current work, proposed a stimulating way to understand linear \textcolor{red}{mixed-effects models} via data simulation.
+A recent paper by DeBruine and Barr (2021), which deeply inspired the current work, proposed a stimulating way to understand linear mixed-effects models via data simulation.
 Simulating data is also a powerful educational tool within this framework.
-For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for \textcolor{red}{equal-effects}, random-effects, and meta-regression models with applications also to statistical power calculation.
+
+For these reasons, this work aims to introduce the basic concepts of meta-analysis and Monte Carlo simulations for equal-effects, random-effects, and meta-regression models with applications also to statistical power calculation.
 In the first section, we will introduce basic concepts of the meta-analysis that are useful for setting up the simulation.
-We will evaluate the effect size, variance calculation, and the \textcolor{red}{equal} vs.~random effects model distinction.
+We will evaluate the effect size, variance calculation, and the equal vs.~random effects model distinction.
 Then we will describe how to simulate data for these models and simulate a meta-regression with categorical and numerical predictors.
 Finally, we will introduce the power analysis extending the previous examples to estimate the statistical power.
-We used the R statistical programming language (version 4.2.3; R Core Team, 2023).
+We used the R statistical programming language (version 4.3.1; R Core Team, 2023).
+
+The aim of the tutorial is not providing a complete theoretical introduction to meta-analysis but presenting the core topics using a simulation-based approach.
+Readers expert in conducting meta-analyses could benefit more from the proposed approach in terms of simulation setup and coding strategies.
+Readers without experience in meta-analysis could benefit both from the theoretical introduction and the simulation approach.
+However, for a better and extensive overview of meta-analysis topics the reader could refer to Harrer, Cuijpers, and Ebert (2019), and Borenstein et al.
+(2009).
+
 We assume the reader is familiar with the basic concepts of R, but core functions will be explained.
 Code and materials are available on the OSF repository (\url{https://osf.io/54djn/}).
-\color{red} A theoretical introduction, simulation examples for multivariate and multilevel models, and more details about the coding approach are available in the supplementary materials.
-\color{black}
+A theoretical introduction, simulation examples for multivariate and multilevel models, and more details about the coding approach are available in the supplementary materials.
 
@@
 
-\color{red} As an example that will be used throughout the paper, we consider the efficacy of memory training in improving memory performance during a cognitive task.
+As an example that will be used throughout the paper, we consider the efficacy of memory training in improving memory performance during a cognitive task.
 The typical primary study will collect data from a group of participants receiving the memory training (\emph{experimental group}) and another group receiving a control treatment (\emph{control group}).
@@
 Despite differences in the type of cognitive task or experimental setup, each primary study collects an \emph{experimental group} (\(n_{T}\)) and \emph{control group} (\(n_{C}\)) and computes the average performance (\emph{experimental group} \(\overline T\) and \emph{control group} \(\overline C\)) and standard deviations (\emph{experimental group} \(s_{T}\) and \emph{control group}
-\(s_{C}\)). \color{black}
+\(s_{C}\)).
 
@@
 This common measure should give an immediate idea of the direction (i.e., the treatment improves or reduces performance) and the size of the effect.
-Standardized effect size measures (see Lakens, 2013 for an overview), such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's \(d\) (Cohen, 1988) or the Pearson correlation coefficient \(\rho\) which could be calculated using the associated sample estimator \emph{r} are commonly used to compare heterogeneous outcome variables\footnote{\color{red} When using bounded effect sizes such as the Pearson correlation, Fisher Z-transformation is commonly used (Borenstein et al., 2009, pp.
-41--43) \color{black}}.
+Standardized effect size measures (see Lakens, 2013 for an overview), such as the Standardized Mean Difference (SMD), which could be estimated by Cohen's \(d\) (Cohen, 1988) or the Pearson correlation coefficient \(\rho\) which could be calculated using the associated sample estimator \emph{r} are commonly used to compare heterogeneous outcome variables\footnote{ When using bounded effect sizes such as the Pearson correlation, Fisher Z-transformation is commonly used (Borenstein et al., 2009, pp.
+41--43)}.
 If all studies used the same raw measure, such as reaction times, it is possible to directly meta-analyze the studies without standardizing e.g.~using the unstandardized mean difference (UMD, Borenstein et al., 2009, pp.
@@
 \subsubsection{Equal vs fixed vs random-effects model}\label{equal-vs-fixed-vs-random-effects-model}}
-
-\color{red}
 
@@
 In the presence of heterogeneity, we can include moderators (e.g., the type of experimental setup) into a meta-regression model to explain the effect size heterogeneity.
-\color{red} An interesting proposal to combine evidence from these different models into a single analysis is called Bayesian model-averaged meta-analysis (Berkhout, Haaf, Gronau, Heck, \& Wagenmakers, 2023; Gronau, Heck, Berkhout, Haaf, \& Wagenmakers, 2021).
+An interesting proposal to combine evidence from these different models into a single analysis is called Bayesian model-averaged meta-analysis (Berkhout, Haaf, Gronau, Heck, \& Wagenmakers, 2023; Gronau, Heck, Berkhout, Haaf, \& Wagenmakers, 2021).
 However, this model is beyond the scope of the tutorial.
-\color{black}
-
-\color{black}
 
@@
 
-\caption{\color{red} The difference between the assumptions of the \emph{equal} and \emph{random-effects model}.
-Each distribution depicts the sampling distribution of \(k = 4\) hypothetical studies (\(i = 1\), \(2\), \ldots, \(4\)) with a certain observed effect size \(y_{i}\) (red squares), sampling variability \(\sigma_{\epsilon_i}^{2}\) and the 95\% confidence interval (black segment).
+\caption{The difference between the assumptions of the \emph{equal} and \emph{random-effects model}.
+Each distribution depicts the sampling distribution of \(k = 4\) hypothetical studies (\(i = 1\), \(2\), \ldots, \(4\)) with a certain observed effect size \(y_{i}\) (pink squares), sampling variability \(\sigma_{\epsilon_i}^{2}\) and the 95\% confidence interval (black segment).
 The \emph{equal-effects} plot on the left suggests that each observed effect size has the same underlying true effect \(\theta\) (each distribution has the same mean) with a different degree of precision (e.g., Study 1 is more precise than Study 3).
@@
 The \emph{random-effects model} on the right suggests that beyond the error term (\(\epsilon_{i}\)), each real effect size is composed of a fixed part (now \(\mu_{\theta}\)) and a random part (\(\delta_{i}\)) sampled from a normal distribution with mean zero and variance \(\tau^{2}\).
-When \(\tau^{2}\) is zero, the random-effects reduces to a equal-effects model.
-\color{black}}\label{fig:img-equal-vs-random}
+When \(\tau^{2}\) is zero, the random-effects reduces to a equal-effects model.}\label{fig:img-equal-vs-random}
 \end{figure}
@@
 
-\color{red}
-
 This tutorial uses several R packages for the simulations, meta-analysis fitting, and figures/tables.
@@
 We set the \texttt{seed} for reproducibility of the simulation environment.
-\color{black}
 
@@
 
-\color{red}
-
 Before diving into the specific simulations, in this section, we define the common aspects of all simulations in the following sections.
@@
 The UMD and the sampling variance are calculated using Equations \eqref{eq:effsize} and \eqref{eq:effsizevar} where \(y\) is the estimated value of \(\Delta\).
-\color{black}
-
-\color{red}
 
@@
 
-\color{black}
-
 We can use the following algorithm implemented in the \texttt{sim\_study()} function to simulate a single study. Before using the \texttt{sim\_study()}
@@
 Table \ref{tab:sim-data-example-tab} depicts an example of the \texttt{make\_data()} output.
-
-\color{red}
 
@@
 For example, after generating data for the two groups the mean difference can be standardized using the pooled standard deviation and applying the appropriate correction (Hedges, 1981; e.g., Hedges, 1989).
-
-\color{black}
 
@@
 
-\color{red}
-
 The \texttt{es} is the true effect size (\(\mu_{\theta}\) for the random-effects model and \(\theta\) for the equal-effects model), \texttt{nc} and \texttt{nt} are the sample size for the control and experimental group, \texttt{aggregate} controls if returning the effect size and the corresponding sampling variance or the participant-level data.
@@
 The \texttt{mapply()} function is clearly explained in the supplementary materials.
-\color{black}
 
@@
 Using the \texttt{sim\_study()} function multiple times (with the appropriate adjustments) we can generate a series of studies simulating a data set for a meta-analysis (using the \texttt{sim\_studies()} function).
-After each example, we will compute the appropriate model (e.g., \emph{equal} or \emph{random-effects}) using the \texttt{metafor} package (Viechtbauer, 2010) to check the recovery of simulated parameters\footnote{\color{red} By default \texttt{metafor} use the Wald \emph{z} test on model parameters but other inference methods are available e.g.~the Knapp and Hartung (2003) method by setting \texttt{test="knha"} within the \texttt{rma()} function.
-This method is useful especially when the number of studies is low (see \href{https://wviechtb.github.io/metafor/reference/misc-recs.html}{wviechtb.github.io/metafor/reference/misc-recs.html} for more details) \color{black}}.
-Table \ref{tab:notation-tab} \textcolor{red}{summarizes the notation used in simulations and code in equations and code.}
+After each example, we will compute the appropriate model (e.g., \emph{equal} or \emph{random-effects}) using the \texttt{metafor} package (Viechtbauer, 2010) to check the recovery of simulated parameters\footnote{ By default \texttt{metafor} use the Wald \emph{z} test on model parameters but other inference methods are available e.g.~the Knapp and Hartung (2003) method by setting \texttt{test="knha"} within the \texttt{rma()} function.
+This method is useful especially when the number of studies is low (see \href{https://wviechtb.github.io/metafor/reference/misc-recs.html}{wviechtb.github.io/metafor/reference/misc-recs.html} for more details)}.
+Table \ref{tab:notation-tab} summarizes the notation used in simulations and code in equations and code.
 
@@
 \normalsize
-
-\color{red}
 
@@
 The most basic model to simulate is the \emph{equal-effects model} (EE).
-As reported in the previous sections, the \emph{equal-effects model} assumes the presence of a single true effect (\(\theta\)) and the \color{red} observed variability is caused by each effect size being a more or less imprecise estimation of the true effect.
-\color{black} In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies (i.e., the sample size).
-Equations \eqref{eq:equal-effects-model1} and \eqref{eq:equal-effects-model2} \textcolor{red}{formalize} the \emph{equal-effects} model.
+As reported in the previous sections, the \emph{equal-effects model} assumes the presence of a single true effect (\(\theta\)) and the observed variability is caused by each effect size being a more or less imprecise estimation of the true effect.
+In other words, the only source of variability is the sampling variability that depends on the variance of the primary studies (i.e., the sample size).
+Equations \eqref{eq:equal-effects-model1} and \eqref{eq:equal-effects-model2} formalize the \emph{equal-effects} model.
 
@@
 
-Each observed effect size \color{red} (\(y_{i}\)) \color{black} is composed of the real effect size \color{red} \(\theta\) \color{black} plus an error term (\(\epsilon_{i}\)) that is sampled from a normal distribution with \(\mu = 0\) and \color{red} \(\sigma^{2} = \sigma_{\epsilon_i}^{2}\) \color{black} (i.e., the known sampling variance of the study \(i\)).
+Each observed effect size (\(y_{i}\)) is composed of the real effect size \(\theta\) plus an error term (\(\epsilon_{i}\)) that is sampled from a normal distribution with \(\mu = 0\) and \(\sigma^{2} = \sigma_{\epsilon_i}^{2}\) (i.e., the known sampling variance of the study \(i\)).
 As demonstrated in Equation \eqref{eq:effsize}, the increase in sample size will decrease the sampling variability.
-\textcolor{red}{A study with a extremely large sample size} will essentially have \(\theta\) as the observed effect size.
+A study with a extremely large sample size will essentially have \(\theta\) as the observed effect size.
 Since we are sampling participants' level data, the error component is already included in the \texttt{sim\_study()} function.
@@
 
-\color{red}
-
 We are using the \(\theta\) parameter to generate random data using the \texttt{sim\_study()} function that will introduce the random error component \(\epsilon_{i}\).
@@
 Increasing the number of studies (\(k\)) and/or the number of participants in each study (\(n\)) will improve estimation reducing the standard error.
-\color{black}
 
@@
 
-\caption{\label{tab:res-equal-effects}\color{red} Summary of the simulated equal-effects model.
+\caption{\label{tab:res-equal-effects}Summary of the simulated equal-effects model.
 The only estimated parameter is the average effect \(\theta\) (\(\beta\)) with the standard error, 95\% confidence interval, and the Wald z-test.
-The z-test evaluates the null hypothesis that the real effect equals zero.\color{black}}
+The z-test evaluates the null hypothesis that the real effect equals zero.}
 \centering
@@
 
-\caption{\color{red} Forest plots of two simulated \emph{equal-effects models}.
+\caption{Forest plots of two simulated \emph{equal-effects models}.
 On the left, the simulated model has \(nt,nc = 30\) for each included study while on the right the sample size for each study is \(nt,nc = 500\).
-Given that the data were generated under a \emph{equal-effects model}, when the sample size is high (on the right) each study is aligned on the real effect size because the error component (\(\epsilon_{i}\)) \textcolor{red}{is close to zero}.
-The average effect size is similar (depending on the random numbers generation) between the two scenarios while the estimation precision (the width of the black diamond) is narrower on the right.
-\color{black}}\label{fig:equal-effects-high-low-precision}
+Given that the data were generated under a \emph{equal-effects model}, when the sample size is high (on the right) each study is aligned on the real effect size because the error component (\(\epsilon_{i}\)) is close to zero.
+The average effect size is similar (depending on the random numbers generation) between the two scenarios while the estimation precision (the width of the black diamond) is narrower on the right.}\label{fig:equal-effects-high-low-precision}
 \end{figure}
@@
 \subsection{Random-effects model}\label{re-model}}
-
-\color{red}
 
@@
 
-\color{black}
-
 \scriptsize
@@
 We are now estimating two parameters: \(\mu_{\theta}\) and \(\tau^{2}\).
-As for the \color{red} \emph{equal-effects model} \color{black}, increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
+As for the \emph{equal-effects model} , increasing the number of studies and/or the number of participants in each study will improve estimation reducing the standard error.
 
@@
 
-\caption{\label{tab:res-random-effects}\color{red} Summary of the simulated random-effects model: Compared to the equal-effects model, there are more parameters.
+\caption{\label{tab:res-random-effects}Summary of the simulated random-effects model: Compared to the equal-effects model, there are more parameters.
 The \(\beta\) is the average effect (\(\mu_{\theta}\)) with the standard error, 95\% confidence interval, and the Wald z-test.
-The \(\tau^{2}\) is the estimated heterogeneity and the \(I^{2}\) (explained in Section \ref{re-model}) represents the percentage of total variability due to between-study heterogeneity.
-\color{black}}
+The \(\tau^{2}\) is the estimated heterogeneity and the \(I^{2}\) (explained in Section \ref{re-model}) represents the percentage of total variability due to between-study heterogeneity.}
 \centering
@@
 An important aspect of the \emph{random-effects model} is the interplay between the heterogeneity (\(\tau^{2}\)) and the sampling variability (\(\sigma_{i}^{2}\)).
-As the number of studies \(k\) increases, the estimation of \(\mu_{\theta}\) and \(\tau^{2}\) becomes more precise \color{red} (Blázquez-Rincón, Sánchez-Meca, Botella, \& Suero, 2023; Rubio-Aparicio et al., 2018) \color{black}.
+As the number of studies \(k\) increases, the estimation of \(\mu_{\theta}\) and \(\tau^{2}\) becomes more precise (Blázquez-Rincón, Sánchez-Meca, Botella, \& Suero, 2023; Rubio-Aparicio et al., 2018) .
 Additionally, as the sample size of each study decreases, each \(\delta_{i}\) will be estimated with higher precision but as long as \(\tau^{2} \neq 0\) there will be variability among effect sizes.
@@
 
-\caption{\color{red} Forest plots of two simulated \emph{random-effects model}s.
+\caption{Forest plots of two simulated \emph{random-effects model}s.
 On the left, the simulated model has \(nt,nc = 30\) for each included study, while on the right, the sample size for each study is \(nt,nc = 500\).
@@
 The estimated average effect is similar between the two scenarios regarding average effect and precision.
-However, compared to the \emph{equal-effects} simulation, increasing the sample size of primary studies only affects the precision without reducing the true heterogeneity (i.e., \(\tau^{2} \neq 0\)).
-\color{black}}\label{fig:random-effects-high-low-precision}
+However, compared to the \emph{equal-effects} simulation, increasing the sample size of primary studies only affects the precision without reducing the true heterogeneity (i.e., \(\tau^{2} \neq 0\)).}\label{fig:random-effects-high-low-precision}
 \end{figure}
@@
 The relationship between the sampling error and the heterogeneity can be expressed using the \(I^{2}\) statistics (e.g., Higgins \& Thompson, 2002) that is the percentage of the total variability \(\tau^{2} + \tilde{v}\) that is attributable to real heterogeneity between studies (\(\tau^{2}\)).
-\color{red} In Equation \eqref{eq:isquared}, \(\tilde{v}\) is the \emph{typical} within-study sampling variability (i.e., a summary statistics of sampling variances of included studies) as proposed by Higgins and Thompson (2002)\footnote{See \url{https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate\#fn__1} for an overview of the \(I^2\) statistics also for complex models.
+In Equation \eqref{eq:isquared}, \(\tilde{v}\) is the \emph{typical} within-study sampling variability (i.e., a summary statistics of sampling variances of included studies) as proposed by Higgins and Thompson (2002)\footnote{See \url{https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate\#fn__1} for an overview of the \(I^2\) statistics also for complex models.
 There are also other proposals to compute the \(\tilde{v}\) statistics (see Takkouche, Cadarso-Suárez, \& Spiegelman, 1999; Takkouche, Khudyakov, Costa-Bouzas, \& Spiegelman, 2013)} and implemented in Equation \eqref{eq:typical-v} with \(w_i = 1/\sigma_{\epsilon_i}^{2}\) and \(k\) is the number of studies.
-\color{black}
 
@@
 
-\color{red}
-
 \begin{align}
@@
 \end{align}
-\color{black}
-
-From Equation \eqref{eq:isquared} and Figure \ref{fig:random-effects-high-low-precision} is clear that if each included study has a considerable sample size \color{red} the sampling variability (\(\tilde{v}\)) will be reduced, and the total variability will be mainly driven by real heterogeneity (\(\tau^2\)) \color{black}.
-This is the crucial difference between the \textcolor{red}{equal-effects} and the random-effects model (see also Borenstein et al., 2009, pp.
+
+From Equation \eqref{eq:isquared} and Figure \ref{fig:random-effects-high-low-precision} is clear that if each included study has a considerable sample size the sampling variability (\(\tilde{v}\)) will be reduced, and the total variability will be mainly driven by real heterogeneity (\(\tau^2\)) .
+This is the crucial difference between the equal-effects and the random-effects model (see also Borenstein et al., 2009, pp.
 117--122).
-
-\color{red}
 
@@
 The only caveat is fixing the \(\tilde{v}\).
-When the sample size of each study is the same\footnote{\color{red} The assumption of studies with homogeneous sample sizes is commonly used to calculate the statistical power \color{black}(see Borenstein et al., 2009, Chapter 29)}, \(\tilde{v} = \sigma_{\epsilon_i}^{2}\).
+When the sample size of each study is the same\footnote{ The assumption of studies with homogeneous sample sizes is commonly used to calculate the statistical power (see Borenstein et al., 2009, Chapter 29)}, \(\tilde{v} = \sigma_{\epsilon_i}^{2}\).
 In the other case, \(\tilde{v}\) needs to be calculated from sampling variances as reported in Higgins and Thompson (2002) and cannot be easily fixed a priori.
 With the assumption of homogeneous sample size across studies\footnote{In case of heterogeneous sample sizes generated from a probability distribution (e.g., Gaussian) the \emph{typical} sampling variance calculated using the expected value of the distribution can be used to somehow fix the (average) \(I^2\) (see the supplementary materials for a simulated example)}, we can solve Equation \eqref{eq:isquared} for \(\tau^{2}\) obtaining the heterogeneity value associated with a certain \(I^{2}\) as reported in Equation \eqref{eq:tau2-from-isquared}.
-\color{black}
-
-\color{red}
 
@@
 For example van Erp and colleagues (2017) estimate an empirical \(\tau^2\) distribution across several published meta-analyses that can be used to simulate a plausible scenario.
-\color{black}
 
@@
 The dotted line is the average effect (i.e., random-effects model without moderators).
-The effect size differs between conditions A and B, including the \emph{condition} moderator explaining part of the total heterogeneity (red plus green segments).
-The green segments depict the explained heterogeneity, and the red segments the residual (unexplained) heterogeneity.
-\color{red} The red squares are simulated observed effect sizes from the sampling distributions.
-\color{black}}\label{fig:img-metaregression-bin}
+The effect size differs between conditions A and B, including the \emph{condition} moderator explaining part of the total heterogeneity (pink plus green segments).
+The green segments depict the explained heterogeneity, and the pink segments the residual (unexplained) heterogeneity.
+The pink squares are simulated observed effect sizes from the sampling distributions.}\label{fig:img-metaregression-bin}
 \end{figure}
@@
 In regression terms (see equation \eqref{eq:meta-regression}), we could use a dummy variable (\(X_{1}\)) that takes the value of 0 for the \emph{lab-based task} (\emph{L}) and a value of 1 for an \emph{online task} (\emph{O})\footnote{The \emph{dummy coding} (also known as treatment coding) is the default in R and \texttt{metafor} but other coding schemes could be used (see Schad, Vasishth, Hohenstein, \& Kliegl, 2020 for an overview of contrast coding schemes).
-\color{red} As in standard linear regression, the contrast coding of categorical factors or the presence/absence of the intercept influences the interpretation of estimated model parameters and omnibus meta-regression tests (see \url{https://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept} for a detailed discussion)} \color{black}.
-Now we fix \(\beta_{1}\) to be the \emph{lab vs.~online effect} (i.e., the expected mean difference between the two groups of studies) and the product between \color{red}\(\beta_{1}X_{1_i}\) \color{black} will consider the \emph{lab vs.~online effect}.
+As in standard linear regression, the contrast coding of categorical factors or the presence/absence of the intercept influences the interpretation of estimated model parameters and omnibus meta-regression tests (see \url{https://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept} for a detailed discussion)} .
+Now we fix \(\beta_{1}\) to be the \emph{lab vs.~online effect} (i.e., the expected mean difference between the two groups of studies) and the product between \(\beta_{1}X_{1_i}\) will consider the \emph{lab vs.~online effect}.
 Crucially, despite \(\tau^{2}\) is still the heterogeneity between effect sizes, now we need to fix \(\tau^{2}\) considering that we included a moderator.
-In other terms, \color{red} \(\delta_i \sim \mathcal{N}(0, \tau_{r}^{2})\) \color{black} where \(\tau^2_r\) is the residual heterogeneity after including the moderator.
+In other terms, \(\delta_i \sim \mathcal{N}(0, \tau_{r}^{2})\) where \(\tau^2_r\) is the residual heterogeneity after including the moderator.
 We can describe our model using equation \eqref{eq:meta-regression-equations} according to the value of \(X_{1_i}\) (L = lab-based experiments, O = online experiments).
@@
 Then the \(\beta_1\) parameter represents the estimated difference in \(y\) between the values of \(X_1\) (i.e., lab-based vs online experiments).
-As said before, \color{red} \(\tau^2_r\) \color{black} is now the residual heterogeneity that is interpreted as the variability between effect sizes after controlling for the moderator \(X_1\).
+As said before, \(\tau^2_r\) is now the residual heterogeneity that is interpreted as the variability between effect sizes after controlling for the moderator \(X_1\).
 
@@
 
-\caption{\label{tab:res-meta-reg-dummy}\color{red} Summary of the \emph{random-effect} model with a categorical predictor (lab vs online experiments).
+\caption{\label{tab:res-meta-reg-dummy}Summary of the \emph{random-effect} model with a categorical predictor (lab vs online experiments).
 The \texttt{intercept} is the average effect for lab-based experiments and \texttt{exponline} is the difference between lab-based and online experiments.
 The \(R^2\) is the percentage of explained heterogeneity and \(\tau^2_r\) is the estimated residual heterogeneity.
-Other parameters are the same as the standard \emph{random-effect} model (see Table \ref{tab:res-random-effects}).
-\color{black}}
+Other parameters are the same as the standard \emph{random-effect} model (see Table \ref{tab:res-random-effects}).}
 \centering
@@
 A first strategy could be to use values estimated from the literature.
-Another approach consists in setting up the model and simulating several expected \color{red} \(y_i\) \color{black} and calculating the range of simulated values.
+Another approach consists in setting up the model and simulating several expected \(y_i\) and calculating the range of simulated values.
 A third possibility is fixing the proportion of explained heterogeneity calculating the \(\beta_1\) value accordingly.
@@
 
-\caption{\color{red} Scatter plots with marginal histograms for the range of simulated \(y_i\) with two \(\beta_1\) values.
+\caption{Scatter plots with marginal histograms for the range of simulated \(y_i\) with two \(\beta_1\) values.
 The x-axis depicts the average age of simulated studies and the y-axis depicts the simulated effect size.
 On the left, the majority of simulated values range between -1.5 and 1.5 thus \(\beta_1 = 0.1\) can be considered a plausible value.
-On the right, (\(\beta = 0.7\)) values range between -10 to 10 that despite theoretically possible can be considered highly implausible values in real meta-analyses for psychological data.
-\color{black}}\label{fig:plot-meta-reg-plausible-implausible}
+On the right, (\(\beta = 0.7\)) values range between -10 to 10 that despite theoretically possible can be considered highly implausible values in real meta-analyses for psychological data.}\label{fig:plot-meta-reg-plausible-implausible}
 \end{figure}
@@
 
-\caption{\label{tab:res-meta-reg-numerical-r2}\color{red} Summary of the \emph{random-effects} model fixing the \(R^{2}\) value.
+\caption{\label{tab:res-meta-reg-numerical-r2}Summary of the \emph{random-effects} model fixing the \(R^{2}\) value.
 The \texttt{intercept} is the effect size for the average \texttt{age} (given that age is mean-centered).
 The \texttt{age0} parameter is the slope between \texttt{age} and the effect size interpreted as an increase in effect size for a unit increase in the average age.
-Other parameters are the same as described in Table \ref{tab:res-random-effects} and \ref{tab:res-meta-reg-dummy}.
-\color{black}}
+Other parameters are the same as described in Table \ref{tab:res-random-effects} and \ref{tab:res-meta-reg-dummy}.}
 \centering
@@
 However, one of the most critical applications is estimating the power of a specific statistical model.
-\color{red} One of the purposes of power analysis by simulations is to estimate the required number of studies to detect an hypothetical effect size when planning a meta-analysis.
+One of the purposes of power analysis by simulations is to estimate the required number of studies to detect an hypothetical effect size when planning a meta-analysis.
 At the same time, the meta-analysis that we presented (\emph{two-level} equal or random-effects model) can be considered as a multi-lab (e.g., Klein et al., 2018) study where experiments are planned and not collected from the literature.
 We can use the same simulation approach to optimize the number of participants and studies when planning a multi-lab project.
-\color{black}
-
-As explained in the introduction, there are several approaches and tools to estimate the power of \emph{equal} and \emph{random-effects models} (see Borenstein et al., 2009, Chapter 29; Harrer, Cuijpers, \& Ebert, 2019, Chapter 14).
+
+As explained in the introduction, there are several approaches and tools to estimate the power of \emph{equal} and \emph{random-effects models} (see Borenstein et al., 2009, Chapter 29; Harrer et al., 2019, Chapter 14).
 These methods are easy to implement but made strong assumptions, such as the homogeneity of sample size, and did not consider the uncertainty in estimating \(\tau^{2}\).
 Jackson and colleagues (2017) partially solve the issue by developing an interesting method that takes into account the uncertainty in estimating \(\tau^{2}\) without using simulations.
-\textcolor{red}{However, complex simulation scenarios can be handled only using Monte Carlo methods.}
+However, complex simulation scenarios can be handled only using Monte Carlo methods.
 
@@
 The x-axis depicts the number of studies (\(k\)), and the y-axis the estimated power.
-The red dotted line is the 80\% power level, usually considered a good value for power analysis.}\label{fig:plot-power-analysis}
+The pink dotted line is the 80\% power level, usually considered a good value for power analysis.}\label{fig:plot-power-analysis}
 \end{figure}
@@
 The present work introduced the basic concepts of the meta-analysis regarding equal-effects, random-effects and meta-regression models with a simulation-based approach.
-\textcolor{red}{We believe the presented examples are useful to implement alternative or more complex models}.
-For example, the \texttt{sim\_study()} function can be easily modified to simulate another effect sizes index, \color{red} such as correlations or odds ratios.
-\color{black} In addition, more complex models, such as \emph{multivariate} or \emph{multilevel} models, can be simulated following a similar approach (see \textcolor{red}{the} supplementary materials).
-The \color{red} \emph{multilevel} (e.g., \emph{three-level}) \color{black} model estimates another heterogeneity component representing the variability of multiple \textcolor{red}{independent} effect sizes \textcolor{red}{within} the same study.
-Similarly, the \emph{multivariate} model \textcolor{red}{includes} the correlation between multiple outcomes and the correlation between sampling errors.
+We believe the presented examples are useful to implement alternative or more complex models.
+For example, the \texttt{sim\_study()} function can be easily modified to simulate another effect sizes index, such as correlations or odds ratios.
+In addition, more complex models, such as \emph{multivariate} or \emph{multilevel} models, can be simulated following a similar approach (see the supplementary materials).
+The \emph{multilevel} (e.g., \emph{three-level}) model estimates another heterogeneity component representing the variability of multiple independent effect sizes within the same study.
+Similarly, the \emph{multivariate} model includes the correlation between multiple outcomes and the correlation between sampling errors.
 
 The present work did have a few limitations.
-\textcolor{red}{Firstly}, we only introduced basic concepts about meta-analysis and Monte Carlo simulations, while setting up complex simulations requires more knowledge and complexity of the simulation setup.
+Firstly, we only introduced basic concepts about meta-analysis and Monte Carlo simulations, while setting up complex simulations requires more knowledge and complexity of the simulation setup.
 We decided to give the foundations to understand meta-analyses with a simulation approach because more complex models are still based on the same principles.
@@
 The downside concerns the efficiency and scalability of the simulation setup.
-For large-scale simulations (e.g., many conditions, iterations, or complex models), simulating from aggregated statistics is probably more efficient (see Heuvel, Almalik, \& Zhan, 2020 for an example) to improve the \textcolor{red}{simulation efficency}\footnote{See \url{https://www.jepusto.com/simulating-correlated-smds/} for a very clear example of the participant-level vs aggregated data simulation approaches}.
+For large-scale simulations (e.g., many conditions, iterations, or complex models), simulating from aggregated statistics is probably more efficient (see Heuvel, Almalik, \& Zhan, 2020 for an example) to improve the simulation efficency\footnote{See \url{https://www.jepusto.com/simulating-correlated-smds/} for a very clear example of the participant-level vs aggregated data simulation approaches}.
 
@@
 We thank the \emph{\href{https://stat.ethz.ch/mailman/listinfo/r-sig-meta-analysis}{R-sig-meta-analysis}} mailing-list.
-\textcolor{red}{Their} suggestions and clarifications significantly improved \textcolor{red}{the} simulations approach and the R code.
+Their suggestions and clarifications significantly improved the simulations approach and the R code.
 
@@
 
-The code to reproduce simulations, figures and tables can be found on Open Science Framework (\url{https://osf.io/54djn/}) \color{red} and Github (\url{https://github.com/shared-research/simulating-meta-analysis}).
-\color{black}
+The code to reproduce simulations, figures and tables can be found on Open Science Framework (\url{https://osf.io/54djn/}) and Github (\url{https://github.com/shared-research/simulating-meta-analysis}).
 
@@
 
-The supplementary materials can be also found on the Open Science Framework repository (\url{https://osf.io/54djn/}) \color{red} and Github (\url{https://github.com/shared-research/simulating-meta-analysis}).
-\color{black}
+The supplementary materials can be also found on the Open Science Framework repository (\url{https://osf.io/54djn/}) and Github (\url{https://github.com/shared-research/simulating-meta-analysis}).
 

### 2023-09-06 documents/paper/paper.Rmd
better plotting settings

@@
                       out.width = "80%",
-                      fig.width = 9,
+                      fig.width = 7,
                       size = "scriptsize",
@@
 })
+
+# base r plot size
+knitr::opts_knit$set(global.par = TRUE)
+mcex <- 1.2
+par(cex.lab = 1.5, cex.axis = 1.5, cex.main = 1.5)
 ```
@@
 # theme for plots
-theme_paper <- function(base_size = 15){
-  ltxplot::theme_latex(base_size = base_size)
+theme_paper <- function(base_size = 18){
+  ltxplot::theme_latex(base_size = base_size) +
+    theme(plot.title = element_text(size = 22))
 }

### 2023-09-06 scripts/r-img.R
better plotting settings

@@ -11,8 +11,9 @@ ltxplot::load_theme_ltx()
 
 # Functions ---------------------------------------------------------------
 
-theme_rfig <- function(size = 15){
-  ltxplot::theme_latex(base_size = size)
+theme_rfig <- function(size = 18){
+  ltxplot::theme_latex(base_size = size) +
+    theme(plot.title = element_text(size = 22))
 }
 
 # color-blind-friendly palettes see http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
@@ -197,20 +198,20 @@ plot_metareg_bin <- ggplot(binary_metareg,
   theme(axis.ticks.x = element_blank()) +
   ylab("Study") +
   geom_label(aes(x = cond_mean, y = 7, label = note), 
-             size = 6, label.size = NA) +
+             size = 6, label.size = NA, family = "lmroman") +
   theme(axis.text.x = element_blank()) +
   theme_rfig() +
   xlab(latex2exp::TeX("$y_i$")) +
   ggtitle("Categorical Predictor") +
   xlim(xlim) +
-  annotate("text", x = 1, y = 1.5, 
+  annotate("text", x = 0.7, y = 1.6666666666666667, 
            label = TeX("\\textbf{Residual $\\tau^2$}"), 
            parse = TRUE, color = cb_red, 
-           family = "bold", size = 8) +
-  annotate("text", x = 1, y = 2, 
+           family = "lmroman", size = 7.5, hjust = 0) +
+  annotate("text", x = 0.7, y = 2.5, 
            label = TeX("\\textbf{Explained $\\tau^2$}"), 
            parse = TRUE, color = cb_green, 
-           family = "bold", size = 8) +
+           family = "lmroman", size = 7.5, hjust = 0) +
   geom_point(aes(x = obs, y = id), color = cb_red, shape = 15, size = 2.5)
 
 # Metaregression with numerical predictor ---------------------------------
@@ -261,14 +262,14 @@ plot_metareg_cont <- ggplot(cont_metareg, aes(x = x, y = yi)) +
   theme_rfig() +
   ylab(latex2exp::TeX("$y_i$")) +
   ggtitle("Numerical Predictor") +
-  annotate("text", x = 6.5, y = 0.25,
+  annotate("text", x = 5, y = 0.20,
            label = TeX("\\textbf{Residual $\\tau^2$}"), 
-           parse = TRUE, color = cb_red, size = 8,
-           family = "lmroman") +
-  annotate("text", x = 6.5, y = 0.3, 
[…]

### 2023-09-06 documents/paper/paper.tex
updating

@@
 {}{\PatchFailed}
-\keywords{meta-analysis, monte carlo simulations, power analysis\newline\indent Word count: 7782}
+\keywords{meta-analysis, monte carlo simulations, power analysis\newline\indent Word count: 7866}
 \usepackage{csquotes}
@@
 Readers without experience in meta-analysis could benefit both from the theoretical introduction and the simulation approach.
-However, for a better and extensive overview of meta-analysis topics the reader could refer to Harrer, Cuijpers, and Ebert (2019), and Borenstein et al.
-(2009).
+However, for a better and extensive overview of meta-analysis topics the reader could refer to meta-analysis textbooks (Borenstein et al., 2009; e.g., Harrer, Cuijpers, \& Ebert, 2019).
 

### 2023-09-11 documents/paper/paper.Rmd
fixing intro and wordcount

@@
 keywords          : "meta-analysis, monte carlo simulations, power analysis"
-wordcount         : "7866"
+wordcount         : 7869
 bibliography      : "../files/references.bib"
@@
 
-The aim of the tutorial is not providing a complete theoretical introduction to meta-analysis but presenting the core topics using a simulation-based approach.
-Readers expert in conducting meta-analyses could benefit more from the proposed approach in terms of simulation setup and coding strategies.
-Readers without experience in meta-analysis could benefit both from the theoretical introduction and the simulation approach.
-However, for a better and extensive overview of meta-analysis topics the reader could refer to meta-analysis textbooks [e.g., @Harrer2019-rl; @Borenstein2009-mo].
+The aim of the tutorial is not to provide a complete theoretical introduction to meta-analysis but rather to present core topics using a simulation-based approach.
+Readers experienced in conducting meta-analyses can benefit from the proposed approach in terms of simulation setup and coding strategies.
+Readers without prior experience in meta-analysis can benefit both from the theoretical introduction and the simulation approach.
+However, for a more comprehensive overview of meta-analysis topics the reader may refer to meta-analysis textbooks [e.g., @Harrer2019-rl; @Borenstein2009-mo]
 

### 2023-09-11 documents/paper/paper.tex
updating

--- 2023-09-06 documents/paper/paper.tex
@@
 {}{\PatchFailed}
-\keywords{meta-analysis, monte carlo simulations, power analysis\newline\indent Word count: 7782}
+\keywords{meta-analysis, monte carlo simulations, power analysis\newline\indent Word count: 7869}
 \usepackage{csquotes}
@@
 
-The aim of the tutorial is not providing a complete theoretical introduction to meta-analysis but presenting the core topics using a simulation-based approach.
-Readers expert in conducting meta-analyses could benefit more from the proposed approach in terms of simulation setup and coding strategies.
-Readers without experience in meta-analysis could benefit both from the theoretical introduction and the simulation approach.
-However, for a better and extensive overview of meta-analysis topics the reader could refer to Harrer, Cuijpers, and Ebert (2019), and Borenstein et al.
-(2009).
+The aim of the tutorial is not to provide a complete theoretical introduction to meta-analysis but rather to present core topics using a simulation-based approach.
+Readers experienced in conducting meta-analyses can benefit from the proposed approach in terms of simulation setup and coding strategies.
+Readers without prior experience in meta-analysis can benefit both from the theoretical introduction and the simulation approach.
+However, for a more comprehensive overview of meta-analysis topics the reader may refer to meta-analysis textbooks (Borenstein et al., 2009; e.g., Harrer, Cuijpers, \& Ebert, 2019)
 

### 2023-09-11 README.Rmd
adding table with links

@@
 
-- Paper [PDF](documents/paper/paper.pdf)
-- Supplementary Materials [PDF](documents/supplementary/supplementary.pdf)
+|               |                                                  |                                                  |
+|---------------|--------------------------------------------------|--------------------------------------------------|
+| **Paper**         | [pdf](documents/paper/paper.pdf)                 | [rmd](documents/paper/paper.Rmd)                 |
+| **Supplementary** | [pdf](documents/supplementary/supplementary.pdf) | [rmd](documents/supplementary/supplementary.Rmd) |
 

### 2023-09-11 documents/paper/supplementary.Rmd
better folder structure

--- 2023-07-24 documents/supplementary/supplementary.Rmd
@@
 We can use the `sim_T()` function that given the number of studies $k$, outcomes $p$, the vector of $\tau^2$ and the correlation $\rho$ generates the random-effect adjustments^[The function allow to specify heterogeneous $\tau^2$ values where the CS structure assume the same $\tau^2$ for each outcome.
-Using a fixed $\rho$ and different $\tau^2$ will create what `rma.mv()` calls a *heteroscedastic compound symmetric structure* thus a matrix where the correlation is the same between outcomes and each outcome has a different $\tau^2$ value].
+Using a fixed $\rho$ and different $\tau^2$ will create what `rma.mv()` calls a *heteroscedastic compound symmetry structure* thus a matrix where the correlation is the same between outcomes and each outcome has a different $\tau^2$ value].
 

### 2023-09-11 README.Rmd
fixing link after folders

--- 2023-07-18 README.Rmd
@@
 
-- Paper [PDF](documents/paper/paper.pdf)
-- Supplementary Materials [PDF](documents/supplementary/supplementary.pdf)
+| **Document** | **Rendered** | **Source** |
+|---|---|---|
+| Paper | [pdf](documents/paper/paper.pdf) | [rmd](documents/paper/paper.Rmd) |
+| Supplementary Materials | [pdf](documents/supplementary/supplementary.pdf) | [rmd](documents/supplementary/supplementary.Rmd) |
 

### 2023-09-11 README.Rmd
fixing suppl link

--- 2023-07-18 README.Rmd
@@
 
-- Paper [PDF](documents/paper/paper.pdf)
-- Supplementary Materials [PDF](documents/supplementary/supplementary.pdf)
+| **Document** | **Rendered** | **Source** |
+|---|---|---|
+| Paper | [pdf](documents/paper/paper.pdf) | [rmd](documents/paper/paper.Rmd) |
+| Supplementary Materials | [pdf](documents/paper/supplementary.pdf) | [rmd](documents/paper/supplementary.Rmd) |
 

### 2023-09-11 README.Rmd
renaming files and links suppl

@@
 | Paper | [pdf](documents/paper/paper.pdf) | [rmd](documents/paper/paper.Rmd) |
-| Supplementary Materials | [pdf](documents/paper/supplementary.pdf) | [rmd](documents/paper/supplementary.Rmd) |
+| Supplementary Materials | [pdf](documents/paper/suppl.pdf) | [rmd](documents/paper/suppl.Rmd) |
 

### 2023-09-11 README.Rmd
fixing typo and better structure

@@
   - `scripts\r-img.R`: create the paper figures.
-Figures that are not created in this script using `ggplot2` are created with the open-source program [*Inkscape*](https://inkscape.org/) and contained in the `../img/` folders.
+Figures that are not created in this script using `ggplot2` are created with the open-source program [*Inkscape*](https://inkscape.org/) and contained in the `**/img/` folder.
 - `objects`: contains the result of objects created by the `scripts/` scripts.

### 2023-09-11 documents/output/paper.tex
fixing typo and better structure

--- 2023-09-11 documents/paper/paper.tex
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/img-equal-vs-random-1}
+{\centering \includegraphics[width=0.8\linewidth]{C:/Users/user/Documents/work/shared-projects/simulating-meta-analysis/documents/output/paper_files/figure-latex/img-equal-vs-random-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/equal-effects-high-low-precision-1}
+{\centering \includegraphics[width=0.8\linewidth]{C:/Users/user/Documents/work/shared-projects/simulating-meta-analysis/documents/output/paper_files/figure-latex/equal-effects-high-low-precision-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/random-effects-high-low-precision-1}
+{\centering \includegraphics[width=0.8\linewidth]{C:/Users/user/Documents/work/shared-projects/simulating-meta-analysis/documents/output/paper_files/figure-latex/random-effects-high-low-precision-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/img-metaregression-bin-1}
+{\centering \includegraphics[width=0.8\linewidth]{C:/Users/user/Documents/work/shared-projects/simulating-meta-analysis/documents/output/paper_files/figure-latex/img-metaregression-bin-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/img-metaregression-num-1}
+{\centering \includegraphics[width=0.8\linewidth]{C:/Users/user/Documents/work/shared-projects/simulating-meta-analysis/documents/output/paper_files/figure-latex/img-metaregression-num-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/plot-meta-reg-plausible-implausible-1}
+{\centering \includegraphics[width=0.8\linewidth]{C:/Users/user/Documents/work/shared-projects/simulating-meta-analysis/documents/output/paper_files/figure-latex/plot-meta-reg-plausible-implausible-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/plot-meta-reg-numerical-r2-1}
+{\centering \includegraphics[width=0.8\linewidth]{C:/Users/user/Documents/work/shared-projects/simulating-meta-analysis/documents/output/paper_files/figure-latex/plot-meta-reg-numerical-r2-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/plot-power-analysis-1}
+{\centering \includegraphics[width=0.8\linewidth]{C:/Users/user/Documents/work/shared-projects/simulating-meta-analysis/documents/output/paper_files/figure-latex/plot-power-analysis-1}
 

### 2023-09-11 documents/render-docs.R

@@ -1,0 +1,12 @@
+# render documents
+
+# paper
+
+rmarkdown::render(input = "documents/paper/paper.Rmd", 
+                  output_dir = "documents/output")
+
+# supplementary
+
+rmarkdown::render(input = "documents/paper/suppl.Rmd", 
+                  output_dir = "documents/output",
+                  output_file = "suppl.pdf")

### 2023-09-11 README.Rmd
fix document paths

@@
 |---|---|---|
-| Paper | [pdf](documents/paper/paper.pdf) | [rmd](documents/paper/paper.Rmd) |
-| Supplementary Materials | [pdf](documents/paper/suppl.pdf) | [rmd](documents/paper/suppl.Rmd) |
+| Paper | [pdf](documents/output/paper.pdf) | [rmd](documents/paper/paper.Rmd) |
+| Supplementary Materials | [pdf](documents/output/suppl.pdf) | [rmd](documents/paper/suppl.Rmd) |
 

### 2023-09-26 R/sim-utils.R
fixing wrong variable

@@ -24,7 +24,7 @@ sim_study <- function(es, nt, nc = NULL, aggregate = TRUE){
   yi <- (mean(yt) - mean(yc))
   
   # sampling variance
-  vi <- var(yt)/nt + var(yt)/nc
+  vi <- var(yt)/nt + var(yc)/nc
   
   if(!aggregate){
     # return raw data

### 2023-09-26 documents/render-docs.R
cleaning extra steps

@@ -10,3 +10,20 @@ rmarkdown::render(input = "documents/paper/paper.Rmd",
 rmarkdown::render(input = "documents/paper/suppl.Rmd", 
                   output_dir = "documents/output",
                   output_file = "suppl.pdf")
+
+# preprint
+
+create_preprint()
+
+# submission
+
+# ## copy the paper
+# fs::file_copy("documents/paper/paper.Rmd",
+#               "documents/submission/paper.Rmd")
+# 
+# ## copy the cache
+# fs::dir_copy("documents/paper/paper_cache/",
+#              "documents/submission/")
+
+rmarkdown::render(input = "documents/submission/paper.Rmd",
+                  params = list(submission = TRUE))

### 2023-09-26 documents/render-docs.R
cleaning

@@ -14,16 +14,3 @@ rmarkdown::render(input = "documents/paper/suppl.Rmd",
 # preprint
 
 create_preprint()
-
-# submission
-
-# ## copy the paper
-# fs::file_copy("documents/paper/paper.Rmd",
-#               "documents/submission/paper.Rmd")
-# 
-# ## copy the cache
-# fs::dir_copy("documents/paper/paper_cache/",
-#              "documents/submission/")
-
-rmarkdown::render(input = "documents/submission/paper.Rmd",
-                  params = list(submission = TRUE))

### 2023-09-26 documents/output/paper.tex
final fix

@@
 \fi
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
   \fi
   % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLBlock}[1]{\hfill\break#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 
-\hypertarget{meta-analysis-introduction}{%
-\subsection{Meta-analysis introduction}\label{meta-analysis-introduction}}
+\subsection{Meta-analysis introduction}\label{meta-analysis-introduction}
 
@@
 
-\hypertarget{effect-size-and-variance}{%
-\subsubsection{Effect size and variance}\label{effect-size-and-variance}}
+\subsubsection{Effect size and variance}\label{effect-size-and-variance}
 
@@
 
-\hypertarget{equal-vs-fixed-vs-random-effects-model}{%
-\subsubsection{Equal vs fixed vs random-effects model}\label{equal-vs-fixed-vs-random-effects-model}}
+\subsubsection{Equal vs fixed vs random-effects model}\label{equal-vs-fixed-vs-random-effects-model}
 
@@
 
-\hypertarget{simulation}{%
-\section{Simulation}\label{simulation}}
-
-\hypertarget{monte-carlo-simulations}{%
-\subsection{Monte Carlo simulations}\label{monte-carlo-simulations}}
+\section{Simulation}\label{simulation}
+
+\subsection{Monte Carlo simulations}\label{monte-carlo-simulations}
 
@@
 
-\hypertarget{simulation-setup}{%
-\subsection{Simulation setup}\label{simulation-setup}}
+\subsection{Simulation setup}\label{simulation-setup}
 
@@
   \CommentTok{\# sampling variance}
-\NormalTok{  vi }\OtherTok{\textless{}{-}} \FunctionTok{var}\NormalTok{(yt)}\SpecialCharTok{/}\NormalTok{nt }\SpecialCharTok{+} \FunctionTok{var}\NormalTok{(yt)}\SpecialCharTok{/}\NormalTok{nc}
+\NormalTok{  vi }\OtherTok{\textless{}{-}} \FunctionTok{var}\NormalTok{(yt)}\SpecialCharTok{/}\NormalTok{nt }\SpecialCharTok{+} \FunctionTok{var}\NormalTok{(yc)}\SpecialCharTok{/}\NormalTok{nc}
 
@@
 
-\hypertarget{equal-effects-model}{%
-\subsection{Equal-effects model}\label{equal-effects-model}}
+\subsection{Equal-effects model}\label{equal-effects-model}
 
@@
 \midrule
-overall & 0.347 (SE = 0.056) & $[0.237, 0.456]$ & 6.223 & < 0.001\\
+overall & 0.339 (SE = 0.057) & $[0.228, 0.451]$ & 5.959 & < 0.001\\
 \bottomrule
@@
 
-\hypertarget{re-model}{%
-\subsection{Random-effects model}\label{re-model}}
+\subsection{Random-effects model}\label{re-model}
 
@@
 \midrule
-overall & 0.394 (SE = 0.110) & $[0.179, 0.609]$ & 3.596 & < 0.001\\
+overall & 0.387 (SE = 0.107) & $[0.177, 0.597]$ & 3.616 & < 0.001\\
 \bottomrule
 \multicolumn{5}{l}{\textsuperscript{} $k = 30$}\\
-\multicolumn{5}{l}{\textsuperscript{} $\tau^2 = 0.262$ ($SE = 0.095$)}\\
-\multicolumn{5}{l}{\textsuperscript{} $I^2 = 73.480\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $\tau^2 = 0.246$ ($SE = 0.090$)}\\
+\multicolumn{5}{l}{\textsuperscript{} $I^2 = 71.925\%$}\\
 \end{tabular}
@@
 \midrule
-overall & 0.300 (SE = 0.071) & $[0.160, 0.440]$ & 4.201 & < 0.001\\
+overall & 0.305 (SE = 0.072) & $[0.164, 0.447]$ & 4.226 & < 0.001\\
 \bottomrule
 \multicolumn{5}{l}{\textsuperscript{} $k = 30$}\\
-\multicolumn{5}{l}{\textsuperscript{} $\tau^2 = 0.087$ ($SE = 0.040$)}\\
-\multicolumn{5}{l}{\textsuperscript{} $I^2 = 57.904\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $\tau^2 = 0.091$ ($SE = 0.041$)}\\
+\multicolumn{5}{l}{\textsuperscript{} $I^2 = 58.479\%$}\\
 \end{tabular}
@@
 
-\hypertarget{metareg}{%
-\subsection{Meta-regression}\label{metareg}}
+\subsection{Meta-regression}\label{metareg}
 
@@
 
-\hypertarget{meta-regression-with-a-categorical-moderator}{%
-\subsubsection{Meta-regression with a categorical moderator}\label{meta-regression-with-a-categorical-moderator}}
+\subsubsection{Meta-regression with a categorical moderator}\label{meta-regression-with-a-categorical-moderator}
 
@@
 \midrule
-intercept & 0.038 (SE = 0.101) & $[-0.161, 0.236]$ & 0.372 & 0.710\\
-exponline & 0.323 (SE = 0.143) & $[0.043, 0.602]$ & 2.263 & 0.024\\
+intercept & 0.043 (SE = 0.103) & $[-0.158, 0.244]$ & 0.42 & 0.674\\
+exponline & 0.323 (SE = 0.144) & $[0.040, 0.606]$ & 2.24 & 0.025\\
 \bottomrule
 \multicolumn{5}{l}{\textsuperscript{} $k = 30$}\\
-\multicolumn{5}{l}{\textsuperscript{} $\tau^2_r = 0.087$ ($SE = 0.041$)}\\
-\multicolumn{5}{l}{\textsuperscript{} $R^2 = 20.659\%$}\\
-\multicolumn{5}{l}{\textsuperscript{} $I^2 = 57.741\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $\tau^2_r = 0.091$ ($SE = 0.042$)}\\
+\multicolumn{5}{l}{\textsuperscript{} $R^2 = 19.824\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $I^2 = 58.323\%$}\\
 \end{tabular}
@@
 
-\hypertarget{meta-regression-with-a-numerical-moderator}{%
-\subsubsection{Meta-regression with a numerical moderator}\label{meta-regression-with-a-numerical-moderator}}
+\subsubsection{Meta-regression with a numerical moderator}\label{meta-regression-with-a-numerical-moderator}
 
@@
 
-\hypertarget{assessing-the-impact-of-beta_1}{%
-\paragraph{\texorpdfstring{Assessing the impact of \(\beta_1\)}{Assessing the impact of \textbackslash beta\_1}}\label{assessing-the-impact-of-beta_1}}
+\paragraph{\texorpdfstring{Assessing the impact of \(\beta_1\)}{Assessing the impact of \textbackslash beta\_1}}\label{assessing-the-impact-of-beta_1}
 
@@
 
-\hypertarget{simulating-using-r2}{%
-\paragraph{\texorpdfstring{Simulating using \(R^2\)}{Simulating using R\^{}2}}\label{simulating-using-r2}}
+\paragraph{\texorpdfstring{Simulating using \(R^2\)}{Simulating using R\^{}2}}\label{simulating-using-r2}
 
@@
 \midrule
-intercept & 0.339 (SE = 0.051) & $[0.239, 0.439]$ & 6.643 & < 0.001\\
-age0 & 0.168 (SE = 0.051) & $[0.067, 0.268]$ & 3.272 & 0.001\\
+intercept & 0.340 (SE = 0.051) & $[0.240, 0.439]$ & 6.680 & < 0.001\\
+age0 & 0.168 (SE = 0.051) & $[0.068, 0.269]$ & 3.296 & < 0.001\\
 \bottomrule
 \multicolumn{5}{l}{\textsuperscript{} $k = 100$}\\
-\multicolumn{5}{l}{\textsuperscript{} $\tau^2_r = 0.197$ ($SE = 0.037$)}\\
-\multicolumn{5}{l}{\textsuperscript{} $R^2 = 11.029\%$}\\
-\multicolumn{5}{l}{\textsuperscript{} $I^2 = 76.611\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $\tau^2_r = 0.194$ ($SE = 0.037$)}\\
+\multicolumn{5}{l}{\textsuperscript{} $R^2 = 11.422\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $I^2 = 75.417\%$}\\
 \end{tabular}
@@
 
-\hypertarget{power-analysis}{%
-\section{Power Analysis}\label{power-analysis}}
+\section{Power Analysis}\label{power-analysis}
 
@@
 
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
+\section{Conclusions}\label{conclusions}
 
@@
 
-\hypertarget{acknowledgments}{%
-\subsubsection*{Acknowledgments}\label{acknowledgments}}
+\subsubsection*{Acknowledgments}\label{acknowledgments}
 \addcontentsline{toc}{subsubsection}{Acknowledgments}
@@
 
-\hypertarget{conflicts-of-interest}{%
-\subsubsection*{Conflicts of Interest}\label{conflicts-of-interest}}
+\subsubsection*{Conflicts of Interest}\label{conflicts-of-interest}
 \addcontentsline{toc}{subsubsection}{Conflicts of Interest}
@@
 
-\hypertarget{data-materials-and-online-resources}{%
-\subsubsection*{Data, materials, and online resources}\label{data-materials-and-online-resources}}
+\subsubsection*{Data, materials, and online resources}\label{data-materials-and-online-resources}
 \addcontentsline{toc}{subsubsection}{Data, materials, and online resources}
@@
 
-\hypertarget{supplemental-material}{%
-\subsubsection*{Supplemental Material}\label{supplemental-material}}
+\subsubsection*{Supplemental Material}\label{supplemental-material}
 \addcontentsline{toc}{subsubsection}{Supplemental Material}
@@
 
-\hypertarget{prior-versions}{%
-\subsubsection*{Prior versions}\label{prior-versions}}
+\subsubsection*{Prior versions}\label{prior-versions}
 \addcontentsline{toc}{subsubsection}{Prior versions}
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
 
-\hypertarget{refs}{}
+\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-R-papaja}{}}%
+\bibitem[\citeproctext]{ref-R-papaja}
 Aust, F., \& Barth, M. (2022). \emph{Papaja: Prepare american psychological association journal articles with r markdown}. Retrieved from \url{https://github.com/crsh/papaja}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Berkhout2023-fj}{}}%
+\bibitem[\citeproctext]{ref-Berkhout2023-fj}
 Berkhout, S.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Blazquez-Rincon2023-kp}{}}%
+\bibitem[\citeproctext]{ref-Blazquez-Rincon2023-kp}
 Blázquez-Rincón, D., Sánchez-Meca, J., Botella, J., \& Suero, M.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Borenstein2009-mo}{}}%
+\bibitem[\citeproctext]{ref-Borenstein2009-mo}
 Borenstein, M., Hedges, L. V., Higgins, J. P. T., \& Rothstein, H. R. (2009). \emph{Introduction to {Meta-Analysis}}. \url{https://doi.org/10.1002/9780470743386}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Cohen1988-ad}{}}%
+\bibitem[\citeproctext]{ref-Cohen1988-ad}
 Cohen, J. (1988). \emph{Statistical power analysis for the behavioral sciences} (2nd ed.). Routledge. \url{https://doi.org/10.4324/9780203771587}
 
-\leavevmode\vadjust pre{\hypertarget{ref-DeBruine2021-id}{}}%
+\bibitem[\citeproctext]{ref-DeBruine2021-id}
 DeBruine, L.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gelman2006-pc}{}}%
+\bibitem[\citeproctext]{ref-Gelman2006-pc}
 Gelman, A., \& Hill, J. (2006). \emph{Data analysis using regression and {Multilevel/Hierarchical} models}. Cambridge University Press. \url{https://doi.org/10.1017/CBO9780511790942}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gelman2020-tg}{}}%
+\bibitem[\citeproctext]{ref-Gelman2020-tg}
 Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other stories}. Cambridge University Press. \url{https://doi.org/10.1017/9781139161879}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gentle2009-cj}{}}%
+\bibitem[\citeproctext]{ref-Gentle2009-cj}
 Gentle, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gronau2021-do}{}}%
+\bibitem[\citeproctext]{ref-Gronau2021-do}
 Gronau, Q.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Harrer2019-rl}{}}%
+\bibitem[\citeproctext]{ref-Harrer2019-rl}
 Harrer, M., Cuijpers, P., \& Ebert, D. (2019). \emph{Doing {Meta-Analysis} in {R}}. \url{https://doi.org/10.5281/zenodo.2551803}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Harrer2021-bz}{}}%
+\bibitem[\citeproctext]{ref-Harrer2021-bz}
 Harrer, M., Cuijpers, P., Furukawa, T. A., \& Ebert, D. D. (2021). \emph{Doing {Meta-Analysis} with r: A {Hands-On} guide}. CRC Press.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Hedges1981-za}{}}%
+\bibitem[\citeproctext]{ref-Hedges1981-za}
 Hedges, L.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Hedges1989-ip}{}}%
+\bibitem[\citeproctext]{ref-Hedges1989-ip}
 Hedges, L. V. (1989). An unbiased correction for sampling error in validity generalization studies. \emph{J. Appl. Psychol.}, \emph{74}(3), 469--477. \url{https://doi.org/10.1037/0021-9010.74.3.469}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Hedges1998-uo}{}}%
+\bibitem[\citeproctext]{ref-Hedges1998-uo}
 Hedges, L. V., \& Vevea, J. L. (1998). Fixed- and random-effects models in meta-analysis. \emph{Psychol. Methods}, \emph{3}(4), 486--504. \url{https://doi.org/10.1037/1082-989X.3.4.486}
 
-\leavevmode\vadjust pre{\hypertarget{ref-van-den-Heuvel2020-rv}{}}%
+\bibitem[\citeproctext]{ref-van-den-Heuvel2020-rv}
 Heuvel, E.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Higgins2002-fh}{}}%
+\bibitem[\citeproctext]{ref-Higgins2002-fh}
 Higgins, J. P. T., \& Thompson, S. G. (2002). Quantifying heterogeneity in a meta-analysis. \emph{Stat. Med.}, \emph{21}(11), 1539--1558. \url{https://doi.org/10.1002/sim.1186}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Ingalls2011-xt}{}}%
+\bibitem[\citeproctext]{ref-Ingalls2011-xt}
 Ingalls, R. G. (2011). Introduction to simulation. \emph{Proceedings of the 2011 Winter Simulation Conference ({WSC})}, 1374--1388. ieeexplore.ieee.org. \url{https://doi.org/10.1109/WSC.2011.6147858}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Jackson2017-dv}{}}%
+\bibitem[\citeproctext]{ref-Jackson2017-dv}
 Jackson, D., \& Turner, R. (2017). Power analysis for random-effects meta-analysis. \emph{Res. Synth. Methods}, \emph{8}(3), 290--302. \url{https://doi.org/10.1002/jrsm.1240}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Klein2018-he}{}}%
+\bibitem[\citeproctext]{ref-Klein2018-he}
 Klein, R.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Knapp2003-qx}{}}%
+\bibitem[\citeproctext]{ref-Knapp2003-qx}
 Knapp, G., \& Hartung, J. (2003). Improved tests for a random effects meta-regression with a single covariate. \emph{Stat. Med.}, \emph{22}(17), 2693--2710. \url{https://doi.org/10.1002/sim.1482}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Laird1990-ro}{}}%
+\bibitem[\citeproctext]{ref-Laird1990-ro}
 Laird, N.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Lakens2013-eq}{}}%
+\bibitem[\citeproctext]{ref-Lakens2013-eq}
 Lakens, D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Lipsey2001-vp}{}}%
+\bibitem[\citeproctext]{ref-Lipsey2001-vp}
 Lipsey, M. W., \& Wilson, D. B. (2001). Practical meta-analysis. \emph{Applied Social Research Methods Series; Vol 49.}, \emph{247}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Lopez-Lopez2014-it}{}}%
+\bibitem[\citeproctext]{ref-Lopez-Lopez2014-it}
 López-López, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-r-lang}{}}%
+\bibitem[\citeproctext]{ref-r-lang}
 R Core Team. (2022). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Rubio-Aparicio2018-pd}{}}%
+\bibitem[\citeproctext]{ref-Rubio-Aparicio2018-pd}
 Rubio-Aparicio, M., López-López, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Schad2020-ht}{}}%
+\bibitem[\citeproctext]{ref-Schad2020-ht}
 Schad, D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Schmid2022-xe}{}}%
+\bibitem[\citeproctext]{ref-Schmid2022-xe}
 Schmid, C. H., Stijnen, T., \& White, I. R. (2022). \emph{Handbook of {Meta-Analysis}}. Taylor \& Francis Limited.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Takkouche1999-sv}{}}%
+\bibitem[\citeproctext]{ref-Takkouche1999-sv}
 Takkouche, B., Cadarso-Suárez, C., \& Spiegelman, D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Takkouche2013-nd}{}}%
+\bibitem[\citeproctext]{ref-Takkouche2013-nd}
 Takkouche, B., Khudyakov, P., Costa-Bouzas, J., \& Spiegelman, D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Van_Erp2017-cf}{}}%
+\bibitem[\citeproctext]{ref-Van_Erp2017-cf}
 Van Erp, S., Verhagen, J., Grasman, R.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2005-yh}{}}%
+\bibitem[\citeproctext]{ref-Viechtbauer2005-yh}
 Viechtbauer, W.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2007-tq}{}}%
+\bibitem[\citeproctext]{ref-Viechtbauer2007-tq}
 Viechtbauer, W.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2010-xz}{}}%
+\bibitem[\citeproctext]{ref-Viechtbauer2010-xz}
 Viechtbauer, W. (2010). Conducting meta-analyses in {R} with the {metafor} package. \url{https://doi.org/10.18637/jss.v036.i03}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2022-rz}{}}%
+\bibitem[\citeproctext]{ref-Viechtbauer2022-rz}
 Viechtbauer, W., \& López-López, J. A. (2022). Location-scale models for meta-analysis. \emph{Res. Synth. Methods}, \emph{13}(6), 697--715. \url{https://doi.org/10.1002/jrsm.1562}
 
-\leavevmode\vadjust pre{\hypertarget{ref-R-tidyverse}{}}%
+\bibitem[\citeproctext]{ref-R-tidyverse}
 Wickham, H. (2023). \emph{Tidyverse: Easily install and load the tidyverse}. Retrieved from \url{https://CRAN.R-project.org/package=tidyverse}
 
-\leavevmode\vadjust pre{\hypertarget{ref-R-ggplot2}{}}%
+\bibitem[\citeproctext]{ref-R-ggplot2}
 Wickham, H., Chang, W., Henry, L., Pedersen, T.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-R-kableExtra}{}}%
+\bibitem[\citeproctext]{ref-R-kableExtra}
 Zhu, H. (2021). \emph{kableExtra: Construct complex table with kable and pipe syntax}. Retrieved from \url{https://CRAN.R-project.org/package=kableExtra}

### 2023-09-26 documents/paper/paper.Rmd
final fix

@@
 figsintext        : yes
-linenumbers       : "`r params$linenumbers`"
+linenumbers       : no
 draft             : no
@@
     latex_engine: xelatex
-params: 
-  linenumbers: "yes"
-  finalsection: TRUE
 ---
@@
 library(ltxplot)
+devtools::load_all()
 
@@
 ltxplot::load_theme_ltx()
-devtools::load_all()
 seed <- 2023 # general seed for all simulations

### 2023-09-27 documents/output/paper.tex
fixing formula

@@
 
-\NormalTok{v }\OtherTok{\textless{}{-}}\NormalTok{ (n }\SpecialCharTok{+}\NormalTok{ n)}\SpecialCharTok{/}\NormalTok{(n }\SpecialCharTok{*}\NormalTok{ n) }\SpecialCharTok{+}\NormalTok{ mu\_theta}\SpecialCharTok{\^{}}\DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\DecValTok{2} \SpecialCharTok{*}\NormalTok{ (n }\SpecialCharTok{+}\NormalTok{ n }\SpecialCharTok{{-}} \DecValTok{2}\NormalTok{)) }\CommentTok{\# typical within study variance}
+\NormalTok{v }\OtherTok{\textless{}{-}} \DecValTok{1}\SpecialCharTok{/}\NormalTok{n }\SpecialCharTok{+} \DecValTok{1}\SpecialCharTok{/}\NormalTok{n }\CommentTok{\# typical within study variance}
 \NormalTok{tau2 }\OtherTok{\textless{}{-}} \SpecialCharTok{{-}}\NormalTok{((I2}\SpecialCharTok{*}\NormalTok{v)}\SpecialCharTok{/}\NormalTok{(I2 }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{))}
@@
 \midrule
-overall & 0.305 (SE = 0.072) & $[0.164, 0.447]$ & 4.226 & < 0.001\\
+overall & 0.305 (SE = 0.072) & $[0.164, 0.447]$ & 4.239 & < 0.001\\
 \bottomrule
 \multicolumn{5}{l}{\textsuperscript{} $k = 30$}\\
-\multicolumn{5}{l}{\textsuperscript{} $\tau^2 = 0.091$ ($SE = 0.041$)}\\
-\multicolumn{5}{l}{\textsuperscript{} $I^2 = 58.479\%$}\\
+\multicolumn{5}{l}{\textsuperscript{} $\tau^2 = 0.090$ ($SE = 0.041$)}\\
+\multicolumn{5}{l}{\textsuperscript{} $I^2 = 58.183\%$}\\
 \end{tabular}

### 2023-09-27 R/sim-utils.R
fix documentation

@@ -47,9 +47,9 @@ sim_study <- function(es, nt, nc = NULL, aggregate = TRUE){
 #' @export
 #'
 #' @examples
-#' sim <- data.frame(k = 1:10, nt = 30, nc = 30)
-#' sim_studies(sim$k, sim$nt, sim$nc, data = sim)
-#' sim_stud
+#' sim <- data.frame(k = 1:10, es = 0.3, nt = 30, nc = 30)
+#' sim_studies(sim$es, sim$nt, sim$nc, data = sim)
+#' 
 sim_studies <- function(..., data = NULL){
   # ... (dots) are the arguments passed to mapply, with the order required by sim_study
   res <- mapply(sim_study, ..., SIMPLIFY = FALSE)

### 2023-09-27 scripts/test.R
fixing typo and more fast tests

@@ -54,7 +54,7 @@ mu <- 0.3 # real effect size
 n <- 1e3 # sample size per group, per study
 I2 <- 0.6 # desired I2 value
 
-v <- (n + n)/(n * n) + mu^2/(2 * (n + n - 2)) # typical within study variance
+v <- 1/n + 1/n # typical within study variance
 tau2 <- -((I2*v)/(I2 - 1))
 
 sim <- make_data(k = k, nc = n, nt = n, mu = mu)
@@ -75,7 +75,7 @@ check_sim(fixed = list(theta = theta,
 # Meta-regression with categorical moderator ------------------------------
 
 set.seed(seed)
-k <- 1e4 # the total number of studies
+k <- 1e3 # the total number of studies
 b0 <- 0.1 # intercept, the effect size of the lab-based studies
 b1 <- 0.2 # the difference between the two levels of the moderator
 tau2r <- 0.1 # the residual heterogeneity
@@ -91,10 +91,10 @@ res <- rma(yi, vi, mods = ~exp, method = "REML", data = sim)
 
 check_sim(fixed = list(b0 = b0,
                        b1 = b1,
-                       tau2 = tau2), 
+                       tau2r = tau2r), 
           results = list(b0 = res$b[[1]],
                          b1 = res$b[[2]],
-                         tau2 = res$tau2),
+                         tau2r = res$tau2),
           name = "Meta-regression with categorical predictor")
 
 # Meta-regression with numerical moderator --------------------------------
@@ -124,13 +124,13 @@ check_sim(fixed = list(b0 = b0,
           results = list(b0 = res$b[[1]],
                          b1 = res$b[[2]],
                          tau2r = res$tau2),
-          name = "Meta-regression with numerical predictor"))
+          name = "Meta-regression with numerical predictor")
 
 
 # Meta-regression with numerical predictor, fixing R2 ---------------------
 
 set.seed(seed)
-k <- 5e3 # the number of studies
+k <- 1e3 # the number of studies
 r2 <- 0.2 # the desired r2 value
 tau2 <- 0.3 # the overall tau2
 b0 <- 0.3 # the intercept i.e., average yi when x1 is 0
@@ -157,13 +157,12 @@ check_sim(fixed = list(b0 = b0,
                          b1 = res$b[[2]],
                          tau2r = res$tau2,
                          r2 = res$R2),
-          name = "Meta-regression fixing R2"))
[…]
